### 2018-02-12 Test_OurBiCopSelect.Rmd
cleaning up and adding ancillary features, and a separate testing doc

---
title: "For testing OurBiCopSelect.R"
author: "Daniel Reuman"
date: "February 12, 2018"
output: pdf_document
---

This is just for testing the `OurBiCopSelect.R` function. We test that
it runs and gives reasonable results by simulating data and seeing if 
you recover what kind of copula was used.

<!--Source in the function and a package, changeable constants-->
```{r}
source("OurBiCopSelect.R")
require(copula)
nbsmall<-8
pthresh<-0.2
nblarge<-15 #for bootstrapping, see below
numdat<-100 #amount of artificial data
```

<!--A function needed for caching-->
```{r echo=F}
mtime <- function(files)
{
  lapply(Sys.glob(files),function(x) X =   
          file.info(x)$mtime)
}
```

If real data come from a Clayton copula, independence should
be rejected, Clayton should be the AIC best, it should pass the
goodness of fit test, and the normal copula should fail the
goodness of fit test:
<!--Test for real data from a clayton-->
```{r cache=T, cache.extra=list(nbsmall,nblarge,pthresh,numdat,mtime("OutBiCopSelect.R"))}
families<-c(1,3:6,13,14,16)

#data are really from a clayton
cop<-claytonCopula(5,2)
dat<-rCopula(numdat,cop)
allresC<-OurBiCopSelect(dat[,1],dat[,2],families,numBSsmall=nbsmall,
                        numBSlarge=nblarge,pthresh=0.2)
allresC
```

Same test, but data now come from a Gumbel copula:
<!--Test for real data from a gumbel-->
```{r cache=T, cache.extra=list(nbsmall,nblarge,pthresh,numdat,mtime("OutBiCopSelect.R"))}
#data are really from a gumbel
cop<-gumbelCopula(4,2)
dat<-rCopula(numdat,cop)
allresG<-OurBiCopSelect(dat[,1],dat[,2],families,numBSsmall=nbsmall,
                        numBSlarge=nblarge,pthresh=0.2)
allresG
```

Same test, but data now come from a Frank copula:
<!--Test for real data from a frank-->
```{r cache=T, cache.extra=list(nbsmall,nblarge,pthresh,numdat,mtime("OutBiCopSelect.R"))}
#data are really from a frank
cop<-frankCopula(5,2)
dat<-rCopula(numdat,cop)
allresF<-OurBiCopSelect(dat[,1],dat[,2],families,numBSsmall=nbsmall,
                        numBSlarge=nblarge,pthresh=0.2)
allresF
```