### 2018-02-13 All_analysis.Rmd
separate the plotter function and make the tester doc into the beginning of a sup mat

---
title: "Exploring the importance of a complete statistical description of dependence in ecology and evolution"
author: "Shyamolina Ghosh and Daniel Reuman"
date: "February 8, 2018"
output: 
  html_document:
    fig_caption: yes
    fig_width: 5
    highlight: zenburn
    keep_tex: yes
    number_sections: yes
    df_print: kable # data frame printing option
fontsize: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
seed<-101
families<-c(1,3:6,14)
source("getcopula.R")
source("OurBiCopSelect.R")
```

Updated on `r Sys.Date()`

# Template analysis

**About the dataset :** These data were simulated from a Clayton copula with parameter 8.

```{r load_data_test, echo=F}
#Load your data - should be a data frame with 2 columns. Can contain NAs.
d<-readRDS("Data/givendata.RDS")
```

**Copula plot of the data :**
```{r rankandplot_test, echo=F, cache=T, cache.extra=list(d)}
#convert to a copula and plot
v<-getcopula(d=d,rankon=F,ploton=T) #with real data, rankon should be T
```

**Bivariate copula model selection and goodness of fit pipeline:**
```{r bivmodselect_test, echo=F, cache=T, cache.extra=list(seed,d)}
x<-fit_BIV(d=d,families)
```

* **Results :**
```{r restest, echo=F}
knitr::kable(x[[1]])
knitr::kable(x[[2]])
```

# Results from dataset1 : Prof. T. Loecke
<mark>*You have to edit this section*</mark>

* **About the dataset :** <mark>*Please provide a brief description*</mark>

```{r loaddata1, echo=F}
# here you can directly load the bivariate data to the variable d
# or you can save the bivariate data to d after some data extraction
d<-readRDS("Data/givendata.RDS")
```

* **Copula plot of the data :**
```{r callingcode1, echo=F, cache=T, cache.extra=list(d)}
#Test the function
cop_rank<-(getcopula(d=d,ploton=T))

families<-c(1,3,5)
x<-fit_BIV(d=d,families)
```

* **Results :**
```{r res1, echo=F}
knitr::kable(x[[1]])
knitr::kable(x[[2]])
```

# Results from dataset2 : Prof. ____________