### 2018-02-14 All_analysis.Rmd
added the analysis of Loecke data

---
title: "Exploring the importance of a complete statistical description of dependence in ecology and evolution"
author: "Shyamolina Ghosh and Daniel Reuman"
date: "February 8, 2018"
output: 
  html_document:
    fig_caption: yes
    fig_width: 5
    highlight: tango
    keep_tex: yes
    number_sections: yes
    df_print: kable # data frame printing option
fontsize: 11pt
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
seed<-101
families<-c(1,3:6,14)
source("getcopula.R")
source("OurBiCopSelect.R")
```

Updated on `r Sys.Date()`

# Template analysis

**About the dataset :**  
These data were simulated from a Clayton copula with parameter 8.

**Plot the raw data:**  
```{r load_data_test, echo=F}
#Load your data - should be a data frame with 2 columns. Can contain NAs.
d<-readRDS("Data/givendata.RDS")
plot(d[,1],d[,2],type="p",col="red",xlab=expression(d[1]),ylab=expression(d[2]),cex.lab=1,
     xlim=c(0,1),ylim=c(0,1),asp=1)
rect(0,0,1,1)
```

**Copula plot of the data:**  
```{r rankandplot_test, echo=F}
#convert to a copula and plot
v<-getcopula(d=d,rankon=F,ploton=T) #with real data, rankon should be T
```

**Bivariate copula model selection and goodness of fit pipeline:**  
```{r bivmodselect_test, echo=F, cache=T, cache.extra=list(seed,v,families)}
set.seed(seed)
BivMS_res_test<-OurBiCopSelect(v[,1],v[,2],families=families,gofnormal=FALSE,status=TRUE)
```

```{r}
BivMS_res_test
```

# Results from dataset1 : T. Loecke

**About the dataset :**  
Carbon and nitrogen in soil at many locations across CONUS (Terry to augment this description).

**Plot the raw data:**  
```{r load_data_Loecke, echo=F}
#Load your data - should be a data frame with 2 columns. Can contain NAs.
d<-readRDS("Data/RaCA_soilorganicC_soiltotalN_stocks100cm.RDS")
d<-d[,c("SOCstock100","TSNstock100")]
plot(d[,1],d[,2],type="p",col="red",xlab=expression(d[1]),ylab=expression(d[2]),cex.lab=1)
rect(0,0,1,1)
```

**Copula plot of the data:**  
```{r rankandplot_Loecke, echo=F}
#convert to a copula and plot
v<-getcopula(d=d,rankon=T,ploton=T) #with real data, rankon should be T
```

**Bivariate copula model selection and goodness of fit pipeline:**  
```{r bivmodselect_Loecke, echo=F, cache=T, cache.extra=list(seed,v)}
set.seed(seed)
families_Loecke<-c(1,3:10,13:14,16:20,104,114,204,214)
BivMS_res_Loecke<-OurBiCopSelect(v[,1],v[,2],families=families_Loecke,
                                 gofnormal=FALSE,status=TRUE)
```

```{r}
BivMS_res_Loecke
```

None of the copulas we tried here was a good fit, but some are much 
better than others, and the normal copula is not close to the best, 
so that shows there is non-trivial (i.e., non-normal) copula structure.

It looks like the data might be asymmetric about the y=x line, which
might require a special copula to model. What is the significance,
biologically?

Strong dependence in the extreme right tails is evident. We could
quantify this, and (probably) show it is significantly more than what one would expect from a null hypothesis of a normal copula with the same kendall as the data. What is the significance, biologically?