### 2019-02-11 SupportingInformation.Rmd
after ncsurrog modification

---
title: "Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields"
author: "Shyamolina Ghosh, Lawrence W. Sheppard, Mark T. Holder, Terrance E. Loecke, P. Chris Reid, Daniel C. Reuman"
geometry: "left=1cm,right=1cm,top=2.2cm,bottom=2.2cm"

output: 
  pdf_document:
    number_sections: yes
    keep_tex: yes
    fig_caption: yes
    includes:
      in_header: head_supp.sty

tables: True
link-citations: True
urlcolor : blue
indent : True

csl: ecology-letters.csl
bibliography: REF_ALL.bib
---

\pagenumbering{roman}
\tableofcontents
\listoffigures
\listoftables

\pagebreak
\pagenumbering{arabic}

<!--Basic setup-->
```{r setup, echo=F}
knitr::opts_chunk$set(echo = TRUE, fig.pos = "H")
seed<-101
# families<-c(1,3:10,13,14,16:20)
source("mtime.R") #A function needed for caching
library(captioner)
```

```{r cop_fig_intro,echo=F,results="hide",cache=T, cache.extra=list(seed,mtime("copsurrognd.R"))}
library(copula)
library(VineCopula)
library(mvtnorm)
set.seed(seed)

source("copsurrognd.R")

#x <- read.table("./Data/dataset/x.txt")
#y <- read.table("./Data/dataset/y.txt")
#mat<-cbind(x,y)
#colnames(mat)<-c("x","y")

#***constants
N<-200 #length of time series
n<-2 #number of sampling locations
numdat<-1 #the number of datasets to get from the normal copula
rho<-0.8 #normal-copula parameter for dependence between locations
#this is a covariance, see below call to rmvnorm

#***generate the starting data, which has normal copula structure
#because it comes originally from a multivariate normal, before 
#transformation
sig<-matrix(rho,n,n)
diag(sig)<-1
m<-rmvnorm(N*numdat,mean=rep(0,n),sigma=sig)
m<-array(m,c(N,numdat,n))
m<-aperm(m,c(1,3,2)) # 2nd dim is normally distributed e.g. hist(m[,1,])
gshape<-7.5
gscale<-1 #gamma parameters

cop_to_marg_trans<-function(x){
  return(qgamma(x,shape=gshape,scale=gscale)) #this is where we choose the marginals,
  #should be something skewed
}
# pnorm(m) makes normal to uniform
# then qgamma(...) makes uniform to gamma distributed
m<-cop_to_marg_trans(pnorm(m))

ccop<-claytonCopula(2,2)
csurrog<-array(NA,dim(m))
csurrog[,,1]<-copsurrognd(m[,,1],targetcop=ccop,numsurrog=1)
mat<-csurrog[,,1]

pdf("./Results/rawdata.pdf",height = 4, width = 4)
scale<-2
xhist <- hist(mat[,1], breaks=30, plot=FALSE)
yhist <- hist(mat[,2], breaks=30, plot=FALSE) 
top <- max(c(xhist$counts, yhist$counts)) 
xrange <- range(mat[,1])
yrange <- range(mat[,2])
nf <- layout(matrix(c(2,0,1,3),2,2,byrow=TRUE), c(4,1), c(1,4), TRUE) 
op<-par(mar=c(3,3,1,1),mgp=c(2, 1, 0))  
plot(mat[,1], mat[,2], xlim=xrange, ylim=yrange, xlab="x", ylab="y", cex.lab=1.5)
legend("bottomright","raw data plot",bty="n",cex=1.5,text.col="magenta")
op<-par(mar=c(0,3,1,1)) 
barplot(xhist$counts/scale, axes=FALSE, ylim=c(0, top), space=0) 
op<-par(mar=c(3,0,1,1)) 
barplot(yhist$counts/scale, axes=FALSE, xlim=c(0, top), space=0, horiz=TRUE)
par(op)
dev.off()

pdf("./Results/copdata.pdf",height = 4, width = 4)
copmat<-pobs(mat)
xhist <- hist(copmat[,1], breaks=30, plot=FALSE)
yhist <- hist(copmat[,2], breaks=30, plot=FALSE) 
top <- max(c(xhist$counts, yhist$counts)) 
xrange <- c(0,1)
yrange <- c(0,1)
nf <- layout(matrix(c(2,0,1,3),2,2,byrow=TRUE), c(4,1), c(1,4), TRUE) 
op<-par(mar=c(3,3,1,1),mgp=c(2, 1, 0)) 
plot(copmat[,1], copmat[,2], xlim=xrange, ylim=yrange, xlab="u=rank(x)/(n+1)", ylab="v=rank(y)/(n+1)",cex.lab=1.5)
legend("bottomright","copula plot",bty="n",cex=1.5,text.col="magenta")
op<-par(mar=c(0,3,1,1)) 
barplot(xhist$counts/scale, axes=FALSE, ylim=c(0, top), space=0) 
op<-par(mar=c(3,0,1,1)) 
barplot(yhist$counts/scale, axes=FALSE, xlim=c(0, top), space=0, horiz=TRUE)
par(op)
dev.off()
```

```{r cop_pedagog_fig,echo=F,results="hide"}
set.seed(seed)
source("./FirstPedagogFig.R")
source("./SecondPedagogFig.R")
source("./ThirdPedagogFig.R")
source("./MakeCopulaPedagogFigs.R")
```

<!--The raw data look like this:-->
```{r load_data_Loecke, echo=F, results='hide'}
#fig.cap="Plot of the raw soil C and N data.\\label{fig_raw_cop_soilCN}(A)", fig.height=3, fig.width=3}
d<-readRDS("Data/dataForShymolinaCoord.rds")
d<-d[,c("SOCstock100","TSNstock100","lat","long")]
d<-na.omit(d)
pdf("./Results/SoilCNraw.pdf")
op<-par(mar=c(6, 6, 0.2, 2.1))
plot(d[,1],d[,2],type="p",col="black",xlab="Soil C",ylab="Soil N",cex.lab=2.5,cex.axis=2.5)
par(op)
dev.off()
```

<!--The copula looks like this:-->
```{r rankandplot_Loecke, echo=F, results='hide'}
source("getcopula.R")
#convert to a copula and plot
pdf("./Results/SoilCNcop.pdf")
v_CN<-getcopula(d=d[,c(1:2)],rankon=T,ploton=T) 
dev.off()
```

```{r plot_SoilCNmapUSA, echo=F, results='hide'}
#---------map of soilCN location on USA------------
library(maps)
pdf("./Results/SoilCNmapUSA.pdf",height=4,width=8)
map("usa")
points(x = d$long, y = d$lat, col="red",pch=16,cex=0.5)
dev.off()
```

```{r read_BMR_birds_and_copulaplot, echo=F, results='hide'}
library(readxl)
set.seed(seed)
raw_b<-readxl::read_xls("./Data/BMR/BirdBodyMassesMetabolicRates/BirdMassesAndMetabolicRatesMcNab2008.xls")
raw_b<-raw_b[,1:3]
raw_b<-na.omit(raw_b)
raw_b<-raw_b[,2:3]
raw_b<-as.data.frame(raw_b)

saveRDS(raw_b,"./Data/BMR/BirdBodyMassesMetabolicRates/my_birdsBMR_data.RDS")

b<-log10(raw_b)
colnames(b)<-c("log10M(g)","log10BMR(KJ/h)")
pdf("./Results/BMR_results/birds/raw_birds_BMR.pdf")
op<-par(mar=c(6, 6, 0.2, 2.1))
plot(b[,1],b[,2],type="p",col="black",xlab="log(Mass)",ylab="log(BMR)",cex.lab=2.5,cex.axis=2.5)
par(op)
dev.off()

source("getcopula.R")
pdf("./Results/BMR_results/birds/copula_birds_BMR.pdf")
cop_b<-getcopula(d=b,rankon=T,ploton=T) #convert to a copula and plot
dev.off()
```

```{r read_BMR_mammals_and_copulaplot, echo=F, results='hide'}
set.seed(seed)
raw_m<-readxl::read_xls("./Data/BMR/MammalianBodyMassesMetabolicRates/MammalianMassesAndMetabolicRatesMcNab2008.xls")
raw_m<-raw_m[,1:3]
raw_m<-na.omit(raw_m)
raw_m<-raw_m[,2:3]

saveRDS(raw_m,"./Data/BMR/MammalianBodyMassesMetabolicRates/my_mammalsBMR_data.RDS")

m<-log10(raw_m)
colnames(m)<-c("log10M(g)","log10BMR(KJ/h)")
m<-as.data.frame(m)

pdf("./Results/BMR_results/mammals/raw_mammals_BMR.pdf")
op<-par(mar=c(6, 6, 0.2, 2.1))
plot(m[,1],m[,2],type="p",col="black",xlab="log(Mass)",ylab="log(BMR)",cex.lab=2.5,cex.axis=2.5)
par(op)
dev.off()

source("getcopula.R")
pdf("./Results/BMR_results/mammals/copula_mammals_BMR.pdf")
cop_m<-getcopula(d=m,rankon=T,ploton=T) #convert to a copula and plot
dev.off()
```

```{r read_ceder_creek_data, echo=F, results="hide"}
set.seed(seed)
#-------------------------------------------------
# This function takes 
# Input :
#       db : a dataframe : smaller subset of ple_120 data to calculate avg. biomass for each plot
#       dp : a dataframe : smaller subset of pce_120 data to calculate avg. shannon's H for those plots
#       year : the year for which you want to calculate biomass and shannon's index data
#       rawplot : logical : T for rawdata plot, F for copula plot
#       ploton : logical for optional plotting

# Output :
#       a list of two data-frame:
#                   tab_dbp : a data-frame with 3 column : (plot no , avg.H, avg.Biomass(g/m2))
#                   cop_tab_dbp : a data-frame with 3 column : (plot no, pobs(avg.H), pobs(avg.Biomass(g/m2)))

cop_HB<-function(db,dp,year,rawplot,ploton){
  # db : biomass dataset for a specific year
  # dp : percent cover dataset for a specific year

  plots<-sort(intersect(db$Plot,dp$Plot)) # common plots for both datasets for a specific year
  db<-subset(db,Plot%in%plots)
  dp<-subset(dp,Plot%in%plots)

  tab_dbp<-data.frame(Plot=plots,avg.H=NA,avg.Biomass=NA)
  for(i in 1:length(plots)){
    # calculate biomass for each plot (averaged over all strips for that plot)
    x<-subset(db,Plot==plots[i])
    tab_dbp$avg.Biomass[i]<-mean(x$Biomass)

    # calculate shannon's index H for each plot (averaged over all Quadrates for that plot)
    x<-subset(dp,Plot==plots[i])
    uniq_sp<-sort(as.character(unique(x$Species)))  # These are the species in a particular plot
    Q<-as.character(unique(x$Quadrat))

    tab_each_sp<-data.frame(sp=uniq_sp,avg_cov=0,rel_avg_cov=0)

    for(sp in 1:length(uniq_sp)){
      my_sp<-uniq_sp[sp]
      tab_temp<-data.frame(Q=Q,abs_cov=0)
      ind<-which(x$Species==my_sp)
      li<-length(ind)
      tab_temp$abs_cov[1:li]<-x$Abscover[ind]
      tab_each_sp$avg_cov[[sp]]<-mean(tab_temp$abs_cov) # avg. over 4Quadrates for a sp.
    }

    tab_each_sp$rel_avg_cov<-tab_each_sp$avg_cov/sum(tab_each_sp$avg_cov)
    tab_dbp$avg.H[i]<- -sum((tab_each_sp$rel_avg_cov)*log(tab_each_sp$rel_avg_cov))

  }

  tab_dbp<-na.omit(tab_dbp)

  cop_tab_dbp<-data.frame(Plot=tab_dbp$Plot,avg.H=NA,avg.Biomass=NA)
  cop_tab_dbp$avg.H<-VineCopula::pobs(tab_dbp$avg.H)
  cop_tab_dbp$avg.Biomass<-VineCopula::pobs(tab_dbp$avg.Biomass)

  if(ploton==T){
    if(rawplot==T){
      op<-par(mar=c(6, 6, 0.2, 2.1))
      plot(tab_dbp$avg.H,tab_dbp$avg.Biomass,xlab="avg. H",ylab="avg. Biomass",pch=1,cex.lab=2.5,cex.axis=2)
      mtext(paste0("Year=",year),cex=1)
      par(op)
    }else{
      op<-par(mar=c(6, 6, 0.2, 2.1))
      plot(cop_tab_dbp$avg.H,cop_tab_dbp$avg.Biomass,xlab="u",ylab="v",pch=1,cex.lab=3,cex.axis=2,
         xlim=c(0,1),ylim=c(0,1),asp=1)
      rect(0,0,1,1)
      lines(c(0,1),c(0,1),type='l',col="grey")
      #mtext(paste0("Year=",year),cex=1)
      par(op)
    }
  }

  return(list(tab_dbp=tab_dbp,
         cop_tab_dbp=cop_tab_dbp))
}
#-----------------------------
r_ple<-read.delim("./Data/Ceder_creek_e120/edited_ple120.txt")
names(r_ple)[36]<-"Biomass"

r_pce<-read.delim("./Data/Ceder_creek_e120/edited_pce120.txt")

y1<-unique(r_ple$Year)
y2<-unique(r_pce$Year)
yr<-sort(intersect(y1,y2))

a<-as.data.frame(table(sort(r_pce$Species)))
b<-as.character(a$Var1)
bad_sp<-c(b[51:53],b[156:158])

ind<-which(r_pce$Species%in%bad_sp)
r_pce_clean<-r_pce[-ind,]
#--------------------------------------------------------------------------------------------
pdf("./Results/Ceder_creek_results/raw_HB_all_yr.pdf",height = 6,width = 10)
op<-par(mfrow=c(2,3),mar=c(3,3,3,3), mgp=c(1.5,0.5,0))
for(i in 1:length(yr)){
  db<-subset(r_ple,Year==yr[i])
  dp<-subset(r_pce_clean,Year==yr[i])

  sr<-cop_HB(db=db,dp=dp,year=yr[i],rawplot=T,ploton = T)
}
par(op)
dev.off()
#-----------------------
#raw data plot for year2000
pdf("./Results/Ceder_creek_results/raw_HB_2000.pdf",height = 6,width = 6)
i<-5
db<-subset(r_ple,Year==yr[i])
dp<-subset(r_pce_clean,Year==yr[i])
sr<-cop_HB(db=db,dp=dp,year=yr[i],rawplot=T,ploton = T)
saveRDS(sr,"./Results/Ceder_creek_results/raw_cop_cc_2000.RDS")
dev.off()
#----------------------------------------------------------------------------------------
cop_HB_all_yr<-vector("list",length(yr))
names(cop_HB_all_yr)<-yr

pdf("./Results/Ceder_creek_results/cop_HB_all_yr.pdf",height = 6,width = 10)
op<-par(mfrow=c(2,3),mar=c(3,3,3,3), mgp=c(1.5,0.5,0))
for(i in 1:length(yr)){
  db<-subset(r_ple,Year==yr[i])
  dp<-subset(r_pce_clean,Year==yr[i])

  s<-cop_HB(db=db,dp=dp,year=yr[i],rawplot=F,ploton = T)
  cop_HB_all_yr[[i]]<-s$cop_tab_dbp[,2:3]
}
par(op)
dev.off()
#-------------------------
#cop data plot for year2000
pdf("./Results/Ceder_creek_results/cop_HB_2000.pdf",height = 6,width = 6)
i<-5
db<-subset(r_ple,Year==yr[i])
dp<-subset(r_pce_clean,Year==yr[i])
s<-cop_HB(db=db,dp=dp,year=yr[i],rawplot=F,ploton = T)
cc_cop_2000<-s$cop_tab_dbp[,2:3]
dev.off()
#------------------

saveRDS(cop_HB_all_yr,"./Results/Ceder_creek_results/cop_HB_all_yr.RDS")
#---------------------------------------------------------------------------------------
cop_HB_all_yr<-readRDS("./Results/Ceder_creek_results/cop_HB_all_yr.RDS")
```

```{r map_plot_aphid_plankton,results="hide",echo=F}
source("aphid_data_calling.R")
source("good_loclist.R")
library(maps)

#----------------- Aphid map--------------------------------------------
d0_count<-as.matrix(read.csv("./Data/Aphid_data/APHID_DATA_RAW/aphidtscount141117.csv",header=F))
data_aphid_count<-d_allsp_data(d0_count)

d0_firstflight<-as.matrix(read.csv("./Data/Aphid_data/APHID_DATA_RAW/aphidtsfirstflight141117.csv",header=F))
data_aphid_ff<-d_allsp_data(d0_firstflight)

selective_loc1<-good_loclist(d_allsp=data_aphid_count,sp=10,data_pt_thrs=30)
selective_loc2<-good_loclist(d_allsp=data_aphid_ff,sp=11,data_pt_thrs=30)
all(selective_loc1==selective_loc2)==T

longs<-c(-4.567,0.57,-3.069,-3.312,-2.637,-1.682,-2.763,-0.356,-3.454,0.427,0.939)
lats<-c(55.477,52.26,56.457,55.949,52.125,55.213,53.854,51.807,50.628,51.733,51.185)

locno<-length(longs)
df_loc_info_org<-data.frame(loc_no=c(1:length(lats)),longs=longs,lats=lats)
df_loc_info<-df_loc_info_org[selective_loc1,]
xlim=c(-10,2)
ylim=c(50, 60)

pdf("./Results/aphid_map_plot.pdf")
map("world", fill=TRUE, col="grey84", bg="white",border=0, xlim=xlim, ylim=ylim)
map.axes(cex.axis=2.5)
points(x = df_loc_info_org$longs, y = df_loc_info_org$lats, col="red",pch=16)
points(x = df_loc_info$longs, y = df_loc_info$lats, col="blue", pch=0, cex=1.5)
text(df_loc_info_org$loc_no, x = df_loc_info_org$longs, y = df_loc_info_org$lats, col="red",  pos=1)
dev.off()

#--------------Plankton map----------------------------

source("plankton_north_sea_data_calling.R")
d0_plankton_north_sea<-as.matrix(read.csv("./Data/Plankton_North_Sea_data/planktontimeseries201117_4.csv",header=F))
data_plankton_north_sea<-d_allsp_data_plankton_north_sea(d0_plankton_north_sea)

longs<-read.csv("./Data/Plankton_North_Sea_data/boxcornerlongs201117.csv",header=F)
lats<-read.csv("./Data/Plankton_North_Sea_data/boxcornerlats201117.csv",header=F)
longs_pns<-longs[[1]]
lats_pns<-lats[[1]]

df_loc_info_org<-data.frame(loc_no=c(1:length(lats_pns)),longs=longs_pns,lats=lats_pns)
selective_loc<-good_loclist(d_allsp=data_plankton_north_sea,sp=16,data_pt_thrs=45)
df_loc_info<-df_loc_info_org[selective_loc,]
xlim=c(-12,10)
ylim=c(48, 60)

pdf("./Results/plankton_map_plot.pdf")
map("world", fill=TRUE, col="grey84", bg="white",border=0, xlim=xlim, ylim=ylim)
map.axes(cex.axis=2.5)
abline(h=seq(from=ylim[1],to=ylim[2],by=2), v=seq(from=xlim[1],to=xlim[2],by=2), col="black", lty=2)
points(x = df_loc_info_org$longs+1, y = df_loc_info_org$lats+1, col="red",pch=16)
points(x = df_loc_info$longs+1, y = df_loc_info$lats+1, col="blue", pch=0, cex=1.5)
text(df_loc_info_org$loc_no, x = df_loc_info_org$longs+1, y = df_loc_info_org$lats+1, col="red",  pos=1)
dev.off()
```

```{r make_data_plot_fig,echo=FALSE,results="hide"}
source("./MakeBivarDataPlot.R")
```

# Transformations used in Fig. \ref{MT-fig_cop_pedag1} \label{pedagog1_details}

The data of Fig. \ref{MT-fig_cop_pedag1}A-B were generated to have standard normal marginals.
Fig. \ref{MT-fig_cop_pedag1}C can be obtained from Fig. \ref{MT-fig_cop_pedag1}A via transformation
of both variables by the cdf of the standard normal distribution, and then by the inverse of 
the cdf of the gamma distribution with shape and scale parameters 2. 
The function `qgamma` in R was used.
Fig. \ref{MT-fig_cop_pedag1}D can be obtained from Fig. \ref{MT-fig_cop_pedag1}B in the same way.
Fig. \ref{MT-fig_cop_pedag1}E can be obtained from Fig. \ref{MT-fig_cop_pedag1}A via transformation
of both variables by the cdf of the standard normal distribution.
Fig. \ref{MT-fig_cop_pedag1}F can be obtained from Fig. \ref{MT-fig_cop_pedag1}B in the same way.

# Details of the data \label{Data_details}

<!--***DAN: This is a dumping ground at the moment, but needs to be organized into a section that gives all the necessary details about the data-->

<!--The CPR is a device which is towed across the ocean behind a ship, filtering plankton from seawater with a continuously
unwinding silk ribbon. The ribbon was cut into 4 inch segments, each corresponding to a particular time and
location in the ocean. Organisms were identified and counted (possibly with sub-sampling) microscopically [@Colebrook1979]. All the samples available for a given variable in a given area in a given month of a given year were
averaged, producing a monthly time series for each year. Where data for a given month was unavailable the missing value was replaced with a median value for that month (drawn from all years in which data was available).-->

 [@Sheppard2017] converted monthly values to yearly values by averaging over all months. But here we only allowed maximum 4 months in a year with missing data which were replaced by median value for that month (drawn from all years in which data was available). We used data from 14 locations which had at least 45 years time series data (see Figure \ref{SM-fig_plankton_map}).

 The Great Miami Wetland Mitigation Bank  was at $36^{\circ}46^{'}51^{''}$ N, $84^{\circ}20^{'}26^{''}$ W) in Trotwood, Montgomery County, Ohio. Additional information on the history of the site and soil types is detailed in @Jarecke2016. Briefly, the closed static flux chamber method was used [@Holland1999] to measure $CH_{4}$ fluxes at 28 locations across a hydrologic gradient (upland to seasonal wetland) at daily to weekly interval from September 2015 to September 2016. **See and Smyth et al, pending for
additional details???** on greenhouse gas flux methods for this site. We considered 13 locations that had at least 50 dates data.

```{r tab_aphid_info, echo=F, results='asis',message=F}
library(tinytex)
library(tibble)
aphid_info <- tibble(c1=c("Apple grass aphid","Bird cherry oat aphid","Black bean aphid","Blackberry cereal aphid","Blackcurrant sowthistle aphid",
                       "Corn leaf aphid","Currant lettuce aphid","Damson hop aphid","Grain aphid","Green spruce aphid",
                       "Leaf-curling plum aphid","Mealy cabbage aphid","Mealy plum aphid","Pea aphid","Peach potato aphid",
                       "Potato aphid","Rose grain aphid","Shallot aphid","Sycamore aphid","Willow carrot aphid"),
                     c2=c("Rhopalosiphum insertum", "Rhopalosiphum padi", "Aphis fabae", "Sitobion fragariae", "Hyperomyzus lactucae",
                       "Rhopalosiphum maidis", "Nasonovia ribisnigri","Phorodon humuli","Sitobion avenae","Elatobium abietinum",
                       "Brachycaudus helichrysi","Brevicoryne brassicae","Hyalopterus pruni","Acyrthosiphon pisum","Myzus persicae",
                       "Macrosiphum euphorbiae","Metopolophium dirhodum","Myzus ascalonicus","Drepanosiphum platanoidis","Cavariella aegopodii")
)
knitr::kable(aphid_info, "latex", booktabs = T, linesep = "",
      caption = "Species names of 20 aphids for which data were available \\label{tab_aphid_info}",
      col.names = c("Common name", "Latin binomial"))
```

```{r tab_plankton_info, echo=F, results='asis',message=F}
library(tibble)
plankton_info <- tibble(c('Calanus I-IV','Para-Pseudocalanus spp','Temora longicornis','Acartia spp',
                          'Centropages typicus',
                          'Oithona spp','Echinoderm larvae','Calanus finmarchicus',
                          'Calanus helgolandicus','Metridia lucens','Decapoda larvae',
                   'Euphausiacea','Thalassiosira spp','Rhizosolenia styliformis',
                   'Ceratium fusus','Ceratium furca','Ceratium tripos','Ceratium macroceros',
                   'Rhizosolenia alata alata','Pseudocalanus elongatus Adult',
                   'Nitzschia delicatissima','Nitzschia seriata')
)
knitr::kable(plankton_info, "latex", booktabs = T, linesep = "",
      caption = "Species names of 22 plankton taxa for which data were available \\label{tab_plankton_info}",
      col.names = c("Plankton taxon name"))
```

```{r read_methane_data, echo=F, results="markup"}
set.seed(seed)
methane_raw<-readRDS("./Data/Methane_data/methane_data_for_Shyamolina.rds")
#class(methane_raw) #it's a data.frame for multivariable data
#knitr::kable(head(methane_raw,3),format='pandoc',
#             caption = "First few rows of raw methane data.\\label{tab_rawdata_methane}",
#             booktabs=T) # display first few rows
```

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm,height=16 cm]{./Results/SoilCNmapUSA.pdf} 
\caption[Sampling locations for C and N data]{Locations of central pedons sampled 
for soil organic carbon and total soil nitrogen (Mg C or N per hectare of soil 
surface) at 5907 sites across the coterminous United States. \label{fig_SoilCNmapUSA}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm,height=16 cm]{./Results/aphid_map_plot.pdf} 
\caption[Sampling locations for aphid data]{Locations of the 
Rothamsted Insect Survey 
suction traps from which data were available are indicated by the red circles, 
out of which only blue boxes had at least 30 years of data and were 
used for our study. \label{fig_aphid_map}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm,height=16 cm]{./Results/plankton_map_plot.pdf} 
\caption[Sampling regions for CPR plankton data]{CPR data were available from 
26 locations ($2^{\circ}$ $\times$ $2^{\circ}$ regions on the map) around UK 
seas (red circles), out of which only blue boxes had at least 45 years of data and
were used for our study. \label{fig_plankton_map}}
\end{center}
\end{figure}

<!--Now the copula figures-->

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Normal.pdf}
\caption[Example N copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example N copulas. 
S is Spearman correlation; p is the value of the parameter for the N 
family (it is a 
one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, respectively.\label{figped_N}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Clayton.pdf}
\caption[Example C copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example C copulas. S is Spearman correlation; p is the value of the parameter for the C family (it is a 
one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively.\label{figped_C}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SurvClayton.pdf}
\caption[Example SC copulas]{Log-transformed pdfs (A-E) and samples (F-J) 
from example SC 
copulas. S is Spearman correlation; p is the value of the parameter for the SC 
family (it is a one-parameter family); and LT and UT are the measures of lower- and 
upper-tail dependence, respectively.\label{figped_SC}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Gumbel.pdf}
\caption[Example G copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example G copulas. S is 
Spearman correlation; p is the value of the parameter for the G family (it is a 
one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively.\label{figped_G}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SurvGumbel.pdf}
\caption[Example SG copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example SG 
copulas. S is Spearman correlation; p is the value of the parameter for the SG 
family (it is a one-parameter family); and LT and UT are the measures of lower- and 
upper-tail dependence, respectively.\label{figped_SG}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Joe.pdf}
\caption[Example J copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example J copulas. K is Kendall correlation; p is the value of the parameter for the J family (it is a 
one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively.\label{figped_J}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SurvJoe.pdf}
\caption[Example SJ copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example SJ 
copulas. K is Spearman correlation; p is the value of the parameter for the SJ 
family (it is a one-parameter family); and LT and UT are the measures of lower- and 
upper-tail dependence, respectively.\label{figped_SJ}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Frank.pdf}
\caption[Example F copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example F copulas. S is Kendall correlation; p is the value of the parameter for the F family (it is a 
one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively.\label{figped_F}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB1.pdf}
\caption[Example BB1 copulas]{Log-transformed pdfs for example BB1 copulas. K is Kendall correlation; p1 and p2
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_BB1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB1.pdf}
\caption[Example SBB1 copulas]{Log-transformed pdfs for example SBB1 copulas. K is Kendall correlation; p1 ad p2 
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_SBB1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB6.pdf}
\caption[Example BB6 copulas]{Log-transformed pdfs for example BB6 copulas. K is Kendall correlation; p1 and p2
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_BB6}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB6.pdf}
\caption[Example SBB6 copulas]{Log-transformed pdfs for example SBB6 copulas. K is Kendall correlation; p1 ad p2 
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_SBB6}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB7.pdf}
\caption[Example BB7 copulas]{Log-transformed pdfs for example BB7 copulas. K is Kendall correlation; p1 and p2
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_BB7}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB7.pdf}
\caption[Example SBB7 copulas]{Log-transformed pdfs for example SBB7 copulas. K is Kendall correlation; p1 ad p2 
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_SBB7}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB8.pdf}
\caption[Example BB8 copulas]{Log-transformed pdfs for example BB8 copulas. K is Kendall correlation; p1 and p2
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_BB8}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB8.pdf}
\caption[Example SBB8 copulas]{Log-transformed pdfs for example SBB8 copulas. K is Kendall correlation; p1 ad p2 
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_SBB8}}
\end{center}
\end{figure}

# Nonparametric statistics \label{nonparam_stats}

We here define $\Ps_{l_b,u_b}$, using the bounds
$u+v=2l_b$ and $u+v=2u_b$ defined in the main text, as follows. For a distance $h$, we define
$S(h)$ to be the number of points $(u_i,v_i)$ within the bounds 
and a distance less than $h$ from the line $v=u$, divided by the 
total number of points within the bounds. This function is
defined for $h$ going from $0$ to a distance $h_{\text{max}}$ which
is half the longer of the two segments obtained by intersecting the
bounds with the unit square. It is easy to see $S(0)=0$ and 
$S(h_{\text{max}})=1$. We define $S_i(h)$, also for $h$ going from $0$
to $h_{\text{max}}$, to be the area within the bounds and the unit square
and within distance $h$ of the line $v=u$, divided by the area
within the bounds and the unit square. This is the expected value of
$S(h)$ for independent data. We then define
\begin{equation}\label{eq.P}
   P_{l_b,u_b} = \int_0^{h_{\text{max}}} ( S(h) - S_i(h) ) dh.
\end{equation}
\noindent As for $\cor_{l_b,u_b}$, larger values of $\Ps_{l_b,u_b}$ 
indicate stronger positive association
between $u$ and $v$ in the region given by the bounds.

\begin{figure}[!h]
\begin{center}
\includegraphics[width=6 cm,height=6 cm]{./stat_image.jpg} \hspace{0.6 cm}
\caption[For definitions of nonparametric statistics]{Schematic diagram supporting definitions of nonparametric tail dependence statistics. The blue and red lines are, respectively, $u+v=2l_b$ and $u+v=2u_b$.\label{fig_stat_image}}
\end{center}
\end{figure}

# Testing the nonparametric statistics \label{Test_npa_stat}

N and F copula families 
are known to have symmetric tail dependence,
i.e., strengths of dependence in the lower and upper tails are the same, for all copulas
in these families [@nelsen2006_copula]. C copulas are known to have stronger lower- than 
upper-tail dependence, and SC copulas, being rotations of C copulas by 
$180^\circ$, have stronger upper- than lower tail dependence [@nelsen2006_copula].
Using the `iRho` function from the `copula` package, for each of the four families N, 
F, C, and SC, we obtained parameters for which the expected value of 
Spearman's $\rho$ was, separately, $0,0.1,0.2,...,0.9$. For each copula family and for each 
of the selected parameter values we generated  $n$ points, 100 times, and thereby 
computed 100 values of the 
statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$. This exercise 
was repeated, separately, 
for $n=1000$ and for $n=35$, these cases being chosen, respectively, to reflect 
asymptotic behavior of the 
statistics and their behavior on the small data sets which are common in some fields
of ecology. The mean and standard error of the three difference statistics were computed for 
each copula family, parameter value, and value of $n$, and results were plotted against 
Spearman's rho (Figs \ref{fig_stat_testing35} and \ref{fig_stat_testing1000}), 
the expectation being that the difference statistics would be positive for 
C copulas, negative for SC copulas (except for the case of $\rho=0$, a 
boundary case of these families corresponding to independence), and not significantly different 
from $0$ for F and N 
copulas.
We used a $t$-test with a significance threshold of $0.05$ to judge significance here.
These expectations were fulfilled (Figs \ref{fig_stat_testing35} and 
\ref{fig_stat_testing1000}).

<!--The message=F and warning=F flags are to suppress warnings that come when you pass a parameter
of 0 to the Frank or Clayton copulas. These are technicallly boundary cases for these families.
They correspond to the independent copula. The packages we use provide the indepdent copula, but
issue a warning. Since anyway we want the independent copula, we can suppress the warning. We ran this
without the warning and it is OK to supress it.-->
<!-- The lines you are talking about appeared as message not as warning(actualy no warning appered for this chunk), so we have to set message=F here-->
```{r stat_testing_save_and_plot_result, message=F, results="hide", echo=F, cache=T, cache.extra=list(seed,mtime("TestStats_multivar_dataset.R"),mtime("CopulaFunctions.R"),mtime("CopulaFunctions_flexible.R"))}
source("TestStats_multivar_dataset.R")
set.seed(seed)

resultsloc<-"./Results/stat_results/stat_testing/"
spearvals<-seq(from=0,to=0.9,by=0.1)
callfn<-coplist_with_params(spearvals=spearvals)
coplist<-callfn$coplist
ncores<-3
res_pt35_rankF<-mclapply(X=coplist,FUN=worker,numpts=35,numsims=100,rank=F,mc.cores=ncores)
res_pt35_rankT<-mclapply(X=coplist,FUN=worker,numpts=35,numsims=100,rank=T,mc.cores=ncores)
res_pt1000_rankF<-mclapply(X=coplist,FUN=worker,numpts=1000,numsims=100,rank=F,mc.cores=ncores)
res_pt1000_rankT<-mclapply(X=coplist,FUN=worker,numpts=1000,numsims=100,rank=T,mc.cores=ncores)
save(res_pt35_rankF,res_pt35_rankT,res_pt1000_rankF,res_pt1000_rankT,file=paste(resultsloc,"StatTestNumericResults.RData",sep=''))

b<-callfn$paramlist
plotter_stat_testing(res_pt35_rankF,"res_pt35_rankF",b,resultsloc,spearvals)
plotter_stat_testing(res_pt35_rankT,"res_pt35_rankT",b,resultsloc,spearvals)
plotter_stat_testing(res_pt1000_rankF,"res_pt1000_rankF",b,resultsloc,spearvals)
plotter_stat_testing(res_pt1000_rankT,"res_pt1000_rankT",b,resultsloc,spearvals)
```

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/stat_results/stat_testing/res_pt35_rankT.pdf}
\caption[Testing nonparametric statistics with 35 data points]{Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the C and SC copulas, and the known symmetry of tail dependence of the N and F copulas. See Appendix  \ref{Test_npa_stat} for details. $n=35$. \label{fig_stat_testing35}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/stat_results/stat_testing/res_pt1000_rankT.pdf}
\caption[Testing nonparametric statistics with 1000 data points]{Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the C and SC copulas, and the known symmetry of tail dependence of the N and F copulas. See Appendix  \ref{Test_npa_stat} for details. $n=1000$.\label{fig_stat_testing1000}}
\end{center}
\end{figure}

<!--This r-chunk generates figure necessary for maintext Paper-->
```{r fig_stat_testing_for_maintext, message=F, results="hide", echo=F, cache=T, cache.extra=list(seed,mtime("TestStats_multivar_dataset.R"),mtime("CopulaFunctions.R"),mtime("CopulaFunctions_flexible.R"),mtime("StatTestNumericResults.RData"))}
set.seed(seed) 
source("fig_Paper_stat_testing.R")
#load("./Results/stat_results/stat_testing/StatTestNumericResults.RData")
resultsloc<-"./Results/stat_results/stat_testing/"

copname<-"C"
reslist<-res_pt35_rankT

statname<-"CorlmCoru"
call_fig_Paper_stat_testing(resultsloc=resultsloc,statname=statname,copname=copname,reslist=reslist, numsims=100)
statname<-"PlmPu"
call_fig_Paper_stat_testing(resultsloc=resultsloc,statname=statname,copname=copname,reslist=reslist, numsims=100)
statname<-"D2umD2l"
call_fig_Paper_stat_testing(resultsloc=resultsloc,statname=statname,copname=copname,reslist=reslist, numsims=100)
```

# Surrogate testing with bivariate datasets for nonparametric statistics \label{surrog_test}

For bivariate data sets, surrogates were produced as follows.

- Given data $(x_i,y_i)$ for $i=1,\ldots,n$ and a 
one-parameter target copula family $\mathcal{B}(\theta)$ (here $\theta$ denotes the parameter), let $\tau$ be the
Kendall correlation of the $x_i$ with the $y_i$ and find the parameter value $\theta_\tau$ for which 
$\mathcal{B}(\theta_\tau)$ has Kendall correlation $\tau$ (this is possible for the one-parameter copula families of this 
study using the `iTau` function of the `VineCopula` package). 
- Generate data $(a_i,b_i)$, $i=1,\ldots,n$ from 
$\mathcal{B}(\theta_\tau)$. 
- Permute the ordered set $(x_1,\ldots,x_n)$ such that the permutation
$(x_{\sigma_x(1)},\ldots,x_{\sigma_x(n)})$ has $\rank(x_{\sigma_x(i)})$ equal to $\rank(a_i)$ for all $i$, where 
$\rank(x_{\sigma_x(i)})$ is the rank of $x_{\sigma_x(i)}$ in the set $(x_{\sigma_x(1)},\ldots,x_{\sigma_x(n)})$, and
$\rank(a_i)$ is the rank of $a_i$ in the set $(a_1,\ldots,a_n)$ (here $\sigma_x$ is a permutation of the 
indices $1,\ldots,n$). 
- Likewise permute $(y_1,\ldots,y_n)$
such that the permutation
$(y_{\sigma_y(1)},\ldots,y_{\sigma_y(n)})$ has $\rank(y_{\sigma_y(i)})$ equal to $\rank(b_i)$ for all $i$ ($\sigma_y$ is 
another permutation of $1,\ldots,n$).
- The surrogate data set is $(x_{\sigma_x(i)},y_{\sigma_y(i)})$ for $i=1,\ldots,n$.

\noindent Our code for this algorithm is `copsurrog2d` in the 
`BIVAN` repository (https://github.com/sghosh89/BIVAN). 
Because $(x_{\sigma_x(1)},\ldots,x_{\sigma_x(n)})$ 
is a permutation of $(x_1,\ldots,x_n)$ and $(y_{\sigma_y(1)},\ldots,y_{\sigma_y(n)})$ is a 
permutation 
of $(y_1,\ldots,y_n)$, the marginal distributions of the surrogate data are exactly the same as 
those of the original data. 
Because ranks of the surrogate data are matched to ranks of the $a_i$ and $b_i$, the
Kendall correlation of the surrogate data set will be the same as that of the $a_i$ and $b_i$, and therefore 
will be close to $\tau$ (differences arising only through sampling variation). If instead of $\theta_\tau$
a parameter value $\theta_\rho$ is used such that the Spearman correlation of $\mathcal{B}(\theta_\rho)$ is the same
as that of the $x_i$ with the $y_i$, then Spearman correlations of surrogates will very nearly equal (except for 
sampling variation) the Spearman correlation of the empirical data.

# Spatial resampling with multivariate datasets for nonparametric statistics \label{spatial_resamp}

The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
widely used spline correlogram method; the same scheme was also used by @walteretal2017 
as part of a method for 
performing model selection among matrix regression models. For each pair of 
locations $i \neq j$, the statistic
of interest, here denoted $s_{ij}$ ($\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, 
$\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, 
$\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$) was computed for those locations. The 
values $s_{ij}$ were then arranged in
a matrix, $S$, with diagonal entries NA. The mean of the non-NA entries of $S$ was computed. For 
$\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, or $\Dsq_{l_b,u_b}$ this mean characterizes the 
average strength of 
dependence, across all pairs of distinct locations, for data in the portion of the distributions 
given by the bounds $l_b$ and $u_b$. For $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$, 
the mean characterizes average
asymmetry of tail dependence across all pairs of locations. 
Confidence intervals for the value of $\mean(S)$ were then computed as quantiles of a 
distribution of bootstrapped values. If $\mean(S)$ was significantly different from 0, as
revealed by the confidence intervals, for the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, 
or $\Dsq_u-\Dsq_l$, then tail dependence was significantly asymmetric, in contrast to a 
N-copula null hypothesis, and answering question Q1 from the Introduction.
Bootstrapping was performed by resampling the locations of measurement, with replacement. 
This corresponds to 
resampling, with replacement, the rows and columns of $S$, and is described in detail elsewhere 
[@bjornstadfalck2001 ; @walteretal2017].

<!--***DAN: text all good up to here, except as marked (the data section). Also
all figs and tables relevant to the stuff this far have been edited and look pretty
good. This is as of 2018 01 10, 1255pm. So start editing from this point.-->

<!--This chunk is for model selection results of soilCN dataset-->
```{r bivmodselect_Loecke, echo=F, cache=T, cache.extra=list(seed,v_CN,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("OurBiCopSelect.R")
set.seed(seed)
families_Loecke<-c(1,3:10,13:14,16:20) 
BivMS_res_Loecke<-OurBiCopSelect(u1=v_CN[,1],u2=v_CN[,2],families=families_Loecke,level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,gofnormal=F,status=F)
saveRDS(BivMS_res_Loecke,file="./Results/fitting_results/BivMS_soilCN.RDS")
```

<!--Display model selection results in a table-->
```{r tab_soilCNfit, echo=F, results="markup"}
#knitr::opts_knit$set(kable.force.latex = TRUE)
h<-BivMS_res_Loecke$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
colnames(h)<-c("copname","AICw","BICw","LT","UT")
knitr::kable(h, #longtable = T, 
             format = "latex",linesep = "",
      caption.short = "Model fitting results for soil C and N dataset",
      caption = "Results of fitting our 16 copula families to the soil C and N dataset. Copula name abbreviations are given in the main text. AICw = AIC weight; BICw = BIC weight; LT = the lower-tail dependence statistic for the indicated copula family, with fitted parameters; UT = the same for upper-tail dependence.\\label{tab_soilCNfit}",
      booktabs = TRUE)
```

<!--This chunk is for npa testing results of soilCN dataset-->
```{r stat_soilCN_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,v_CN,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
source("bivfunctionplot.R")
set.seed(seed) 
stat_soilCN_ga<-bivfunctionplot(v=v_CN,resloc="./Results/stat_results/stat_soilCN/",nametag="soilCN",numbin=10)
```

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/stat_results/stat_soilCN/soilCN_bivfunctionplot.pdf}
\caption[Nonparametric tests against an N-copula null, soil C and N data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for soil C and N data. As described in the main text, the values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data (red crosses) were compared to distributions of their values on 1000 Kendall- or Spearman-preserving N surrogates of the data (blue and green x's show 0.025 and 0.975 quantiles), separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$. Whenever the red cross was outside the range given by the x's, text at the top of panels indicates the number of surrogate values the real-data value was greater than or less than. For instance, a value $>N$ (respectively, $<N$) means the value of the statistic on real data was greater than (respectively, less than) its value on $N$ surrogates. When the statistic was greater than 975 or less than 975 surrogate values, it indicates significance ($95\%$ confidence level). When $\cor$ or $\Ps$ values (respectively, $\Dsq$ values) were greater than surrogates, it means dependence in that part of the distributions was stronger than (respectively, weaker than) expected from a normal-copula null hypothesis. \label{fig_soilCN_nonparam}}
\end{center}
\end{figure}

<!--Now make and display a table, this one about asymmetry of tail dependence-->
```{r make_table_stat_fn, echo=F}

fractxt<-function(fracval,numsurrog){
  ptxt<-''
  if (fracval>.5*numsurrog){                        # ?DCR why 0.5 factor?
    ptxt<-paste(">",fracval,sep='')
  } 
  if (fracval<=.5*numsurrog){
    ptxt<-paste("<",numsurrog-fracval,sep='')
  }
  return(ptxt)
}

make_table_stat_fn<-function(data){

 #atdres<-data.frame(Statistic=c("$\\cor_{0,0.1}-\\cor_{0.9,1}$",
 #                               "$\\Ps_{0,0.1}-\\Ps_{0.9,1}$",
 #                               "$\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$"),
 atdres<-data.frame(Statistic=c("cor$_{FS}$-cor$_{LS}$",
                               "P$_{FS}$-P$_{LS}$",
                                "D$_{LS}^2$-D$_{FS}^2$"),
                   Kendall=rep("",3),
                   Spearman=rep("",3),stringsAsFactors = F)

 atdres[1,2]<-fractxt(data$corlmcoru_frac_K,numsurrog = data$numsurrog_success_K)
 atdres[1,3]<-fractxt(data$corlmcoru_frac_S,numsurrog = data$numsurrog_success_S)
 atdres[2,2]<-fractxt(data$PlmPu_frac_K,numsurrog = data$numsurrog_success_K)
 atdres[2,3]<-fractxt(data$PlmPu_frac_S,numsurrog = data$numsurrog_success_S)
 atdres[3,2]<-fractxt(data$D2umD2l_frac_K,numsurrog = data$numsurrog_success_K)
 atdres[3,3]<-fractxt(data$D2umD2l_frac_S,numsurrog = data$numsurrog_success_S)

 return(atdres)

}
```

```{r tab_soilCN_asym,echo=F,results="markup"}
#knitr::opts_knit$set(kable.force.latex = TRUE)
tab_soilCN_asym<-make_table_stat_fn(data=stat_soilCN_ga)
#tab_soilCN_asym[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
knitr::kable(tab_soilCN_asym, booktabs = T, escape=F,
             format = "latex",linesep = "",
      caption.short = "Nonparametric asymmetry tests against an N-copula null, soil C and N",
      caption = "Soil C and N dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis. The values of the listed statistics (cor$_{FS}$-cor$_{LS}$ = $\\cor_{0,0.1}-\\cor_{0.9,1}$, P$_{FS}$-P$_{LS}$ = $\\Ps_{0,0.1}-\\Ps_{0.9,1}$, D$_{LS}^2$-D$_{FS}^2$ = $\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$; here FS stands for `first slice' and LS for `last slice') for real data were compared to their values for 1000 Kendall- or Spearman-preserving N surrogates, in seperate comparisons. A table entry $<X$ indicates the value of the given statistic on the data was less than its value on $X$ of the surrogates, so entries of the form $<X$ for $X$ equal to $975$ or above indicate that upper-tail dependence was significantly stronger than lower-tail dependence.\\label{tab_soilCN_asym}")
```

```{r bivmodselect_birds_BMR, echo=F, cache=T, cache.extra=list(seed,cop_b,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("OurBiCopSelect.R")
set.seed(seed)
BivMS_birds<-OurBiCopSelect(u1=cop_b[,1],u2=cop_b[,2],families=c(1,3:10,13:14,16:20),
                              level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,
                              gofnormal=F,status=F)
saveRDS(BivMS_birds,file="./Results/BMR_results/birds/BivMS_birds.RDS")
```

<!--Display model selection results for BMR:birds data in a table-->
```{r tab_bmr_birds_fit, echo=F, results='markup'}
h<-BivMS_birds$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
colnames(h)<-c("copname","AICw","BICw","LT","UT")
knitr::kable(h,
             format='latex',linesep = "",
             caption.short = "Model fitting results for the bird body mass and BMR dataset",
             caption = "Results of fitting our 16 copula families to the bird body mass and BMR dataset. Analogous to Fig. \\ref{tab_soilCNfit}, see caption of that figure for details. \\label{tab_bmr_birds_fit}",
             booktabs=T)
```

```{r stat_bmr_birds_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,cop_b,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
source("bivfunctionplot.R")
set.seed(seed)
stat_birds_ga<-bivfunctionplot(v=cop_b,resloc="./Results/BMR_results/birds/",nametag="birds_numbin_10",numbin = 10)
```

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/BMR_results/birds/birds_numbin_10_bivfunctionplot.pdf}
\caption[Nonparametric tests against an N-copula null, bird body mass and BMR data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for bird body mass and BMR data. Same format as Fig. \ref{fig_soilCN_nonparam}, see caption of that figure. \label{fig_bmr_birds_nonparam}}
\end{center}
\end{figure}

```{r tab_bmr_birds_asym, echo=F, results="markup"}
tab_bmr_birds_asym<-make_table_stat_fn(data=stat_birds_ga)
#tab_bmr_birds_asym[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
knitr::kable(tab_bmr_birds_asym,
             format="latex",linesep = "",escape=F,
             caption.short = "Nonparametric asymmetry tests against an N-copula null, bird body mass and BMR",
             caption = "Bird body mass and BMR dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis. Same format as Table \\ref{tab_soilCN_asym}, see caption of that figure.\\label{tab_bmr_birds_asym}",
             booktabs=T)
```

```{r bivmodselect_mammals_BMR, echo=F, cache=T, cache.extra=list(seed,cop_m,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("OurBiCopSelect.R")
set.seed(seed)
BivMS_mammals<-OurBiCopSelect(u1=cop_m[,1],u2=cop_m[,2],families=c(1,3:10,13:14,16:20),
                              level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,
                              gofnormal=F,status=F)
saveRDS(BivMS_mammals,file="./Results/BMR_results/mammals/BivMS_mammals.RDS")
```

<!--Display model selection results for BMR:mammals data in a table-->
```{r tab_bmr_mammals_fit, echo=F, results='markup'}
h<-BivMS_mammals$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
colnames(h)<-c("copname","AICw","BICw","LT","UT")
knitr::kable(h,
             format='latex',linesep = "",
             caption.short = "Model fitting results for the mammal body mass and BMR dataset",
             caption = "Results of fitting our 16 copula families to the mammal body mass and BMR dataset. Analogous to Fig. \\ref{tab_soilCNfit}, see caption of that figure for details. \\label{tab_bmr_mammals_fit}",
             booktabs=T)
```

```{r stat_bmr_mammals_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,cop_m,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
source("bivfunctionplot.R")
set.seed(seed)
stat_mammals_ga<-bivfunctionplot(v=cop_m,resloc="./Results/BMR_results/mammals/",nametag="mammals_numbin_10",numbin = 10)
```

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/BMR_results/mammals/mammals_numbin_10_bivfunctionplot.pdf}
\caption[Nonparametric tests against an N-copula null, mammal body mass and BMR data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for mammal body mass and BMR data. Same format as Fig. \ref{fig_soilCN_nonparam}, see caption of that figure. \label{fig_bmr_mammals_nonparam}}
\end{center}
\end{figure}

```{r tab_bmr_mammals_asym, echo=F, results="markup"}
tab_bmr_mammals_asym<-make_table_stat_fn(data=stat_mammals_ga)
#tab_bmr_mammals_asym[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
knitr::kable(tab_bmr_mammals_asym,
             format="latex",linesep = "",escape=F,
             caption.short = "Nonparametric asymmetry tests against an N-copula null, mammal body mass and BMR",
             caption = "Mammal body mass and BMR dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis. Same format as Table \\ref{tab_soilCN_asym}, see caption of that figure. \\label{tab_bmr_mammals_asym}",
             booktabs=T)
```

```{r bivmodselect_cc, echo=F, cache=T, cache.extra=list(seed,cop_HB_all_yr,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("OurBiCopSelect.R")
set.seed(seed)

BivMS_res_cc<-vector("list",length(cop_HB_all_yr))
names(BivMS_res_cc)<-names(cop_HB_all_yr)

for(i in 1:length(BivMS_res_cc)){
  BivMS_res_cc[[i]]<-OurBiCopSelect(u1=cop_HB_all_yr[[i]]$avg.H,u2=cop_HB_all_yr[[i]]$avg.Biomass,
                                    families=c(1,3:10,13:14,16:20),
                                    level=0.05,AICBIC="AIC",
                                    numBSsmall=100,pthresh=0.2,numBSlarge=1000,gofnormal=F,status=F)
}
saveRDS(BivMS_res_cc,file="./Results/Ceder_creek_results/BivMS_res_cc.RDS")
BivMS_res_cc<-readRDS("./Results/Ceder_creek_results/BivMS_res_cc.RDS")
```

<!--Display model selection results for Ceder-creek data for year 2000 in a table-->
```{r tab_cc2000fit, echo=F, results='markup'}
h<-BivMS_res_cc$`2000`$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
colnames(h)<-c("copname","AICw","BICw","LT","UT")
knitr::kable(h,
             format='latex',linesep = "",
             caption.short = "Model fitting results for the Cedar Creek dataset",
             caption = "Results of fitting our 16 copula families to the Cedar Creek dataset. Analogous to Fig. \\ref{tab_soilCNfit}, see caption of that figure for details. \\label{tab_cc2000fit}",
             booktabs=T)
```

```{r stat_cc_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,cop_HB_all_yr,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
source("bivfunctionplot.R")
set.seed(seed)

stat_cc_ga<-vector("list",length(cop_HB_all_yr))
names(stat_cc_ga)<-names(cop_HB_all_yr)

for(i in 1:length(stat_cc_ga)){
  temp<-bivfunctionplot(v=cop_HB_all_yr[[i]],
                             resloc="./Results/Ceder_creek_results/",
                             nametag=paste0(names(cop_HB_all_yr)[i],"cc_numbin_10",sep=""),
                             numbin = 10)
  stat_cc_ga[[i]]<-temp
}

saveRDS(stat_cc_ga,"./Results/Ceder_creek_results/stat_cc_ga.RDS")
```

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/Ceder_creek_results/2000cc_numbin_10_bivfunctionplot.pdf}
\caption[Nonparametric tests against an N-copula null, Cedar Creek data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for Cedar Creek data. Same format as Fig. \ref{fig_soilCN_nonparam}, see caption of that figure.\label{fig_cc2000_nonparam}}
\end{center}
\end{figure}

```{r tab_cc_asym_2000, echo=F, results="markup"}
tab_cc_asym_2000<-make_table_stat_fn(data=stat_cc_ga$`2000`)
#tab_cc_asym_2000[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
knitr::kable(tab_cc_asym_2000,
             format="latex",linesep = "",escape=F,
             caption.short = "Nonparametric asymmetry tests against an N-copula null, Cedar Creek data",
             caption = "Cedar Creek dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis. Same format as Table \\ref{tab_soilCN_asym}, see caption of that figure.\\label{tab_cc_asym_2000}",
             booktabs=T)
```

<!--Load the raw data-->
```{r read_aphid_data, echo=F}
source("aphid_data_calling.R")

# Aphid raw data for count
d0_count<-as.matrix(read.csv("./Data/Aphid_data/APHID_DATA_RAW/aphidtscount141117.csv",header=F))
data_aphid_count<-d_allsp_data(d0_count)

# Aphid raw data for first flight
d0_firstflight<-as.matrix(read.csv("./Data/Aphid_data/APHID_DATA_RAW/aphidtsfirstflight141117.csv",header=F))
data_aphid_ff<-d_allsp_data(d0_firstflight)

# Aphid raw data for flight duration   
#d0_flightduration<-as.matrix(read.csv("./Data/Aphid_data/APHID_DATA_RAW/aphidtsflightduration141117.csv",header=F))

# Time series analysis aphids winter avg temp (here taken as -ve)
temp_aphid_raw<-as.matrix(read.csv("./Data/Aphid_data/APHID_DATA_RAW/aphidmonthlytempdata.csv",header = F))

temp_aphid<-vector("list",dim(temp_aphid_raw)[1])
names(temp_aphid)<-paste("loc",c(1:dim(temp_aphid_raw)[1]),sep="")

for(il in 1:length(temp_aphid)){

  s<-matrix(temp_aphid_raw[il,],36,12,byrow = T)
  rownames(s)<-c(1975:2010)
  colnames(s)<-c("jan","feb","mar","apr","may","jun","jul","aug","sep","oct","nov","dec")

  temp_aphid[[il]]<-s
}

winter_avg_temp<-matrix(NA,11,35)
rownames(winter_avg_temp)<-paste("loc",c(1:dim(temp_aphid_raw)[1]),sep="")
colnames(winter_avg_temp)<-c(1976:2010)

for(il in 1:length(temp_aphid)){
  temploc<-temp_aphid[[il]]
  for(i in 2:dim(temploc)[1]){
    t<-c(temploc[i-1,12],temploc[i,1:3]) # previous yr dec to present year march averaging temp.
    winter_avg_temp[il,i-1]<-mean(t,na.rm = T)
  }
}

aphid_winter_neg_temp<-vector("list",11) #initializing
names(aphid_winter_neg_temp)<-rownames(winter_avg_temp)

for(i in 1:11){
  aphid_winter_neg_temp[[i]]<-data.frame(Year=c(1976:2010),Dat=-winter_avg_temp[i,])
}

temp_aphid<-list(aphid_winter_neg_temp)
names(temp_aphid)<-"aphid_winter_neg_temp"
saveRDS(temp_aphid,"./Results/temp_aphid_winter_neg_temp_avg_data.RDS")
```

```{r modelsel_and_npa_aphid_count_allsp, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"),mtime("NonParamStat.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}

set.seed(seed)
library(parallel)
source("FittingCopula_selective_loc.R")
RES_all_sp_aphid_count<-mclapply(X=c(1:20),FUN=RES_single_sp,d_allsp=data_aphid_count,
                                 families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=30,
                                 good_loc_given=F,good_loc=NA,mc.cores=3)
saveRDS(RES_all_sp_aphid_count,"./Results/fitting_results/RES_all_sp_aphid_count.RDS")

source("good_loclist.R")
source("NonParamStat.R")
resloc<-"./Results/stat_results/stat_aphid_count/"
longs<-c(-4.567,0.57,-3.069,-3.312,-2.637,-1.682,-2.763,-0.356,-3.454,0.427,0.939)
lats<-c(55.477,52.26,56.457,55.949,52.125,55.213,53.854,51.807,50.628,51.733,51.185)

stat_aphid_count_allsp<-vector("list",20)

for(sp in 1:20){
  nm<-paste("sp_",sp,sep="")
  resloc2<-paste(resloc,nm,sep="")
   if (!dir.exists(resloc2)){
    dir.create(resloc2)
  }
  good_loc<-good_loclist(d_allsp=data_aphid_count,sp=sp,data_pt_thrs=30)

  temp<-multcall(d_allsp=data_aphid_count,sp=sp,lats=lats,longs=longs,
                 pfname=paste(resloc2,"/","Sp_",sp,sep=''),good_loc=good_loc)
  stat_aphid_count_allsp[[sp]]<-temp
}
saveRDS(stat_aphid_count_allsp,paste(resloc,file="stat_aphid_count_all_sp.RDS",sep=''))

summary_aphid_count_allsp<-as.data.frame(matrix(NA,nrow=20,ncol=8))
colnames(summary_aphid_count_allsp)<-c("species","#LTmUT>0","#LTmUT<0","val_LTmUT>0","val_LTmUT<0",
                                     "CorlmCoru_mean_CI","PlmPu_mean_CI","D2umD2l_mean_CI")

for(sp in c(1:20)){
  #cat("sp=",sp,"\n")
  summary_aphid_count_allsp$species[sp]<-sp
  LTmUT<-(RES_all_sp_aphid_count[[sp]]$LTdep_AICw - RES_all_sp_aphid_count[[sp]]$UTdep_AICw) 
  summary_aphid_count_allsp$`#LTmUT>0`[sp]<-sum(LTmUT>0,na.rm=T)
  summary_aphid_count_allsp$`#LTmUT<0`[sp]<-sum(LTmUT<0,na.rm=T)
  ind_LTmUT_pos<-which(LTmUT>0,arr.ind = T)
  summary_aphid_count_allsp$`val_LTmUT>0`[sp]<-sum(LTmUT[ind_LTmUT_pos])
  ind_LTmUT_neg<-which(LTmUT<0,arr.ind = T)
  summary_aphid_count_allsp$`val_LTmUT<0`[sp]<-sum(LTmUT[ind_LTmUT_neg])

  summary_aphid_count_allsp$CorlmCoru_mean_CI[sp]<-list(
    c(unname(round(stat_aphid_count_allsp[[sp]]$numericdf[9,c(2:4)],3))))
  summary_aphid_count_allsp$PlmPu_mean_CI[sp]<-list(
    c(unname(round(stat_aphid_count_allsp[[sp]]$numericdf[10,c(2:4)],3))))
  summary_aphid_count_allsp$D2umD2l_mean_CI[sp]<-list(
    c(unname(round(stat_aphid_count_allsp[[sp]]$numericdf[11,c(2:4)],3))))
}
saveRDS(summary_aphid_count_allsp,"./Results/summary_aphid_count_allsp.RDS")
```

<!--Do the model selection for aphid count data and then showing results in Tables-->
```{r RES_aphid_count, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}

source("FittingCopula_selective_loc.R")  # top most hierarchy source file
set.seed(seed)
#cat(paste("start-time: ",Sys.time(),"\n")) 
RES_aphid_count<-RES_single_sp(d_allsp=data_aphid_count,sp=10,families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=30,
                               good_loc_given=F,good_loc=NA)
#cat(paste("stop-time: ",Sys.time(),"\n"))
resloc<-"./Results/fitting_results/"
saveRDS(RES_aphid_count,paste(resloc,file="AphidCopulaFit_selecloc_count_species_10.RDS",sep=""))
```

```{r tab_count_bestcop, echo=F, results="markup"}
#library(dplyr)
#knitr::opts_knit$set(kable.force.latex = TRUE)
knitr::kable(RES_aphid_count$info_ord_copname[,,1], #longtable = T,
             format='latex',linesep = "",
             caption.short = "Best-fitting copulas, green spruce aphid abundance data",
             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for green spruce aphid abundance data. Copulas 
considered were the normal, Clayton, Gumbel, Frank, Joe, BB1, BB6, BB7, BB8, survival Clayton, survival Gumbel, 
survival Joe, survival BB1, survival BB6, survival BB7, and survival BB8 copulas, with abbreviations in the methods for Q1. NA occurs along the diagonal of the matrix because the diagonal represents a sampling location compared to itself.\\label{tab_count_bestcop}",
             booktabs=TRUE
             )#%>% # only with 'latex'/'html' format
 # kableExtra::kable_styling(latex_options = c("striped","HOLD_position"))
```

```{r tab_count_AICw, echo=F, results='markup'}
knitr::kable(RES_aphid_count$info_ord_AICw[,,1], 
             format='latex',linesep = "",digits=3,
             caption.short = "AIC weights, best-fitting copulas, green spruce aphid abundance data",
             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for green spruce aphid abundance data.\\label{tab_count_AICw}",
             booktabs=TRUE
             )
```

```{r tab_count_AICw_normal, echo=F, results='markup'}
source("normal_cop_AICw.R")
knitr::kable(normal_cop_AICw_matrix(r=RES_aphid_count),
             format='latex',linesep = "",digits=3,
             caption.short = "AIC weights, N copulas, green spruce aphid abundance data",
             caption = "$\\text{AIC}_{\\text{w}}$ for the N copula family for each location pair for green spruce aphid abundance data, for comparison with Table \\ref{tab_count_AICw}. \\label{tab_count_AICw_normal}",
             booktabs=TRUE
             )
```

```{r tab_count_p_CvM, echo=F, results='markup'}
knitr::kable(RES_aphid_count$gfc_p_CvM,
             format='latex',linesep = "",digits=3,
             caption.short="Goodness of fit of best-fitting copulas, green spruce aphid abundance data",
             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for green spruce abundance data.\\label{tab_count_p_CvM}",
             booktabs=TRUE
             )
```

```{r tab_count_p_KS, echo=F, results='markup'}
knitr::kable(RES_aphid_count$gfc_p_KS,
             format='latex',linesep = "",digits=3,
             caption.short="Goodness of fit of best-fitting copulas, green spruce aphid abundance data",
             caption = "$P$-value results of a Kolmogorov-Smirnov-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for green spruce abundance data.\\label{tab_count_p_KS}",
             booktabs=TRUE
             )
```

<!--\newpage-->
```{r tab_count_LTdep_AICw, echo=F, results='markup'}
knitr::kable(RES_aphid_count$LTdep_AICw,
             format='latex',linesep = "",digits=3,
             caption.short="Lower-tail dependence, green spruce aphid abundance data",
             caption = "Model-averaged lower-tail dependence statistics for each location pair for green spruce aphid abundance data.\\label{tab_count_LTdep_AICw}",
             booktabs=TRUE
             )
```

<!--\newpage-->
```{r tab_count_UTdep_AICw, echo=F, results='markup'}
knitr::kable(RES_aphid_count$UTdep_AICw,
             format='latex',linesep = "",digits=3,
             caption.short="Upper-tail dependence, green spruce aphid abundance data",
             caption = "Model-averaged upper-tail dependence statistics for each location pair for green spruce aphid abundance data.\\label{tab_count_UTdep_AICw}",
             booktabs=TRUE
             )
```

<!--\newpage-->
```{r tab_count_LTmUT, echo=F, results='markup'}
knitr::kable(RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw,
             digits = 3,linesep = "",
             format='latex',
             caption.short="Lower- minus upper-tail dependence, green spruce aphid abundance data",
             caption = "Model-averaged lower-tail dependence statistics minus model-averaged upper-tail dependence statistics for each location pair for green spruce aphid abundance data.\\label{tab_count_LTmUT}",
             booktabs=TRUE
             )
```

<!--Do the main computations for nonparametric analysis of aphid count data-->
```{r stat_aphid_count, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("good_loclist.R"),mtime("NonParamStat.R"),mtime("vivj_matrix.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}

source("good_loclist.R")
source("NonParamStat.R")
set.seed(seed) 
resloc<-"./Results/stat_results/stat_aphid_count/"

sp<-10
good_loc <- good_loclist(d_allsp=data_aphid_count,sp=sp,data_pt_thrs=30)
longs<-c(-4.567,0.57,-3.069,-3.312,-2.637,-1.682,-2.763,-0.356,-3.454,0.427,0.939)
lats<-c(55.477,52.26,56.457,55.949,52.125,55.213,53.854,51.807,50.628,51.733,51.185)

stat_aphid_count<-multcall(d_allsp=data_aphid_count,sp=sp,lats=lats,longs=longs,pfname=paste(resloc,"Sp_",sp,sep=''),good_loc=good_loc)
saveRDS(stat_aphid_count,paste(resloc,file="stat_aphid_count_sp_10.RDS",sep='')) 
```

<!--Show the statistical results in a table-->
```{r tab_count_resamp,echo=F,results="markup"}
s<-stat_aphid_count$numericdf[,c(3,2,4)]
sc<-as.data.frame(c("Spearman","Kendall","cor$_{l}$","cor$_{u}$","P$_{l}$","P$_{u}$","D$_{l}^2$","D$_{u}^2$",
                    "cor$_{l}$ - cor$_{u}$","P$_{l}$ - P$_{u}$","D$_{u}^2$ - D$_{l}^2$"))
s<-cbind(sc,s)
knitr::kable(s, digits = 3,
             col.names=c("","2.5$^{th}$ quantile", "Mean", "97.5$^{th}$ quantile"),
             format = "latex",linesep = "", escape = F,align="c",
             caption.short="Results for nonparametric statistics, green spruce aphid abundance data",
             caption = "Average values of statistics across all locations pairs and confidence intervals based on spatial resampling for green spruce aphid abundance data.\\label{tab_count_resamp}",
             booktabs=TRUE)
```

\begin{figure}[!h]
\textbf{ \hspace{3.4 cm} (A) \hspace{3.2 cm} (B) \hspace{3.2 cm} (C) \hspace{3.4 cm} (D)} \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_count/Sp_10_Spearman_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_count/Sp_10_Kendall_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_count/Sp_10_Corl_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_count/Sp_10_Coru_vs_D.pdf}\\
\end{center}
\textbf{ \hspace{3.4 cm} (E) \hspace{3.2 cm} (F) \hspace{3.2 cm} (G) \hspace{3.4 cm} (H)} \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_count/Sp_10_Pl_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_count/Sp_10_Pu_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_count/Sp_10_D2u_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_count/Sp_10_D2l_vs_D.pdf}\\
\end{center}
\textbf{ \hspace{5.4 cm} (I) \hspace{3.4 cm} (J) \hspace{3.2 cm} (K) } \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_count/Sp_10_Corl-Coru_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_count/Sp_10_Pl-Pu_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_count/Sp_10_D2u-D2l_vs_D.pdf}
\caption[Nonparametric tail statistics versus distance between locations for green spruce aphid abundance data]{Nonparametric tail statistics versus great-circle distance, $D$, between sampling locations for green spruce aphid abundance data. \label{fig_aphid_count}}
\end{center}
\end{figure}

```{r modelsel_and_npa_aphid_ff_allsp, echo=F, cache=T, cache.extra=list(seed,data_aphid_ff,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"),mtime("NonParamStat.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}

set.seed(seed)
library(parallel)
source("FittingCopula_selective_loc.R")
RES_all_sp_aphid_ff<-mclapply(X=c(1:20),FUN=RES_single_sp,d_allsp=data_aphid_ff,
                                 families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=30,
                                 good_loc_given=F,good_loc=NA,mc.cores=3)
saveRDS(RES_all_sp_aphid_ff,"./Results/fitting_results/RES_all_sp_aphid_ff.RDS")

source("good_loclist.R")
source("NonParamStat.R")
resloc<-"./Results/stat_results/stat_aphid_ff/"
longs<-c(-4.567,0.57,-3.069,-3.312,-2.637,-1.682,-2.763,-0.356,-3.454,0.427,0.939)
lats<-c(55.477,52.26,56.457,55.949,52.125,55.213,53.854,51.807,50.628,51.733,51.185)

stat_aphid_ff_allsp<-vector("list",20)

for(sp in 1:20){
  nm<-paste("sp_",sp,sep="")
  resloc2<-paste(resloc,nm,sep="")
   if (!dir.exists(resloc2)){
    dir.create(resloc2)
  }
  good_loc<-good_loclist(d_allsp=data_aphid_ff,sp=sp,data_pt_thrs=30)

  temp<-multcall(d_allsp=data_aphid_ff,sp=sp,lats=lats,longs=longs,
                 pfname=paste(resloc2,"/","Sp_",sp,sep=''),good_loc=good_loc)
  stat_aphid_ff_allsp[[sp]]<-temp
}
saveRDS(stat_aphid_ff_allsp,paste(resloc,file="stat_aphid_ff_all_sp.RDS",sep=''))

summary_aphid_ff_allsp<-as.data.frame(matrix(NA,nrow=20,ncol=8))
colnames(summary_aphid_ff_allsp)<-c("species","#LTmUT>0","#LTmUT<0","val_LTmUT>0","val_LTmUT<0",
                                     "CorlmCoru_mean_CI","PlmPu_mean_CI","D2umD2l_mean_CI")

for(sp in c(1:20)){
  #cat("sp=",sp,"\n")
  summary_aphid_ff_allsp$species[sp]<-sp
  LTmUT<-(RES_all_sp_aphid_ff[[sp]]$LTdep_AICw - RES_all_sp_aphid_ff[[sp]]$UTdep_AICw) 
  summary_aphid_ff_allsp$`#LTmUT>0`[sp]<-sum(LTmUT>0,na.rm=T)
  summary_aphid_ff_allsp$`#LTmUT<0`[sp]<-sum(LTmUT<0,na.rm=T)
  ind_LTmUT_pos<-which(LTmUT>0,arr.ind = T)
  summary_aphid_ff_allsp$`val_LTmUT>0`[sp]<-sum(LTmUT[ind_LTmUT_pos])
  ind_LTmUT_neg<-which(LTmUT<0,arr.ind = T)
  summary_aphid_ff_allsp$`val_LTmUT<0`[sp]<-sum(LTmUT[ind_LTmUT_neg])

  summary_aphid_ff_allsp$CorlmCoru_mean_CI[sp]<-list(
    c(unname(round(stat_aphid_ff_allsp[[sp]]$numericdf[9,c(2:4)],3))))
  summary_aphid_ff_allsp$PlmPu_mean_CI[sp]<-list(
    c(unname(round(stat_aphid_ff_allsp[[sp]]$numericdf[10,c(2:4)],3))))
  summary_aphid_ff_allsp$D2umD2l_mean_CI[sp]<-list(
    c(unname(round(stat_aphid_ff_allsp[[sp]]$numericdf[11,c(2:4)],3))))
}
saveRDS(summary_aphid_ff_allsp,"./Results/summary_aphid_ff_allsp.RDS")
```

<!--Do the model selection for aphid ff data and then showing results in Tables-->
```{r RES_aphid_ff, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}

source("FittingCopula_selective_loc.R")  # top most hierarchy source file
set.seed(seed)
#cat(paste("start-time: ",Sys.time(),"\n"))
RES_aphid_ff<-RES_single_sp(d_allsp=data_aphid_ff,sp=11,families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=30,
                            good_loc_given=F,good_loc=NA)
#cat(paste("stop-time: ",Sys.time(),"\n")) 
resloc<-"./Results/fitting_results/"
saveRDS(RES_aphid_ff,paste(resloc,file="AphidCopulaFit_selecloc_firstflight_species_11.RDS",sep=""))
```

```{r tab_ff_bestcop, echo=F, results="markup"}
knitr::kable(RES_aphid_ff$info_ord_copname[,,1], #longtable = T,
             format='latex',linesep = "",
             caption.short = "Best-fitting copulas, leaf-curling plum aphid first flight data",
             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for leaf-curling plum aphid first flight data. Copulas considered were as in Table \\ref{tab_count_bestcop}, with abbreviations in the methods for Q1. NA occurs along the diagonal of the matrix because the diagonal represents a sampling location compared to itself.  \\label{tab_ff_bestcop}",
             booktabs=TRUE
             )
```

```{r tab_ff_AICw, echo=F, results='markup'}
knitr::kable(RES_aphid_ff$info_ord_AICw[,,1], 
             format='latex',linesep = "",digits=3,
             caption.short="AIC weights, best-fitting copulas, leaf-curling plum aphid first flight data",
             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for leaf-curling plum aphid first flight data.\\label{tab_ff_AICw}",
             booktabs=TRUE
             )
```

<!--\newpage-->
```{r tab_ff_AICw_normal, echo=F, results='markup'}
source("normal_cop_AICw.R")
knitr::kable(normal_cop_AICw_matrix(r=RES_aphid_ff),
             format='latex',linesep = "",digits=3,
             caption.short="AIC weights, N copulas, leaf-curling plum aphid first flight data",
             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for leaf-curlung plum aphid first flight data, for comparison with Table \\ref{tab_ff_AICw}.\\label{tab_ff_AICw_normal}",
             booktabs=TRUE
             )
```

```{r tab_ff_p_CvM, echo=F, results='markup'}
knitr::kable(RES_aphid_ff$gfc_p_CvM,
             format='latex',linesep = "",digits=3,
             caption.short="Goodness of fit of best-fitting copulas, leaf-curling plum aphid first flight data",
             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for leaf-curling plum aphid first flight data.\\label{tab_ff_p_CvM}",
             booktabs=TRUE
             )
```

```{r tab_ff_p_KS, echo=F, results='markup'}
knitr::kable(RES_aphid_ff$gfc_p_KS,
             format='latex',linesep = "",digits=3,
             caption.short="Goodness of fit of best-fitting copulas, leaf-curling plum aphid first flight data",
             caption = "$P$-value results of a Kolmogorov-Smirnov-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for leaf-curling plum aphid first flight data.\\label{tab_ff_p_KS}",
             booktabs=TRUE
             )
```

```{r tab_ff_LTdep_AICw, echo=F, results='markup'}
knitr::kable(RES_aphid_ff$LTdep_AICw,
             format='latex',linesep = "",digits=3,
             caption.short="Lower-tail dependence, leaf-curling plum aphid first flight data",
             caption = "Model-averaged lower-tail dependence statistics for each location pair for leaf-curling plum aphid first flight data.\\label{tab_ff_LTdep_AICw}",
             booktabs=TRUE
             )
```

```{r tab_ff_UTdep_AICw, echo=F, results='markup'}
knitr::kable(RES_aphid_ff$UTdep_AICw,
             format='latex',linesep = "",digits=3,
             caption.short="Upper-tail dependence, leaf-curling plum aphid first flight data",
             caption = "Model-averaged upper-tail dependence statistics for each location pair for leaf-curling plum aphid first flight data.\\label{tab_ff_UTdep_AICw}",
             booktabs=TRUE
             )
```

```{r tab_ff_LTmUT, echo=F, results='markup'}
knitr::kable(RES_aphid_ff$LTdep_AICw-RES_aphid_ff$UTdep_AICw,
             digits = 3,
             format='latex',linesep = "",
             caption.short="Lower- minus upper-tail dependence, leaf-curling plum aphid first flight data",
             caption = "Model-averaged lower-tail dependence statistics minus model-averaged upper-tail dependence statistics for each location pair for leaf-curling plum aphid first flight data.\\label{tab_ff_LTmUT}",
             booktabs=TRUE
             )
```

<!--Do the main computations for nonparametric analysis of aphid ff data-->
```{r stat_aphid_ff, echo=F, cache=T,  cache.extra=list(seed,data_aphid_ff,mtime("good_loclist.R"),mtime("NonParamStat.R"),mtime("vivj_matrix.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}

source("good_loclist.R")
source("NonParamStat.R")
set.seed(seed)
resloc<-"./Results/stat_results/stat_aphid_ff/"

sp<-11
good_loc<-good_loclist(d_allsp=data_aphid_ff,sp=sp,data_pt_thrs=30)
longs<-c(-4.567,0.57,-3.069,-3.312,-2.637,-1.682,-2.763,-0.356,-3.454,0.427,0.939)
lats<-c(55.477,52.26,56.457,55.949,52.125,55.213,53.854,51.807,50.628,51.733,51.185)

stat_aphid_ff<-multcall(d_allsp=data_aphid_ff,sp=sp,lats=lats,longs=longs,pfname=paste(resloc,"Sp_",sp,sep=''),good_loc=good_loc) 
saveRDS(stat_aphid_ff,paste(resloc,file="stat_aphid_ff_sp_11.RDS",sep='')) 
```

<!--display the main statistical results in a table-->
```{r tab_ff_resamp,echo=F,results="markup"}
s<-stat_aphid_ff$numericdf[,c(3,2,4)]
sc<-as.data.frame(c("Spearman","Kendall","cor$_{l}$","cor$_{u}$","P$_{l}$","P$_{u}$","D$_{l}^2$","D$_{u}^2$",
                    "cor$_{l}$ - cor$_{u}$","P$_{l}$ - P$_{u}$","D$_{u}^2$ - D$_{l}^2$"))
s<-cbind(sc,s)
knitr::kable(s, digits = 3,
             col.names=c("","2.5$^{th}$ quantile", "Mean", "97.5$^{th}$ quantile"),
             format = "latex",linesep = "", escape = F,align="c",
             caption.short="Results for nonparametric statistics, leaf-curling plum aphid first flight data",
             caption = "Average values of statistics across all location pairs and confidence intervals based on spatial resampling for leaf-curling plum aphid first flight data.\\label{tab_ff_resamp}",
             booktabs=TRUE)
```

\begin{figure}[!h]
\textbf{ \hspace{3.4 cm} (A) \hspace{3.2 cm} (B) \hspace{3.2 cm} (C) \hspace{3.4 cm} (D)} \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_ff/Sp_11_Spearman_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_ff/Sp_11_Kendall_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_ff/Sp_11_Corl_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_ff/Sp_11_Coru_vs_D.pdf}\\
\end{center}
\textbf{ \hspace{3.4 cm} (E) \hspace{3.2 cm} (F) \hspace{3.2 cm} (G) \hspace{3.4 cm} (H)} \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_ff/Sp_11_Pl_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_ff/Sp_11_Pu_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_ff/Sp_11_D2u_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_ff/Sp_11_D2l_vs_D.pdf}\\
\end{center}
\textbf{ \hspace{5.4 cm} (I) \hspace{3.4 cm} (J) \hspace{3.2 cm} (K) } \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_ff/Sp_11_Corl-Coru_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_ff/Sp_11_Pl-Pu_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_ff/Sp_11_D2u-D2l_vs_D.pdf}
\caption[Nonparametric tail statistics versus distance between locations for leaf-curling plum aphid first flight data]{Nonparametric tail statistics versus great-circle distance, $D$, between sampling locations for leaf-curling plum aphid first flight data. \label{fig_aphid_ff}}
\end{center}
\end{figure}

<!--Load the data-->
```{r read_plankton_north_sea_data, echo=F}
source("plankton_north_sea_data_calling.R")
d0_plankton_north_sea<-as.matrix(read.csv("./Data/Plankton_North_Sea_data/planktontimeseries201117_4.csv",header=F))
data_plankton_north_sea<-d_allsp_data_plankton_north_sea(d0_plankton_north_sea)
```

```{r modelsel_and_npa_pns_allsp, echo=F, cache=T, cache.extra=list(seed,data_plankton_north_sea,longs_pns,lats_pns,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"),mtime("NonParamStat.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}

set.seed(seed)
library(parallel)
source("FittingCopula_selective_loc.R")
RES_all_sp_pns<-mclapply(X=c(1:22),FUN=RES_single_sp,d_allsp=data_plankton_north_sea,
                              families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=45,
                              good_loc_given=F,good_loc=NA,mc.cores=3)
saveRDS(RES_all_sp_pns,"./Results/fitting_results/RES_all_sp_pns.RDS")

source("good_loclist.R")
source("NonParamStat.R")
resloc<-"./Results/stat_results/stat_plankton_north_sea/"

stat_plankton_north_sea_allsp<-vector("list",22)

for(sp in 1:22){
  nm<-paste("sp_",sp,sep="")
  resloc2<-paste(resloc,nm,sep="")
   if (!dir.exists(resloc2)){
    dir.create(resloc2)
  }
  good_loc<-good_loclist(d_allsp=data_plankton_north_sea,sp=sp,data_pt_thrs=45)

  temp<-multcall(d_allsp=data_plankton_north_sea,sp=sp,lats=lats_pns,longs=longs_pns,
                                      pfname=paste(resloc2,"/","Sp_",sp,sep=''),good_loc=good_loc)
  stat_plankton_north_sea_allsp[[sp]]<-temp
}
saveRDS(stat_plankton_north_sea_allsp,paste(resloc,file="stat_plankton_north_sea_allsp.RDS",sep=''))

summary_pns_allsp<-as.data.frame(matrix(NA,nrow=22,ncol=8))
colnames(summary_pns_allsp)<-c("species","#LTmUT>0","#LTmUT<0","val_LTmUT>0","val_LTmUT<0",
                                     "CorlmCoru_mean_CI","PlmPu_mean_CI","D2umD2l_mean_CI")

for(sp in c(1:22)){
  #cat("sp=",sp,"\n")
  summary_pns_allsp$species[sp]<-sp
  LTmUT<-(RES_all_sp_pns[[sp]]$LTdep_AICw - RES_all_sp_pns[[sp]]$UTdep_AICw) 
  summary_pns_allsp$`#LTmUT>0`[sp]<-sum(LTmUT>0,na.rm=T)
  summary_pns_allsp$`#LTmUT<0`[sp]<-sum(LTmUT<0,na.rm=T)
  ind_LTmUT_pos<-which(LTmUT>0,arr.ind = T)
  summary_pns_allsp$`val_LTmUT>0`[sp]<-sum(LTmUT[ind_LTmUT_pos])
  ind_LTmUT_neg<-which(LTmUT<0,arr.ind = T)
  summary_pns_allsp$`val_LTmUT<0`[sp]<-sum(LTmUT[ind_LTmUT_neg])

  summary_pns_allsp$CorlmCoru_mean_CI[sp]<-list(
    c(unname(round(stat_plankton_north_sea_allsp[[sp]]$numericdf[9,c(2:4)],3))))
  summary_pns_allsp$PlmPu_mean_CI[sp]<-list(
    c(unname(round(stat_plankton_north_sea_allsp[[sp]]$numericdf[10,c(2:4)],3))))
  summary_pns_allsp$D2umD2l_mean_CI[sp]<-list(
    c(unname(round(stat_plankton_north_sea_allsp[[sp]]$numericdf[11,c(2:4)],3))))
}
saveRDS(summary_pns_allsp,"./Results/summary_pns_allsp.RDS")
```

<!--Do the model selection for plankton data and then showing results in Tables-->
```{r RES_plankton_north_sea, echo=F, cache=T, results="hide", cache.extra=list(seed,data_plankton_north_sea,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}

source("FittingCopula_selective_loc.R")  # top most hierarchy source file

set.seed(seed)
#cat(paste("start-time: ",Sys.time(),"\n")) 
RES_plankton_north_sea<-RES_single_sp(d_allsp=data_plankton_north_sea,sp=16,families=c(1,3:10,13,14,16:20),level=0.05,
                                      data_pt_thrs=45,good_loc_given=F,good_loc=NA)
#cat(paste("stop-time: ",Sys.time(),"\n"))    
resloc<-"./Results/fitting_results/"
saveRDS(RES_plankton_north_sea,paste(resloc,file="Plankton_North_Sea_CopulaFit_selecloc_species_16.RDS",sep=""))
```

```{r tab_pns_bestcop, echo=F, results="markup"}
knitr::kable(RES_plankton_north_sea$info_ord_copname[,,1], #longtable=T,
             format='latex', linesep = "",
             caption.short = "Best-fitting copulas, \\textit{Ceratium furca} abundance data",
             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for \\textit{Ceratium furca} abundance data. See methods for Q1 for copula abbreviations. Indep.=Independent. NA occurs along the diagonal because diagonal entries represent a sampling location compared to itself.\\label{tab_pns_bestcop}",
             booktabs=TRUE
             )
```

```{r tab_pns_AICw,echo=F,results="markup"}
h<-RES_plankton_north_sea$info_ord_AICw[,,1]
h[is.infinite(h)]<-NA
knitr::kable(h,
             format="latex",
             digits=3,linesep = "",
             caption.short="AIC weights, best-fitting copulas, \\textit{Ceratium furca} abundance data",
             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for \\textit{Ceratium furca} abundance data.\\label{tab_pns_AICw}",
             booktabs=TRUE)
```

```{r tab_pns_AICw_normal,echo=F,results="markup"}
source("normal_cop_AICw.R")
h<-normal_cop_AICw_matrix(r=RES_plankton_north_sea)
h[is.infinite(h)]<-NA
knitr::kable(h,
             format="latex",linesep = "",
             digits=3,
             caption.short="AIC weights, N copulas, \\textit{Ceratium furca} abundance data",
             caption = "$\\text{AIC}_{\\text{w}}$ for the N copula family for each location pair for \\textit{Ceratium furca} abundance data, for comparison with Table \\ref{tab_pns_AICw}.\\label{tab_pns_AICw_normal}",
             booktabs=TRUE)
```

```{r tab_pns_p_CvM,echo=F,results="markup"}
h<-RES_plankton_north_sea$gfc_p_CvM
h[is.infinite(h)]<-NA
knitr::kable(h,
             format="latex",linesep = "",
             digits=3,
             caption.short="Goodness of fit of best-fitting copulas, \\textit{Ceratium furca} abundance data",
             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for \\textit{Ceratium furca} abundance data.\\label{tab_pns_p_CvM}",
             booktabs=TRUE)
```

```{r tab_pns_p_KS,echo=F,results="markup"}
h<-RES_plankton_north_sea$gfc_p_KS
h[is.infinite(h)]<-NA
knitr::kable(h,
             format="latex",linesep = "",
             digits=3,
             caption.short="Goodness of fit of best-fitting copulas, \\textit{Ceratium furca} abundance data",
             caption = "$P$-value results of a Kolmogorov-Smirnov-based test of
the goodness of fit of the lowest-AIC copula family for each pair of
locations for \\textit{Ceratium furca} abundance data.\\label{tab_pns_p_KS}",
             booktabs=TRUE)
```

```{r tab_pns_LTdep_AICw,echo=F,results="markup"}
h<-RES_plankton_north_sea$LTdep_AICw
h[is.infinite(h)]<-NA
knitr::kable(h,
             format="latex",linesep = "",
             digits=3,
             caption.short="Lower-tail dependence, \\textit{Ceratium furca} abundance data",
             caption = "Model-averaged lower-tail dependence statistics for each location pair for \\textit{Ceratium furca} abundance data.\\label{tab_pns_LTdep_AICw}",
             booktabs=TRUE)
```

```{r tab_pns_UTdep_AICw,echo=F,results="markup"}
h<-RES_plankton_north_sea$UTdep_AICw
h[is.infinite(h)]<-NA
knitr::kable(h,
             format="latex",linesep = "",
             digits=3,
             caption.short="Upper-tail dependence, \\textit{Ceratium furca} abundance data",
             caption = "Model-averaged upper-tail dependence statistics for each location pair for \\textit{Ceratium furca} abundance data.\\label{tab_pns_UTdep_AICw}",
             booktabs=TRUE)
```

```{r tab_pns_LTmUT,echo=F,results="markup"}
h<-RES_plankton_north_sea$LTdep_AICw-RES_plankton_north_sea$UTdep_AICw
h[is.nan(h)]<-NA
knitr::kable(h,
             format="latex",linesep = "",
             digits=3,
             caption.short="Lower- minus upper-tail dependence, \\textit{Ceratium furca} abundance data",
             caption = "Model-averaged lower-tail dependence statistics minus model-averaged upper-tail dependence statistics for each location pair for \\textit{Ceratium furca} abundance data.\\label{tab_pns_LTmUT}",
             booktabs=TRUE)
```

<!--Do the main computations for nonparametric analysis of plankton data-->
```{r stat_plankton_north_sea, echo=F, cache=T, cache=T,  cache.extra=list(seed,data_plankton_north_sea,longs_pns,lats_pns,mtime("good_loclist.R"),mtime("NonParamStat.R"),mtime("vivj_matrix.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}

source("good_loclist.R")
source("NonParamStat.R")
set.seed(seed)
resloc<-"./Results/stat_results/stat_plankton_north_sea/"

sp<-16
good_loc<-good_loclist(d_allsp=data_plankton_north_sea,sp=sp,data_pt_thrs=45)

stat_plankton_north_sea<-multcall(d_allsp=data_plankton_north_sea,sp=sp,lats=lats_pns,longs=longs_pns,
                                  pfname=paste(resloc,"Sp_",sp,sep=''),good_loc=good_loc) 
saveRDS(stat_plankton_north_sea,paste(resloc,file="stat_plankton_north_sea_sp_16.RDS",sep=''))
```

<!--display the main statistical results in a table-->
```{r tab_pns_resamp,echo=F,results="markup"}
s<-stat_plankton_north_sea$numericdf[,c(3,2,4)]
sc<-as.data.frame(c("Spearman","Kendall","cor$_{l}$","cor$_{u}$","P$_{l}$","P$_{u}$","D$_{l}^2$","D$_{u}^2$",
                    "cor$_{l}$ - cor$_{u}$","P$_{l}$ - P$_{u}$","D$_{u}^2$ - D$_{l}^2$"))
s<-cbind(sc,s)
knitr::kable(s, digits = 3,
             col.names=c("","2.5$^{th}$ quantile", "Mean", "97.5$^{th}$ quantile"),
             format = "latex",linesep = "", escape = F,align="c",
             caption.short="Results for nonparametric statistics, \\textit{Ceratium furca} abundance data",
             caption = "Average values of statistics across all locations pairs and confidence intervals based on spatial resampling for \\textit{Ceratium furca} abundance data.\\label{tab_pns_resamp}",
             booktabs=TRUE)
```

\begin{figure}[!h]
\textbf{ \hspace{3.4 cm} (A) \hspace{3.2 cm} (B) \hspace{3.2 cm} (C) \hspace{3.4 cm} (D)} \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Spearman_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Kendall_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Corl_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Coru_vs_D.pdf}\\
\end{center}
\textbf{ \hspace{3.4 cm} (E) \hspace{3.2 cm} (F) \hspace{3.2 cm} (G) \hspace{3.4 cm} (H)} \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Pl_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Pu_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_plankton_north_sea/Sp_16_D2u_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_plankton_north_sea/Sp_16_D2l_vs_D.pdf}\\
\end{center}
\textbf{ \hspace{5.4 cm} (I) \hspace{3.4 cm} (J) \hspace{3.2 cm} (K) } \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Corl-Coru_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Pl-Pu_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_plankton_north_sea/Sp_16_D2u-D2l_vs_D.pdf}
\caption[Nonparametric tail statistics versus distance between locations for \textit{Ceratium furca} abundance data]{Nonparametric tail statistics versus great-circle distance, $D$, between sampling locations for \textit{Ceratium furca} abundance data. \label{fig_pns}}
\end{center}
\end{figure}

<!--Load and format the data-->
```{r read_format_methane_data, echo=F}

methane_raw<-readRDS("./Data/Methane_data/methane_data_for_Shyamolina.rds")

myls <- vector("list", length = dim(methane_raw)[2]-1)
for(i in 2:dim(methane_raw)[2]){
  temp<-methane_raw[,c(1,i)]
  colnames(temp)<-c("Year","Dat")
  myls[[i-1]]<-temp
}
methane_data<-list(myls=myls)
```

<!--Do the main analysis-->
```{r RES_methane, echo=F, cache=T, results="hide", cache.extra=list(seed,methane_data,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}

source("FittingCopula_selective_loc.R")  # top most hierarchy source file

set.seed(seed)
#cat(paste("start-time: ",Sys.time(),"\n")) 
RES_methane<-RES_single_sp(d_allsp=methane_data,sp=1,families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=50,
                           good_loc_given=F,good_loc=NA)
#cat(paste("stop-time: ",Sys.time(),"\n"))    
resloc<-"./Results/fitting_results/"
saveRDS(RES_methane,paste(resloc,file="MethaneFluxFit_selecloc_data_pt_thrs_50.RDS",sep=""))
```

```{r tab_methane_bestcop, echo=F, results="markup"}
knitr::kable(RES_methane$info_ord_copname[,,1], #longtable=T,
             format='latex',linesep = "",
             caption.short = "Best-fitting copulas, methane flux data",
             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for methane flux data. See methods for Q1 for copula abbreviations. Indep.=Independent. NA occurs along the diagonal because diagonal entries represent a sampling location compared to itself.\\label{tab_methane_bestcop}",
             booktabs=TRUE
             )
```

```{r tab_methane_AICw,echo=F,results="markup"}
h<-RES_methane$info_ord_AICw[,,1]
h[is.infinite(h)]<-NA
knitr::kable(h,
             format="latex",linesep = "",
             digits=3,
             caption.short="AIC weights, best-fitting copulas, methane flux data",
             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for methane flux data.\\label{tab_methane_AICw}",
             booktabs=TRUE)
```

```{r tab_methane_AICw_normal,echo=F,results="markup"}
source("normal_cop_AICw.R")
h<-normal_cop_AICw_matrix(r=RES_methane)
h[is.infinite(h)]<-NA
knitr::kable(h,
             format="latex",linesep = "",
             digits=3,
             caption.short="AIC weights, N copulas, methane flux data",
             caption = "$\\text{AIC}_{\\text{w}}$ for the N copula family for each location pair for methane flux data, for comparison with Table \\ref{tab_methane_AICw}.\\label{tab_methane_AICw_normal}",
             booktabs=TRUE)
```

```{r tab_methane_p_CvM,echo=F,results="markup"}
h<-RES_methane$gfc_p_CvM
h[is.infinite(h)]<-NA
knitr::kable(h,
             format="latex",linesep = "",
             digits=3,
             caption.short="Goodness of fit of best-fitting copulas, methane flux data",
             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for methane flux data.\\label{tab_methane_p_CvM}",
             booktabs=TRUE)
```

```{r tab_methane_p_KS,echo=F,results="markup"}
h<-RES_methane$gfc_p_KS
h[is.infinite(h)]<-NA
knitr::kable(h,
             format="latex",linesep = "",
             digits=3,
             caption.short="Goodness of fit of best-fitting copulas, methane flux data",
             caption = "$P$-value results of a Kolmogorov-Smirnov-based test of
the goodness of fit of the lowest-AIC copula family for each pair of
locations for methane flux data.\\label{tab_methane_p_KS}",
             booktabs=TRUE)
```

```{r tab_methane_LTdep_AICw,echo=F,results="markup"}
h<-RES_methane$LTdep_AICw
h[is.infinite(h)]<-NA
knitr::kable(h,
             format="latex",linesep = "",
             digits=3,
             caption.short="Lower-tail dependence, methane flux data",
             caption = "Model-averaged lower-tail dependence statistics for each location pair for methane flux data.\\label{tab_methane_LTdep_AICw}",
             booktabs=TRUE)
```

```{r tab_methane_UTdep_AICw,echo=F,results="markup"}
h<-RES_methane$UTdep_AICw
h[is.infinite(h)]<-NA
knitr::kable(h,
             format="latex",linesep = "",
             digits=3,
             caption.short="Upper-tail dependence, methane flux data",
             caption = "Model-averaged upper-tail dependence statistics for each location pair for methane flux data.\\label{tab_methane_UTdep_AICw}",
             booktabs=TRUE)
```

```{r tab_methane_LTmUT,echo=F,results="markup"}
h<-RES_methane$LTdep_AICw-RES_methane$UTdep_AICw
h[is.nan(h)]<-NA
knitr::kable(h,
             format="latex",linesep = "",
             digits=3,
             caption.short="Lower- minus upper-tail dependence, methane flux data",
             caption = "Model-averaged lower-tail dependence statistics minus model-averaged upper-tail dependence statistics for each location pair for methane flux data.\\label{tab_methane_LTmUT}",
             booktabs=TRUE)
```

<!--Do the main computations for nonparametric analysis of methane-flux data-->
```{r stat_methane, echo=F, cache=T, cache=T,  cache.extra=list(seed,methane_data,mtime("good_loclist.R"),mtime("NonParamStat.R"),mtime("vivj_matrix.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}

source("good_loclist.R")
source("NonParamStat.R")
set.seed(seed)
resloc<-"./Results/stat_results/stat_methane/"

data_pt_thrs<-50
good_loc <- good_loclist(d_allsp=methane_data,sp=1,data_pt_thrs=data_pt_thrs)

latlon<-readRDS("./Data/Methane_data/chamberLocations.rds")
chamber_name<-colnames(methane_raw[,c(2:dim(methane_raw)[2])])
ndim<-length(chamber_name)

latlon_arranged<-data.frame(chamber_name=chamber_name,lat=NA*numeric(ndim),lon=NA*numeric(ndim))
for(i in 1:ndim){
  if (chamber_name[i] %in% latlon$chamber){
    ind<-which(latlon$chamber==chamber_name[i])
    latlon_arranged[i,2:3]<-unname(latlon[ind,2:3])
  }
}

stat_methane<-multcall(d_allsp=methane_data,sp=1,lats=latlon_arranged$lat,longs=latlon_arranged$lon,
                           pfname=paste(resloc,"data_pt_thrs",data_pt_thrs,sep=''),good_loc=good_loc)

saveRDS(stat_methane,paste(resloc,file="stat_methane.RDS",sep='')) 
```

```{r read_stat_methane,echo=F}
stat_methane<-readRDS("./Results/stat_results/stat_methane/stat_methane.RDS")
```

<!--display the main statistical results in a table-->
```{r tab_methane_resamp,echo=F,results="markup"}
s<-stat_methane$numericdf[,c(3,2,4)]

#rownames(s)<-c("Spearman","Kendall","cor_l","cor_u","P_l","P_u","D^2_l","D^2_u","cor_l - cor_u","P_l - P_u","D^2_u - D^2_l")

sc<-as.data.frame(c("Spearman","Kendall","cor$_{l}$","cor$_{u}$","P$_{l}$","P$_{u}$","D$_{l}^2$","D$_{u}^2$",
                    "cor$_{l}$ - cor$_{u}$","P$_{l}$ - P$_{u}$","D$_{u}^2$ - D$_{l}^2$"))
s<-cbind(sc,s)
knitr::kable(s,digits = 3,
             format = "latex",linesep = "", escape = F,align="c",
             col.names=c("","2.5$^{th}$ quantile", "Mean", "97.5$^{th}$ quantile"),
             caption.short="Results for nonparametric statistics, methane flux data",
             caption = "Average values of statistics across all locations pairs and confidence intervals based on spatial resampling for methane flux data.\\label{tab_methane_resamp}",
             booktabs=TRUE)
```

\begin{figure}[!h]
\textbf{ \hspace{3.4 cm} (A) \hspace{3.2 cm} (B) \hspace{3.2 cm} (C) \hspace{3.4 cm} (D)} \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/stat_results/stat_methane/data_pt_thrs50_Spearman_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_methane/data_pt_thrs50_Kendall_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_methane/data_pt_thrs50_Corl_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_methane/data_pt_thrs50_Coru_vs_D.pdf}\\
\end{center}
\textbf{ \hspace{3.4 cm} (E) \hspace{3.2 cm} (F) \hspace{3.2 cm} (G) \hspace{3.4 cm} (H)} \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/stat_results/stat_methane/data_pt_thrs50_Pl_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_methane/data_pt_thrs50_Pu_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_methane/data_pt_thrs50_D2u_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_methane/data_pt_thrs50_D2l_vs_D.pdf}\\
\end{center}
\textbf{ \hspace{5.4 cm} (I) \hspace{3.4 cm} (J) \hspace{3.2 cm} (K) } \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/stat_results/stat_methane/data_pt_thrs50_Corl-Coru_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_methane/data_pt_thrs50_Pl-Pu_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_methane/data_pt_thrs50_D2u-D2l_vs_D.pdf}
\caption[Nonparametric tail statistics versus distance between locations for methane-flux data]{Nonparametric tail statistics versus great-circle distance, $D$, between sampling locations for methane-flux data. \label{fig_methane}}
\end{center}
\end{figure}

```{r bivar_tab_summary_fn,echo=F}
bivar_tab_summary_fn<-function(r,rs){

  m<-matrix(NA,1,10)
  colnames(m)<-c("p.Independence.test","minimumAIC.bestfit.copula","Bestfit.copula.AICw",
                 "Normal.copula.AICw","p.CvM","p.KS",
                 "LT.model.avg","UT.model.avg","(LT-UT).model.avg",
                 "Rank.(Pl-Pu)")
  m<-as.data.frame(m)

  m$p.Independence.test<-round(r$IndepTestRes,3)
  m$minimumAIC.bestfit.copula<-r$InfCritRes$copname[which.min(r$InfCritRes$AIC)]
  m$Bestfit.copula.AICw<-round(max(r$InfCritRes$AICw),3)
  m$Normal.copula.AICw<-round(r$InfCritRes$AICw[1],3)
  m$p.CvM<-r$GofRes_CvM
  m$p.KS<-r$GofRes_KS
  m$LT.model.avg<-round(r$relLTdep_AICw,3)
  m$UT.model.avg<-round(r$relUTdep_AICw,3)
  m$`(LT-UT).model.avg`<-m$LT.model.avg-m$UT.model.avg
  m$`Rank.(Pl-Pu)`<-rs$Rank_PlmPu_K
  mt<-as.data.frame(t(m))
  return(mt)

}
#----------------------------------
#calling the above function on bivariable datasets
m1<-bivar_tab_summary_fn(r=BivMS_res_Loecke,rs=stat_soilCN_ga)
m2<-bivar_tab_summary_fn(r=BivMS_birds,rs=stat_birds_ga)
m3<-bivar_tab_summary_fn(r=BivMS_mammals,rs=stat_mammals_ga)
m4<-bivar_tab_summary_fn(r=BivMS_res_cc$`2000`,rs=stat_cc_ga$`2000`)
tab_bivar_summary<-cbind(m1,m2,m3,m4)
colnames(tab_bivar_summary)<-c("SoilCN","Birds' BMR","Mammals' BMR","Ceder-creek")
saveRDS(tab_bivar_summary,"./Results/tab_bivar_summary.RDS")
```

```{r multivar_tab_summary_fn,echo=F}
source("normal_cop_AICw.R")

multivar_tab_summary_fn<-function(r,rs){

  m<-matrix(NA,1,17)
  colnames(m)<-c("num.pairs","num.non-indep","num.non-N",
                 "Avg.best.AICw","Avg.N.AICw","num.good.fit",
                 "0.025-CI.LT","0.975-CI.LT","0.025-CI.UT","0.975-CI.UT",
                 "percent.LT-UT>0",
                 "0.025-CI.Pl","0.975-CI-Pl","0.025-CI.Pu","0.975-CI.Pu","0.025-CI.Pl-Pu","0.975-CI.Pl-Pu")
  m<-as.data.frame(m)

  normal_AICw<-normal_cop_AICw_matrix(r=r)

  d<-dim(r$gfc_numBS)[1]
  m$num.pairs<-(d^2)-d 
  m$`num.non-indep`<- (d^2)-d-2*(r$num_indep_loc_pair+r$num_neg_cor_loc_pair)
  m$`num.non-N`<-sum(r$info_ord_copcode[,,1]!=1,na.rm = T)
  r$info_ord_AICw[r$info_ord_AICw==Inf] <- NaN
  m$Avg.best.AICw<-round(mean(r$info_ord_AICw[,,1],na.rm = T),3)
  normal_AICw[normal_AICw==Inf]<-NaN
  m$Avg.N.AICw<-round(mean(normal_AICw,na.rm=T),3)
  pcvm<-r$gfc_p_CvM
  pcvm[pcvm==Inf]<-NaN
  pks<-r$gfc_p_KS
  pks[pks==Inf]<-NaN
  temp<-which((pcvm>0.01) & (pks>0.01),arr.ind=T)
  m$num.good.fit<-dim(temp)[1]
  t<-r$LTdep_AICw[is.finite(r$LTdep_AICw)]
  CI_LT<-unname(quantile(t,probs=c(0.025,0.975),na.rm = T))
  m$`0.025-CI.LT`<-round(CI_LT[1],3)
  m$`0.975-CI.LT`<-round(CI_LT[2],3)
  t<-r$UTdep_AICw[is.finite(r$UTdep_AICw)]
  CI_UT<-unname(quantile(t,probs=c(0.025,0.975),na.rm = T))
  m$`0.025-CI.UT`<-round(CI_UT[1],3)
  m$`0.975-CI.UT`<-round(CI_UT[2],3)
  m$`percent.LT-UT>0`<-round((sum(r$LTdep_AICw-r$UTdep_AICw>0,na.rm=T)/((d^2)-d-2*(r$num_indep_loc_pair+r$num_neg_cor_loc_pair)))*100,2)
  m$`0.025-CI.Pl`<-round(rs$numericdf[5,3],3)
  m$`0.975-CI-Pl`<-round(rs$numericdf[5,4],3)
  m$`0.025-CI.Pu`<-round(rs$numericdf[6,3],3)
  m$`0.975-CI.Pu`<-round(rs$numericdf[6,4],3)
  m$`0.025-CI.Pl-Pu`<-round(rs$numericdf[10,3],3)
  m$`0.975-CI.Pl-Pu`<-list(round(rs$numericdf[10,4],3))
  #m$CI_Pl<-list(c(round(rs$numericdf[5,3],3),round(rs$numericdf[5,4],3)))
  #m$CI_Pu<-list(c(round(rs$numericdf[6,3],3),round(rs$numericdf[6,4],3)))
  #m$`CI_Pl-Pu`<-list(c(round(rs$numericdf[10,3],3),round(rs$numericdf[10,4],3)))

  mt<-as.data.frame(t(m))
  return(mt)

}
#----------------------------------
#calling the above function on 4 multivariable datasets
m1<-multivar_tab_summary_fn(r=RES_aphid_count,rs=stat_aphid_count)
m2<-multivar_tab_summary_fn(r=RES_aphid_ff,rs=stat_aphid_ff)
m3<-multivar_tab_summary_fn(r=RES_plankton_north_sea,rs=stat_plankton_north_sea)
m4<-multivar_tab_summary_fn(r=RES_methane,rs=stat_methane)
tab_multivar_summary<-cbind(m1,m2,m3,m4)
colnames(tab_multivar_summary)<-c("Green-spruce aphid_count","Leaf-plum curling aphid_firstflight","C. furca_count","Methane-flux")
saveRDS(tab_multivar_summary,"./Results/tab_multivar_summary.RDS")
```

```{r sample_plot_aphid_count_ff,echo=F,results="hide"}
#-------------------------------------------------------------------------------
# code to plot raw data and copula plot for aphid count @location pair (4,11)
# bestfit cop : SJ

d1<-data_aphid_count$`Green spruce aphid`[[4]]
d2<-data_aphid_count$`Green spruce aphid`[[11]]
pdf("./Results/Aphid_count_raw_loc_4_11.pdf",height = 8,width=8)
op<-par(mar=c(10,10, 0.2, 2.1),mgp=c(3,1,0))
plot(d1$Dat,d2$Dat,type="p",xlab="",
     ylab="Green-spruce aphid \ncount (loc. 11)",cex.lab=3.5,cex.axis=2.5)
mtext(paste("Green-spruce aphid \ncount (loc. 4)"),line=-38,cex=3.5)
par(op)
dev.off()

source("./vivj_matrix.R")
m<-vivj_matrix(d_allsp = data_aphid_count,sp=10,i=4,j=11)
pdf("./Results/Aphid_count_cop_loc_4_11.pdf",height = 8,width=8)
op<-par(mar=c(6,6, 0.2, 2.1),mgp=c(3,1,0))
plot(m[,1],m[,2],type="p",xlab="u",ylab="v",
     cex.lab=4,cex.axis=3,
     xlim=c(0,1),ylim=c(0,1),asp=1)
rect(0,0,1,1)
lines(c(0,1),c(0,1),type='l',col="grey")
par(op)
dev.off()

# code to plot raw data and copula plot for aphid ff @location pair (2,5)
# bestfit cop : J
d1<-data_aphid_ff$`Leaf-curling plum aphid`[[2]]
d2<-data_aphid_ff$`Leaf-curling plum aphid`[[5]]
pdf("./Results/Aphid_ff_raw_loc_2_5.pdf",height = 8,width=8)
op<-par(mar=c(10,10, 0.2, 2.1),mgp=c(3,1,0))
plot(d1$Dat,d2$Dat,type="p",xlab="",ylab="Leaf-plum curling's \nfirst-flight (loc. 5)",cex.lab=3.5,cex.axis=2.5)
mtext(paste("Leaf-plum curling's \nfirst-flight (loc. 2)"),line=-38,cex=3.5)
par(op)
dev.off()

m<-vivj_matrix(d_allsp = data_aphid_ff,sp=11,i=2,j=5)
pdf("./Results/Aphid_ff_cop_loc_2_5.pdf",height = 8,width=8)
op<-par(mar=c(6,6, 0.2, 2.1),mgp=c(3,1,0))
plot(m[,1],m[,2],type="p",xlab="u",ylab="v",
     cex.lab=4,cex.axis=3,
     xlim=c(0,1),ylim=c(0,1),asp=1)
rect(0,0,1,1)
lines(c(0,1),c(0,1),type='l',col="grey")
par(op)
dev.off()
```

```{r cause4copula_pGOF_500, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
#source("Cause4copula.R")
#-----------------
set.seed(seed)
#---------------------------------------
#pdf("./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_500.pdf", width=8, height=3)
#op<-par(mfrow=c(1,2))

#Plotter_Cause4copula_GOF(N=5000,fcode=3,method = "spearman",num_keep_last=500,BS=1000)
#Plotter_Cause4copula_GOF(N=5000,fcode=13,method="spearman",num_keep_last=500,BS=1000)
#Plotter_Cause4copula_GOF(N=5000,fcode=4,method = "spearman",num_keep_last=500,BS=1000)
#Plotter_Cause4copula_GOF(N=5000,fcode=14,method="spearman",num_keep_last=500,BS=1000)

#Plotter_Cause4copula_GOF(N=5000,fcode=3,method = "kendall",num_keep_last=500,BS=1000)
#Plotter_Cause4copula_GOF(N=5000,fcode=13,method="kendall",num_keep_last=500,BS=1000)
#Plotter_Cause4copula_GOF(N=5000,fcode=4,method = "kendall",num_keep_last=500,BS=1000)
#Plotter_Cause4copula_GOF(N=5000,fcode=14,method="kendall",num_keep_last=500,BS=1000)

#par(op)

#op<-par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 1, 0, 0), new = TRUE)
#plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
#legend("top", c("noise copula,", "population copula,",  "Gof test : p-value(CvM),", "Gof test : p-value(KS)"), col = c("red","blue", #"magenta", "green2"),
#       cex = 0.9, pch = c(2, 6, 16, 16), xpd = TRUE, horiz = TRUE, inset = c(0,0),
#       bty = "n") 
#par(op)
#dev.off()
```

```{r cause4copula_pGOF_2500, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
#source("Cause4copula.R")
#-----------------
#set.seed(seed)
#---------------------------------------

#pdf("./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_2500.pdf", width=8, height=3)
#op<-par(mfrow=c(1,2))

#Plotter_Cause4copula_GOF(N=5000,fcode=3,method = "spearman",num_keep_last=2500,BS=1000)
#Plotter_Cause4copula_GOF(N=5000,fcode=13,method="spearman",num_keep_last=2500,BS=1000)
#Plotter_Cause4copula_GOF(N=5000,fcode=4,method = "spearman",num_keep_last=2500,BS=1000)
#Plotter_Cause4copula_GOF(N=5000,fcode=14,method="spearman",num_keep_last=2500,BS=1000)

#Plotter_Cause4copula_GOF(N=5000,fcode=3,method = "kendall",num_keep_last=2500,BS=1000)
#Plotter_Cause4copula_GOF(N=5000,fcode=13,method="kendall",num_keep_last=2500,BS=1000)
#Plotter_Cause4copula_GOF(N=5000,fcode=4,method = "kendall",num_keep_last=2500,BS=1000)
#Plotter_Cause4copula_GOF(N=5000,fcode=14,method="kendall",num_keep_last=2500,BS=1000)

#op<-par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 1, 0, 0), new = TRUE)
#plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
#legend("top", c("noise copula,", "population copula,",  "Gof test : p-value(CvM),", "Gof test : p-value(KS)"), col = c("red","blue", #"magenta", "green2"),
#       cex = 0.9, pch = c(2, 6, 16, 16), xpd = TRUE, horiz = TRUE, inset = c(0,0),
#       bty = "n") 
#par(op)
#dev.off()
```

```{r cause4copula_stat_ar1_clayton, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("Cause4copula.R")
#-----------------
set.seed(seed)
#---------------------------------------
# This chunk stores results for ar1 model : clayton

model<-"ar1"
fcode<-3
nsd<-1

# for positive beta
params<-c(0.5) # this parameter is beta in ar1 model
resloc<-"./Results/Cause4copula_results/"

tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")

if (!dir.exists(tempo)){
  dir.create(tempo)
}
resloc<-paste(tempo,"/",sep="")
# for clayton with spearman and kendall method
ar1_C_spear_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
                                           num_keep_last=500,resloc=resloc,
                                           params=params,p0=c(0,0),model=model)

saveRDS(ar1_C_spear_beta0.5,paste(resloc,"ar1_C_spear_beta0.5.RDS",sep=""))

ar1_C_kend_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
                                          num_keep_last=500,resloc=resloc,
                                          params=params,p0=c(0,0),model=model)

saveRDS(ar1_C_kend_beta0.5,paste(resloc,"ar1_C_kend_beta0.5.RDS",sep=""))

# for negative beta
resloc<-"./Results/Cause4copula_results/"
params<-c(-0.5) # this parameter is beta in ar1 model
tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")

if (!dir.exists(tempo)){
  dir.create(tempo)
}

resloc<-paste(tempo,"/",sep="")
# for clayton with spearman and kendall method
ar1_C_spear_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
                                              num_keep_last=500,resloc=resloc,
                                              params=params,p0=c(0,0),model=model)

saveRDS(ar1_C_spear_negbeta0.5,paste(resloc,"ar1_C_spear_negbeta0.5.RDS",sep=""))

ar1_C_kend_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
                                               num_keep_last=500,resloc=resloc,
                                               params=params,p0=c(0,0),model=model)

saveRDS(ar1_C_kend_negbeta0.5,paste(resloc,"ar1_C_kend_negbeta0.5.RDS",sep=""))
```

```{r plot_cause4copula_stat_ar1_clayton, echo=F, results="hide", cache=T, cache.extra=list(seed,ar1_C_spear_beta0.5,ar1_C_spear_negbeta0.5,ar1_C_kend_beta0.5,ar1_C_kend_negbeta0.5,mtime("Plotter_Cause4copula_stat.R"))}

set.seed(seed)
source("Plotter_Cause4copula_stat.R")
#---------------------------------------
# This chunk plots results for ar1 model : clayton

model<-"ar1"
fcode<-3
resloc0<-"./Results/Cause4copula_results/"

# for positive beta
params<-c(0.5) # this parameter is beta in ar1 model

tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
resloc<-paste(tempo,"/",sep="")

readres<-readRDS(paste(resloc,"ar1_C_spear_beta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)
readres<-readRDS(paste(resloc,"ar1_C_kend_beta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)

# for negative beta
params<-c(-0.5) # this parameter is beta in ar1 model
tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
resloc<-paste(tempo,"/",sep="")

readres<-readRDS(paste(resloc,"ar1_C_spear_negbeta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)
readres<-readRDS(paste(resloc,"ar1_C_kend_negbeta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)
```

```{r cause4copula_stat_ar1_survivalclayton, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("Cause4copula.R")
#-----------------
set.seed(seed)
#---------------------------------------
# This chunk stores results for ar1 model : survival clayton
model<-"ar1"
fcode<-13
nsd<-1

# for positive beta
params<-c(0.5) # this parameter is beta in ar1 model

resloc<-"./Results/Cause4copula_results/"
tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")

if (!dir.exists(tempo)){
  dir.create(tempo)
}
resloc<-paste(tempo,"/",sep="")
# for survival clayton with spearman and kendall method
ar1_SC_spear_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
                                            num_keep_last=500,resloc=resloc,
                                            params=params,p0=c(0,0),model=model)

saveRDS(ar1_SC_spear_beta0.5,paste(resloc,"ar1_SC_spear_beta0.5.RDS",sep=""))

ar1_SC_kend_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
                                           num_keep_last=500,resloc=resloc,
                                           params=params,p0=c(0,0),model=model)

saveRDS(ar1_SC_kend_beta0.5,paste(resloc,"ar1_SC_kend_beta0.5.RDS",sep=""))
# for negative beta
params<-c(-0.5) # this parameter is beta in ar1 model

resloc<-"./Results/Cause4copula_results/"
tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")

if (!dir.exists(tempo)){
  dir.create(tempo)
}

resloc<-paste(tempo,"/",sep="")
# for survival clayton with spearman and kendall method
ar1_SC_spear_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
                                               num_keep_last=500,resloc=resloc,
                                               params=params,p0=c(0,0),model=model)
saveRDS(ar1_SC_spear_negbeta0.5,paste(resloc,"ar1_SC_spear_negbeta0.5.RDS",sep=""))

ar1_SC_kend_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
                                              num_keep_last=500,resloc=resloc,
                                              params=params,p0=c(0,0),model=model)
saveRDS(ar1_SC_kend_negbeta0.5,paste(resloc,"ar1_SC_kend_negbeta0.5.RDS",sep=""))
```

```{r plot_cause4copula_stat_ar1_survivalclayton, echo=F, results="hide", cache=T, cache.extra=list(seed,ar1_SC_spear_beta0.5,ar1_SC_spear_negbeta0.5,ar1_SC_kend_beta0.5,ar1_SC_kend_negbeta0.5,mtime("Plotter_Cause4copula_stat.R"))}

set.seed(seed)
source("Plotter_Cause4copula_stat.R")
#---------------------------------------
# This chunk plots results for ar1 model : clayton

model<-"ar1"
fcode<-13
resloc0<-"./Results/Cause4copula_results/"

# for positive beta
params<-c(0.5) # this parameter is beta in ar1 model

tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
resloc<-paste(tempo,"/",sep="")

readres<-readRDS(paste(resloc,"ar1_SC_spear_beta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)
readres<-readRDS(paste(resloc,"ar1_SC_kend_beta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)

# for negative beta
params<-c(-0.5) # this parameter is beta in ar1 model
tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
resloc<-paste(tempo,"/",sep="")

readres<-readRDS(paste(resloc,"ar1_SC_spear_negbeta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)
readres<-readRDS(paste(resloc,"ar1_SC_kend_negbeta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)
```

```{r cause4copula_stat_ar1_gumbel, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("Cause4copula.R")
#-----------------
set.seed(seed)
#---------------------------------------
# This chunk stores results for ar1 model : clayton

model<-"ar1"
fcode<-4
nsd<-1

# for positive beta
params<-c(0.5) # this parameter is beta in ar1 model
resloc<-"./Results/Cause4copula_results/"

tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")

if (!dir.exists(tempo)){
  dir.create(tempo)
}
resloc<-paste(tempo,"/",sep="")
# for gumbel with spearman and kendall method
ar1_G_spear_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
                                              num_keep_last=500,resloc=resloc,
                                              params=params,p0=c(0,0),model=model)
saveRDS(ar1_G_spear_beta0.5,paste(resloc,"ar1_G_spear_beta0.5.RDS",sep=""))

ar1_G_kend_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
                                             num_keep_last=500,resloc=resloc,
                                             params=params,p0=c(0,0),model=model)

saveRDS(ar1_G_kend_beta0.5,paste(resloc,"ar1_G_kend_beta0.5.RDS",sep=""))

# for negative beta
resloc<-"./Results/Cause4copula_results/"
params<-c(-0.5) # this parameter is beta in ar1 model
tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")

if (!dir.exists(tempo)){
  dir.create(tempo)
}

resloc<-paste(tempo,"/",sep="")
# for gumbel with spearman and kendall method
ar1_G_spear_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
                                              num_keep_last=500,resloc=resloc,
                                              params=params,p0=c(0,0),model=model)
saveRDS(ar1_G_spear_negbeta0.5,paste(resloc,"ar1_G_spear_negbeta0.5.RDS",sep=""))

ar1_G_kend_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
                                             num_keep_last=500,resloc=resloc,
                                             params=params,p0=c(0,0),model=model)

saveRDS(ar1_G_kend_negbeta0.5,paste(resloc,"ar1_G_kend_negbeta0.5.RDS",sep=""))
```

```{r plot_cause4copula_stat_ar1_gumbel, echo=F, results="hide", cache=T, cache.extra=list(seed,ar1_G_spear_beta0.5,ar1_G_spear_negbeta0.5,ar1_G_kend_beta0.5,ar1_G_kend_negbeta0.5,mtime("Plotter_Cause4copula_stat.R"))}

set.seed(seed)
source("Plotter_Cause4copula_stat.R")
#---------------------------------------
# This chunk plots results for ar1 model : clayton

model<-"ar1"
fcode<-4
resloc0<-"./Results/Cause4copula_results/"

# for positive beta
params<-c(0.5) # this parameter is beta in ar1 model

tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
resloc<-paste(tempo,"/",sep="")

readres<-readRDS(paste(resloc,"ar1_G_spear_beta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)
readres<-readRDS(paste(resloc,"ar1_G_kend_beta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)

# for negative beta
params<-c(-0.5) # this parameter is beta in ar1 model
tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
resloc<-paste(tempo,"/",sep="")

readres<-readRDS(paste(resloc,"ar1_G_spear_negbeta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)
readres<-readRDS(paste(resloc,"ar1_G_kend_negbeta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)
```

```{r cause4copula_stat_ar1_survivalgumbel, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("Cause4copula.R")
#-----------------
set.seed(seed)
#---------------------------------------
# This chunk stores results for ar1 model : survival clayton
model<-"ar1"
fcode<-14
nsd<-1

# for positive beta
params<-c(0.5) # this parameter is beta in ar1 model

resloc<-"./Results/Cause4copula_results/"
tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")

if (!dir.exists(tempo)){
  dir.create(tempo)
}
resloc<-paste(tempo,"/",sep="")
# for survival gumbel with spearman and kendall method
ar1_SG_spear_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
                                              num_keep_last=500,resloc=resloc,
                                              params=params,p0=c(0,0),model=model)
saveRDS(ar1_SG_spear_beta0.5,paste(resloc,"ar1_SG_spear_beta0.5.RDS",sep=""))

ar1_SG_kend_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
                                             num_keep_last=500,resloc=resloc,
                                             params=params,p0=c(0,0),model=model)

saveRDS(ar1_SG_kend_beta0.5,paste(resloc,"ar1_SG_kend_beta0.5.RDS",sep=""))

# for negative beta
params<-c(-0.5) # this parameter is beta in ar1 model

resloc<-"./Results/Cause4copula_results/"
tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")

if (!dir.exists(tempo)){
  dir.create(tempo)
}

resloc<-paste(tempo,"/",sep="")
# for survival gumbel with spearman and kendall method
ar1_SG_spear_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
                                              num_keep_last=500,resloc=resloc,
                                              params=params,p0=c(0,0),model=model)
saveRDS(ar1_SG_spear_negbeta0.5,paste(resloc,"ar1_SG_spear_negbeta0.5.RDS",sep=""))

ar1_SG_kend_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
                                             num_keep_last=500,resloc=resloc,
                                             params=params,p0=c(0,0),model=model)

saveRDS(ar1_SG_kend_negbeta0.5,paste(resloc,"ar1_SG_kend_negbeta0.5.RDS",sep=""))
```

```{r plot_cause4copula_stat_ar1_survivalgumbel, echo=F, results="hide", cache=T, cache.extra=list(seed,ar1_SG_spear_beta0.5,ar1_SG_spear_negbeta0.5,ar1_SG_kend_beta0.5,ar1_SG_kend_negbeta0.5,mtime("Plotter_Cause4copula_stat.R"))}

set.seed(seed)
source("Plotter_Cause4copula_stat.R")
#---------------------------------------
# This chunk plots results for ar1 model : clayton

model<-"ar1"
fcode<-14
resloc0<-"./Results/Cause4copula_results/"

# for positive beta
params<-c(0.5) # this parameter is beta in ar1 model

tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
resloc<-paste(tempo,"/",sep="")

readres<-readRDS(paste(resloc,"ar1_SG_spear_beta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)
readres<-readRDS(paste(resloc,"ar1_SG_kend_beta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)

# for negative beta
params<-c(-0.5) # this parameter is beta in ar1 model
tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
resloc<-paste(tempo,"/",sep="")

readres<-readRDS(paste(resloc,"ar1_SG_spear_negbeta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)
readres<-readRDS(paste(resloc,"ar1_SG_kend_negbeta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)
```

```{r cause4copula_stat_ar1_joe, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("Cause4copula.R")
#-----------------
set.seed(seed)
#---------------------------------------
# This chunk stores results for ar1 model : clayton

model<-"ar1"
fcode<-6
nsd<-1

# for positive beta
params<-c(0.5) # this parameter is beta in ar1 model
resloc<-"./Results/Cause4copula_results/"

tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")

if (!dir.exists(tempo)){
  dir.create(tempo)
}
resloc<-paste(tempo,"/",sep="")
# for joe with kendall method only

ar1_J_kend_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
                                             num_keep_last=500,resloc=resloc,
                                             params=params,p0=c(0,0),model=model)

saveRDS(ar1_J_kend_beta0.5,paste(resloc,"ar1_J_kend_beta0.5.RDS",sep=""))

# for negative beta
resloc<-"./Results/Cause4copula_results/"
params<-c(-0.5) # this parameter is beta in ar1 model
tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")

if (!dir.exists(tempo)){
  dir.create(tempo)
}

resloc<-paste(tempo,"/",sep="")
# for joe with kendall method only

ar1_J_kend_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
                                             num_keep_last=500,resloc=resloc,
                                             params=params,p0=c(0,0),model=model)

saveRDS(ar1_J_kend_negbeta0.5,paste(resloc,"ar1_J_kend_negbeta0.5.RDS",sep=""))
```

```{r plot_cause4copula_stat_ar1_joe, echo=F, results="hide", cache=T, cache.extra=list(seed,ar1_J_kend_beta0.5,ar1_J_kend_negbeta0.5,mtime("Plotter_Cause4copula_stat.R"))}

set.seed(seed)
source("Plotter_Cause4copula_stat.R")
#---------------------------------------
# This chunk plots results for ar1 model : clayton

model<-"ar1"
fcode<-6
resloc0<-"./Results/Cause4copula_results/"

# for positive beta
params<-c(0.5) # this parameter is beta in ar1 model

tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
resloc<-paste(tempo,"/",sep="")

readres<-readRDS(paste(resloc,"ar1_J_kend_beta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)

# for negative beta
params<-c(-0.5) # this parameter is beta in ar1 model
tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
resloc<-paste(tempo,"/",sep="")

readres<-readRDS(paste(resloc,"ar1_J_kend_negbeta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)
```

```{r cause4copula_stat_ar1_survivaljoe, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("Cause4copula.R")
#-----------------
set.seed(seed)
#---------------------------------------
# This chunk stores results for ar1 model : survival clayton
model<-"ar1"
fcode<-16
nsd<-1

# for positive beta
params<-c(0.5) # this parameter is beta in ar1 model

resloc<-"./Results/Cause4copula_results/"
tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")

if (!dir.exists(tempo)){
  dir.create(tempo)
}
resloc<-paste(tempo,"/",sep="")
# for survival joe with kendall method only

ar1_SJ_kend_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
                                             num_keep_last=500,resloc=resloc,
                                             params=params,p0=c(0,0),model=model)

saveRDS(ar1_SJ_kend_beta0.5,paste(resloc,"ar1_SJ_kend_beta0.5.RDS",sep=""))

# for negative beta
params<-c(-0.5) # this parameter is beta in ar1 model

resloc<-"./Results/Cause4copula_results/"
tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")

if (!dir.exists(tempo)){
  dir.create(tempo)
}

resloc<-paste(tempo,"/",sep="")
# for survival joe with kendall method only

ar1_SJ_kend_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
                                             num_keep_last=500,resloc=resloc,
                                             params=params,p0=c(0,0),model=model)

saveRDS(ar1_SJ_kend_negbeta0.5,paste(resloc,"ar1_SJ_kend_negbeta0.5.RDS",sep=""))
```

```{r plot_cause4copula_stat_ar1_survivaljoe, echo=F, results="hide", cache=T, cache.extra=list(seed,ar1_SJ_kend_beta0.5,ar1_SJ_kend_negbeta0.5,mtime("Plotter_Cause4copula_stat.R"))}

set.seed(seed)
source("Plotter_Cause4copula_stat.R")
#---------------------------------------
# This chunk plots results for ar1 model : clayton

model<-"ar1"
fcode<-16
resloc0<-"./Results/Cause4copula_results/"

# for positive beta
params<-c(0.5) # this parameter is beta in ar1 model

tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
resloc<-paste(tempo,"/",sep="")

readres<-readRDS(paste(resloc,"ar1_SJ_kend_beta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)

# for negative beta
params<-c(-0.5) # this parameter is beta in ar1 model
tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
resloc<-paste(tempo,"/",sep="")

readres<-readRDS(paste(resloc,"ar1_SJ_kend_negbeta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)
```

```{r cause4copula_stat_ricker_clayton, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("Cause4copula.R")
#-----------------
set.seed(seed)
#---------------------------------------
# This chunk stores results for ar1 model : clayton
tag_ricker_C<-"999C" #this is a caching tag used later to cache output from present chunk

model<-"ricker"
fcode<-3

nsdlist<-c(0.1,1)
Klist<-c(100)

# for varying r (@r=1 transition from under- to over-compensatory dynamics )
rlist<-seq(from=0.1,to=1.3,by=0.2)

for(insd in nsdlist){
  for(iK in Klist){
   for(i in rlist){
    params<-c(i,iK) # this parameter is a vector with r and K
    p0<-c(params[2],params[2])

    resloc<-"./Results/Cause4copula_results/"

    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")

    if (!dir.exists(tempo)){
      dir.create(tempo)
    }
    resloc<-paste(tempo,"/",sep="")

  # for clayton with spearman and kendall method
    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="spearman",lb=0,ub=0.1,
                                           num_keep_last=500,resloc=resloc,
                                           params=params,p0=p0,model=model)

    saveRDS(myres,paste(resloc,"ricker_C_spear.RDS",sep=""))

    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,
                                          num_keep_last=500,resloc=resloc,
                                          params=params,p0=p0,model=model)

    saveRDS(myres,paste(resloc,"ricker_C_kend.RDS",sep=""))

  }
 }
}
```

```{r plot_cause4copula_stat_ricker_C, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_C,mtime("Plotter_Cause4copula.R"))}

set.seed(seed)
source("Plotter_Cause4copula_stat.R")
#---------------------------------------
# This chunk plots results for ar1 model : clayton

model<-"ricker"
fcode<-3
resloc0<-"./Results/Cause4copula_results/"

nsdlist<-c(0.1,1)
Klist<-c(100)

# for varying r (@r=1 transition from under- to over-compensatory dynamics )
rlist<-seq(from=0.1,to=1.3,by=0.2)

for(insd in nsdlist){
  for(iK in Klist){
   for(i in rlist){
    params<-c(i,iK) # this parameter is a vector with r and K
    p0<-c(params[2],params[2])

    resloc<-"./Results/Cause4copula_results/"

    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")

    resloc<-paste(tempo,"/",sep="")

    readres<-readRDS(paste(resloc,"ricker_C_spear.RDS",sep=""))
    Plotter_Cause4copula_stat(res=readres)
    readres<-readRDS(paste(resloc,"ricker_C_kend.RDS",sep=""))
    Plotter_Cause4copula_stat(res=readres)

  }
 }
}
```

```{r cause4copula_stat_ricker_survivalclayton, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("Cause4copula.R")
#-----------------
set.seed(seed)
#---------------------------------------
tag_ricker_SC<-"999SC" #this is a caching tag used later to cache output from present chunk

# This chunk stores results for ar1 model : survival clayton

model<-"ricker"
fcode<-13
nsdlist<-c(0.1,1)
Klist<-c(100)
# for varying r (@r=1 transition from under- to over-compensatory dynamics )
rlist<-seq(from=0.1,to=1.3,by=0.2)

for(insd in nsdlist){
  for(iK in Klist){
   for(i in rlist){
    params<-c(i,iK) # this parameter is a vector with r and K
    p0<-c(params[2],params[2])

    resloc<-"./Results/Cause4copula_results/"

    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")

    if (!dir.exists(tempo)){
      dir.create(tempo)
    }
    resloc<-paste(tempo,"/",sep="")

  # for survival clayton with spearman and kendall method
    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="spearman",lb=0,ub=0.1,
                                           num_keep_last=500,resloc=resloc,
                                           params=params,p0=p0,model=model)

    saveRDS(myres,paste(resloc,"ricker_SC_spear.RDS",sep=""))

    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,
                                          num_keep_last=500,resloc=resloc,
                                          params=params,p0=p0,model=model)

    saveRDS(myres,paste(resloc,"ricker_SC_kend.RDS",sep=""))
  }
 }
}
```

```{r plot_cause4copula_stat_ricker_SC, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_SC,mtime("Plotter_Cause4copula.R"))}

set.seed(seed)
source("Plotter_Cause4copula_stat.R")
#---------------------------------------

model<-"ricker"
fcode<-13
resloc0<-"./Results/Cause4copula_results/"

nsdlist<-c(0.1,1)
Klist<-c(100)

# for varying r (@r=1 transition from under- to over-compensatory dynamics )
rlist<-seq(from=0.1,to=1.3,by=0.2)

for(insd in nsdlist){
  for(iK in Klist){
   for(i in rlist){
    params<-c(i,iK) # this parameter is a vector with r and K
    p0<-c(params[2],params[2])

    resloc<-"./Results/Cause4copula_results/"

    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")

    resloc<-paste(tempo,"/",sep="")

    readres<-readRDS(paste(resloc,"ricker_SC_spear.RDS",sep=""))
    Plotter_Cause4copula_stat(res=readres)
    readres<-readRDS(paste(resloc,"ricker_SC_kend.RDS",sep=""))
    Plotter_Cause4copula_stat(res=readres)

  }
 }
}
```

```{r cause4copula_stat_ricker_gumbel, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("Cause4copula.R")
#-----------------
set.seed(seed)
#---------------------------------------
tag_ricker_G<-"999G" #this is a caching tag used later to cache output from present chunk

# This chunk stores results for ar1 model : gumbel

model<-"ricker"
fcode<-4
nsdlist<-c(0.1,1)
Klist<-c(100)
# for varying r (@r=1 transition from under- to over-compensatory dynamics )
rlist<-seq(from=0.1,to=1.3,by=0.2)

for(insd in nsdlist){
  for(iK in Klist){
   for(i in rlist){
    params<-c(i,iK) # this parameter is a vector with r and K
    p0<-c(params[2],params[2])

    resloc<-"./Results/Cause4copula_results/"

    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")

    if (!dir.exists(tempo)){
      dir.create(tempo)
    }
    resloc<-paste(tempo,"/",sep="")

  # for gumbel with spearman and kendall method
    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="spearman",lb=0,ub=0.1,
                                           num_keep_last=500,resloc=resloc,
                                           params=params,p0=p0,model=model)

    saveRDS(myres,paste(resloc,"ricker_G_spear.RDS",sep=""))

    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,
                                          num_keep_last=500,resloc=resloc,
                                          params=params,p0=p0,model=model)

    saveRDS(myres,paste(resloc,"ricker_G_kend.RDS",sep=""))
  }
 }
}
```

```{r plot_cause4copula_stat_ricker_G, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_G,mtime("Plotter_Cause4copula.R"))}

set.seed(seed)
source("Plotter_Cause4copula_stat.R")
#---------------------------------------

model<-"ricker"
fcode<-4
resloc0<-"./Results/Cause4copula_results/"

nsdlist<-c(0.1,1)
Klist<-c(100)

# for varying r (@r=1 transition from under- to over-compensatory dynamics )
rlist<-seq(from=0.1,to=1.3,by=0.2)

for(insd in nsdlist){
  for(iK in Klist){
   for(i in rlist){
    params<-c(i,iK) # this parameter is a vector with r and K
    p0<-c(params[2],params[2])

    resloc<-"./Results/Cause4copula_results/"

    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")

    resloc<-paste(tempo,"/",sep="")

    readres<-readRDS(paste(resloc,"ricker_G_spear.RDS",sep=""))
    Plotter_Cause4copula_stat(res=readres)
    readres<-readRDS(paste(resloc,"ricker_G_kend.RDS",sep=""))
    Plotter_Cause4copula_stat(res=readres)

  }
 }
}
```

```{r cause4copula_stat_ricker_survivalgumbel, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("Cause4copula.R")
#-----------------
set.seed(seed)
#---------------------------------------
tag_ricker_SG<-"999SG" #this is a caching tag used later to cache output from present chunk

# This chunk stores results for ar1 model : survival gumbel

model<-"ricker"
fcode<-14
nsdlist<-c(0.1,1)
Klist<-c(100)
# for varying r (@r=1 transition from under- to over-compensatory dynamics )
rlist<-seq(from=0.1,to=1.3,by=0.2)

for(insd in nsdlist){
  for(iK in Klist){
   for(i in rlist){
    params<-c(i,iK) # this parameter is a vector with r and K
    p0<-c(params[2],params[2])

    resloc<-"./Results/Cause4copula_results/"

    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")

    if (!dir.exists(tempo)){
      dir.create(tempo)
    }
    resloc<-paste(tempo,"/",sep="")

  # for survival gumbel with spearman and kendall method
    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="spearman",lb=0,ub=0.1,
                                           num_keep_last=500,resloc=resloc,
                                           params=params,p0=p0,model=model)

    saveRDS(myres,paste(resloc,"ricker_SG_spear.RDS",sep=""))

    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,
                                          num_keep_last=500,resloc=resloc,
                                          params=params,p0=p0,model=model)

    saveRDS(myres,paste(resloc,"ricker_SG_kend.RDS",sep=""))
  }
 }
}
```

```{r plot_cause4copula_stat_ricker_SG, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_SG,mtime("Plotter_Cause4copula.R"))}

set.seed(seed)
source("Plotter_Cause4copula_stat.R")
#---------------------------------------

model<-"ricker"
fcode<-14
resloc0<-"./Results/Cause4copula_results/"

nsdlist<-c(0.1,1)
Klist<-c(100)

# for varying r (@r=1 transition from under- to over-compensatory dynamics )
rlist<-seq(from=0.1,to=1.3,by=0.2)

for(insd in nsdlist){
  for(iK in Klist){
   for(i in rlist){
    params<-c(i,iK) # this parameter is a vector with r and K
    p0<-c(params[2],params[2])

    resloc<-"./Results/Cause4copula_results/"

    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")

    resloc<-paste(tempo,"/",sep="")

    readres<-readRDS(paste(resloc,"ricker_SG_spear.RDS",sep=""))
    Plotter_Cause4copula_stat(res=readres)
    readres<-readRDS(paste(resloc,"ricker_SG_kend.RDS",sep=""))
    Plotter_Cause4copula_stat(res=readres)

  }
 }
}
```

```{r cause4copula_stat_ricker_joe, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("Cause4copula.R")
#-----------------
set.seed(seed)
#---------------------------------------
tag_ricker_J<-"999J" #this is a caching tag used later to cache output from present chunk

# This chunk stores results for ar1 model : joe

model<-"ricker"
fcode<-6
nsdlist<-c(0.1,1)
Klist<-c(100)
# for varying r (@r=1 transition from under- to over-compensatory dynamics )
rlist<-seq(from=0.1,to=1.3,by=0.2)

for(insd in nsdlist){
  for(iK in Klist){
   for(i in rlist){
    params<-c(i,iK) # this parameter is a vector with r and K
    p0<-c(params[2],params[2])

    resloc<-"./Results/Cause4copula_results/"

    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")

    if (!dir.exists(tempo)){
      dir.create(tempo)
    }
    resloc<-paste(tempo,"/",sep="")

  # for joe with spearman and kendall method

    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,
                                          num_keep_last=500,resloc=resloc,
                                          params=params,p0=p0,model=model)

    saveRDS(myres,paste(resloc,"ricker_J_kend.RDS",sep=""))
  }
 }
}
```

```{r plot_cause4copula_stat_ricker_J, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_J,mtime("Plotter_Cause4copula.R"))}

set.seed(seed)
source("Plotter_Cause4copula_stat.R")
#---------------------------------------

model<-"ricker"
fcode<-6
resloc0<-"./Results/Cause4copula_results/"

nsdlist<-c(0.1,1)
Klist<-c(100)

# for varying r (@r=1 transition from under- to over-compensatory dynamics )
rlist<-seq(from=0.1,to=1.3,by=0.2)

for(insd in nsdlist){
  for(iK in Klist){
   for(i in rlist){
    params<-c(i,iK) # this parameter is a vector with r and K
    p0<-c(params[2],params[2])

    resloc<-"./Results/Cause4copula_results/"

    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")

    resloc<-paste(tempo,"/",sep="")

    readres<-readRDS(paste(resloc,"ricker_J_kend.RDS",sep=""))
    Plotter_Cause4copula_stat(res=readres)

  }
 }
}
```

```{r cause4copula_stat_ricker_survivaljoe, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("Cause4copula.R")
#-----------------
set.seed(seed)
#---------------------------------------
tag_ricker_SJ<-"999SJ" #this is a caching tag used later to cache output from present chunk

# This chunk stores results for ar1 model : survival joe

model<-"ricker"
fcode<-16
nsdlist<-c(0.1,1)
Klist<-c(100)
# for varying r (@r=1 transition from under- to over-compensatory dynamics )
rlist<-seq(from=0.1,to=1.3,by=0.2)

for(insd in nsdlist){
  for(iK in Klist){
   for(i in rlist){
    params<-c(i,iK) # this parameter is a vector with r and K
    p0<-c(params[2],params[2])

    resloc<-"./Results/Cause4copula_results/"

    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")

    if (!dir.exists(tempo)){
      dir.create(tempo)
    }
    resloc<-paste(tempo,"/",sep="")

  # for survival joe with spearman and kendall method

    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,
                                          num_keep_last=500,resloc=resloc,
                                          params=params,p0=p0,model=model)

    saveRDS(myres,paste(resloc,"ricker_SJ_kend.RDS",sep=""))
  }
 }
}
```

```{r plot_cause4copula_stat_ricker_SJ, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_SJ,mtime("Plotter_Cause4copula.R"))}

set.seed(seed)
source("Plotter_Cause4copula_stat.R")
#---------------------------------------

model<-"ricker"
fcode<-16
resloc0<-"./Results/Cause4copula_results/"

nsdlist<-c(0.1,1)
Klist<-c(100)

# for varying r (@r=1 transition from under- to over-compensatory dynamics )
rlist<-seq(from=0.1,to=1.3,by=0.2)

for(insd in nsdlist){
  for(iK in Klist){
   for(i in rlist){
    params<-c(i,iK) # this parameter is a vector with r and K
    p0<-c(params[2],params[2])

    resloc<-"./Results/Cause4copula_results/"

    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")

    resloc<-paste(tempo,"/",sep="")

    readres<-readRDS(paste(resloc,"ricker_SJ_kend.RDS",sep=""))
    Plotter_Cause4copula_stat(res=readres)

  }
 }
}
```
<!--
\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_2500.pdf}
\caption[Comparing noise-copula parameters and estimated population copula parameters with $2500$ sampling points]{Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last \textbf{2500} points for both noise and population data for comparison). p-values $> 0.05$ (above horizontal dashed line) indicate good fit of population copula with same noise cpoula family. \label{fig_Cause4copula_params_GOF_2500}}
\end{center}
\end{figure}
-->

\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2u-D2l_vs_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, AR(1) model,
C copula.]{Comparison of copula structure of noise inputs and model 
outputs, AR(1) model, C copula. Red and blue points give 
means over 50 replicate simulations of the statistics listed on the left axes, 
red points are values for noise inputs and blue points are for
model outputs. Error bars are standard errors. 
$P$-values (purple points, right axis) are for a 
paired $t$-test of the null hypothesis that the distributions of 50 noise and 
output values have the same mean; $p$-values less than $0.05$ indicate that 
copula structure as measured by the listed statistic differs significant
between noise inputs and model outputs. See methods section for Q2 
for additional details.\label{fig_Cause4copula_C_kend_ar1}}
\end{center}
\end{figure}
<!--***DAN: Shya, can you please change the top legend of this and the similar-format
figures so instead of saying "noise copula" for the red points and "population copula" 
for the blue points it says "input noise copula" for the red points and "model output 
copula" for the blue ones? This is because one of our models is much more general than
just a population model, so I have changed the langauge in the text to reflect that.
Please be aware that when you make this change it will mean the next knit will take 
several hours, unless you go to the effort of reorganizing your code to put the sims
and the plots in separate chunks. So maybe make this change right before you go home,
so you can knit over night. Also please let me know, so I can knit overnight, too.

At the same time that you do that above, please see if you can fix the alignment
between the bottom three panels of Fig S32 (and subsequent analogous figures) so they
line up with the panels above that.
***Shya: Done.-->

\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_D2u-D2l_vs_Kendall's Tau.pdf}\\
\hspace{-0.8cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, AR(1) model, 
SC copula.]{Comparison of copula structure of noise inputs and model outputs, AR(1) 
model, SC copula. Same format as Figs \ref{fig_Cause4copula_C_kend_ar1} and 
\ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, see captions of those figures for 
details.\label{fig_Cause4copula_SC_kend_ar1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_D2u-D2l_vs_Kendall's Tau.pdf}\\
\hspace{-0.8cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, AR(1) model, 
G copula.]{Comparison of copula structure of noise inputs and model outputs, AR(1) 
model, G copula. Same format as Figs \ref{fig_Cause4copula_C_kend_ar1} and 
\ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, see captions of those figures for 
details.\label{fig_Cause4copula_G_kend_ar1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_D2u-D2l_vs_Kendall's Tau.pdf}\\
\hspace{-0.8cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, AR(1) model, 
SG copula.]{Comparison of copula structure of noise inputs and model outputs, AR(1) 
model, SG copula. Same format as Figs \ref{fig_Cause4copula_C_kend_ar1} and 
\ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, see captions of those figures for 
details.\label{fig_Cause4copula_SG_kend_ar1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_D2u-D2l_vs_Kendall's Tau.pdf}\\
\hspace{-0.8cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_scatter_D2umD2l_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, AR(1) model, 
J copula.]{Comparison of copula structure of noise inputs and model outputs, AR(1) 
model, J copula. Same format as Figs \ref{fig_Cause4copula_C_kend_ar1} and 
\ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, see captions of those figures for 
details.\label{fig_Cause4copula_J_kend_ar1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_D2u-D2l_vs_Kendall's Tau.pdf}\\
\hspace{-0.8cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_scatter_D2umD2l_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, AR(1) model, 
SJ copula.]{Comparison of copula structure of noise inputs and model outputs, AR(1) 
model, SJ copula. Same format as Figs \ref{fig_Cause4copula_C_kend_ar1} and 
\ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, see captions of those figures for 
details.\label{fig_Cause4copula_SJ_kend_ar1}}
\end{center}
\end{figure}

<!--adding results for stochastic ricker with weak noise : nsd=0.1-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_D2u-D2l_vs_Kendall's Tau.pdf}\\
\hspace{-0.8cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
model, weak noise, C copula.]{Comparison of copula structure of noise inputs and model 
outputs, stochastic Ricker model, weak noise, C copula. Same format as Figs 
\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
see captions of those figures for details.\label{fig_Cause4copula_C_kend_sRicker_nsd_0.1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_D2u-D2l_vs_Kendall's Tau.pdf}\\
\hspace{-0.8cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
model, weak noise, SC copula.]{Comparison of copula structure of noise inputs and model 
outputs, stochastic Ricker model, weak noise, SC copula. Same format as Figs 
\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
see captions of those figures for details. \label{fig_Cause4copula_SC_kend_sRicker_nsd_0.1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_D2u-D2l_vs_Kendall's Tau.pdf}\\
\hspace{-0.8cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
model, weak noise, G copula.]{Comparison of copula structure of noise inputs and model 
outputs, stochastic Ricker model, weak noise, G copula. Same format as Figs 
\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
see captions of those figures for details.\label{fig_Cause4copula_G_kend_sRicker_nsd_0.1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_D2u-D2l_vs_Kendall's Tau.pdf}\\
\hspace{-0.8cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
model, weak noise, SG copula.]{Comparison of copula structure of noise inputs and model 
outputs, stochastic Ricker model, weak noise, SG copula. Same format as Figs 
\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
see captions of those figures for details. \label{fig_Cause4copula_SG_kend_sRicker_nsd_0.1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_D2u-D2l_vs_Kendall's Tau.pdf}\\
\hspace{-0.8cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_scatter_D2umD2l_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
model, weak noise, J copula.]{Comparison of copula structure of noise inputs and model 
outputs, stochastic Ricker model, weak noise, J copula. Same format as Figs 
\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
see captions of those figures for details.\label{fig_Cause4copula_J_kend_sRicker_nsd_0.1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_D2u-D2l_vs_Kendall's Tau.pdf}\\
\hspace{-0.8cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_scatter_D2umD2l_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
model, weak noise, SJ copula.]{Comparison of copula structure of noise inputs and model 
outputs, stochastic Ricker model, weak noise, SJ copula. Same format as Figs 
\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
see captions of those figures for details.\label{fig_Cause4copula_SJ_kend_sRicker_nsd_0.1}}
\end{center}
\end{figure}

<!--adding results for stochastic ricker with strong noise : nsd=1-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_D2u-D2l_vs_Kendall's Tau.pdf}\\
\hspace{-0.8cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
model, strong noise, C copula.]{Comparison of copula structure of noise inputs and model 
outputs, stochastic Ricker model, strong noise, C copula. Same format as Figs 
\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
see captions of those figures for details.\label{fig_Cause4copula_C_kend_sRicker_nsd_1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_D2u-D2l_vs_Kendall's Tau.pdf}\\
\hspace{-0.8cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
model, strong noise, SC copula.]{Comparison of copula structure of noise inputs and model 
outputs, stochastic Ricker model, strong noise, SC copula. Same format as Figs 
\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
see captions of those figures for details. \label{fig_Cause4copula_SC_kend_sRicker_nsd_1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_D2u-D2l_vs_Kendall's Tau.pdf}\\
\hspace{-0.8cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
model, strong noise, G copula.]{Comparison of copula structure of noise inputs and model 
outputs, stochastic Ricker model, strong noise, G copula. Same format as Figs 
\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
see captions of those figures for details.\label{fig_Cause4copula_G_kend_sRicker_nsd_1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_D2u-D2l_vs_Kendall's Tau.pdf}\\
\hspace{-0.8cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
model, strong noise, SG copula.]{Comparison of copula structure of noise inputs and model 
outputs, stochastic Ricker model, strong noise, SG copula. Same format as Figs 
\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
see captions of those figures for details. \label{fig_Cause4copula_SG_kend_sRicker_nsd_1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_D2u-D2l_vs_Kendall's Tau.pdf}\\
\hspace{-0.8cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_scatter_D2umD2l_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
model, strong noise, J copula.]{Comparison of copula structure of noise inputs and model 
outputs, stochastic Ricker model, strong noise, J copula. Same format as Figs 
\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
see captions of those figures for details.\label{fig_Cause4copula_J_kend_sRicker_nsd_1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_D2u-D2l_vs_Kendall's Tau.pdf}\\
\hspace{-0.8cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_scatter_D2umD2l_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
model, strong noise, SJ copula.]{Comparison of copula structure of noise inputs and model 
outputs, stochastic Ricker model, strong noise, SJ copula. Same format as Figs 
\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
see captions of those figures for details.\label{fig_Cause4copula_SJ_kend_sRicker_nsd_1}}
\end{center}
\end{figure}

```{r moran_effect_on_aphid_ff, echo=F, results="hide", cache=T, cache.extra=list(seed,temp_aphid,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
set.seed(seed)
source("FittingCopula_selective_loc.R")  # top most hierarchy source file

#cat(paste("start-time: ",Sys.time(),"\n"))
RES_temp_aphid<-RES_single_sp(d_allsp=temp_aphid,sp=1,families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=30,
                              good_loc_given=T,good_loc=c(2:11))
#cat(paste("stop-time: ",Sys.time(),"\n")) 
resloc<-"./Results/fitting_results/"
saveRDS(RES_temp_aphid,paste(resloc,file="AphidCopulaFit_selecloc_winter_neg_temp.RDS",sep=""))

LTmUT_temp<-(RES_temp_aphid$LTdep_AICw - RES_temp_aphid$UTdep_AICw)

sum(LTmUT_temp>0,na.rm = T)
sum(LTmUT_temp<0,na.rm = T) #66 out of 90 : suggests UT in leaf-plum curling aphids firstflight data (90 out of 90 UT)
```

<!--
```{r tab_temp_aphid_bestcop, echo=F, results="markup"}
knitr::kable(RES_temp_aphid$info_ord_copname[,,1], #longtable = T,
             format='latex',
             #caption.short = "Best-fitting copulas for each location pair from among those considered, according to AIC, for first flight data for the leaf-curling plum aphid.",
             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for **negative average winter temperature data (1958-2013)**. Copulas' abbreviations were given in \\nameref{MT-Methods}. NA occurs along the diagonal of the matrix because the diagonal represents a sampling location compared to itself.  \\label{tab_temp_aphid_bestcop}",
             booktabs=TRUE
             )
```

```{r tab_temp_aphid_AICw, echo=F, results='markup'}
knitr::kable(RES_temp_aphid$info_ord_AICw[,,1], 
             format='latex',
             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for **negative average winter temperature data**.\\label{tab_temp_aphid_AICw}",
             booktabs=TRUE
             )
```

```{r tab_temp_aphid_AICw_normal, echo=F, results='markup'}
source("normal_cop_AICw.R")
knitr::kable(normal_cop_AICw_matrix(r=RES_temp_aphid),
             format='latex',
             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for **negative average winter temperature data**, for comparison with Table \\ref{tab_ff_AICw}.\\label{tab_temp_aphid_AICw_normal}",
             booktabs=TRUE
             )
```

<!--\newpage-->
```{r tab_temp_aphid_p_CvM, echo=F, results='markup'}
knitr::kable(RES_temp_aphid$gfc_p_CvM,
             format='latex',
             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for **negative average winter temperature data**. Tests were based on bootstrapping as described in Methods.\\label{tab_temp_aphid_p_CvM}",
             booktabs=TRUE
             )
```

<!--\newpage-->
```{r tab_temp_aphid_p_KS, echo=F, results='markup'}
knitr::kable(RES_temp_aphid$gfc_p_KS,
             format='latex',
             caption = "$P$-value results of a Kolmogorov-Smirnov-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for **negative average winter temperature data**. Tests were based on bootstrapping as described in Methods.\\label{tab_temp_aphid_p_KS}",
             booktabs=TRUE
             )
```

<!--\newpage-->
```{r tab_temp_aphid_LTdep_AICw, echo=F, results='markup'}
knitr::kable(RES_temp_aphid$LTdep_AICw,
             format='latex',
             caption = "Model-averaged lower tail dependence statistic for each location pair for **negative average winter temperature data**.\\label{tab_temp_aphid_LTdep_AICw}",
             booktabs=TRUE
             )
```

<!--\newpage-->
```{r tab_temp_aphid_UTdep_AICw, echo=F, results='markup'}
knitr::kable(RES_temp_aphid$UTdep_AICw,
             format='latex',
             caption = "Model-averaged upper tail dependence statistic for each location pair for **negative average winter temperature data**.\\label{tab_temp_aphid_UTdep_AICw}",
             booktabs=TRUE
             )
```

<!--\newpage-->
```{r tab_temp_aphid_LTmUT, echo=F, results='markup'}
knitr::kable(RES_temp_aphid$LTdep_AICw - RES_temp_aphid$UTdep_AICw,
             digits = 5,
             format='latex',
             caption = "Model averaged lower tail dependence minus model averaged upper tail dependence statistic for each location pair for **negative average winter temperature data**.\\label{tab_temp_aphid_LTmUT}",
             booktabs=TRUE
             )
```

<!--# Cause of non-Normal copula structure : the Moran effect and copula structure in populations \label{Cause_Moran}-->

<!--
## Asymmetric sensitivity and copula structure in populations
           ----moved to Paper.Rmd----
-->

```{r asym_sens_b_variation, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"),mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
set.seed(seed)
#-----------------------------------------------------------------------------------
source("OurBiCopSelect.R")
source("bivfunctionplot.R")
source("asym_sensitivity.R")
#-----------------------------------------------------------------------------------
# This function takes 
# Input :
#       mg : global noise matrix (dim = N by 2)
#       ml : local noise matrix (dim = N by 2)
#       a : a number : local noise coeeficient
#       b : growth factor of pop
#       decrease : logical (controls the nature of global noise)
#       numkeep_last : number of points you want to keep for the pop matrix from the end
#       ploton : logical for optional plot

# Output :
#     a list of 4 pop matrices 
#                         1. (for b=+ve, decrease=F)
#                         2. (for b=+ve, decrease=T)
#                         3. (for b=-ve, decrease=F)
#                         4. (for b=-ve, decrease=T)

plotter_asym_sens_b_varn<-function(mg,ml,a,b,numkeep_last,ploton,resloc){

  pop_g<-Simulator_asym_sens(mg=mg,ml=ml,a=a,b=b,p0=c(0,0),decrease=F,numkeep_last=numkeep_last)
  pop_d_g<-Simulator_asym_sens(mg=mg,ml=ml,a=a,b=b,p0=c(0,0),decrease=T,numkeep_last=numkeep_last)

  if(ploton==T){
    pdf(paste0(resloc,"a_",a,"_b_",b,"_r_",r,"_asym_sens.pdf",sep=""),height=3,width=6)
    op<-par(mfrow=c(1,2),mar=c(4,4.2,1.5,1),mgp=c(2.5,0.6,0))
    #plot(pop[,1],pop[,2],col="red",main = bquote(beta ==.(beta)))
    pop<-pop_g$pop
    geps<-pop_g$g_noise
    #plot(mg[,1],geps[,1],cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[1])),main="location=1")
    #plot(mg[,2],geps[,2],cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[1])),main="location=2")
    plot(pop[,1],pop[,2],xlab=expression(paste(P[1])), ylab=expression(paste(P[2])),cex.lab=2,cex.axis=2,col="grey",pch=20,cex=1.4)
    #title(expression(paste("g = ",g[1]," , ")), adj=0.06,cex.main=1.5)
    #mtext(paste("a = ",a,", b = ",b),side = 3, line=0.2, adj=0.9)

    pop_d<-pop_d_g$pop
    geps<-pop_d_g$g_noise
    #plot(mg[,1],geps[,1],cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[2])),main="location=1")
    #plot(mg[,2],geps[,2],cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[2])),main="location=2")
    plot(pop_d[,1],pop_d[,2],xlab=expression(paste(P[1])), ylab=expression(paste(P[2])),cex.lab=2,cex.axis=2,col="grey",pch=20,cex=1.4)
    #title(expression(paste("g = ",g[2]," , ")), adj=0.06,cex.main=1.5)
    #mtext(paste("a = ",a,", b = ",b),side = 3, line=0.2, adj=0.9)

    par(op)
    dev.off()
  }

  return(list(pop=pop,
              pop_d=pop_d))

}
#----------------------------------------------------------
N<-5000
r<-0.8
a<-0.2
numkeep_last<-500
temp<-get_noise_glob_loc_mat(N=N,r=r)
mg<-temp$mg
ml<-temp$ml

resloc<-"./Results/asym_sens_results/"
ans_b_0.1<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=0.1,numkeep_last=numkeep_last,ploton=T,resloc=resloc)
ans_b_neg0.1<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=-0.1,numkeep_last=numkeep_last,ploton=T,resloc=resloc)
ans_b_0.9<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=0.9,numkeep_last=numkeep_last,ploton=T,resloc=resloc)
ans_b_neg0.9<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=-0.9,numkeep_last=numkeep_last,ploton=T,resloc=resloc)

vlist_b_0.1_b_0.9<-c(ans_b_0.1,ans_b_neg0.1,ans_b_0.9,ans_b_neg0.9) # a list of 8 matrices

summary_stat<-as.data.frame(matrix(NA,7,8),stringsAsFactors = F) # Table for model selection summary
rownames(summary_stat)<-c("bestcop_AICw","N_AICw","relLTdep_AICw","relUTdep_AICw","p_CvM","p_KS","BS_success%")
#bg <- intToUtf8(0x03B2) 
cnm<-c("b_0.1_g~1~","b_0.1_g~2~","b_-0.1_g~1~","b_-0.1_g~2~","b_0.9_g~1~","b_0.9_g~2~","b_-0.9_g~1~","b_-0.9_g~2~")
#cname<-paste0(bg,cnm)
colnames(summary_stat)<-cnm

ms_asym_b_0.1_b_0.9<-vector(mode = "list", length = 8) # initializing list to store model selection results for all possible b variation
names(ms_asym_b_0.1_b_0.9)<-cnm

stat_npa_asym_b_0.1_b_0.9<-vector(mode = "list", length = 8) # initializing list to store NPA results for all possible b variation
names(stat_npa_asym_b_0.1_b_0.9)<-cnm

for(ic in c(1:8)){
  v<-vlist_b_0.1_b_0.9[[ic]]
  res<-OurBiCopSelect(u1=v[,1],u2=v[,2],families=c(1,3:10,13:14,16:20),level=0.05,
                      AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,
                      gofnormal=F,status=F)

  ms_asym_b_0.1_b_0.9[[ic]]<-res

  summary_stat[1,ic]<-max(res$InfCritRes$AICw)
  summary_stat[2,ic]<-res$InfCritRes$AICw[1]
  summary_stat[3,ic]<-res$relLTdep_AICw
  summary_stat[4,ic]<-res$relUTdep_AICw
  summary_stat[5,ic]<-res$GofRes_CvM
  summary_stat[6,ic]<-res$GofRes_KS
  summary_stat[7,ic]<-(res$Numboot_success/res$Numboot)*100

  stat_npa_asym_b_0.1_b_0.9[[ic]]<-bivfunctionplot(v=v,resloc="./Results/asym_sens_results/",nametag=cnm[ic],numbin=10)

}

saveRDS(summary_stat,"./Results/asym_sens_results/b_variation_summary_stat.RDS")
saveRDS(ms_asym_b_0.1_b_0.9,"./Results/asym_sens_results/ms_asym_b_0.1_b_0.9.RDS")
saveRDS(stat_npa_asym_b_0.1_b_0.9,"./Results/asym_sens_results/stat_npa_asym_b_0.1_b_0.9.RDS")
```

```{r overwritting_plots_asym_cop,echo=F,results="hide"}
pdf("./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf",height=3,width=6)
op<-par(mfrow=c(1,2),mar=c(3.5,4,1,1),mgp=c(2.2,0.5,0))
pop<-ans_b_0.1$pop
plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,cex.axis=1,col="grey",pch=20,cex=0.6)
text(0.1,0.95,"A",cex=2)
pop<-ans_b_0.1$pop_d
plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,cex.axis=1,col="grey",pch=20,cex=0.6)
text(0.1,0.95,"B",cex=2)
par(op)
dev.off()

pdf("./Results/asym_sens_results/a_0.2_b_-0.1_r_0.8_asym_sens.pdf",height=3,width=6)
op<-par(mfrow=c(1,2),mar=c(3.5,4,1,1),mgp=c(2.2,0.5,0))
pop<-ans_b_neg0.1$pop
plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,cex.axis=1,col="grey",pch=20,cex=0.6)
text(0.1,0.95,"A",cex=2)
pop<-ans_b_neg0.1$pop_d
plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,cex.axis=1,col="grey",pch=20,cex=0.6)
text(0.1,0.95,"B",cex=2)
par(op)
dev.off()

pdf("./Results/asym_sens_results/a_0.2_b_0.9_r_0.8_asym_sens.pdf",height=3,width=6)
op<-par(mfrow=c(1,2),mar=c(3.5,4,1,1),mgp=c(2.2,0.5,0))
pop<-ans_b_0.9$pop
plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,cex.axis=1,col="grey",pch=20,cex=0.6)
text(0.1,0.95,"A",cex=2)
pop<-ans_b_0.9$pop_d
plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,cex.axis=1,col="grey",pch=20,cex=0.6)
text(0.1,0.95,"B",cex=2)
par(op)
dev.off()

pdf("./Results/asym_sens_results/a_0.2_b_-0.9_r_0.8_asym_sens.pdf",height=3,width=6)
op<-par(mfrow=c(1,2),mar=c(3.5,4,1,1),mgp=c(2.2,0.5,0))
pop<-ans_b_neg0.9$pop
plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,cex.axis=1,col="grey",pch=20,cex=0.6)
text(0.1,0.95,"A",cex=2)
pop<-ans_b_neg0.9$pop_d
plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,cex.axis=1,col="grey",pch=20,cex=0.6)
text(0.1,0.95,"B",cex=2)
par(op)
dev.off()
```

```{r tab_summary_stat_asym_sens,echo=F,results="markup"}
summary_stat<-summary_stat[c(1:6),]
rownames(summary_stat)<-c("Best-fit copula AICw","N-copula AICw","Model-avg. LT","Model-avg. UT","p, CvM","p, KS")
knitr::kable(summary_stat, booktabs = T, escape=F,
             format = "latex", linesep = "",
             digits=3, col.names=c("b=0.1, g$_{1}$","b=0.1, g$_{2}$",
                                   "b=-0.1, g$_{1}$","b=-0.1, g$_{2}$",
                                   "b=0.9, g$_{1}$","b=0.9, g$_{2}$",
                                   "b=-0.9, g$_{1}$","b=-0.9, g$_{2}$"),
      caption.short="Asymmetric sensitivity model, model selection results",
      caption="Asymmetric sensitivity model, model selection results. 
The output of simulations from the second model detailed in 
the methods section for Q2 (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
\\ref{fig_ms_asym_b_-0.1}-\\ref{fig_ms_asym_b_-0.9}) were subjected to the model 
selection procedures detailed in the 
methods section for Q1. Best-fit copula AICw: the AIC weight of the best-fitting copula family;
N-copula AICw: the AIC weight of the N copula family; Model-avg. LT: the lower-tail 
statistic, weighted-averaged across models using AICw; Model-avg. UT: same, but for the 
upper-tail statistic; p, CvM: p-value for goodness of fit test of the best-fitting copula
family using the test based on the Cramer-von Mises statistic; p, KS: p-value for goodness of fit 
test of the best-fitting copula family using the test based on the Kolmogorov-Smirnov 
statistic.\\label{tab_summary_stat_asym_sens}")
```

```{r read_tab_stat_npa_asym,echo=F,results="asis"}
#knitr::opts_knit$set(kable.force.latex = TRUE)
#tab_asym<-readRDS("./Results/asym_sens_results/stat_npa_asym_b_0.1_b_0.9.RDS")

#for(i in c(1:8)){
#  temp<-make_table_stat_fn(data=tab_asym[[i]])
#  temp2<-knitr::kable(temp, booktabs = T, 
#             format = "latex",
      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
#      caption =paste(cnm[i]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence."))
#  print(temp2)
#  cat("\\newpage")
#}
```

```{r tab_stat_npa_asym,echo=F,results="markup"}
library(kableExtra)

tab_asym<-readRDS("./Results/asym_sens_results/stat_npa_asym_b_0.1_b_0.9.RDS")
tab_stat_npa_asym<-c()
for(i in c(1:8)){
  temp<-make_table_stat_fn(data=tab_asym[[i]])
  tab_stat_npa_asym<-rbind(tab_stat_npa_asym,temp)
}
gtext<-c("b=0.1, g$_{1}$","b=0.1, g$_{2}$","b=-0.1, g$_{1}$","b=-0.1, g$_{2}$","b=0.9, g$_{1}$","b=0.9, g$_{2}$","b=-0.9, g$_{1}$","b=-0.9, g$_{2}$")
tab_stat_npa_asym<-cbind(b=rep(c(0.1,0.1,-0.1,-0.1,0.9,0.9,-0.9,-0.9),each=3),
  g=rep(c("g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$"),each=3),tab_stat_npa_asym)
knitr::kable(tab_stat_npa_asym, booktabs = T, longtable=T,
             format = "latex",linesep = "",escape=F,
             caption.short = "Asymmetric sensitivity model, nonparametric statistics results",
             caption ="Asymmetric sensitivity model, nonparametric statistics results. 
The output of simulations from the second model detailed in 
the methods section for Q2 (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
\\ref{fig_ms_asym_b_-0.1}-\\ref{fig_ms_asym_b_-0.9}) were subjected to nonparamtric statistical 
analyses decribed in the methods section for Q1. The listed statistic was computed for the
output of the model with the indicated b and g, and the value was compared to 1000 values
of the same statistic computed on surrogate time series randomized to have N-copula structure
(see methods for Q1). A table entry $<X$ (respectively, $>X$) indicates the value of the 
given statistic on the data was less (respectively, more) than its value on $X$ of the 
surrogates, so entries of the form $<X$ (respectively, $>X$) for $X$ equal to $975$ or above 
indicate that upper-tail (respectively, lower-tail) dependence was significantly stronger than 
lower-tail dependence compared to an N-copula null hypothesis. `FS' stands for `first slice'
and refers to the bounds 0 to 0.1; `LS' stands for `last slice' and refers to the bounds
0.9 to 1.\\label{tab_stat_npa_asym}")#%>%
 #            kable_styling()%>%
 #            group_rows(index=setNames(c(3,3,3,3,3,3,3,3), gtext))
 #            group_rows("b=0.1, g$_{1}$", 2,4, escape=F) %>%
 #            group_rows("b=0.1, g$_{2}$", 5,7, escape=F) %>%
 #            group_rows("b=-0.1, g$_{1}$", 8,10, escape=F) %>%
 #            group_rows("b=-0.1, g$_{2}$", 11,13, escape=F) %>%
 #            group_rows("b=0.9, g$_{1}$", 14,16, escape=F) %>%
 #            group_rows("b=0.9, g$_{2}$", 17,19, escape=F) %>%
 #            group_rows("b=-0.9, g$_{1}$", 20,22, escape=F) %>%
 #            group_rows("b=-0.9, g$_{2}$", 23,25, escape=F)
```

\begin{figure}[!h]
\begin{center}
\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_-0.1_r_0.8_asym_sens.pdf}
\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=-0.1$.\label{fig_ms_asym_b_-0.1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_0.9_r_0.8_asym_sens.pdf}
\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=0.9$.\label{fig_ms_asym_b_0.9}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_-0.9_r_0.8_asym_sens.pdf}
\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=-0.9$.\label{fig_ms_asym_b_-0.9}}
\end{center}
\end{figure}

```{r skewness_aphidcount_allsp_plot, echo=F, results="hide", warning=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}

# For sp =1, sp=4 and sp=6 : kendall preserving surrogates gives warning : Error in ncsurrog: ncov is not positive semidefinite
# also it can't handle ties for "kendall" and code crashes
# you can see those msg by deleting results="hide" and warning=F option in r chunk

source("good_loclist.R")
source("skewness_multivar_dataset.R")
set.seed(seed)

numsurrog<-10000

splist<-c(1:20)
skw_aphid_count_spearman<-as.data.frame(matrix(NA,nrow=length(splist),ncol=3))
colnames(skw_aphid_count_spearman)<-c("realskw","p_left","p_right")
rownames(skw_aphid_count_spearman)<-paste("sp",splist,sep="")
pdf("./Results/skewness_results/skewness_aphid_count/aphid_count_allsp_skewness_spearman.pdf",height=25,width=25)
op<-par(mfrow=c(4,5),mar=c(5,5,1,1),mgp=c(3,1,0))
cat("Now for spearman preserving surrogs : aphid count","\n")
for(i in splist){
  cat("sp=",i,"\n")
  loclist<-good_loclist(d_allsp=data_aphid_count,sp=i,data_pt_thrs=30)
  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_aphid_count),loclist=loclist,
                        numsurrog=numsurrog,
                        ploton=T,
                        corpres="spearman") 
  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
  p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
  skw_aphid_count_spearman$realskw[i]<-ans$realskw
  skw_aphid_count_spearman$p_left[i]<-p_left
  skw_aphid_count_spearman$p_right[i]<-p_right
  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
}
par(op)
dev.off()
saveRDS(skw_aphid_count_spearman,"./Results/skewness_results/skewness_aphid_count/skw_aphid_count_spearman.RDS")

#skw_aphid_count_kendall<-as.data.frame(matrix(NA,nrow=length(splist),ncol=3))
#colnames(skw_aphid_count_kendall)<-c("realskw","p_left","p_right")
#rownames(skw_aphid_count_kendall)<-paste("sp",splist,sep="")
#pdf("./Results/skewness_results/skewness_aphid_count/aphid_count_allsp_skewness_kendall.pdf",height=25,width=25)
#op<-par(mfrow=c(4,5),mar=c(5,5,1,1),mgp=c(3,1,0))
#cat("Now for kendall preserving surrogs : aphid count","\n")
#for(i in splist){
#  cat("sp=",i,"\n")
#  loclist<-good_loclist(d_allsp=data_aphid_count,sp=i,data_pt_thrs=30)
#  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_aphid_count),loclist=loclist,
#                        numsurrog=numsurrog,
#                        ploton=T,
#                        corpres="kendall") 
#  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
# p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
#  skw_aphid_count_kendall$realskw[i]<-ans$realskw
#  skw_aphid_count_kendall$p_left[i]<-p_left
#  skw_aphid_count_kendall$p_right[i]<-p_right
#  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
#}
#par(op)
#dev.off()
#saveRDS(skw_aphid_count_kendall,"./Results/skewness_results/skewness_aphid_count/skw_aphid_count_kendall.RDS")
```

```{r skewness_aphidff_allsp_plot, echo=F, results="hide", warning=F, cache=T, cache.extra=list(seed,data_aphid_ff,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}

# For sp =12 and sp=16 : kendall preserving surrogates gives warning : Error in ncsurrog: ncov is not positive semidefinite,
# also it can't handle ties for "kendall" and code crashes
# you can see those msg by deleting results="hide" and warning=F option in r chunk

source("good_loclist.R")
source("skewness_multivar_dataset.R")
set.seed(seed)

numsurrog<-10000

splist<-c(1:20)
skw_aphid_ff_spearman<-as.data.frame(matrix(NA,nrow=length(splist),ncol=3))
colnames(skw_aphid_ff_spearman)<-c("realskw","p_left","p_right")
rownames(skw_aphid_ff_spearman)<-paste("sp",splist,sep="")
pdf("./Results/skewness_results/skewness_aphid_ff/aphid_ff_allsp_skewness_spearman.pdf",height=25,width=25)
op<-par(mfrow=c(4,5),mar=c(5,5,1,1),mgp=c(3,1,0))
cat("Now for spearman preserving surrogs : aphid ff","\n")
for(i in splist){
  cat("sp=",i,"\n")
  loclist<-good_loclist(d_allsp=data_aphid_ff,sp=i,data_pt_thrs=30)
  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_aphid_ff),loclist=loclist,
                        numsurrog=numsurrog,
                        ploton=T,
                        corpres="spearman") 
  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
  p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
  skw_aphid_ff_spearman$realskw[i]<-ans$realskw
  skw_aphid_ff_spearman$p_left[i]<-p_left
  skw_aphid_ff_spearman$p_right[i]<-p_right
  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
}
par(op)
dev.off()
saveRDS(skw_aphid_ff_spearman,"./Results/skewness_results/skewness_aphid_ff/skw_aphid_ff_spearman.RDS")

#skw_aphid_ff_kendall<-as.data.frame(matrix(NA,nrow=length(splist),ncol=3))
#colnames(skw_aphid_ff_kendall)<-c("realskw","p_left","p_right")
#rownames(skw_aphid_ff_kendall)<-paste("sp",splist,sep="")
#pdf("./Results/skewness_results/skewness_aphid_ff/aphid_ff_allsp_skewness_kendall.pdf",height=25,width=25)
#op<-par(mfrow=c(4,5),mar=c(5,5,1,1),mgp=c(3,1,0))
#cat("Now for kendall preserving surrogs : aphid ff","\n")
#for(i in splist){
#  cat("sp=",i,"\n")
#  loclist<-good_loclist(d_allsp=data_aphid_ff,sp=i,data_pt_thrs=30)
#  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_aphid_ff),loclist=loclist,
#                        numsurrog=numsurrog,
#                        ploton=T,
#                        corpres="kendall") 
#  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
#  p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
#  skw_aphid_ff_kendall$realskw[i]<-ans$realskw
#  skw_aphid_ff_kendall$p_left[i]<-p_left
#  skw_aphid_ff_kendall$p_right[i]<-p_right
#  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
#}
#par(op)
#dev.off()
#saveRDS(skw_aphid_ff_kendall,"./Results/skewness_results/skewness_aphid_ff/skw_aphid_ff_kendall.RDS")
```

```{r skewness_plankton_allsp_plot, echo=F, results="hide", cache=T, cache.extra=list(seed,data_plankton_north_sea,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}

# There is no error msg or warning coming from ncsurrog : ncov semidefinite matrix for any species with spearman or kendall preserving surrogates.
# But there may be error handling ties with "kendall"
# So,for safety, now we will only analyze with "spearman"

source("good_loclist.R")
source("skewness_multivar_dataset.R")
set.seed(seed)

splist<-c(1:22)
skw_plankton_spearman<-as.data.frame(matrix(NA,nrow=length(splist),ncol=3))
colnames(skw_plankton_spearman)<-c("realskw","p_left","p_right")
rownames(skw_plankton_spearman)<-paste("sp",splist,sep="")
#skw_plankton_kendall<-skw_plankton_spearman
pdf("./Results/skewness_results/skewness_plankton_north_sea/plankton_allsp_skewness_spearman.pdf",height=30,width=30)
op<-par(mfrow=c(4,6),mar=c(5,5,1,1),mgp=c(3,1,0))
#cat("Now for spearman preserving surrogs : pns","\n")
for(i in splist){
  #cat("sp=",i,"\n")
  loclist<-good_loclist(d_allsp=data_plankton_north_sea,sp=i,data_pt_thrs=45)
  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_plankton_north_sea),loclist=loclist,
                        numsurrog=numsurrog,
                        ploton=T,
                        corpres="spearman") 
  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
  p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
  skw_plankton_spearman$realskw[i]<-ans$realskw
  skw_plankton_spearman$p_left[i]<-p_left
  skw_plankton_spearman$p_right[i]<-p_right
  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
}
par(op)
dev.off()
saveRDS(skw_plankton_spearman,"./Results/skewness_results/skewness_plankton_north_sea/skw_plankton_spearman.RDS")

#pdf("./Results/skewness_results/skewness_plankton_north_sea/plankton_allsp_skewness_kendall.pdf",height=30,width=30)
#op<-par(mfrow=c(4,6),mar=c(5,5,1,1),mgp=c(3,1,0))
#cat("Now for kendall preserving surrogs : pns","\n")
#for(i in splist){
  #cat("sp=",i,"\n")
 # loclist<-good_loclist(d_allsp=data_plankton_north_sea,sp=i,data_pt_thrs=45)
#  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_plankton_north_sea),loclist=loclist,
#                        numsurrog=numsurrog,
#                        ploton=T,
#                        corpres="kendall") 
#  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
#  p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
#  skw_plankton_kendall$realskw[i]<-ans$realskw
#  skw_plankton_kendall$p_left[i]<-p_left
#  skw_plankton_kendall$p_right[i]<-p_right
#  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
#}
#par(op)
#dev.off()
#saveRDS(skw_plankton_kendall,"./Results/skewness_results/skewness_plankton_north_sea/skw_plankton_kendall.RDS")
```

```{r skewness_count_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}

# Only done for "Spearman" method

source("good_loclist.R")
source("skewness_multivar_dataset.R")
set.seed(seed)

numsurrog<-10000
loclist<-good_loclist(d_allsp=data_aphid_count,sp=10,data_pt_thrs=30)

#pdf("./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10_kendall.pdf",height=6,width=8)
#op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
#answer1k<-skewness_testing(ts_matrix=sp_data(sp=10,d_allsp=data_aphid_count),loclist=loclist,numsurrog=numsurrog,ploton=T,
#                          corpres="kendall") 
#p_left<-sum(answer1k$surrogskw < answer1k$realskw)/numsurrog
#mtext(paste0("p = ",p_left),side=3,line=-0.5,cex = 2)
#par(op)
#dev.off()

pdf("./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10_spearman.pdf",height=6,width=8)
op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
answer1s<-skewness_testing(ts_matrix=sp_data(sp=10,d_allsp=data_aphid_count),loclist=loclist,numsurrog=numsurrog,ploton=T,
                          corpres="spearman")

p_left<-sum(answer1s$surrogskw < answer1s$realskw)/numsurrog
mtext(paste0("p = ",p_left),side=3,line=-0.5,cex = 2)
par(op)
dev.off()
```

```{r skewness_ff_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,data_aphid_ff,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}

# Only done for "Spearman" method

source("good_loclist.R")
source("skewness_multivar_dataset.R")
set.seed(seed)

numsurrog<-10000
loclist<-good_loclist(d_allsp=data_aphid_ff,sp=11,data_pt_thrs=30)

#pdf("./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11_kendall.pdf",height=6,width=8)
#op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
#answer2k<-skewness_testing(ts_matrix=sp_data(sp=11,d_allsp=data_aphid_ff),loclist=loclist,numsurrog=numsurrog,ploton=T,
#                          corpres="kendall") 
#p_right<-sum(answer2k$surrogskw > answer2k$realskw)/numsurrog
#mtext(paste0("p = ",p_right),side=3,line=-0.5,cex = 2)
#par(op)
#dev.off()

pdf("./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11_spearman.pdf",height=6,width=8)
op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
answer2s<-skewness_testing(ts_matrix=sp_data(sp=11,d_allsp=data_aphid_ff),loclist=loclist,numsurrog=numsurrog,ploton=T,
                          corpres="spearman") 
p_right<-sum(answer2s$surrogskw > answer2s$realskw)/numsurrog
mtext(paste0("p = ",p_right),side=3,line=-0.5,cex = 2)
par(op)
dev.off()
```

```{r skewness_pns_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,data_plankton_north_sea,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}

# Only done for "Spearman" method

source("good_loclist.R")
source("skewness_multivar_dataset.R")
set.seed(seed)

numsurrog<-10000
loclist<-good_loclist(d_allsp=data_plankton_north_sea,sp=16,data_pt_thrs=45)

#pdf("./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16_kendall.pdf",height=6,width=8)
#op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
#answer3k<-skewness_testing(ts_matrix=sp_data(sp=16,d_allsp=data_plankton_north_sea),loclist=loclist,numsurrog=numsurrog,ploton=T,
#                          corpres="kendall") 
#p_left<-sum(answer3k$surrogskw < answer3k$realskw)/numsurrog
#mtext(paste0("p = ",p_left),side=3,line=-0.5,cex = 2)
#par(op)
#dev.off()

pdf("./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16_spearman.pdf",height=6,width=8)
op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
answer3s<-skewness_testing(ts_matrix=sp_data(sp=16,d_allsp=data_plankton_north_sea),loclist=loclist,numsurrog=numsurrog,ploton=T,
                          corpres="spearman") 
p_left<-sum(answer3s$surrogskw < answer3s$realskw)/numsurrog
mtext(paste0("p = ",p_left),side=3,line=-0.5,cex = 2)
par(op)
dev.off()
```

```{r skewness_methane_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,data_plankton_north_sea,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}

# Only done for "Spearman" method

source("good_loclist.R")
source("skewness_multivar_dataset.R")
set.seed(seed)

numsurrog<-10000
loclist<-good_loclist(d_allsp=methane_data,sp=1,data_pt_thrs=50)

#pdf("./Results/skewness_results/skewness_methane/skewness_methane_kendall.pdf",height=6,width=8)
#op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
#answer1k<-skewness_testing(ts_matrix=sp_data(sp=1,d_allsp=methane_data),loclist=loclist,numsurrog=numsurrog,ploton=T,
#                           corpres="kendall") 
#p_left<-sum(answer1k$surrogskw < answer1k$realskw)/numsurrog
#mtext(paste0("p = ",p_left),side=3,line=-0.5,cex = 2)
#par(op)
#dev.off()

pdf("./Results/skewness_results/skewness_methane/skewness_methane_spearman.pdf",height=6,width=8)
op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
answer1s<-skewness_testing(ts_matrix=sp_data(sp=1,d_allsp=methane_data),loclist=loclist,numsurrog=numsurrog,ploton=T,
                           corpres="spearman") 
p_left<-sum(answer1s$surrogskw < answer1s$realskw)/numsurrog
mtext(paste0("p = ",p_left),side=3,line=-0.5,cex = 2)
par(op)
dev.off()
```

<!-- Taylor's law section moved to Paper.Rmd -->

<!--make surrogate datasets for TL analysis to follow-->
```{r taylors_law_make_surrogs, echo=F, cache=T, cache.extra=list(seed,mtime("copsurrognd.R"))}
source("copsurrognd.R")
set.seed(seed)

#***constants
N<-50 #length of time series
n<-25 #number of sampling locations
#numdat<-10 #the number of datasets to get from the normal copula
numdat<-1000 #the number of datasets to get from the normal copula
rho<-0.7 #normal-copula parameter for dependence between locations
          #this is a covariance, see below call to rmvnorm

#***generate the starting data, which has normal copula structure
#because it comes originally from a multivariate normal, before 
#transformation
sig<-matrix(rho,n,n)
diag(sig)<-1
m<-rmvnorm(N*numdat,mean=rep(0,n),sigma=sig) # each column of m has normal histogram
m<-array(m,c(N,numdat,n))
m<-aperm(m,c(1,3,2)) # 2nd dim is normally distributed e.g. hist(m[,1,])
gshape<-7.5
gscale<-1 #gamma parameters

cop_to_marg_trans<-function(x){
  return(qgamma(x,shape=gshape,scale=gscale)) #this is where we choose the marginals,
                                  #should be something skewed
}

marg_to_cop_trans<-function(x){ #transforms to copula space
  return(pgamma(x,shape=gshape,scale=gscale))
}

# pnorm(m) makes normal to uniform
# then qgamma(...) makes uniform to gamma distributed
m<-cop_to_marg_trans(pnorm(m))

#***surrogates using a Clayton copula, kendall preserving

#get the Clayton parameter to use
ncop<-normalCopula(rho,2)
ccop<-claytonCopula(2,2)
cparam<-iTau(ccop,tau(ncop))
ccop<-claytonCopula(cparam,n)

#generate the surrogates
csurrog<-array(NA,dim(m))
for (counter in 1:dim(m)[3]){
  csurrog[,,counter]<-copsurrognd(m[,,counter],targetcop=ccop,numsurrog=1)
}

# #test to see whether kendalls are similar between m and csurrog - 
#only works at all as a test if N is large (I used 1000)
# cor(m[,1,1],m[,2,1],method='kendall')
# cor(csurrog[,1,1],csurrog[,2,1],method='kendall')
# cor(m[,3,1],m[,4,1],method='kendall')
# cor(csurrog[,3,1],csurrog[,4,1],method='kendall')
# cor(m[,5,1],m[,6,1],method='kendall')
# cor(csurrog[,5,1],csurrog[,6,1],method='kendall')
# 
# cor(m[,1,2],m[,2,2],method='kendall')
# cor(csurrog[,1,2],csurrog[,2,2],method='kendall')
# cor(m[,3,2],m[,4,2],method='kendall')
# cor(csurrog[,3,2],csurrog[,4,2],method='kendall')
# cor(m[,5,2],m[,6,2],method='kendall')
# cor(csurrog[,5,2],csurrog[,6,2],method='kendall')

#***surrogates using a Frank copula, kendall preserving
#get the Frank parameter to use
fcop<-frankCopula(4,2)
fparam<-iTau(fcop,tau(ncop))
fcop<-frankCopula(fparam,n)

#generate the surrogates
fsurrog<-array(NA,dim(m))
for (counter in 1:dim(m)[3]){
  fsurrog[,,counter]<-copsurrognd(m[,,counter],targetcop=fcop,numsurrog=1)
}

#***surrogates using a survival Clayton, kendall preserving
#take the clayton surrogates and transform them to the 
#copula space, then flip, then transform back
scsurrog<-cop_to_marg_trans((1-marg_to_cop_trans(csurrog)))

# check : plot(csurrog[,1,1],csurrog[,2,1]) and plot(scsurrog[,1,1],scsurrog[,2,1])

# #test: should have identical taus as csurrog
# cor(csurrog[,1,1],csurrog[,2,1],meth='kendall')
# cor(scsurrog[,1,1],scsurrog[,2,1],meth='kendall')
# cor(csurrog[,1,2],csurrog[,2,2],meth='kendall')
# cor(scsurrog[,1,2],scsurrog[,2,2],meth='kendall')

kend.surrog<-list(cla=csurrog,nor=m,fra=fsurrog,scl=scsurrog)

#***surrogates using a Clayton copula, spearman preserving
#get the Clayton parameter to use
ccop<-claytonCopula(2,2)
cparam<-iRho(ccop,rho(ncop))
ccop<-claytonCopula(cparam,n)

#generate the surrogates
csurrog<-array(NA,dim(m))
for (counter in 1:dim(m)[3]){
  csurrog[,,counter]<-copsurrognd(m[,,counter],targetcop=ccop,numsurrog=1)
}

#***surrogates using a Frank copula, spearman preserving
#get the Frank parameter to use
fcop<-frankCopula(4,2)
fparam<-iRho(fcop,rho(ncop))
fcop<-frankCopula(fparam,n)

#generate the surrogates
fsurrog<-array(NA,dim(m))
for (counter in 1:dim(m)[3]){
  fsurrog[,,counter]<-copsurrognd(m[,,counter],targetcop=fcop,numsurrog=1)
}

#***surrogates using a survival Clayton, spearman preserving
#take the clayton surrogates and transform them to the 
#copula space, then flip, then transform back
scsurrog<-cop_to_marg_trans((1-marg_to_cop_trans(csurrog)))

spea.surrog<-list(cla=csurrog,nor=m,fra=fsurrog,scl=scsurrog)
```

<!--do a TL analysis on the data just created-->
```{r tl_analysis, echo=F, cache=T, cache.extra=list(kend.surrog,spea.surrog,mtime("mncurvat.R"),mtime("TaylorsLawTools.R"),seed,mtime("copsurrognd.R"))}
source("copsurrognd.R")
source("TaylorsLawTools.R")

#We are only doing spatial taylor's law because we know a priori temporal
#won't give anything interesting for the surrogates we use

kend.stats.res<-list()
kend.points.res<-list()
spea.stats.res<-list()
spea.points.res<-list()
for (counter in 1:length(kend.surrog)){
 ks<-kend.surrog[[counter]]
 ss<-spea.surrog[[counter]]
 kend.stats.res[[counter]]<-matrix(NA,7,numdat) # to keep spatial TL stats
 spea.stats.res[[counter]]<-matrix(NA,7,numdat)
 kend.points.res[[counter]]<-array(NA,c(N,2,numdat)) # to keep log(mu), log(var) of spatial TL as a dataframe
 spea.points.res[[counter]]<-array(NA,c(N,2,numdat))
 for (sc in 1:numdat){
   khold<-tl_stats(ks[,,sc])
   kend.stats.res[[counter]][,sc]<-khold$s.stats #spatial TL only
   rownames(kend.stats.res[[counter]])<-names(khold$s.stats)
   kend.points.res[[counter]][,,sc]<-as.matrix(khold$s.points)

   shold<-tl_stats(ss[,,sc])
   spea.stats.res[[counter]][,sc]<-shold$s.stats #spatial TL only
   rownames(spea.stats.res[[counter]])<-names(shold$s.stats)
   spea.points.res[[counter]][,,sc]<-as.matrix(shold$s.points)
 }
}
names(kend.stats.res)<-names(kend.surrog)
names(kend.points.res)<-names(kend.surrog)
names(spea.stats.res)<-names(spea.surrog)
names(spea.points.res)<-names(spea.surrog)
```

<!--Now make some plots, these one for kendall surrogates-->
```{r taylor_plots_kendall, echo=F, results="hide"}
xlimits<-range(kend.points.res[[1]][,1,],kend.points.res[[2]][,1,],kend.points.res[[3]][,1,],kend.points.res[[4]][,1,])
ylimits<-range(kend.points.res[[1]][,2,],kend.points.res[[2]][,2,],kend.points.res[[3]][,2,],kend.points.res[[4]][,2,])

pdf("./Results/taylorslaw_results/SpatialTL_CNFSC_kendall.pdf",height=4,width=16)
op<-op<-par(mfrow=c(1,4),mar=c(4,5,1,1),mgp=c(3,1,0))
##plot TL for clayton
plot(kend.points.res[[1]][,1,],kend.points.res[[1]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
     cex.axis=1.5,cex.lab=1.5)  
points(kend.points.res[[1]][,1,1],kend.points.res[[1]][,2,1],type='p',pch=20,cex=.75,col='red')
points(kend.points.res[[1]][,1,2],kend.points.res[[1]][,2,2],type='p',pch=20,cex=.75,col='green')
mtext(paste("Clayton"),side = 3, line=-2, adj=0.1, col="black")

##plot TL for normal copula
plot(kend.points.res[[2]][,1,],kend.points.res[[2]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
     cex.axis=1.5,cex.lab=1.5)  
points(kend.points.res[[2]][,1,1],kend.points.res[[2]][,2,1],type='p',pch=20,cex=.75,col='red')
points(kend.points.res[[2]][,1,2],kend.points.res[[2]][,2,2],type='p',pch=20,cex=.75,col='green')
mtext(paste("Normal"),side = 3, line=-2, adj=0.1, col="black")

#plot TL for frank
plot(kend.points.res[[3]][,1,],kend.points.res[[3]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
     cex.axis=1.5,cex.lab=1.5)  
points(kend.points.res[[3]][,1,1],kend.points.res[[3]][,2,1],type='p',pch=20,cex=.75,col='red')
points(kend.points.res[[3]][,1,2],kend.points.res[[3]][,2,2],type='p',pch=20,cex=.75,col='green')
mtext(paste("Frank"),side = 3, line=-2, adj=0.1, col="black")

#plot TL for survival clayton
plot(kend.points.res[[4]][,1,],kend.points.res[[4]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
     cex.axis=1.5,cex.lab=1.5)  
points(kend.points.res[[4]][,1,1],kend.points.res[[4]][,2,1],type='p',pch=20,cex=.75,col='red')
points(kend.points.res[[4]][,1,2],kend.points.res[[4]][,2,2],type='p',pch=20,cex=.75,col='green')
mtext(paste("Survival Clayton"),side=3, line=-2, adj=0.1, col="black")

par(op)
dev.off()
#---------------------------------------------------
#prepare for plots coming below
s.lin<-cbind(kend.stats.res[[1]][1,],kend.stats.res[[2]][1,],kend.stats.res[[3]][1,],kend.stats.res[[4]][1,])
s.hom<-cbind(kend.stats.res[[1]][2,],kend.stats.res[[2]][2,],kend.stats.res[[3]][2,],kend.stats.res[[4]][2,])
s.rmse<-cbind(kend.stats.res[[1]][3,],kend.stats.res[[2]][3,],kend.stats.res[[3]][3,],kend.stats.res[[4]][3,])
s.int<-cbind(kend.stats.res[[1]][4,],kend.stats.res[[2]][4,],kend.stats.res[[3]][4,],kend.stats.res[[4]][4,])
s.slope<-cbind(kend.stats.res[[1]][5,],kend.stats.res[[2]][5,],kend.stats.res[[3]][5,],kend.stats.res[[4]][5,])
s.quadcoeff<-cbind(kend.stats.res[[1]][6,],kend.stats.res[[2]][6,],kend.stats.res[[3]][6,],kend.stats.res[[4]][6,])
s.mncurv<-cbind(kend.stats.res[[1]][7,],kend.stats.res[[2]][7,],kend.stats.res[[3]][7,],kend.stats.res[[4]][7,])

pdf("./Results/taylorslaw_results/SpatialTL_7stats_kendall.pdf",height=3,width=21)
op<-par(mfrow=c(1,7),mar=c(4.2,5,1,1),mgp=c(3,1,0))
#***DAN: x axis needs to be redone to show the copula names used in the surrogates

#x<-1:4
x2<-c("C", "N", "F", "SC")
#plot(x,s.lin[1,],type="l",col="grey",ylim=range(s.lin),xlab="Set of surrogates",ylab="Linearity test p-vals",
#     cex.axis=1.5,cex.lab=1.5,xaxt="n")
#points(x,s.lin[1,],col="red")
#axis(1, at = c(1:4), x2, cex.axis=1.5)
#for (counter in 2:dim(s.lin)[1]){
#  lines(x,s.lin[counter,],col="grey")
#  points(x,s.lin[counter,],col="red")
#}
boxplot(s.lin , col=rainbow(4), xlab="Copula family", ylab="Linearity test p-vals", cex.axis=1.5,cex.lab=1.5,xaxt="n")
axis(1, at = c(1:4), x2, cex.axis=1.5)

#plot s.hom distr for each set of surrogates
boxplot(s.hom , col=rainbow(4), xlab="Copula family", ylab="Homoskedasticity test p-vals", cex.axis=1.5,cex.lab=1.5,xaxt="n")
axis(1, at = c(1:4), x2, cex.axis=1.5)

#plot s.rmse distr for each set of surrogates
boxplot(s.rmse , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL rmse", cex.axis=1.5,cex.lab=1.5,xaxt="n")
axis(1, at = c(1:4), x2, cex.axis=1.5)

#plot s.int distr for each set of surrogates
boxplot(s.int , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL intercept", cex.axis=1.5,cex.lab=1.5,xaxt="n")
axis(1, at = c(1:4), x2, cex.axis=1.5)

#plot s.slope distr for each set of surrogates
boxplot(s.slope , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL slope", cex.axis=1.5,cex.lab=1.5,xaxt="n")
axis(1, at = c(1:4), x2, cex.axis=1.5)

#plot s.quadcoeff distr for each set of surrogates
boxplot(s.quadcoeff , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL quadratic coefficient", cex.axis=1.5,cex.lab=1.5,xaxt="n")
axis(1, at = c(1:4), x2, cex.axis=1.5)

#plot s.mncurv distr for each set of surrogates
boxplot(s.mncurv , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL mean curvature", cex.axis=1.5,cex.lab=1.5,xaxt="n")
axis(1, at = c(1:4), x2, cex.axis=1.5)

par(op)
dev.off()
```

<!--Now make some plots, these one for spearman surrogates-->
```{r taylor_plots_spearman, echo=F, results="hide"}
xlimits<-range(spea.points.res[[1]][,1,],spea.points.res[[2]][,1,],spea.points.res[[3]][,1,],spea.points.res[[4]][,1,])
ylimits<-range(spea.points.res[[1]][,2,],spea.points.res[[2]][,2,],spea.points.res[[3]][,2,],spea.points.res[[4]][,2,])

pdf("./Results/taylorslaw_results/SpatialTL_CNFSC_spearman.pdf",height=4,width=16)
op<-op<-par(mfrow=c(1,4),mar=c(4,5,1,1),mgp=c(3,1,0))
##plot TL for clayton
plot(spea.points.res[[1]][,1,],spea.points.res[[1]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
     cex.axis=1.5,cex.lab=1.5)  
points(spea.points.res[[1]][,1,1],spea.points.res[[1]][,2,1],type='p',pch=20,cex=.75,col='red')
points(spea.points.res[[1]][,1,2],spea.points.res[[1]][,2,2],type='p',pch=20,cex=.75,col='green')
mtext(paste("Clayton"),side = 3, line=-2, adj=0.1, col="black")

##plot TL for normal copula
plot(spea.points.res[[2]][,1,],spea.points.res[[2]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
     cex.axis=1.5,cex.lab=1.5)  
points(spea.points.res[[2]][,1,1],spea.points.res[[2]][,2,1],type='p',pch=20,cex=.75,col='red')
points(spea.points.res[[2]][,1,2],spea.points.res[[2]][,2,2],type='p',pch=20,cex=.75,col='green')
mtext(paste("Normal"),side = 3, line=-2, adj=0.1, col="black")

#plot TL for frank
plot(spea.points.res[[3]][,1,],spea.points.res[[3]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
     cex.axis=1.5,cex.lab=1.5)  
points(spea.points.res[[3]][,1,1],spea.points.res[[3]][,2,1],type='p',pch=20,cex=.75,col='red')
points(spea.points.res[[3]][,1,2],spea.points.res[[3]][,2,2],type='p',pch=20,cex=.75,col='green')
mtext(paste("Frank"),side = 3, line=-2, adj=0.1, col="black")

#plot TL for survival clayton
plot(spea.points.res[[4]][,1,],spea.points.res[[4]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
     cex.axis=1.5,cex.lab=1.5)  
points(spea.points.res[[4]][,1,1],spea.points.res[[4]][,2,1],type='p',pch=20,cex=.75,col='red')
points(spea.points.res[[4]][,1,2],spea.points.res[[4]][,2,2],type='p',pch=20,cex=.75,col='green')
mtext(paste("Survival Clayton"),side=3, line=-2, adj=0.1, col="black")

par(op)
dev.off()
#---------------------------------------------------
#prepare for plots coming below
s.lin<-cbind(spea.stats.res[[1]][1,],spea.stats.res[[2]][1,],spea.stats.res[[3]][1,],spea.stats.res[[4]][1,])
s.hom<-cbind(spea.stats.res[[1]][2,],spea.stats.res[[2]][2,],spea.stats.res[[3]][2,],spea.stats.res[[4]][2,])
s.rmse<-cbind(spea.stats.res[[1]][3,],spea.stats.res[[2]][3,],spea.stats.res[[3]][3,],spea.stats.res[[4]][3,])
s.int<-cbind(spea.stats.res[[1]][4,],spea.stats.res[[2]][4,],spea.stats.res[[3]][4,],spea.stats.res[[4]][4,])
s.slope<-cbind(spea.stats.res[[1]][5,],spea.stats.res[[2]][5,],spea.stats.res[[3]][5,],spea.stats.res[[4]][5,])
s.quadcoeff<-cbind(spea.stats.res[[1]][6,],spea.stats.res[[2]][6,],spea.stats.res[[3]][6,],spea.stats.res[[4]][6,])
s.mncurv<-cbind(spea.stats.res[[1]][7,],spea.stats.res[[2]][7,],spea.stats.res[[3]][7,],spea.stats.res[[4]][7,])

pdf("./Results/taylorslaw_results/SpatialTL_7stats_spearman.pdf",height=3,width=21)
op<-par(mfrow=c(1,7),mar=c(4.2,5,1,1),mgp=c(3,1,0))
x2<-c("C", "N", "F", "SC")

#plot s.lin distr for each set of surrogates
boxplot(s.lin , col=rainbow(4), xlab="Copula family", ylab="Linearity test p-vals", cex.axis=1.5,cex.lab=1.5,xaxt="n")
axis(1, at = c(1:4), x2, cex.axis=1.5)

#plot s.hom distr for each set of surrogates
boxplot(s.hom , col=rainbow(4), xlab="Copula family", ylab="Homoskedasticity test p-vals", cex.axis=1.5,cex.lab=1.5,xaxt="n")
axis(1, at = c(1:4), x2, cex.axis=1.5)

#plot s.rmse distr for each set of surrogates
boxplot(s.rmse , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL rmse", cex.axis=1.5,cex.lab=1.5,xaxt="n")
axis(1, at = c(1:4), x2, cex.axis=1.5)

#plot s.int distr for each set of surrogates
boxplot(s.int , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL intercept", cex.axis=1.5,cex.lab=1.5,xaxt="n")
axis(1, at = c(1:4), x2, cex.axis=1.5)

#plot s.slope distr for each set of surrogates
boxplot(s.slope , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL slope", cex.axis=1.5,cex.lab=1.5,xaxt="n")
axis(1, at = c(1:4), x2, cex.axis=1.5)

#plot s.quadcoeff distr for each set of surrogates
boxplot(s.quadcoeff , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL quadratic coefficient", cex.axis=1.5,cex.lab=1.5,xaxt="n")
axis(1, at = c(1:4), x2, cex.axis=1.5)

#plot s.mncurv distr for each set of surrogates
boxplot(s.mncurv , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL mean curvature", cex.axis=1.5,cex.lab=1.5,xaxt="n")
axis(1, at = c(1:4), x2, cex.axis=1.5)

par(op)
dev.off()
```

\begin{figure}[!h]
\textbf{ \hspace{2.8 cm} (A) \hspace{3.6 cm} (B) \hspace{3.8 cm} (C) \hspace{3.6 cm} (D)} \\
\vspace{-0.5 cm} 
\begin{center}
\includegraphics[width=18 cm]{./Results/taylorslaw_results/SpatialTL_CNFSC_spearman.pdf}\\
\end{center}
\textbf{ \hspace{1.8 cm} (E) \hspace{1.6 cm} (F) \hspace{1.8 cm} (G) \hspace{1.8 cm} (H) \hspace{1.8 cm} (I) \hspace{1.8 cm} (J) \hspace{1.8 cm} (K)}\\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=18 cm]{./Results/taylorslaw_results/SpatialTL_7stats_spearman.pdf}
\caption[Spatial Taylor's Law results]{Spatial Taylor's Law results: log(spatial variance) vs. log(spatial mean) for $ns=1000$ number of (A) Clayton, (B) Normal, (C) Frank, (D) Survival Clayton surrogates; (E) p-value for test of linearity, (F) p-value for test of homoskedasticity, (G) root mean square error, (H) intercept, (I) slope, (J) quadratic coefficient, (K) mean curvature of regression line from first panel-plots. \label{fig_spTL}}
\end{center}
\end{figure}

```{r spatialTL_on_realdata_aphid_count, results="hide", echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("good_loclist.R"),mtime("TaylorsLawTools.R"),mtime("mncurvat.R"),mtime("ncsurrog.R"))}

set.seed(seed)
source("ncsurrog.R")
source("TaylorsLawTools.R")
source("good_loclist.R")

TL_on_realdata<-function(data,numsurrog=1000,sp,data_pt_thrs,ploton,tag,labelon=T){

  loclist<-good_loclist(d_allsp=data,sp=sp,data_pt_thrs=data_pt_thrs)
  data<-data[[sp]]
  df<-c()
  for(i in loclist){
    tem<-data[[i]]$Dat
    df<-cbind(df,tem)
  }
  colnames(df)<-paste("loc",loclist,sep="")
  df<-na.omit(df)

  mat_real<-as.matrix(df)
  check_rep<-function(y){all( abs(y - mean(y)) < 1e-8 )} #This fn checks for repeatation of same values in a vector
  call_check_rep<-apply(mat_real,1,check_rep) # checks the same for mat_real matrix along all row

  if(all(call_check_rep==F)==T){

    realTL_stats<-tl_stats(m = mat_real)
    real_spatialTLslope<-unname(realTL_stats$s.stats[5])
    real_sTL_quadcoeff<-unname(realTL_stats$s.stats[6])
    real_sTL_mncurv<-unname(realTL_stats$s.stats[7])

    #generate the normal surrogates with spearman correlation preserving
    array_surrog<-ncsurrog(m=mat_real,corpres="spearman",numsurrog=numsurrog)

    bad_id<-c()
    for(i in 1:numsurrog){
      m<-array_surrog[,,i]
      mean_m<-apply(X=m, MARGIN = 1, FUN = mean)
      var_m<-apply(X=m, MARGIN = 1, FUN = var)
      if(any(mean_m==0) | any(var_m==0)){
        bad_id<-c(bad_id,i)
      }
    }

    m<-array_surrog
    if(length(bad_id)!=0){
      m<-array_surrog[,,-bad_id]
    }
    ts<-apply(m,3,tl_stats)

    good_surrog_slope<-c()
    good_surrog_quadcoeff<-c()
    good_surrog_mncurv<-c()
    for(i in c(1:length(ts))){
      good_surrog_slope<-c(good_surrog_slope,unname(ts[[i]]$s.stats[5]))
      good_surrog_quadcoeff<-c(good_surrog_quadcoeff,unname(ts[[i]]$s.stats[6]))
      good_surrog_mncurv<-c(good_surrog_mncurv,unname(ts[[i]]$s.stats[7]))
    }

    numsurrog_good<- numsurrog - length(bad_id)

    p_left<-sum(good_surrog_slope < real_spatialTLslope)/numsurrog_good
    p_right<-sum(good_surrog_slope > real_spatialTLslope)/numsurrog_good

    p_left_quadcoeff<-sum(good_surrog_quadcoeff < real_sTL_quadcoeff)/numsurrog_good
    p_right_quadcoeff<-sum(good_surrog_quadcoeff > real_sTL_quadcoeff)/numsurrog_good

    p_left_mncurv<-sum(good_surrog_mncurv < real_sTL_mncurv)/numsurrog_good
    p_right_mncurv<-sum(good_surrog_mncurv > real_sTL_mncurv)/numsurrog_good

    if(ploton==T & tag=="slope"){
      hist(good_surrog_slope,breaks=100,main="",xlab="surrogate_TL_slope",cex.axis=1.5,cex.lab=1.5)
      points(x=real_spatialTLslope,y=0,col="red",pch=15)
      if(labelon==T){
        mtext(paste0("sp= ",sp,", p_left= ",round(p_left,3),", p_right=",round(p_right,3),sep=""),
              side = 3, line=-2, adj=0, col="blue")
      }
    }

    if(ploton==T & tag=="quadcoeff"){
      hist(good_surrog_quadcoeff,breaks=100,main="",xlab="surrogate_TL_quadcoeff",cex.axis=1.5,cex.lab=1.5)
      points(x=real_sTL_quadcoeff,y=0,col="red",pch=15)
      if(labelon==T){
        mtext(paste0("sp= ",sp,", p_left= ",round(p_left_quadcoeff,3),", p_right=",round(p_right_quadcoeff,3),sep=""),
              side = 3, line=-2, adj=0, col="blue")
      }
    }

    if(ploton==T & tag=="mncurv"){
      hist(good_surrog_mncurv,breaks=100,main="",xlab="surrogate_TL_meanCurvature",cex.axis=1.5,cex.lab=1.5)
      points(x=real_sTL_mncurv,y=0,col="red",pch=15)
      if(labelon==T){
        mtext(paste0("sp= ",sp,", p_left= ",round(p_left_mncurv,3),", p_right=",round(p_right_mncurv,3),sep=""),
              side = 3, line=-2, adj=0, col="blue")
      }
    }

  }else{
    real_spatialTLslope<-NA
    p_left<-NA
    p_right<-NA
    numsurrog_good<-NA
    real_sTL_quadcoeff<-NA
    p_left_quadcoeff<-NA
    p_right_quadcoeff<-NA
    real_sTL_mncurv<-NA
    p_left_mncurv<-NA
    p_right_mncurv<-NA
  }
  return(list(realTL_stats=realTL_stats,
              real_spatialTLslope=real_spatialTLslope,
              p_left=p_left,
              p_right=p_right,
              numsurrog_good=numsurrog_good,
              real_sTL_quadcoeff=real_sTL_quadcoeff,
              p_left_quadcoeff=p_left_quadcoeff,
              p_right_quadcoeff=p_right_quadcoeff,
              real_sTL_mncurv=real_sTL_mncurv,
              p_left_mncurv=p_left_mncurv,
              p_right_mncurv=p_right_mncurv,
              surrogTL_stats=ts,
              good_surrog_mncurv=good_surrog_mncurv))
}

#call on real data
# For all sp
splist<-c(1:20)
tl_aphid_count<-as.data.frame(matrix(NA,nrow=length(splist),ncol=10))
colnames(tl_aphid_count)<-c("real_sTL_slope","p_left","p_right","numsurrog_good",
                            "real_sTL_quadcoeff","p_left_quadcoeff","p_right_quadcoeff",
                            "real_sTL_mncurv","p_left_mncurv","p_right_mncurv")
rownames(tl_aphid_count)<-paste("sp",splist,sep="")
#pdf("aphid_count_allsp_spTLslope.pdf",height=20,width=20)
#pdf("./Results/taylorslaw_results/aphid_count_allsp_sTL_quadcoeff.pdf",height=20,width=20)
pdf("./Results/taylorslaw_results/aphid_count_allsp_sTL_mncurv.pdf",height=20,width=20)
op<-par(mfrow=c(4,5),mar=c(4.2,5,1,1),mgp=c(3,1,0))
for(i in splist){
  #cat("sp=",i,"\n")
  #ans<-TL_on_realdata(data=data_aphid_count,numsurrog=10000,sp=i,data_pt_thrs = 30,ploton=T,tag="slope",labelon=T)
  #ans<-TL_on_realdata(data=data_aphid_count,numsurrog=10000,sp=i,data_pt_thrs = 30,ploton=T,tag="quadcoeff",labelon=T)
  ans<-TL_on_realdata(data=data_aphid_count,numsurrog=10000,sp=i,data_pt_thrs = 30,ploton=T,tag="mncurv",labelon=T)
  tl_aphid_count$real_sTL_slope[i]<-ans$real_spatialTLslope
  tl_aphid_count$p_left[i]<-ans$p_left
  tl_aphid_count$p_right[i]<-ans$p_right
  tl_aphid_count$numsurrog_good[i]<-ans$numsurrog_good
  tl_aphid_count$real_sTL_quadcoeff[i]<-ans$real_sTL_quadcoeff
  tl_aphid_count$p_left_quadcoeff[i]<-ans$p_left_quadcoeff
  tl_aphid_count$p_right_quadcoeff[i]<-ans$p_right_quadcoeff
  tl_aphid_count$real_sTL_mncurv[i]<-ans$real_sTL_mncurv
  tl_aphid_count$p_left_mncurv[i]<-ans$p_left_mncurv
  tl_aphid_count$p_right_mncurv[i]<-ans$p_right_mncurv
}
par(op)
dev.off()
saveRDS(tl_aphid_count,"./Results/taylorslaw_results/tl_aphid_count.RDS")
```

```{r spatialTL_on_realdata_aphid_count_sp6, results="hide", echo=F, cache=T, cache.extra=list(seed,data_aphid_count,TL_on_realdata,mtime("good_loclist.R"),mtime("TaylorsLawTools.R"),mtime("mncurvat.R"),mtime("ncsurrog.R"))}

set.seed(seed)

#Now for sp=6
sp<-6
pdf("./Results/taylorslaw_results/aphid_count_sp6_sTL_mncurv.pdf",height=5,width=5)
op<-par(mar=c(4,5,1,1),mgp=c(3,1,0))
ans<-TL_on_realdata(data=data_aphid_count,numsurrog=10000,sp=sp,data_pt_thrs = 30,ploton=T,tag="mncurv",labelon = F)
saveRDS(ans,"./Results/taylorslaw_results/tl_aphid_count_sp6.RDS")
mtext(paste0("p= ",ans$p_left_mncurv),side = 3, line=-2, adj=0.5, col="black",cex=2)
par(op)
dev.off()

pdf("./Results/taylorslaw_results/aphid_count_sp6_logv_vs_logm.pdf",height=5,width=10)
op<-op<-par(mfrow=c(1,2),mar=c(4,5,1,1),mgp=c(3,1,0))
plot(ans$realTL_stats$s.points,xlab=expression(log[10]("spatial mean")),
     ylab=expression(log[10]("spatial variance")),col="black",
     cex.axis=1.5,cex.lab=1.5)
mtext(paste0("mean curvature= ",round(unname(ans$realTL_stats$s.stats[7]),3)),side = 3, line=-2, adj=0, col="black",cex=2)

target_mncurv<-0.1
id<-which(abs(ans$good_surrog_mncurv-target_mncurv)<1e-4)[1]
ans_s<-ans$surrogTL_stats
plot(ans_s[[id]]$s.points,xlab=expression(log[10]("spatial mean")),
     ylab=expression(log[10]("spatial variance")),
     cex.axis=1.5,cex.lab=1.5)
mtext(paste0("mean curvature= ",round(ans$good_surrog_mncurv[id],3)),side = 3, line=-2, adj=0, col="black",cex=2)

par(op)
dev.off()
```

```{r fig_ext_risk, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}

#source("mydispersal.R")
#set.seed(seed)

#rlist<-c(0,0.05,0.25,0.5,0.7,0.75,0.9,1.0,1.25,1.5,1.75,1.95)

# plot for local dispersal in D matrix ; r=0 for LC model, r!=0 for Ricker model
#pdf("./Results/ext_risk_local_disp.pdf",heigh=6,width=10)
#op<-par(mfrow=c(3,4),mar=c(5,4,1,1) + 0.2)
#for(r in rlist){
    #cat("r=",r,"\n")
#    s<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=r,K=50,disp_everywhere=F,ploton=F)
#    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
#    lines(s$d_seq,s$risk_right,type="b",col="blue")
#    legend(x=-0.06,y=1.1, c("noise : left","noise : right"), fill=c('red', 'blue'), horiz=T, bty='n')
#    mtext(paste0("r = ", r))
#}
#par(op)
#dev.off()

# plot for global dispersal in D matrix ; r=0 for LC model, r!=0 for Ricker model
#pdf("./Results/ext_risk_global_disp.pdf",heigh=6,width=10)
#op<-par(mfrow=c(3,4),mar=c(5,4,1,1) + 0.2)
#for(r in rlist){
  #cat("r=",r,"\n")
#  s<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=r,K=50,disp_everywhere=T,ploton=F)
#  plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
#  lines(s$d_seq,s$risk_right,type="b",col="blue")
#  legend(x=-0.06,y=1.1, c("noise : left","noise : right"), fill=c('red', 'blue'), horiz=T, bty='n')
#  mtext(paste0("r = ", r))
#}
#par(op)
#dev.off()

# plot to show the effect of increasing numlocs for r=0 (LC model), r=0.25 (Ricker model)
#pdf("./Results/ext_risk_numlocs_effect_local_disp.pdf",heigh=6,width=6)
#op<-par(mfrow=c(2,2),mar=c(4,4,1,1) + 0.1)
#for(r in c(0,0.25)){
  #cat("r=",r,"\n")
#  s5<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=r,K=50,disp_everywhere=F,ploton=F)
#  s25<-varying_d(numsims=10000,numsteps =25,numlocs=25,r=r,K=50,disp_everywhere=F,ploton=F)

#  plot(s5$d_seq,s5$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
#  lines(s25$d_seq,s25$risk_left,type="b",col="orange")
#  legend("topleft", title=paste0("r=",r," , left tail dep. in noise"), c("numlocs=5","numlocs=25"), fill=c('red', 'orange'),  horiz=F, bty='n')

#  plot(s5$d_seq,s5$risk_right,xlim=c(0,1),ylim=c(0,1),type="b",col="blue",
#       panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
#  lines(s25$d_seq,s25$risk_right,type="b",col="skyblue")
#  legend("topleft",title=paste0("r=",r," , right tail dep. in noise"), c("numlocs=5","numlocs=25"), fill=c('blue', 'skyblue'),  horiz=F, bty='n')
  #mtext(paste0("B",1))
#}
#par(op)
#dev.off()

# plot to show the effect of increasing numlocs for r=0 (LC model), r=0.25 (Ricker model)
#pdf("./Results/ext_risk_numlocs_effect_global_disp.pdf",heigh=6,width=6)
#op<-par(mfrow=c(2,2),mar=c(4,4,1,1) + 0.1)
#for(r in c(0,0.25)){
  #cat("r=",r,"\n")
#  s5<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=r,K=50,disp_everywhere=T,ploton=F)
#  s25<-varying_d(numsims=10000,numsteps =25,numlocs=25,r=r,K=50,disp_everywhere=T,ploton=F)

#  plot(s5$d_seq,s5$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
#  lines(s25$d_seq,s25$risk_left,type="b",col="orange")
#  legend("topleft", title=paste0("r=",r," , left tail dep. in noise"), c("numlocs=5","numlocs=25"), fill=c('red', 'orange'),  horiz=F, bty='n')

#  plot(s5$d_seq,s5$risk_right,xlim=c(0,1),ylim=c(0,1),type="b",col="blue",
#       panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
#  lines(s25$d_seq,s25$risk_right,type="b",col="skyblue")
#  legend("topleft",title=paste0("r=",r," , right tail dep. in noise"), c("numlocs=5","numlocs=25"), fill=c('blue', 'skyblue'),  horiz=F, bty='n')
  #mtext(paste0("B",1))
#}
#par(op)
#dev.off()

# plot to see the effect of K for r=0.25 (Ricker model)
#pdf("./Results/ext_risk_K_effect.pdf",heigh=6,width=6)
#op<-par(mfrow=c(2,2),mar=c(4,4,1,1) + 0.1)

#s1<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=0.25,K=50,disp_everywhere=F,ploton=F)
#s2<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=0.25,K=200,disp_everywhere=F,ploton=F)

#s3<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=0.25,K=50,disp_everywhere=T,ploton=F)
#s4<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=0.25,K=200,disp_everywhere=T,ploton=F)

#plot(s1$d_seq,s1$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
#lines(s2$d_seq,s2$risk_left,type="b",col="orange")
#legend("topleft",title=("Noise : left tail dep., local disp."), c("K=50","K=200"), fill=c('red', 'orange'),  horiz=F, bty='n')

#plot(s1$d_seq,s1$risk_right,xlim=c(0,1),ylim=c(0,1),type="b",col="blue",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
#lines(s2$d_seq,s2$risk_right,type="b",col="skyblue")
#legend("topleft",title=("Noise : right tail dep., local disp."),c("K=50","K=200"), fill=c('blue', 'skyblue'),  horiz=F, bty='n')

#plot(s3$d_seq,s3$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
#lines(s4$d_seq,s4$risk_left,type="b",col="orange")
#legend("topleft",title= ("Noise : left tail dep., global disp."),c("K=50","K=200"), fill=c('red', 'orange'),  horiz=F, bty='n')

#plot(s3$d_seq,s3$risk_right,xlim=c(0,1),ylim=c(0,1),type="b",col="blue",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
#lines(s4$d_seq,s4$risk_right,type="b",col="skyblue")
#legend("topleft",title=("Noise : right tail dep., global disp."),c("K=50","K=200"), fill=c('blue', 'skyblue'),  horiz=F, bty='n')

#par(op)
#dev.off()
```

<!-- Extinction risk results are not shown currently, it is introduced as a future idea in Discussion-->
<!--
![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Lewton-Cohen model : $r=0$, Ricker model : $r\neq0$) with local dispersion (see Eq. \ref{eq.extrisk_model}); numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp}](./Results/ext_risk_local_disp.pdf){ width=40%, height=40% }

![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Lewton-Cohen model : $r=0$, Ricker model : $r\neq0$) with global dispersion (see Eq. \ref{eq.extrisk_model}); numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp}](./Results/ext_risk_global_disp.pdf){ width=40%, height=40% }

![Increasing patch number allows metapopulation to disperse more efficiently and thereby decresing extinction risk calculated after 25 years; models with local dispersion with numsims=10000 (see Eq. \ref{eq.extrisk_model}).\label{fig_risk_numlocs_effect_local_disp}](./Results/ext_risk_numlocs_effect_local_disp.pdf){ width=40%, height=40% }

![Increasing patch number allows metapopulation to disperse more efficiently and thereby decresing extinction risk calculated after 25 years; models with global dispersion with numsims=10000 (see Eq. \ref{eq.extrisk_model}).\label{fig_risk_numlocs_effect_global_disp}](./Results/ext_risk_numlocs_effect_global_disp.pdf){ width=40%, height=40% }

![Extinction risk as calculated after 25 years on Ricker model (r=0.25, numsims=10000) decreases as carrying capacity $K$ increases for both local and global dispersion (see Eq. \ref{eq.extrisk_model}).\label{fig_risk_K_effect}](./Results/ext_risk_K_effect.pdf){ width=40%, height=40% }
-->

# Surrogate testing with multivariate datasets\label{Surrog_multi}

For multivariate datasets, surrogates were produced as follows:

- Given data $x_i(t)$ for $i=1,\ldots,N$ and $t=1,\ldots,T$, let $\tau_{ij}$ be the Spearman (or Kendall) 
correlation of $x_i(t)$ and $x_j(t)$.
- Calculate the covariance matrix of a multivariate normal distribution with standard-normal
marginals and with pairwise Spearman (or Kendall) correlations equal to the $\tau_{ij}$. This is possible
using the `iRho` function (or the `iTau function) of the `VineCopula` package
- Generate data $a_i(t)$ for $i=1,\ldots,N$ and $t=1,\ldots,T$ by taking $T$ independent draws
from the the multivariate normal distribution of the previous step.
- For any $i,t$ for which the datum $x_i(t)$ was missing, replace $a_i(t)$ by `NA`.
- For each $k$ and $i$ for which $a_i(t)$ is not `NA`, replace the 
$k$th-smallest (non-`NA`) element of the time series 
$a_i(t)$, $t=1,\ldots,T$ by the $k$th smallest (non-`NA`) 
element of $x_i(t)$, $t=1,\ldots,T$.
Call the result $b_i(t)$. This is the surrogate dataset.

\noindent Our code for this algorithm is `ncsurrog` in the `BIVAN` repository 
(\url{https://github.com/sghosh89/BIVAN}). Because, for each $i$, the final time series 
$b_i(t)$ is a permuted version of $x_i(t)$ (and with `NA`s for the same $t$), 
the marginal distributions of the surrogate dataset (i.e., the distributions of 
values for each sampling location, $i$) are exactly the same as for the original data.
Because ranks are preserved in passing from $a_i(t)$ to $b_i(t)$, pairwise Spearman 
(or Kendall) 
correlations for the $b_i(t)$ will be the same as those of the $a_i(t)$. 
Discrepancies between the $\tau_{ij}$ and the pairwise Spearman (or Kendall) 
correlations of the
$b_i(t)$ will therefore come about only from the presence of NAs and from sampling variation. 
These discrepancies should be minor if the 
numbers of NAs are not too large and $T$ is not too small. The Kendall version of the
algorithm throws an error if there are ties in any of the time series. The algorithm
also throws an error in the event that
the matrix generated in the second step is not positive semidefinite.

# Consequence of non-Normal copula structure : Effect on Taylor's law \label{Consequence_TL_simulation}

Here, we examined the nature of the slope and how it changed if population time series (taken as $50$ years) from different locations (say, $numloc=25$) had some certain type of copula structure (namely, \textbf{C}, \textbf{F} and \textbf{SC}) other than the $\textbf{N}$ormal copula. First, we simulated (N $\times$ n $\times$ nr) dimensional normal copula with co-variance $\rho =0.7$, where $(N=50,n=25,nr=10)$ are the length of time series, number of sampling locations and number of replica for that copula, respectively. Then we generated Kendall-correlation preserving \textbf{C}layton, \textbf{F}rank and $\textbf{S}$urvival \textbf{C}layton surrogates with same dimension of \textbf{N}ormal. With these four different types of copula we computed `log`(spatial variance) and `log`(mean) and plotted in Fig.\ref{fig_spTL}(A-D). We also tested statistically that whether linearity, homoskedasticity is maintained in each copula data set or not and plotted our results along with root mean square error, intercept and slope in Fig.\ref{fig_spTL}(E-I). We did a quadratic regression with each simulated copula for all types and measured quadratic coefficient and mean curvature for the regression line which is plotted in Fig.\ref{fig_spTL}(J-K) (For code see [`copsurrognd.R`](https://github.com/sghosh89/BIVAN/blob/master/copsurrognd.R) and [`TaylorsLawTools.R`](https://github.com/sghosh89/BIVAN/blob/master/TaylorsLawTools.R)).

<!-- Extinction risk results are not shown currently, it is introduced as a future idea in Discussion------

# Consequence of non-Normal copula structure : effect on extinction risk \label{Consequence_ext_risk}

In order to study the effects of non-Normal copula structure on extinction risk for metapopulations with different dispersal rules, we examined the following generalized multiple patch model where local population dynamics at each patch were governed by the Lewton-Cohen model ($r=0$) or Ricker model ($r \neq 0$) as the case may be;

\begin{equation}\label{eq.extrisk_model}
P(t+1) = \exp\left [  r\left ( 1-\frac{P(t)}{K} \right ) + \epsilon(t) \right ] P(t)
\end{equation}

$r$ is the liner growth rate and $K$ is the carrying capacity of the environment. $\epsilon$ is the environmental noise which might take stronger right tail (means stronger upper tail)  or stronger left tail (means stronger lower tail) dependence. We then considered each population $P(t+1)$ was updated depending upon $D$, the dispersal matrix, which could take either forms; one case when each population patch on a linear chain interacted with its two nearest neighbors (we called it local dispersion) or when any patch was allowed to interact with the rest of all (we called it as global dispersion). We simulated Eq. \ref{eq.extrisk_model} for $10000$ simulations and calculated extinction risk after $25$ years with different possibilities (for code see [mydispersal.R](https://github.com/sghosh89/BIVAN/blob/master/mydispersal.R)) and summarized our findings as follows;

- LC model (r=0) :

    * For LC model, as expected, when extinction risk calculated after $25$ years, was plotted against each varying $d$ (=degree of dispersion), we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one. [see first subplot of Fig. \ref{fig_risk_local_disp} and Fig. \ref{fig_risk_global_disp}]

- Ricker model (r $\neq$ 0):

    * For ricker model when extinction risk calculated after $25$ years, was plotted against each varying $d$ (=degree of dispersion), we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one upto $r=0.7$. For $r>0.7$, extinction risk is greater for data came from right tail dependent environmental fluctuations. [see Fig. \ref{fig_risk_local_disp} - \ref{fig_risk_global_disp}]

    * With increasing $K$ from $50$ to $200$, extinction risk in both case gets decreased as expected. [see Figure \ref{fig_risk_K_effect}]

With increasing number of patches from $5$ to $25$, extinction risk decreased [for a fixed $d$] ; [see Fig. \ref{fig_risk_numlocs_effect_local_disp} - Fig. \ref{fig_risk_numlocs_effect_global_disp}] for both model. -->

\clearpage

#  References

<!--
<div id="refs"></div>
-->