### 2019-03-26 SupportingInformation.Rmd
modified r chunk for asym sens results with b=+-0.1, +-0.5

---
title: "Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields"
author: "Shyamolina Ghosh, Lawrence W. Sheppard, Mark T. Holder, Terrance E. Loecke, P. Chris Reid, Daniel C. Reuman"
geometry: "left=1cm,right=1cm,top=2.2cm,bottom=2.2cm"

output: 
  pdf_document:
    number_sections: yes
    keep_tex: yes
    fig_caption: yes
    includes:
      in_header: head_supp.sty

tables: True
link-citations: True
urlcolor : blue
indent : True

csl: ecology-letters.csl
bibliography: REF_ALL.bib
---

\pagenumbering{roman}
\tableofcontents
\listoffigures
\listoftables

\pagebreak
\pagenumbering{arabic}

<!--Basic setup-->
```{r setup, echo=F}
knitr::opts_chunk$set(echo = TRUE, fig.pos = "H")
seed<-101
# families<-c(1,3:10,13,14,16:20)
source("mtime.R") #A function needed for caching
library(captioner)
```

```{r cop_fig_intro,echo=F,results="hide",cache=T, cache.extra=list(seed,mtime("copsurrognd.R"))}
library(copula)
library(VineCopula)
library(mvtnorm)
set.seed(seed)

source("copsurrognd.R")

#x <- read.table("./Data/dataset/x.txt")
#y <- read.table("./Data/dataset/y.txt")
#mat<-cbind(x,y)
#colnames(mat)<-c("x","y")

#***constants
N<-200 #length of time series
n<-2 #number of sampling locations
numdat<-1 #the number of datasets to get from the normal copula
rho<-0.8 #normal-copula parameter for dependence between locations
#this is a covariance, see below call to rmvnorm

#***generate the starting data, which has normal copula structure
#because it comes originally from a multivariate normal, before 
#transformation
sig<-matrix(rho,n,n)
diag(sig)<-1
m<-rmvnorm(N*numdat,mean=rep(0,n),sigma=sig)
m<-array(m,c(N,numdat,n))
m<-aperm(m,c(1,3,2)) # 2nd dim is normally distributed e.g. hist(m[,1,])
gshape<-7.5
gscale<-1 #gamma parameters

cop_to_marg_trans<-function(x){
  return(qgamma(x,shape=gshape,scale=gscale)) #this is where we choose the marginals,
  #should be something skewed
}
# pnorm(m) makes normal to uniform
# then qgamma(...) makes uniform to gamma distributed
m<-cop_to_marg_trans(pnorm(m))

ccop<-claytonCopula(2,2)
csurrog<-array(NA,dim(m))
csurrog[,,1]<-copsurrognd(m[,,1],targetcop=ccop,numsurrog=1)
mat<-csurrog[,,1]

pdf("./Results/rawdata.pdf",height = 4, width = 4)
scale<-2
xhist <- hist(mat[,1], breaks=30, plot=FALSE)
yhist <- hist(mat[,2], breaks=30, plot=FALSE) 
top <- max(c(xhist$counts, yhist$counts)) 
xrange <- range(mat[,1])
yrange <- range(mat[,2])
nf <- layout(matrix(c(2,0,1,3),2,2,byrow=TRUE), c(4,1), c(1,4), TRUE) 
op<-par(mar=c(3,3,1,1),mgp=c(2, 1, 0))  
plot(mat[,1], mat[,2], xlim=xrange, ylim=yrange, xlab="x", ylab="y", cex.lab=1.5)
legend("bottomright","raw data plot",bty="n",cex=1.5,text.col="magenta")
op<-par(mar=c(0,3,1,1)) 
barplot(xhist$counts/scale, axes=FALSE, ylim=c(0, top), space=0) 
op<-par(mar=c(3,0,1,1)) 
barplot(yhist$counts/scale, axes=FALSE, xlim=c(0, top), space=0, horiz=TRUE)
par(op)
dev.off()

pdf("./Results/copdata.pdf",height = 4, width = 4)
copmat<-pobs(mat)
xhist <- hist(copmat[,1], breaks=30, plot=FALSE)
yhist <- hist(copmat[,2], breaks=30, plot=FALSE) 
top <- max(c(xhist$counts, yhist$counts)) 
xrange <- c(0,1)
yrange <- c(0,1)
nf <- layout(matrix(c(2,0,1,3),2,2,byrow=TRUE), c(4,1), c(1,4), TRUE) 
op<-par(mar=c(3,3,1,1),mgp=c(2, 1, 0)) 
plot(copmat[,1], copmat[,2], xlim=xrange, ylim=yrange, xlab="u=rank(x)/(n+1)", ylab="v=rank(y)/(n+1)",cex.lab=1.5)
legend("bottomright","copula plot",bty="n",cex=1.5,text.col="magenta")
op<-par(mar=c(0,3,1,1)) 
barplot(xhist$counts/scale, axes=FALSE, ylim=c(0, top), space=0) 
op<-par(mar=c(3,0,1,1)) 
barplot(yhist$counts/scale, axes=FALSE, xlim=c(0, top), space=0, horiz=TRUE)
par(op)
dev.off()
```

```{r cop_pedagog_fig,echo=F,results="hide"}
set.seed(seed)
source("./FirstPedagogFig.R")
source("./SecondPedagogFig.R")
source("./ThirdPedagogFig.R")
source("./MakeCopulaPedagogFigs.R")
```

<!--The raw data look like this:-->
```{r load_data_Loecke, echo=F, results='hide'}
#fig.cap="Plot of the raw soil C and N data.\\label{fig_raw_cop_soilCN}(A)", fig.height=3, fig.width=3}
d<-readRDS("Data/dataForShymolinaCoord.rds")
d<-d[,c("SOCstock100","TSNstock100","lat","long")]
d<-na.omit(d)
pdf("./Results/SoilCNraw.pdf")
op<-par(mar=c(6, 6, 0.2, 2.1))
plot(d[,1],d[,2],type="p",col="black",xlab="Soil C",ylab="Soil N",cex.lab=2.5,cex.axis=2.5)
par(op)
dev.off()
```

<!--The copula looks like this:-->
```{r rankandplot_Loecke, echo=F, results='hide'}
source("getcopula.R")
#convert to a copula and plot
pdf("./Results/SoilCNcop.pdf")
v_CN<-getcopula(d=d[,c(1:2)],rankon=T,ploton=T) 
dev.off()
```

```{r plot_SoilCNmapUSA, echo=F, results='hide'}
#---------map of soilCN location on USA------------
library(maps)
pdf("./Results/SoilCNmapUSA.pdf",height=4,width=8)
map("usa")
points(x = d$long, y = d$lat, col="red",pch=16,cex=0.5)
dev.off()
```

```{r read_BMR_birds_and_copulaplot, echo=F, results='hide'}
library(readxl)
set.seed(seed)
raw_b<-readxl::read_xls("./Data/BMR/BirdBodyMassesMetabolicRates/BirdMassesAndMetabolicRatesMcNab2008.xls")
raw_b<-raw_b[,1:3]
raw_b<-na.omit(raw_b)
raw_b<-raw_b[,2:3]
raw_b<-as.data.frame(raw_b)

saveRDS(raw_b,"./Data/BMR/BirdBodyMassesMetabolicRates/my_birdsBMR_data.RDS")

b<-log10(raw_b)
colnames(b)<-c("log10M(g)","log10BMR(KJ/h)")
pdf("./Results/BMR_results/birds/raw_birds_BMR.pdf")
op<-par(mar=c(6, 6, 0.2, 2.1))
plot(b[,1],b[,2],type="p",col="black",xlab="log(Mass)",ylab="log(BMR)",cex.lab=2.5,cex.axis=2.5)
par(op)
dev.off()

source("getcopula.R")
pdf("./Results/BMR_results/birds/copula_birds_BMR.pdf")
cop_b<-getcopula(d=b,rankon=T,ploton=T) #convert to a copula and plot
dev.off()
```

```{r read_BMR_mammals_and_copulaplot, echo=F, results='hide'}
set.seed(seed)
raw_m<-readxl::read_xls("./Data/BMR/MammalianBodyMassesMetabolicRates/MammalianMassesAndMetabolicRatesMcNab2008.xls")
raw_m<-raw_m[,1:3]
raw_m<-na.omit(raw_m)
raw_m<-raw_m[,2:3]

saveRDS(raw_m,"./Data/BMR/MammalianBodyMassesMetabolicRates/my_mammalsBMR_data.RDS")

m<-log10(raw_m)
colnames(m)<-c("log10M(g)","log10BMR(KJ/h)")
m<-as.data.frame(m)

pdf("./Results/BMR_results/mammals/raw_mammals_BMR.pdf")
op<-par(mar=c(6, 6, 0.2, 2.1))
plot(m[,1],m[,2],type="p",col="black",xlab="log(Mass)",ylab="log(BMR)",cex.lab=2.5,cex.axis=2.5)
par(op)
dev.off()

source("getcopula.R")
pdf("./Results/BMR_results/mammals/copula_mammals_BMR.pdf")
cop_m<-getcopula(d=m,rankon=T,ploton=T) #convert to a copula and plot
dev.off()
```

```{r read_ceder_creek_data, echo=F, results="hide"}
set.seed(seed)
#-------------------------------------------------
# This function takes 
# Input :
#       db : a dataframe : smaller subset of ple_120 data to calculate avg. biomass for each plot
#       dp : a dataframe : smaller subset of pce_120 data to calculate avg. shannon's H for those plots
#       year : the year for which you want to calculate biomass and shannon's index data
#       rawplot : logical : T for rawdata plot, F for copula plot
#       ploton : logical for optional plotting

# Output :
#       a list of two data-frame:
#                   tab_dbp : a data-frame with 3 column : (plot no , avg.H, avg.Biomass(g/m2))
#                   cop_tab_dbp : a data-frame with 3 column : (plot no, pobs(avg.H), pobs(avg.Biomass(g/m2)))

cop_HB<-function(db,dp,year,rawplot,ploton){
  # db : biomass dataset for a specific year
  # dp : percent cover dataset for a specific year

  plots<-sort(intersect(db$Plot,dp$Plot)) # common plots for both datasets for a specific year
  db<-subset(db,Plot%in%plots)
  dp<-subset(dp,Plot%in%plots)

  tab_dbp<-data.frame(Plot=plots,avg.H=NA,avg.Biomass=NA)
  for(i in 1:length(plots)){
    # calculate biomass for each plot (averaged over all strips for that plot)
    x<-subset(db,Plot==plots[i])
    tab_dbp$avg.Biomass[i]<-mean(x$Biomass)

    # calculate shannon's index H for each plot (averaged over all Quadrates for that plot)
    x<-subset(dp,Plot==plots[i])
    uniq_sp<-sort(as.character(unique(x$Species)))  # These are the species in a particular plot
    Q<-as.character(unique(x$Quadrat))

    tab_each_sp<-data.frame(sp=uniq_sp,avg_cov=0,rel_avg_cov=0)

    for(sp in 1:length(uniq_sp)){
      my_sp<-uniq_sp[sp]
      tab_temp<-data.frame(Q=Q,abs_cov=0)
      ind<-which(x$Species==my_sp)
      li<-length(ind)
      tab_temp$abs_cov[1:li]<-x$Abscover[ind]
      tab_each_sp$avg_cov[[sp]]<-mean(tab_temp$abs_cov) # avg. over 4Quadrates for a sp.
    }

    tab_each_sp$rel_avg_cov<-tab_each_sp$avg_cov/sum(tab_each_sp$avg_cov)
    tab_dbp$avg.H[i]<- -sum((tab_each_sp$rel_avg_cov)*log(tab_each_sp$rel_avg_cov))

  }

  tab_dbp<-na.omit(tab_dbp)

  cop_tab_dbp<-data.frame(Plot=tab_dbp$Plot,avg.H=NA,avg.Biomass=NA)
  cop_tab_dbp$avg.H<-VineCopula::pobs(tab_dbp$avg.H)
  cop_tab_dbp$avg.Biomass<-VineCopula::pobs(tab_dbp$avg.Biomass)

  if(ploton==T){
    if(rawplot==T){
      op<-par(mar=c(6, 6, 0.2, 2.1))
      plot(tab_dbp$avg.H,tab_dbp$avg.Biomass,xlab="avg. H",ylab="avg. Biomass",pch=1,cex.lab=2.5,cex.axis=2)
      mtext(paste0("Year=",year),cex=1)
      par(op)
    }else{
      op<-par(mar=c(6, 6, 0.2, 2.1))
      plot(cop_tab_dbp$avg.H,cop_tab_dbp$avg.Biomass,xlab="u",ylab="v",pch=1,cex.lab=3,cex.axis=2,
         xlim=c(0,1),ylim=c(0,1),asp=1)
      rect(0,0,1,1)
      lines(c(0,1),c(0,1),type='l',col="grey")
      #mtext(paste0("Year=",year),cex=1)
      par(op)
    }
  }

  return(list(tab_dbp=tab_dbp,
         cop_tab_dbp=cop_tab_dbp))
}
#-----------------------------
r_ple<-read.delim("./Data/Ceder_creek_e120/edited_ple120.txt")
names(r_ple)[36]<-"Biomass"

r_pce<-read.delim("./Data/Ceder_creek_e120/edited_pce120.txt")

y1<-unique(r_ple$Year)
y2<-unique(r_pce$Year)
yr<-sort(intersect(y1,y2))

a<-as.data.frame(table(sort(r_pce$Species)))
b<-as.character(a$Var1)
bad_sp<-c(b[51:53],b[156:158])

ind<-which(r_pce$Species%in%bad_sp)
r_pce_clean<-r_pce[-ind,]
#--------------------------------------------------------------------------------------------
pdf("./Results/Ceder_creek_results/raw_HB_all_yr.pdf",height = 6,width = 10)
op<-par(mfrow=c(2,3),mar=c(3,3,3,3), mgp=c(1.5,0.5,0))
for(i in 1:length(yr)){
  db<-subset(r_ple,Year==yr[i])
  dp<-subset(r_pce_clean,Year==yr[i])

  sr<-cop_HB(db=db,dp=dp,year=yr[i],rawplot=T,ploton = T)
}
par(op)
dev.off()
#-----------------------
#raw data plot for year2000
pdf("./Results/Ceder_creek_results/raw_HB_2000.pdf",height = 6,width = 6)
i<-5
db<-subset(r_ple,Year==yr[i])
dp<-subset(r_pce_clean,Year==yr[i])
sr<-cop_HB(db=db,dp=dp,year=yr[i],rawplot=T,ploton = T)
saveRDS(sr,"./Results/Ceder_creek_results/raw_cop_cc_2000.RDS")
dev.off()
#----------------------------------------------------------------------------------------
cop_HB_all_yr<-vector("list",length(yr))
names(cop_HB_all_yr)<-yr

pdf("./Results/Ceder_creek_results/cop_HB_all_yr.pdf",height = 6,width = 10)
op<-par(mfrow=c(2,3),mar=c(3,3,3,3), mgp=c(1.5,0.5,0))
for(i in 1:length(yr)){
  db<-subset(r_ple,Year==yr[i])
  dp<-subset(r_pce_clean,Year==yr[i])

  s<-cop_HB(db=db,dp=dp,year=yr[i],rawplot=F,ploton = T)
  cop_HB_all_yr[[i]]<-s$cop_tab_dbp[,2:3]
}
par(op)
dev.off()
#-------------------------
#cop data plot for year2000
pdf("./Results/Ceder_creek_results/cop_HB_2000.pdf",height = 6,width = 6)
i<-5
db<-subset(r_ple,Year==yr[i])
dp<-subset(r_pce_clean,Year==yr[i])
s<-cop_HB(db=db,dp=dp,year=yr[i],rawplot=F,ploton = T)
cc_cop_2000<-s$cop_tab_dbp[,2:3]
dev.off()
#------------------

saveRDS(cop_HB_all_yr,"./Results/Ceder_creek_results/cop_HB_all_yr.RDS")
#---------------------------------------------------------------------------------------
cop_HB_all_yr<-readRDS("./Results/Ceder_creek_results/cop_HB_all_yr.RDS")
```

```{r map_plot_aphid_plankton,results="hide",echo=F}
source("aphid_data_calling.R")
source("good_loclist.R")
library(maps)

#----------------- Aphid map--------------------------------------------
d0_count<-as.matrix(read.csv("./Data/Aphid_data/APHID_DATA_RAW/aphidtscount141117.csv",header=F))
data_aphid_count<-d_allsp_data(d0_count)

d0_firstflight<-as.matrix(read.csv("./Data/Aphid_data/APHID_DATA_RAW/aphidtsfirstflight141117.csv",header=F))
data_aphid_ff<-d_allsp_data(d0_firstflight)

selective_loc1<-good_loclist(d_allsp=data_aphid_count,sp=10,data_pt_thrs=30)
selective_loc2<-good_loclist(d_allsp=data_aphid_ff,sp=11,data_pt_thrs=30)
all(selective_loc1==selective_loc2)==T

longs<-c(-4.567,0.57,-3.069,-3.312,-2.637,-1.682,-2.763,-0.356,-3.454,0.427,0.939)
lats<-c(55.477,52.26,56.457,55.949,52.125,55.213,53.854,51.807,50.628,51.733,51.185)

locno<-length(longs)
df_loc_info_org<-data.frame(loc_no=c(1:length(lats)),longs=longs,lats=lats)
df_loc_info<-df_loc_info_org[selective_loc1,]
xlim=c(-10,2)
ylim=c(50, 60)

pdf("./Results/aphid_map_plot.pdf")
map("world", fill=TRUE, col="grey84", bg="white",border=0, xlim=xlim, ylim=ylim)
map.axes(cex.axis=2.5)
points(x = df_loc_info_org$longs, y = df_loc_info_org$lats, col="red",pch=16)
points(x = df_loc_info$longs, y = df_loc_info$lats, col="blue", pch=0, cex=1.5)
text(df_loc_info_org$loc_no, x = df_loc_info_org$longs, y = df_loc_info_org$lats, col="red",  pos=1)
dev.off()

#--------------Plankton map----------------------------

source("plankton_north_sea_data_calling.R")
d0_plankton_north_sea<-as.matrix(read.csv("./Data/Plankton_North_Sea_data/planktontimeseries201117_4.csv",header=F))
data_plankton_north_sea<-d_allsp_data_plankton_north_sea(d0_plankton_north_sea)

longs<-read.csv("./Data/Plankton_North_Sea_data/boxcornerlongs201117.csv",header=F)
lats<-read.csv("./Data/Plankton_North_Sea_data/boxcornerlats201117.csv",header=F)
longs_pns<-longs[[1]]
lats_pns<-lats[[1]]

df_loc_info_org<-data.frame(loc_no=c(1:length(lats_pns)),longs=longs_pns,lats=lats_pns)
selective_loc<-good_loclist(d_allsp=data_plankton_north_sea,sp=16,data_pt_thrs=45)
df_loc_info<-df_loc_info_org[selective_loc,]
xlim=c(-12,10)
ylim=c(48, 60)

pdf("./Results/plankton_map_plot.pdf")
map("world", fill=TRUE, col="grey84", bg="white",border=0, xlim=xlim, ylim=ylim)
map.axes(cex.axis=2.5)
abline(h=seq(from=ylim[1],to=ylim[2],by=2), v=seq(from=xlim[1],to=xlim[2],by=2), col="black", lty=2)
points(x = df_loc_info_org$longs+1, y = df_loc_info_org$lats+1, col="red",pch=16)
points(x = df_loc_info$longs+1, y = df_loc_info$lats+1, col="blue", pch=0, cex=1.5)
text(df_loc_info_org$loc_no, x = df_loc_info_org$longs+1, y = df_loc_info_org$lats+1, col="red",  pos=1)
dev.off()
```

```{r make_data_plot_fig,echo=FALSE,results="hide"}
source("./MakeBivarDataPlot.R")
```

# Transformations used in Fig. \ref{MT-fig_cop_pedag1} \label{pedagog1_details}

The data of Fig. \ref{MT-fig_cop_pedag1}A-B were generated to have standard normal marginals.
Fig. \ref{MT-fig_cop_pedag1}C can be obtained from Fig. \ref{MT-fig_cop_pedag1}A via transformation
of both variables by the cdf of the standard normal distribution, and then by the inverse of 
the cdf of the gamma distribution with shape and scale parameters 2. 
The function `qgamma` in R was used.
Fig. \ref{MT-fig_cop_pedag1}D can be obtained from Fig. \ref{MT-fig_cop_pedag1}B in the same way.
Fig. \ref{MT-fig_cop_pedag1}E can be obtained from Fig. \ref{MT-fig_cop_pedag1}A via transformation
of both variables by the cdf of the standard normal distribution.
Fig. \ref{MT-fig_cop_pedag1}F can be obtained from Fig. \ref{MT-fig_cop_pedag1}B in the same way.

\begin{figure}
\begin{center}
\includegraphics[width=16 cm,height=8 cm]{./Results/copula_pedagog_fig_BB1.pdf}
\caption[Lower- and upper-tail dependence can vary independently of correlation]{Lower- and upper-tail dependence can vary independently of correlation. (A-C) Data generated using the BB1 copula (see section on methods for Q1) 
with three different 
sets of parameters chosen to produce a variety of 
lower-tail (LT) and upper-tail (UT) dependence strengths (methods for Q1), but the same 
Kendall correlation (K) (to two digits) in all cases. (D-F) Similar to A-C, but
with stronger correlation. Asymmetry of tail dependence (LT-UT) is the same 
within each column.\label{fig_cop_pedag}}
\end{center}
\end{figure}

# Details of the data \label{Data_details}

<!--***DAN: This is a dumping ground at the moment, but needs to be organized into a section that gives all the necessary details about the data-->

<!--The CPR is a device which is towed across the ocean behind a ship, filtering plankton from seawater with a continuously
unwinding silk ribbon. The ribbon was cut into 4 inch segments, each corresponding to a particular time and
location in the ocean. Organisms were identified and counted (possibly with sub-sampling) microscopically [@Colebrook1979]. All the samples available for a given variable in a given area in a given month of a given year were
averaged, producing a monthly time series for each year. Where data for a given month was unavailable the missing value was replaced with a median value for that month (drawn from all years in which data was available).-->

 [@Sheppard2017] converted monthly values to yearly values by averaging over all months. But here we only allowed maximum 4 months in a year with missing data which were replaced by median value for that month (drawn from all years in which data was available). We used data from 14 locations which had at least 45 years time series data (see Figure \ref{SM-fig_plankton_map}).

 The Great Miami Wetland Mitigation Bank  was at $36^{\circ}46^{'}51^{''}$ N, $84^{\circ}20^{'}26^{''}$ W) in Trotwood, Montgomery County, Ohio. Additional information on the history of the site and soil types is detailed in @Jarecke2016. Briefly, the closed static flux chamber method was used [@Holland1999] to measure $CH_{4}$ fluxes at 28 locations across a hydrologic gradient (upland to seasonal wetland) at daily to weekly interval from September 2015 to September 2016. **See and Smyth et al, pending for
additional details???** on greenhouse gas flux methods for this site. We considered 13 locations that had at least 50 dates data.

```{r tab_aphid_info, echo=F, results='asis',message=F}
library(tinytex)
library(tibble)
library(kableExtra)
library(dplyr)
aphid_info <- tibble(c1=c("Apple grass aphid","Bird cherry oat aphid","Black bean aphid","Blackberry cereal aphid","Blackcurrant sowthistle aphid",
                       "Corn leaf aphid","Currant lettuce aphid","Damson hop aphid","Grain aphid","Green spruce aphid",
                       "Leaf-curling plum aphid","Mealy cabbage aphid","Mealy plum aphid","Pea aphid","Peach potato aphid",
                       "Potato aphid","Rose grain aphid","Shallot aphid","Sycamore aphid","Willow carrot aphid"),
                     c2=c("Rhopalosiphum insertum", "Rhopalosiphum padi", "Aphis fabae", "Sitobion fragariae", "Hyperomyzus lactucae",
                       "Rhopalosiphum maidis", "Nasonovia ribisnigri","Phorodon humuli","Sitobion avenae","Elatobium abietinum",
                       "Brachycaudus helichrysi","Brevicoryne brassicae","Hyalopterus pruni","Acyrthosiphon pisum","Myzus persicae",
                       "Macrosiphum euphorbiae","Metopolophium dirhodum","Myzus ascalonicus","Drepanosiphum platanoidis","Cavariella aegopodii")
)

saveRDS(aphid_info,"./Results/aphid_info.RDS")
knitr::kable(aphid_info, "latex", booktabs = T, linesep = "",
      caption = "Species names of 20 aphids for which data were available \\label{tab_aphid_info}",
      col.names = NULL)%>%
     column_spec(column = 2,italic=T)%>%add_header_above(header=c("Common name" = 1, "Latin binomial" = 1))
```

```{r tab_plankton_info, echo=F, results='asis',message=F}
library(tibble)
plankton_info <- tibble(c1=c('Calanus stages I-IV','Para-pseudocalanus spp.','Temora longicornis','Acartia sp.',
                          'Centropages typicus',
                          'Oithona sp.','Echinoderm larvae','Calanus finmarchicus',
                          'Calanus helgolandicus','Metridia lucens','Decapoda larvae',
                   'Euphausiids','Thalassiosira sp.','Rhizosolenia styliformis',
                   'Ceratium fusus','Ceratium furca','Ceratium tripos','Ceratium macroceros',
                   'Proboscia alata','Pseudocalanus sp.',
                   'Pseudo-nitzschia delicatissima','Pseudo-nitzschia seriata')
)
cns<-c('Echinoderm larvae','Euphausiids','Decapoda larvae')
'%ni%' <- Negate('%in%')
idn<-which(plankton_info$c1%ni%cns)
knitr::kable(plankton_info, "latex", booktabs = T, linesep = "",
      caption = "Species names of 22 plankton taxa for which data were available \\label{tab_plankton_info}",
      col.names = c("Plankton taxon name"))%>%row_spec(row=idn,italic=T)
```
<!--***DAN: These names differ from those used in Emma's 2016 Glob Change Biol paper.
Need to check with Chris Reid. Have to get this right.-->

```{r read_methane_data, echo=F, results="markup"}
set.seed(seed)
methane_raw<-readRDS("./Data/Methane_data/methane_data_for_Shyamolina.rds")
#class(methane_raw) #it's a data.frame for multivariable data
#knitr::kable(head(methane_raw,3),format='pandoc',
#             caption = "First few rows of raw methane data.\\label{tab_rawdata_methane}",
#             booktabs=T) # display first few rows
```

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm,height=16 cm]{./Results/SoilCNmapUSA.pdf} 
\caption[Sampling locations for C and N data]{Locations of central pedons sampled 
for soil organic carbon and total soil nitrogen (Mg C or N per hectare of soil 
surface) at 5907 sites across the coterminous United States. \label{fig_SoilCNmapUSA}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm,height=16 cm]{./Results/aphid_map_plot.pdf} 
\caption[Sampling locations for aphid data]{Locations of the 
Rothamsted Insect Survey 
suction traps from which data were available are indicated by the red circles, 
out of which only blue boxes had at least 30 years of data and were 
used for our study. \label{fig_aphid_map}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm,height=16 cm]{./Results/plankton_map_plot.pdf} 
\caption[Sampling regions for CPR plankton data]{CPR data were available from 
26 locations ($2^{\circ}$ $\times$ $2^{\circ}$ regions on the map) around UK 
seas (red circles), out of which only blue boxes had at least 45 years of data and
were used for our study. \label{fig_plankton_map}}
\end{center}
\end{figure}

<!--Now the copula figures-->

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Normal.pdf}
\caption[Example N copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example N copulas. 
S is Spearman correlation; p is the value of the parameter for the N 
family (it is a 
one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, respectively.\label{figped_N}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Clayton.pdf}
\caption[Example C copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example C copulas. S is Spearman correlation; p is the value of the parameter for the C family (it is a 
one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively.\label{figped_C}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SurvClayton.pdf}
\caption[Example SC copulas]{Log-transformed pdfs (A-E) and samples (F-J) 
from example SC 
copulas. S is Spearman correlation; p is the value of the parameter for the SC 
family (it is a one-parameter family); and LT and UT are the measures of lower- and 
upper-tail dependence, respectively.\label{figped_SC}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Gumbel.pdf}
\caption[Example G copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example G copulas. S is 
Spearman correlation; p is the value of the parameter for the G family (it is a 
one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively.\label{figped_G}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SurvGumbel.pdf}
\caption[Example SG copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example SG 
copulas. S is Spearman correlation; p is the value of the parameter for the SG 
family (it is a one-parameter family); and LT and UT are the measures of lower- and 
upper-tail dependence, respectively.\label{figped_SG}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Joe.pdf}
\caption[Example J copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example J copulas. K is Kendall correlation; p is the value of the parameter for the J family (it is a 
one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively.\label{figped_J}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SurvJoe.pdf}
\caption[Example SJ copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example SJ 
copulas. K is Spearman correlation; p is the value of the parameter for the SJ 
family (it is a one-parameter family); and LT and UT are the measures of lower- and 
upper-tail dependence, respectively.\label{figped_SJ}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Frank.pdf}
\caption[Example F copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example F copulas. S is Kendall correlation; p is the value of the parameter for the F family (it is a 
one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
respectively.\label{figped_F}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB1.pdf}
\caption[Example BB1 copulas]{Log-transformed pdfs for example BB1 copulas. K is Kendall correlation; p1 and p2
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_BB1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB1.pdf}
\caption[Example SBB1 copulas]{Log-transformed pdfs for example SBB1 copulas. K is Kendall correlation; p1 ad p2 
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_SBB1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB6.pdf}
\caption[Example BB6 copulas]{Log-transformed pdfs for example BB6 copulas. K is Kendall correlation; p1 and p2
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_BB6}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB6.pdf}
\caption[Example SBB6 copulas]{Log-transformed pdfs for example SBB6 copulas. K is Kendall correlation; p1 ad p2 
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_SBB6}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB7.pdf}
\caption[Example BB7 copulas]{Log-transformed pdfs for example BB7 copulas. K is Kendall correlation; p1 and p2
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_BB7}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB7.pdf}
\caption[Example SBB7 copulas]{Log-transformed pdfs for example SBB7 copulas. K is Kendall correlation; p1 ad p2 
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_SBB7}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB8.pdf}
\caption[Example BB8 copulas]{Log-transformed pdfs for example BB8 copulas. K is Kendall correlation; p1 and p2
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_BB8}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB8.pdf}
\caption[Example SBB8 copulas]{Log-transformed pdfs for example SBB8 copulas. K is Kendall correlation; p1 ad p2 
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively.\label{figped_SBB8}}
\end{center}
\end{figure}

# Nonparametric statistics \label{nonparam_stats}

\noindent We here define
\begin{equation}\label{eq.Cor}
   \cor_{l_b,u_b}(u,v) = \frac{\sum 
   (u_i-\mean(u)) (v_i-\mean(v))}{(n-1)\sqrt{\var(u)\var(v)}}.
\end{equation}
\noindent Here means and sample variances are computed using all $n$ data 
points, but the
sum is over only the indices $i$ for which $u_i+v_i > 2l_b$ and 
$u_i+v_i < 2u_b$. 
The sum of 
$\cor_{0,0.5}(u,v)$ and $\cor_{0.5,1}(u,v)$ (or some other choice of 
$\cor_{l_{b_k},u_{b_k}}(u,v)$ for bounds $l_{b_k},u_{b_k}$ that partition the
interval $(0,1)$) equals the standard Spearman correlation.

We also define $\Ps_{l_b,u_b}$, using the bounds
$u+v=2l_b$ and $u+v=2u_b$ defined in the main text, as follows. For a distance $h$, we define
$S(h)$ to be the number of points $(u_i,v_i)$ within the bounds 
and a distance less than $h$ from the line $v=u$, divided by the 
total number of points within the bounds. This function is
defined for $h$ going from $0$ to a distance $h_{\text{max}}$ which
is half the longer of the two segments obtained by intersecting the
bounds with the unit square. It is easy to see $S(0)=0$ and 
$S(h_{\text{max}})=1$. We define $S_i(h)$, also for $h$ going from $0$
to $h_{\text{max}}$, to be the area within the bounds and the unit square
and within distance $h$ of the line $v=u$, divided by the area
within the bounds and the unit square. This is the expected value of
$S(h)$ for independent data. We then define
\begin{equation}\label{eq.P}
   P_{l_b,u_b} = \int_0^{h_{\text{max}}} ( S(h) - S_i(h) ) dh.
\end{equation}
\noindent As for $\cor_{l_b,u_b}$, larger values of $\Ps_{l_b,u_b}$ 
indicate stronger positive association
between $u$ and $v$ in the region given by the bounds.
We use both statistics 
because $\cor_{l_{b_k},u_{b_k}}(u,v)$ is more familiar, but 
in some instances $\Ps_{l_b,u_b}$ appears to have more power to
reveal tail dependence.

\begin{figure}[!h]
\begin{center}
\includegraphics[width=6 cm,height=6 cm]{./stat_image.jpg} \hspace{0.6 cm}
\caption[For definitions of nonparametric statistics]{Schematic diagram supporting definitions of nonparametric tail dependence statistics. The blue and red lines are, respectively, $u+v=2l_b$ and $u+v=2u_b$.\label{fig_stat_image}}
\end{center}
\end{figure}

# Testing the nonparametric statistics \label{Test_npa_stat}

N and F copula families 
are known to have symmetric tail dependence,
i.e., strengths of dependence in the lower and upper tails are the same, for all copulas
in these families [@nelsen2006_copula]. C copulas are known to have stronger lower- than 
upper-tail dependence, and SC copulas, being rotations of C copulas by 
$180^\circ$, have stronger upper- than lower tail dependence [@nelsen2006_copula].
Using the `iRho` function from the `copula` package, for each of the four families N, 
F, C, and SC, we obtained parameters for which the expected value of 
Spearman's $\rho$ was, separately, $0,0.1,0.2,...,0.9$. For each copula family and for each 
of the selected parameter values we generated  $n$ points, 100 times, and thereby 
computed 100 values of the 
statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$. This exercise 
was repeated, separately, 
for $n=1000$ and for $n=35$, these cases being chosen, respectively, to reflect 
asymptotic behavior of the 
statistics and their behavior on the small datasets which are common in some fields
of ecology. The mean and standard error of the three difference statistics were computed for 
each copula family, parameter value, and value of $n$, and results were plotted against 
Spearman's rho (Figs \ref{fig_stat_testing35} and \ref{fig_stat_testing1000}), 
the expectation being that the difference statistics would be positive for 
C copulas, negative for SC copulas (except for the case of $\rho=0$, a 
boundary case of these families corresponding to independence), and not significantly different 
from $0$ for F and N 
copulas.
We used a $t$-test with a significance threshold of $0.05$ to judge significance here.
These expectations were fulfilled (Figs \ref{fig_stat_testing35} and 
\ref{fig_stat_testing1000}).

<!--The message=F and warning=F flags are to suppress warnings that come when you pass a parameter
of 0 to the Frank or Clayton copulas. These are technicallly boundary cases for these families.
They correspond to the independent copula. The packages we use provide the indepdent copula, but
issue a warning. Since anyway we want the independent copula, we can suppress the warning. We ran this
without the warning and it is OK to supress it.-->
<!-- The lines you are talking about appeared as message not as warning(actualy no warning appered for this chunk), so we have to set message=F here-->
```{r stat_testing_save_and_plot_result, message=F, results="hide", echo=F, cache=T, cache.extra=list(seed,mtime("TestStats_multivar_dataset.R"),mtime("CopulaFunctions.R"),mtime("CopulaFunctions_flexible.R"))}
source("TestStats_multivar_dataset.R")
set.seed(seed)

resultsloc<-"./Results/stat_results/stat_testing/"
spearvals<-seq(from=0,to=0.9,by=0.1)
callfn<-coplist_with_params(spearvals=spearvals)
coplist<-callfn$coplist
ncores<-3
res_pt35_rankF<-mclapply(X=coplist,FUN=worker,numpts=35,numsims=100,rank=F,mc.cores=ncores)
res_pt35_rankT<-mclapply(X=coplist,FUN=worker,numpts=35,numsims=100,rank=T,mc.cores=ncores)
res_pt1000_rankF<-mclapply(X=coplist,FUN=worker,numpts=1000,numsims=100,rank=F,mc.cores=ncores)
res_pt1000_rankT<-mclapply(X=coplist,FUN=worker,numpts=1000,numsims=100,rank=T,mc.cores=ncores)
save(res_pt35_rankF,res_pt35_rankT,res_pt1000_rankF,res_pt1000_rankT,file=paste(resultsloc,"StatTestNumericResults.RData",sep=''))

b<-callfn$paramlist
plotter_stat_testing(res_pt35_rankF,"res_pt35_rankF",b,resultsloc,spearvals)
plotter_stat_testing(res_pt35_rankT,"res_pt35_rankT",b,resultsloc,spearvals)
plotter_stat_testing(res_pt1000_rankF,"res_pt1000_rankF",b,resultsloc,spearvals)
plotter_stat_testing(res_pt1000_rankT,"res_pt1000_rankT",b,resultsloc,spearvals)
```

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/stat_results/stat_testing/res_pt35_rankT.pdf}
\caption[Testing nonparametric statistics with 35 data points]{Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the C and SC copulas, and the known symmetry of tail dependence of the N and F copulas. See Appendix  \ref{Test_npa_stat} for details. $n=35$. \label{fig_stat_testing35}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/stat_results/stat_testing/res_pt1000_rankT.pdf}
\caption[Testing nonparametric statistics with 1000 data points]{Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the C and SC copulas, and the known symmetry of tail dependence of the N and F copulas. See Appendix  \ref{Test_npa_stat} for details. $n=1000$.\label{fig_stat_testing1000}}
\end{center}
\end{figure}

<!--This r-chunk generates figure necessary for maintext Paper-->
```{r fig_stat_testing_for_maintext, message=F, results="hide", echo=F, cache=T, cache.extra=list(seed,mtime("TestStats_multivar_dataset.R"),mtime("CopulaFunctions.R"),mtime("CopulaFunctions_flexible.R"),mtime("StatTestNumericResults.RData"))}
set.seed(seed) 
source("fig_Paper_stat_testing.R")
#load("./Results/stat_results/stat_testing/StatTestNumericResults.RData")
resultsloc<-"./Results/stat_results/stat_testing/"

copname<-"C"
reslist<-res_pt35_rankT

statname<-"CorlmCoru"
call_fig_Paper_stat_testing(resultsloc=resultsloc,statname=statname,copname=copname,reslist=reslist, numsims=100)
statname<-"PlmPu"
call_fig_Paper_stat_testing(resultsloc=resultsloc,statname=statname,copname=copname,reslist=reslist, numsims=100)
statname<-"D2umD2l"
call_fig_Paper_stat_testing(resultsloc=resultsloc,statname=statname,copname=copname,reslist=reslist, numsims=100)
```

# Verifying Q1 results for bivariate datasets using nonparametric statistics \label{Q1bivarnonparam}

For bivariate datasets, we compared values 
of the statistics $\cor_{l_b,u_b}$,
$\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, $\cor_{0,u_b}-\cor_{1-u_b,1}$, 
$\Ps_{0,u_b}-\Ps_{1-u_b,1}$ and
$\Dsq_{1-u_b,1}-\Dsq_{0,u_b}$ to 
distributions of values of the same statistics computed on surrogate datasets that were produced from 
the empirical data by randomizing it in a special way to have no 
tail dependence (Appendix \ref{surrog_test}). 
The surrogate/randomized datasets had exactly 
the same marginal distributions as the empirical data and had very similar Kendall or Spearman 
correlation (the sorrogate algorithm had two versions, one for preserving each correlation coefficient), 
but had normal copula structure. Thus our comparisons tested the null hypothesis 
that our statistics took values on the empirical data no different from what would have been expected 
if the copula structure of the data were normal, but the data were otherwise 
statistically unchanged. The comparison of one of the
statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, as computed on the 
empirical data, to the distribution
of its values computed on many surrogate datasets 
provides a test of whether dependence in the part of the distribution specified 
by $l_b$ and $u_b$ is different from
what would be expected from a null hypothesis of normal copula structure but 
unchanged marginals and similar 
Kendall (or Spearman)
correlation. Thus the comparison helps answer the first two parts of Q1
for bivariate datasets. 
Significant deviations correspond to deviations from normal 
copula structure. In particular, deviations 
using $l_b=0$ and $u_b$ small (say, $0.1$) correspond to lower tail dependence 
different from that of a normal copula;
likewise, using $l_b=0.9$ and $u_b=1$ tests for upper tail dependence different 
from that of a normal copula.
The comparison of one of the statistics 
$\cor_{0,u_b}-\cor_{1-u_b,1}$, 
$\Ps_{0,u_b}-\Ps_{1-u_b,1}$ and
$\Dsq_{1-u_b,1}-\Dsq_{0,u_b}$,
as computed on the empirical data, to the distribution of its values on many surrogate datasets 
provides a test of asymmetry of dependence. Thus this comparison helps answer the
third part of Q1 for bivariate 
datasets.

The values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for the soil
C and N data
were compared to distributions of their values on 1000 Kendall- or Spearman-preserving N surrogates
of the data,
separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$ 
(Fig. \ref{fig_soilCN_nonparam}). Results show that tail dependence of data is stronger in 
both the lower and upper tails than would be expected under a null hypothesis of a normal copula.
The values of the statistics $\cor_{0,0.1}-\cor_{0.9,1}$, $\Ps_{0,0.1}-\Ps_{0.9,1}$ and
$\Dsq_{0.9,1}-\Dsq_{0,0.1}$ for the soil C and N data were also compared to distributions of their values
on 1000 Kendall- or Spearman-preserving N surrogates, separately in two 
comparisons (Table \ref{tab_soilCN_asym}). Results showed that
upper-tail dependence was significantly stronger than lower-tail dependence.

For all four bivariate datasets the statistic $P_{0,0.1}-P_{0.9,1}$ was computed for real data and 
for 1000 Kendall-preserving surrogate datasets with N copulas, and the rank of the real-data value in 
the distribution of surrogate values was computed. If ranks were $>975$ or $<25$, asymmetry of tail 
dependence was significant compared to a N-copula null hypothesis. Ranks were 1, 1, 1, 998
<!--***DAN: Shya, please make the quantities in the previous line automatic links-->
for, respectively, the soil C and N data, bird masses and metabolic rates, mammal masses
and metabolic rates, and the Cedar Creek data. Thus the first three datasets showed stronger 
upper- than lower-tail dependence, and the Cedar Creek data showed the reverse, confirming the 
results based on model selection presented in the main text.

# Surrogate testing with bivariate datasets for nonparametric statistics \label{surrog_test}

For bivariate datasets, surrogates were produced as follows.

- Given data $(x_i,y_i)$ for $i=1,\ldots,n$ and a 
one-parameter target copula family $\mathcal{B}(\theta)$ (here $\theta$ denotes the parameter), let $\tau$ be the
Kendall correlation of the $x_i$ with the $y_i$ and find the parameter value $\theta_\tau$ for which 
$\mathcal{B}(\theta_\tau)$ has Kendall correlation $\tau$ (this is possible for the one-parameter copula families of this 
study using the `iTau` function of the `copula` package). 
- Generate data $(a_i,b_i)$, $i=1,\ldots,n$ from 
$\mathcal{B}(\theta_\tau)$. 
- Permute the ordered set $(x_1,\ldots,x_n)$ such that the permutation
$(x_{\sigma_x(1)},\ldots,x_{\sigma_x(n)})$ has $\rank(x_{\sigma_x(i)})$ equal to $\rank(a_i)$ for all $i$, where 
$\rank(x_{\sigma_x(i)})$ is the rank of $x_{\sigma_x(i)}$ in the set $(x_{\sigma_x(1)},\ldots,x_{\sigma_x(n)})$, and
$\rank(a_i)$ is the rank of $a_i$ in the set $(a_1,\ldots,a_n)$ (here $\sigma_x$ is a permutation of the 
indices $1,\ldots,n$). 
- Likewise permute $(y_1,\ldots,y_n)$
such that the permutation
$(y_{\sigma_y(1)},\ldots,y_{\sigma_y(n)})$ has $\rank(y_{\sigma_y(i)})$ equal to $\rank(b_i)$ for all $i$ ($\sigma_y$ is 
another permutation of $1,\ldots,n$).
- The surrogate dataset is $(x_{\sigma_x(i)},y_{\sigma_y(i)})$ for $i=1,\ldots,n$.

\noindent Our code for this algorithm is `copsurrog2d` in the 
`BIVAN` repository (https://github.com/sghosh89/BIVAN). 
Because $(x_{\sigma_x(1)},\ldots,x_{\sigma_x(n)})$ 
is a permutation of $(x_1,\ldots,x_n)$ and $(y_{\sigma_y(1)},\ldots,y_{\sigma_y(n)})$ is a 
permutation 
of $(y_1,\ldots,y_n)$, the marginal distributions of the surrogate data are exactly the same as 
those of the original data. 
Because ranks of the surrogate data are matched to ranks of the $a_i$ and $b_i$, the
Kendall correlation of the surrogate dataset will be the same as that of the $a_i$ and $b_i$, and therefore 
will be close to $\tau$ (differences arising only through sampling variation). If instead of $\theta_\tau$
a parameter value $\theta_\rho$ is used such that the Spearman correlation of $\mathcal{B}(\theta_\rho)$ is the same
as that of the $x_i$ with the $y_i$, then Spearman correlations of surrogates will very nearly equal (except for 
sampling variation) the Spearman correlation of the empirical data.

# Verifying Q1 results for multivariate datasets using nonparametric statistics \label{Q1multvarnonparam}

For multivariate datasets, we calculated Spearman and Kendall correlations
and the statistics 
$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$,
$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$
for all pairs of sampling locations, and we then computed means across the pairs
for each statistic. We used a spatial resampling scheme 
(Appendix \ref{spatial_resamp}) to calculate confidence intervals of these means.

For the green spruce aphid abundance data, mean values across all pairs of 
locations of the Spearman and Kendall correlations and the statistics 
$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ were positive and
confidence intervals excluded $0$ (Table \ref{tab_count_resamp}). 
Mean values of the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$
also were always positive and had confidence intervals that excluded $0$ 
(Table \ref{tab_count_resamp}), confirming the results in the main text on 
asymmetry of tail dependence and 
stronger lower- than upper-tail dependence for this dataset. Each of the statistics 
was plotted against geographic (great-circle) distances between sampling locations, 
across all pairs of sampling locations (Fig. \ref{fig_aphid_count}).

Asymmetry of tail dependence was also verified by our nonparametric methods for
the other multivariate datasets. For instance, the $95\%$ confidence intervals of the
mean over pairs of sampling locations of the statistic $\Ps_l-\Ps_u$ were
$(-0.085,-0.041)$ for the leaf-curling plum aphid first flight data, 
indicating greater upper-tail dependence, and consistent with
<!--***DAN: Shya, pls make above results autolink-->
the results of Table \ref{MT-tab_multivar_summary} (row 9). 
For the *Ceratium furca* abundance data, confidence intervals were
$(0.011,0.044)$, indicating greater lower-tail dependence, and again
<!--***DAN: Shya, pls make above results autolink-->
consistent with the results of Table \ref{MT-tab_multivar_summary} (row 9).
For the methane flux data, confidence intervals were 
$(-0.008,0.019)$.

# Spatial resampling with multivariate datasets for nonparametric statistics \label{spatial_resamp}

The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
widely used spline correlogram method; the same scheme was also used by @walteretal2017 
as part of a method for 
performing model selection among matrix regression models. For each pair of 
locations $i \neq j$, the statistic
of interest, here denoted $s_{ij}$ ($\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, 
$\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, 
$\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$) was computed for those locations. The 
values $s_{ij}$ were then arranged in
a matrix, $S$, with diagonal entries NA. The mean of the non-NA entries of $S$ was computed. For 
$\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, or $\Dsq_{l_b,u_b}$ this mean characterizes the 
average strength of 
dependence, across all pairs of distinct locations, for data in the portion of the distributions 
given by the bounds $l_b$ and $u_b$. For $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$, 
the mean characterizes average
asymmetry of tail dependence across all pairs of locations. 
Confidence intervals for the value of $\mean(S)$ were then computed as quantiles of a 
distribution of bootstrapped values. If $\mean(S)$ was significantly different from 0, as
revealed by the confidence intervals, for the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, 
or $\Dsq_u-\Dsq_l$, then tail dependence was significantly asymmetric, in contrast to a 
N-copula null hypothesis, and answering question Q1 from the Introduction.
Bootstrapping was performed by resampling the locations of measurement, with replacement. 
This corresponds to 
resampling, with replacement, the rows and columns of $S$, and is described in detail elsewhere 
[@bjornstadfalck2001 ; @walteretal2017].

<!--This chunk is for model selection results of soilCN dataset-->
```{r bivmodselect_Loecke, echo=F, cache=T, cache.extra=list(seed,v_CN,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("OurBiCopSelect.R")
set.seed(seed)
families_Loecke<-c(1,3:10,13:14,16:20) 
BivMS_res_Loecke<-OurBiCopSelect(u1=v_CN[,1],u2=v_CN[,2],families=families_Loecke,level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,gofnormal=F,status=F)
saveRDS(BivMS_res_Loecke,file="./Results/fitting_results/BivMS_soilCN.RDS")
```

<!--Display model selection results in a table-->
```{r tab_soilCNfit, echo=F, results="markup"}
#knitr::opts_knit$set(kable.force.latex = TRUE)
h<-BivMS_res_Loecke$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
colnames(h)<-c("copname","AICw","BICw","LT","UT")
knitr::kable(h, #longtable = T, 
             format = "latex",linesep = "",
      caption.short = "Model fitting results for soil C and N dataset",
      caption = "Results of fitting our 16 copula families to the soil C and N dataset. Copula name abbreviations are given in the main text. AICw = AIC weight; BICw = BIC weight; LT = the lower-tail dependence statistic for the indicated copula family, with fitted parameters; UT = the same for upper-tail dependence.\\label{tab_soilCNfit}",
      booktabs = TRUE)
```

<!--This chunk is for npa testing results of soilCN dataset-->
```{r stat_soilCN_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,v_CN,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
source("bivfunctionplot.R")
set.seed(seed) 
stat_soilCN_ga<-bivfunctionplot(v=v_CN,resloc="./Results/stat_results/stat_soilCN/",nametag="soilCN",numbin=10)
```

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/stat_results/stat_soilCN/soilCN_bivfunctionplot.pdf}
\caption[Nonparametric tests against an N-copula null, soil C and N data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for soil C and N data. As described in the main text, the values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data (red crosses) were compared to distributions of their values on 1000 Kendall- or Spearman-preserving N surrogates of the data (blue and green x's show 0.025 and 0.975 quantiles), separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$. Whenever the red cross was outside the range given by the x's, text at the top of panels indicates the number of surrogate values the real-data value was greater than or less than. For instance, a value $>N$ (respectively, $<N$) means the value of the statistic on real data was greater than (respectively, less than) its value on $N$ surrogates. When the statistic was greater than 975 or less than 975 surrogate values, it indicates significance ($95\%$ confidence level). When $\cor$ or $\Ps$ values (respectively, $\Dsq$ values) were greater than surrogates, it means dependence in that part of the distributions was stronger than (respectively, weaker than) expected from a normal-copula null hypothesis. \label{fig_soilCN_nonparam}}
\end{center}
\end{figure}

<!--Now make and display a table, this one about asymmetry of tail dependence-->
```{r make_table_stat_fn, echo=F}

fractxt<-function(fracval,numsurrog){
  ptxt<-''
  if (fracval>.5*numsurrog){                        # ?DCR why 0.5 factor?
    ptxt<-paste(">",fracval,sep='')
  } 
  if (fracval<=.5*numsurrog){
    ptxt<-paste("<",numsurrog-fracval,sep='')
  }
  return(ptxt)
}

make_table_stat_fn<-function(data){

 #atdres<-data.frame(Statistic=c("$\\cor_{0,0.1}-\\cor_{0.9,1}$",
 #                               "$\\Ps_{0,0.1}-\\Ps_{0.9,1}$",
 #                               "$\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$"),
 atdres<-data.frame(Statistic=c("cor$_{FS}$-cor$_{LS}$",
                               "P$_{FS}$-P$_{LS}$",
                                "D$_{LS}^2$-D$_{FS}^2$"),
                   Kendall=rep("",3),
                   Spearman=rep("",3),stringsAsFactors = F)

 atdres[1,2]<-fractxt(data$corlmcoru_frac_K,numsurrog = data$numsurrog_success_K)
 atdres[1,3]<-fractxt(data$corlmcoru_frac_S,numsurrog = data$numsurrog_success_S)
 atdres[2,2]<-fractxt(data$PlmPu_frac_K,numsurrog = data$numsurrog_success_K)
 atdres[2,3]<-fractxt(data$PlmPu_frac_S,numsurrog = data$numsurrog_success_S)
 atdres[3,2]<-fractxt(data$D2umD2l_frac_K,numsurrog = data$numsurrog_success_K)
 atdres[3,3]<-fractxt(data$D2umD2l_frac_S,numsurrog = data$numsurrog_success_S)

 return(atdres)

}
```

```{r tab_soilCN_asym,echo=F,results="markup"}
#knitr::opts_knit$set(kable.force.latex = TRUE)
tab_soilCN_asym<-make_table_stat_fn(data=stat_soilCN_ga)
#tab_soilCN_asym[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
knitr::kable(tab_soilCN_asym, booktabs = T, escape=F,
             format = "latex",linesep = "",
      caption.short = "Nonparametric asymmetry tests against an N-copula null, soil C and N",
      caption = "Soil C and N dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis. The values of the listed statistics (cor$_{FS}$-cor$_{LS}$ = $\\cor_{0,0.1}-\\cor_{0.9,1}$, P$_{FS}$-P$_{LS}$ = $\\Ps_{0,0.1}-\\Ps_{0.9,1}$, D$_{LS}^2$-D$_{FS}^2$ = $\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$; here FS stands for `first slice' and LS for `last slice') for real data were compared to their values for 1000 Kendall- or Spearman-preserving N surrogates, in seperate comparisons. A table entry $<X$ indicates the value of the given statistic on the data was less than its value on $X$ of the surrogates, so entries of the form $<X$ for $X$ equal to $975$ or above indicate that upper-tail dependence was significantly stronger than lower-tail dependence.\\label{tab_soilCN_asym}")
```

```{r bivmodselect_birds_BMR, echo=F, cache=T, cache.extra=list(seed,cop_b,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("OurBiCopSelect.R")
set.seed(seed)
BivMS_birds<-OurBiCopSelect(u1=cop_b[,1],u2=cop_b[,2],families=c(1,3:10,13:14,16:20),
                              level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,
                              gofnormal=F,status=F)
saveRDS(BivMS_birds,file="./Results/BMR_results/birds/BivMS_birds.RDS")
```

<!--Display model selection results for BMR:birds data in a table-->
```{r tab_bmr_birds_fit, echo=F, results='markup'}
h<-BivMS_birds$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
colnames(h)<-c("copname","AICw","BICw","LT","UT")
knitr::kable(h,
             format='latex',linesep = "",
             caption.short = "Model fitting results for the bird body mass and BMR dataset",
             caption = "Results of fitting our 16 copula families to the bird body mass and BMR dataset. Analogous to Fig. \\ref{tab_soilCNfit}, see caption of that figure for details. \\label{tab_bmr_birds_fit}",
             booktabs=T)
```

```{r stat_bmr_birds_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,cop_b,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
source("bivfunctionplot.R")
set.seed(seed)
stat_birds_ga<-bivfunctionplot(v=cop_b,resloc="./Results/BMR_results/birds/",nametag="birds_numbin_10",numbin = 10)
```

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/BMR_results/birds/birds_numbin_10_bivfunctionplot.pdf}
\caption[Nonparametric tests against an N-copula null, bird body mass and BMR data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for bird body mass and BMR data. Same format as Fig. \ref{fig_soilCN_nonparam}, see caption of that figure. \label{fig_bmr_birds_nonparam}}
\end{center}
\end{figure}

```{r tab_bmr_birds_asym, echo=F, results="markup"}
tab_bmr_birds_asym<-make_table_stat_fn(data=stat_birds_ga)
#tab_bmr_birds_asym[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
knitr::kable(tab_bmr_birds_asym,
             format="latex",linesep = "",escape=F,
             caption.short = "Nonparametric asymmetry tests against an N-copula null, bird body mass and BMR",
             caption = "Bird body mass and BMR dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis. Same format as Table \\ref{tab_soilCN_asym}, see caption of that figure.\\label{tab_bmr_birds_asym}",
             booktabs=T)
```

```{r bivmodselect_mammals_BMR, echo=F, cache=T, cache.extra=list(seed,cop_m,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("OurBiCopSelect.R")
set.seed(seed)
BivMS_mammals<-OurBiCopSelect(u1=cop_m[,1],u2=cop_m[,2],families=c(1,3:10,13:14,16:20),
                              level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,
                              gofnormal=F,status=F)
saveRDS(BivMS_mammals,file="./Results/BMR_results/mammals/BivMS_mammals.RDS")
```

<!--Display model selection results for BMR:mammals data in a table-->
```{r tab_bmr_mammals_fit, echo=F, results='markup'}
h<-BivMS_mammals$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
colnames(h)<-c("copname","AICw","BICw","LT","UT")
knitr::kable(h,
             format='latex',linesep = "",
             caption.short = "Model fitting results for the mammal body mass and BMR dataset",
             caption = "Results of fitting our 16 copula families to the mammal body mass and BMR dataset. Analogous to Fig. \\ref{tab_soilCNfit}, see caption of that figure for details. \\label{tab_bmr_mammals_fit}",
             booktabs=T)
```

```{r stat_bmr_mammals_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,cop_m,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
source("bivfunctionplot.R")
set.seed(seed)
stat_mammals_ga<-bivfunctionplot(v=cop_m,resloc="./Results/BMR_results/mammals/",nametag="mammals_numbin_10",numbin = 10)
```

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/BMR_results/mammals/mammals_numbin_10_bivfunctionplot.pdf}
\caption[Nonparametric tests against an N-copula null, mammal body mass and BMR data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for mammal body mass and BMR data. Same format as Fig. \ref{fig_soilCN_nonparam}, see caption of that figure. \label{fig_bmr_mammals_nonparam}}
\end{center}
\end{figure}

```{r tab_bmr_mammals_asym, echo=F, results="markup"}
tab_bmr_mammals_asym<-make_table_stat_fn(data=stat_mammals_ga)
#tab_bmr_mammals_asym[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
knitr::kable(tab_bmr_mammals_asym,
             format="latex",linesep = "",escape=F,
             caption.short = "Nonparametric asymmetry tests against an N-copula null, mammal body mass and BMR",
             caption = "Mammal body mass and BMR dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis. Same format as Table \\ref{tab_soilCN_asym}, see caption of that figure. \\label{tab_bmr_mammals_asym}",
             booktabs=T)
```

```{r bivmodselect_cc, echo=F, cache=T, cache.extra=list(seed,cop_HB_all_yr,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("OurBiCopSelect.R")
set.seed(seed)

BivMS_res_cc<-vector("list",length(cop_HB_all_yr))
names(BivMS_res_cc)<-names(cop_HB_all_yr)

for(i in 1:length(BivMS_res_cc)){
  BivMS_res_cc[[i]]<-OurBiCopSelect(u1=cop_HB_all_yr[[i]]$avg.H,u2=cop_HB_all_yr[[i]]$avg.Biomass,
                                    families=c(1,3:10,13:14,16:20),
                                    level=0.05,AICBIC="AIC",
                                    numBSsmall=100,pthresh=0.2,numBSlarge=1000,gofnormal=F,status=F)
}
saveRDS(BivMS_res_cc,file="./Results/Ceder_creek_results/BivMS_res_cc.RDS")
BivMS_res_cc<-readRDS("./Results/Ceder_creek_results/BivMS_res_cc.RDS")
```

<!--Display model selection results for Ceder-creek data for year 2000 in a table-->
```{r tab_cc2000fit, echo=F, results='markup'}
h<-BivMS_res_cc$`2000`$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
colnames(h)<-c("copname","AICw","BICw","LT","UT")
knitr::kable(h,
             format='latex',linesep = "",
             caption.short = "Model fitting results for the Cedar Creek dataset",
             caption = "Results of fitting our 16 copula families to the Cedar Creek dataset. Analogous to Fig. \\ref{tab_soilCNfit}, see caption of that figure for details. \\label{tab_cc2000fit}",
             booktabs=T)
```

```{r stat_cc_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,cop_HB_all_yr,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
source("bivfunctionplot.R")
set.seed(seed)

stat_cc_ga<-vector("list",length(cop_HB_all_yr))
names(stat_cc_ga)<-names(cop_HB_all_yr)

for(i in 1:length(stat_cc_ga)){
  temp<-bivfunctionplot(v=cop_HB_all_yr[[i]],
                             resloc="./Results/Ceder_creek_results/",
                             nametag=paste0(names(cop_HB_all_yr)[i],"cc_numbin_10",sep=""),
                             numbin = 10)
  stat_cc_ga[[i]]<-temp
}

saveRDS(stat_cc_ga,"./Results/Ceder_creek_results/stat_cc_ga.RDS")
```

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/Ceder_creek_results/2000cc_numbin_10_bivfunctionplot.pdf}
\caption[Nonparametric tests against an N-copula null, Cedar Creek data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for Cedar Creek data. Same format as Fig. \ref{fig_soilCN_nonparam}, see caption of that figure.\label{fig_cc2000_nonparam}}
\end{center}
\end{figure}

```{r tab_cc_asym_2000, echo=F, results="markup"}
tab_cc_asym_2000<-make_table_stat_fn(data=stat_cc_ga$`2000`)
#tab_cc_asym_2000[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
knitr::kable(tab_cc_asym_2000,
             format="latex",linesep = "",escape=F,
             caption.short = "Nonparametric asymmetry tests against an N-copula null, Cedar Creek data",
             caption = "Cedar Creek dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis. Same format as Table \\ref{tab_soilCN_asym}, see caption of that figure.\\label{tab_cc_asym_2000}",
             booktabs=T)
```

<!--Load the raw data-->
```{r read_aphid_data, echo=F}
source("aphid_data_calling.R")

# Aphid raw data for count
d0_count<-as.matrix(read.csv("./Data/Aphid_data/APHID_DATA_RAW/aphidtscount141117.csv",header=F))
data_aphid_count<-d_allsp_data(d0_count)

# Aphid raw data for first flight
d0_firstflight<-as.matrix(read.csv("./Data/Aphid_data/APHID_DATA_RAW/aphidtsfirstflight141117.csv",header=F))
data_aphid_ff<-d_allsp_data(d0_firstflight)

# Aphid raw data for flight duration   
#d0_flightduration<-as.matrix(read.csv("./Data/Aphid_data/APHID_DATA_RAW/aphidtsflightduration141117.csv",header=F))

# Time series analysis aphids winter avg temp (here taken as -ve)
temp_aphid_raw<-as.matrix(read.csv("./Data/Aphid_data/APHID_DATA_RAW/aphidmonthlytempdata.csv",header = F))

temp_aphid<-vector("list",dim(temp_aphid_raw)[1])
names(temp_aphid)<-paste("loc",c(1:dim(temp_aphid_raw)[1]),sep="")

for(il in 1:length(temp_aphid)){

  s<-matrix(temp_aphid_raw[il,],36,12,byrow = T)
  rownames(s)<-c(1975:2010)
  colnames(s)<-c("jan","feb","mar","apr","may","jun","jul","aug","sep","oct","nov","dec")

  temp_aphid[[il]]<-s
}

winter_avg_temp<-matrix(NA,11,35)
rownames(winter_avg_temp)<-paste("loc",c(1:dim(temp_aphid_raw)[1]),sep="")
colnames(winter_avg_temp)<-c(1976:2010)

for(il in 1:length(temp_aphid)){
  temploc<-temp_aphid[[il]]
  for(i in 2:dim(temploc)[1]){
    t<-c(temploc[i-1,12],temploc[i,1:3]) # previous yr dec to present year march averaging temp.
    winter_avg_temp[il,i-1]<-mean(t,na.rm = T)
  }
}

aphid_winter_neg_temp<-vector("list",11) #initializing
names(aphid_winter_neg_temp)<-rownames(winter_avg_temp)

for(i in 1:11){
  aphid_winter_neg_temp[[i]]<-data.frame(Year=c(1976:2010),Dat=-winter_avg_temp[i,])
}

temp_aphid<-list(aphid_winter_neg_temp)
names(temp_aphid)<-"aphid_winter_neg_temp"
saveRDS(temp_aphid,"./Results/temp_aphid_winter_neg_temp_avg_data.RDS")
```

```{r modelsel_and_npa_aphid_count_allsp, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"),mtime("NonParamStat.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}

set.seed(seed)
library(parallel)
source("FittingCopula_selective_loc.R")
RES_all_sp_aphid_count<-mclapply(X=c(1:20),FUN=RES_single_sp,d_allsp=data_aphid_count,
                                 families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=30,
                                 good_loc_given=F,good_loc=NA,mc.cores=3)
saveRDS(RES_all_sp_aphid_count,"./Results/fitting_results/RES_all_sp_aphid_count.RDS")

source("good_loclist.R")
source("NonParamStat.R")
resloc<-"./Results/stat_results/stat_aphid_count/"
longs<-c(-4.567,0.57,-3.069,-3.312,-2.637,-1.682,-2.763,-0.356,-3.454,0.427,0.939)
lats<-c(55.477,52.26,56.457,55.949,52.125,55.213,53.854,51.807,50.628,51.733,51.185)

stat_aphid_count_allsp<-vector("list",20)

for(sp in 1:20){
  nm<-paste("sp_",sp,sep="")
  resloc2<-paste(resloc,nm,sep="")
   if (!dir.exists(resloc2)){
    dir.create(resloc2)
  }
  good_loc<-good_loclist(d_allsp=data_aphid_count,sp=sp,data_pt_thrs=30)

  temp<-multcall(d_allsp=data_aphid_count,sp=sp,lats=lats,longs=longs,
                 pfname=paste(resloc2,"/","Sp_",sp,sep=''),good_loc=good_loc)
  stat_aphid_count_allsp[[sp]]<-temp
}
saveRDS(stat_aphid_count_allsp,paste(resloc,file="stat_aphid_count_all_sp.RDS",sep=''))

summary_aphid_count_allsp<-as.data.frame(matrix(NA,nrow=20,ncol=8))
colnames(summary_aphid_count_allsp)<-c("species","#LTmUT>0","#LTmUT<0","val_LTmUT>0","val_LTmUT<0",
                                     "CorlmCoru_mean_CI","PlmPu_mean_CI","D2umD2l_mean_CI")

for(sp in c(1:20)){
  #cat("sp=",sp,"\n")
  summary_aphid_count_allsp$species[sp]<-sp
  LTmUT<-(RES_all_sp_aphid_count[[sp]]$LTdep_AICw - RES_all_sp_aphid_count[[sp]]$UTdep_AICw) 
  summary_aphid_count_allsp$`#LTmUT>0`[sp]<-sum(LTmUT>0,na.rm=T)
  summary_aphid_count_allsp$`#LTmUT<0`[sp]<-sum(LTmUT<0,na.rm=T)
  ind_LTmUT_pos<-which(LTmUT>0,arr.ind = T)
  summary_aphid_count_allsp$`val_LTmUT>0`[sp]<-sum(LTmUT[ind_LTmUT_pos])
  ind_LTmUT_neg<-which(LTmUT<0,arr.ind = T)
  summary_aphid_count_allsp$`val_LTmUT<0`[sp]<-sum(LTmUT[ind_LTmUT_neg])

  summary_aphid_count_allsp$CorlmCoru_mean_CI[sp]<-list(
    c(unname(round(stat_aphid_count_allsp[[sp]]$numericdf[9,c(2:4)],3))))
  summary_aphid_count_allsp$PlmPu_mean_CI[sp]<-list(
    c(unname(round(stat_aphid_count_allsp[[sp]]$numericdf[10,c(2:4)],3))))
  summary_aphid_count_allsp$D2umD2l_mean_CI[sp]<-list(
    c(unname(round(stat_aphid_count_allsp[[sp]]$numericdf[11,c(2:4)],3))))
}
saveRDS(summary_aphid_count_allsp,"./Results/summary_aphid_count_allsp.RDS")
```

<!--Do the model selection for aphid count data and then showing results in Tables-->
```{r RES_aphid_count, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}

source("FittingCopula_selective_loc.R")  # top most hierarchy source file
set.seed(seed)
#cat(paste("start-time: ",Sys.time(),"\n")) 
RES_aphid_count<-RES_single_sp(d_allsp=data_aphid_count,sp=10,families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=30,
                               good_loc_given=F,good_loc=NA)
#cat(paste("stop-time: ",Sys.time(),"\n"))
resloc<-"./Results/fitting_results/"
saveRDS(RES_aphid_count,paste(resloc,file="AphidCopulaFit_selecloc_count_species_10.RDS",sep=""))
```

```{r tab_count_bestcop, echo=F, results="markup"}
#library(dplyr)
#knitr::opts_knit$set(kable.force.latex = TRUE)
knitr::kable(RES_aphid_count$info_ord_copname[,,1], #longtable = T,
             format='latex',linesep = "",
             caption.short = "Best-fitting copulas, green spruce aphid abundance data",
             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for green spruce aphid abundance data. Copulas 
considered were the normal, Clayton, Gumbel, Frank, Joe, BB1, BB6, BB7, BB8, survival Clayton, survival Gumbel, 
survival Joe, survival BB1, survival BB6, survival BB7, and survival BB8 copulas, with abbreviations in the methods for Q1. NA occurs along the diagonal of the matrix because the diagonal represents a sampling location compared to itself.\\label{tab_count_bestcop}",
             booktabs=TRUE
             )#%>% # only with 'latex'/'html' format
 # kableExtra::kable_styling(latex_options = c("striped","HOLD_position"))
```

```{r tab_count_AICw, echo=F, results='markup'}
knitr::kable(RES_aphid_count$info_ord_AICw[,,1], 
             format='latex',linesep = "",digits=3,
             caption.short = "AIC weights, best-fitting copulas, green spruce aphid abundance data",
             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for green spruce aphid abundance data.\\label{tab_count_AICw}",
             booktabs=TRUE
             )
```

```{r tab_count_AICw_normal, echo=F, results='markup'}
source("normal_cop_AICw.R")
knitr::kable(normal_cop_AICw_matrix(r=RES_aphid_count),
             format='latex',linesep = "",digits=3,
             caption.short = "AIC weights, N copulas, green spruce aphid abundance data",
             caption = "$\\text{AIC}_{\\text{w}}$ for the N copula family for each location pair for green spruce aphid abundance data, for comparison with Table \\ref{tab_count_AICw}. \\label{tab_count_AICw_normal}",
             booktabs=TRUE
             )
```

```{r tab_count_p_CvM, echo=F, results='markup'}
knitr::kable(RES_aphid_count$gfc_p_CvM,
             format='latex',linesep = "",digits=3,
             caption.short="Goodness of fit of best-fitting copulas, green spruce aphid abundance data",
             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for green spruce abundance data.\\label{tab_count_p_CvM}",
             booktabs=TRUE
             )
```

```{r tab_count_p_KS, echo=F, results='markup'}
knitr::kable(RES_aphid_count$gfc_p_KS,
             format='latex',linesep = "",digits=3,
             caption.short="Goodness of fit of best-fitting copulas, green spruce aphid abundance data",
             caption = "$P$-value results of a Kolmogorov-Smirnov-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for green spruce abundance data.\\label{tab_count_p_KS}",
             booktabs=TRUE
             )
```

<!--\newpage-->
```{r tab_count_LTdep_AICw, echo=F, results='markup'}
knitr::kable(RES_aphid_count$LTdep_AICw,
             format='latex',linesep = "",digits=3,
             caption.short="Lower-tail dependence, green spruce aphid abundance data",
             caption = "Model-averaged lower-tail dependence statistics for each location pair for green spruce aphid abundance data.\\label{tab_count_LTdep_AICw}",
             booktabs=TRUE
             )
```

<!--\newpage-->
```{r tab_count_UTdep_AICw, echo=F, results='markup'}
knitr::kable(RES_aphid_count$UTdep_AICw,
             format='latex',linesep = "",digits=3,
             caption.short="Upper-tail dependence, green spruce aphid abundance data",
             caption = "Model-averaged upper-tail dependence statistics for each location pair for green spruce aphid abundance data.\\label{tab_count_UTdep_AICw}",
             booktabs=TRUE
             )
```

<!--\newpage-->
```{r tab_count_LTmUT, echo=F, results='markup'}
knitr::kable(RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw,
             digits = 3,linesep = "",
             format='latex',
             caption.short="Lower- minus upper-tail dependence, green spruce aphid abundance data",
             caption = "Model-averaged lower-tail dependence statistics minus model-averaged upper-tail dependence statistics for each location pair for green spruce aphid abundance data.\\label{tab_count_LTmUT}",
             booktabs=TRUE
             )
```

<!--Do the main computations for nonparametric analysis of aphid count data-->
```{r stat_aphid_count, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("good_loclist.R"),mtime("NonParamStat.R"),mtime("vivj_matrix.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}

source("good_loclist.R")
source("NonParamStat.R")
set.seed(seed) 
resloc<-"./Results/stat_results/stat_aphid_count/"

sp<-10
good_loc <- good_loclist(d_allsp=data_aphid_count,sp=sp,data_pt_thrs=30)
longs<-c(-4.567,0.57,-3.069,-3.312,-2.637,-1.682,-2.763,-0.356,-3.454,0.427,0.939)
lats<-c(55.477,52.26,56.457,55.949,52.125,55.213,53.854,51.807,50.628,51.733,51.185)

stat_aphid_count<-multcall(d_allsp=data_aphid_count,sp=sp,lats=lats,longs=longs,pfname=paste(resloc,"Sp_",sp,sep=''),good_loc=good_loc)
saveRDS(stat_aphid_count,paste(resloc,file="stat_aphid_count_sp_10.RDS",sep='')) 
```

<!--Show the statistical results in a table-->
```{r tab_count_resamp,echo=F,results="markup"}
s<-stat_aphid_count$numericdf[,c(3,2,4)]
sc<-as.data.frame(c("Spearman","Kendall","cor$_{l}$","cor$_{u}$","P$_{l}$","P$_{u}$","D$_{l}^2$","D$_{u}^2$",
                    "cor$_{l}$ - cor$_{u}$","P$_{l}$ - P$_{u}$","D$_{u}^2$ - D$_{l}^2$"))
s<-cbind(sc,s)
knitr::kable(s, digits = 3,
             col.names=c("","2.5$^{th}$ quantile", "Mean", "97.5$^{th}$ quantile"),
             format = "latex",linesep = "", escape = F,align="c",
             caption.short="Results for nonparametric statistics, green spruce aphid abundance data",
             caption = "Average values of statistics across all locations pairs and confidence intervals based on spatial resampling for green spruce aphid abundance data.\\label{tab_count_resamp}",
             booktabs=TRUE)
```

\begin{figure}[!h]
\textbf{ \hspace{3.4 cm} (A) \hspace{3.2 cm} (B) \hspace{3.2 cm} (C) \hspace{3.4 cm} (D)} \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_count/Sp_10_Spearman_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_count/Sp_10_Kendall_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_count/Sp_10_Corl_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_count/Sp_10_Coru_vs_D.pdf}\\
\end{center}
\textbf{ \hspace{3.4 cm} (E) \hspace{3.2 cm} (F) \hspace{3.2 cm} (G) \hspace{3.4 cm} (H)} \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_count/Sp_10_Pl_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_count/Sp_10_Pu_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_count/Sp_10_D2u_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_count/Sp_10_D2l_vs_D.pdf}\\
\end{center}
\textbf{ \hspace{5.4 cm} (I) \hspace{3.4 cm} (J) \hspace{3.2 cm} (K) } \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_count/Sp_10_Corl-Coru_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_count/Sp_10_Pl-Pu_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_count/Sp_10_D2u-D2l_vs_D.pdf}
\caption[Nonparametric tail statistics versus distance between locations for green spruce aphid abundance data]{Nonparametric tail statistics versus great-circle distance, $D$, between sampling locations for green spruce aphid abundance data. \label{fig_aphid_count}}
\end{center}
\end{figure}

```{r modelsel_and_npa_aphid_ff_allsp, echo=F, cache=T, cache.extra=list(seed,data_aphid_ff,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"),mtime("NonParamStat.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}

set.seed(seed)
library(parallel)
source("FittingCopula_selective_loc.R")
RES_all_sp_aphid_ff<-mclapply(X=c(1:20),FUN=RES_single_sp,d_allsp=data_aphid_ff,
                                 families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=30,
                                 good_loc_given=F,good_loc=NA,mc.cores=3)
saveRDS(RES_all_sp_aphid_ff,"./Results/fitting_results/RES_all_sp_aphid_ff.RDS")

source("good_loclist.R")
source("NonParamStat.R")
resloc<-"./Results/stat_results/stat_aphid_ff/"
longs<-c(-4.567,0.57,-3.069,-3.312,-2.637,-1.682,-2.763,-0.356,-3.454,0.427,0.939)
lats<-c(55.477,52.26,56.457,55.949,52.125,55.213,53.854,51.807,50.628,51.733,51.185)

stat_aphid_ff_allsp<-vector("list",20)

for(sp in 1:20){
  nm<-paste("sp_",sp,sep="")
  resloc2<-paste(resloc,nm,sep="")
   if (!dir.exists(resloc2)){
    dir.create(resloc2)
  }
  good_loc<-good_loclist(d_allsp=data_aphid_ff,sp=sp,data_pt_thrs=30)

  temp<-multcall(d_allsp=data_aphid_ff,sp=sp,lats=lats,longs=longs,
                 pfname=paste(resloc2,"/","Sp_",sp,sep=''),good_loc=good_loc)
  stat_aphid_ff_allsp[[sp]]<-temp
}
saveRDS(stat_aphid_ff_allsp,paste(resloc,file="stat_aphid_ff_all_sp.RDS",sep=''))

summary_aphid_ff_allsp<-as.data.frame(matrix(NA,nrow=20,ncol=8))
colnames(summary_aphid_ff_allsp)<-c("species","#LTmUT>0","#LTmUT<0","val_LTmUT>0","val_LTmUT<0",
                                     "CorlmCoru_mean_CI","PlmPu_mean_CI","D2umD2l_mean_CI")

for(sp in c(1:20)){
  #cat("sp=",sp,"\n")
  summary_aphid_ff_allsp$species[sp]<-sp
  LTmUT<-(RES_all_sp_aphid_ff[[sp]]$LTdep_AICw - RES_all_sp_aphid_ff[[sp]]$UTdep_AICw) 
  summary_aphid_ff_allsp$`#LTmUT>0`[sp]<-sum(LTmUT>0,na.rm=T)
  summary_aphid_ff_allsp$`#LTmUT<0`[sp]<-sum(LTmUT<0,na.rm=T)
  ind_LTmUT_pos<-which(LTmUT>0,arr.ind = T)
  summary_aphid_ff_allsp$`val_LTmUT>0`[sp]<-sum(LTmUT[ind_LTmUT_pos])
  ind_LTmUT_neg<-which(LTmUT<0,arr.ind = T)
  summary_aphid_ff_allsp$`val_LTmUT<0`[sp]<-sum(LTmUT[ind_LTmUT_neg])

  summary_aphid_ff_allsp$CorlmCoru_mean_CI[sp]<-list(
    c(unname(round(stat_aphid_ff_allsp[[sp]]$numericdf[9,c(2:4)],3))))
  summary_aphid_ff_allsp$PlmPu_mean_CI[sp]<-list(
    c(unname(round(stat_aphid_ff_allsp[[sp]]$numericdf[10,c(2:4)],3))))
  summary_aphid_ff_allsp$D2umD2l_mean_CI[sp]<-list(
    c(unname(round(stat_aphid_ff_allsp[[sp]]$numericdf[11,c(2:4)],3))))
}
saveRDS(summary_aphid_ff_allsp,"./Results/summary_aphid_ff_allsp.RDS")
```

<!--Do the model selection for aphid ff data and then showing results in Tables-->
```{r RES_aphid_ff, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}

source("FittingCopula_selective_loc.R")  # top most hierarchy source file
set.seed(seed)
#cat(paste("start-time: ",Sys.time(),"\n"))
RES_aphid_ff<-RES_single_sp(d_allsp=data_aphid_ff,sp=11,families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=30,
                            good_loc_given=F,good_loc=NA)
#cat(paste("stop-time: ",Sys.time(),"\n")) 
resloc<-"./Results/fitting_results/"
saveRDS(RES_aphid_ff,paste(resloc,file="AphidCopulaFit_selecloc_firstflight_species_11.RDS",sep=""))
```

```{r tab_ff_bestcop, echo=F, results="markup"}
knitr::kable(RES_aphid_ff$info_ord_copname[,,1], #longtable = T,
             format='latex',linesep = "",
             caption.short = "Best-fitting copulas, leaf-curling plum aphid first flight data",
             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for leaf-curling plum aphid first flight data. Copulas considered were as in Table \\ref{tab_count_bestcop}, with abbreviations in the methods for Q1. NA occurs along the diagonal of the matrix because the diagonal represents a sampling location compared to itself.  \\label{tab_ff_bestcop}",
             booktabs=TRUE
             )
```

```{r tab_ff_AICw, echo=F, results='markup'}
knitr::kable(RES_aphid_ff$info_ord_AICw[,,1], 
             format='latex',linesep = "",digits=3,
             caption.short="AIC weights, best-fitting copulas, leaf-curling plum aphid first flight data",
             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for leaf-curling plum aphid first flight data.\\label{tab_ff_AICw}",
             booktabs=TRUE
             )
```

<!--\newpage-->
```{r tab_ff_AICw_normal, echo=F, results='markup'}
source("normal_cop_AICw.R")
knitr::kable(normal_cop_AICw_matrix(r=RES_aphid_ff),
             format='latex',linesep = "",digits=3,
             caption.short="AIC weights, N copulas, leaf-curling plum aphid first flight data",
             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for leaf-curlung plum aphid first flight data, for comparison with Table \\ref{tab_ff_AICw}.\\label{tab_ff_AICw_normal}",
             booktabs=TRUE
             )
```

```{r tab_ff_p_CvM, echo=F, results='markup'}
knitr::kable(RES_aphid_ff$gfc_p_CvM,
             format='latex',linesep = "",digits=3,
             caption.short="Goodness of fit of best-fitting copulas, leaf-curling plum aphid first flight data",
             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for leaf-curling plum aphid first flight data.\\label{tab_ff_p_CvM}",
             booktabs=TRUE
             )
```

```{r tab_ff_p_KS, echo=F, results='markup'}
knitr::kable(RES_aphid_ff$gfc_p_KS,
             format='latex',linesep = "",digits=3,
             caption.short="Goodness of fit of best-fitting copulas, leaf-curling plum aphid first flight data",
             caption = "$P$-value results of a Kolmogorov-Smirnov-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for leaf-curling plum aphid first flight data.\\label{tab_ff_p_KS}",
             booktabs=TRUE
             )
```

```{r tab_ff_LTdep_AICw, echo=F, results='markup'}
knitr::kable(RES_aphid_ff$LTdep_AICw,
             format='latex',linesep = "",digits=3,
             caption.short="Lower-tail dependence, leaf-curling plum aphid first flight data",
             caption = "Model-averaged lower-tail dependence statistics for each location pair for leaf-curling plum aphid first flight data.\\label{tab_ff_LTdep_AICw}",
             booktabs=TRUE
             )
```

```{r tab_ff_UTdep_AICw, echo=F, results='markup'}
knitr::kable(RES_aphid_ff$UTdep_AICw,
             format='latex',linesep = "",digits=3,
             caption.short="Upper-tail dependence, leaf-curling plum aphid first flight data",
             caption = "Model-averaged upper-tail dependence statistics for each location pair for leaf-curling plum aphid first flight data.\\label{tab_ff_UTdep_AICw}",
             booktabs=TRUE
             )
```

```{r tab_ff_LTmUT, echo=F, results='markup'}
knitr::kable(RES_aphid_ff$LTdep_AICw-RES_aphid_ff$UTdep_AICw,
             digits = 3,
             format='latex',linesep = "",
             caption.short="Lower- minus upper-tail dependence, leaf-curling plum aphid first flight data",
             caption = "Model-averaged lower-tail dependence statistics minus model-averaged upper-tail dependence statistics for each location pair for leaf-curling plum aphid first flight data.\\label{tab_ff_LTmUT}",
             booktabs=TRUE
             )
```

<!--Do the main computations for nonparametric analysis of aphid ff data-->
```{r stat_aphid_ff, echo=F, cache=T,  cache.extra=list(seed,data_aphid_ff,mtime("good_loclist.R"),mtime("NonParamStat.R"),mtime("vivj_matrix.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}

source("good_loclist.R")
source("NonParamStat.R")
set.seed(seed)
resloc<-"./Results/stat_results/stat_aphid_ff/"

sp<-11
good_loc<-good_loclist(d_allsp=data_aphid_ff,sp=sp,data_pt_thrs=30)
longs<-c(-4.567,0.57,-3.069,-3.312,-2.637,-1.682,-2.763,-0.356,-3.454,0.427,0.939)
lats<-c(55.477,52.26,56.457,55.949,52.125,55.213,53.854,51.807,50.628,51.733,51.185)

stat_aphid_ff<-multcall(d_allsp=data_aphid_ff,sp=sp,lats=lats,longs=longs,pfname=paste(resloc,"Sp_",sp,sep=''),good_loc=good_loc) 
saveRDS(stat_aphid_ff,paste(resloc,file="stat_aphid_ff_sp_11.RDS",sep='')) 
```

<!--display the main statistical results in a table-->
```{r tab_ff_resamp,echo=F,results="markup"}
s<-stat_aphid_ff$numericdf[,c(3,2,4)]
sc<-as.data.frame(c("Spearman","Kendall","cor$_{l}$","cor$_{u}$","P$_{l}$","P$_{u}$","D$_{l}^2$","D$_{u}^2$",
                    "cor$_{l}$ - cor$_{u}$","P$_{l}$ - P$_{u}$","D$_{u}^2$ - D$_{l}^2$"))
s<-cbind(sc,s)
knitr::kable(s, digits = 3,
             col.names=c("","2.5$^{th}$ quantile", "Mean", "97.5$^{th}$ quantile"),
             format = "latex",linesep = "", escape = F,align="c",
             caption.short="Results for nonparametric statistics, leaf-curling plum aphid first flight data",
             caption = "Average values of statistics across all location pairs and confidence intervals based on spatial resampling for leaf-curling plum aphid first flight data.\\label{tab_ff_resamp}",
             booktabs=TRUE)
```

\begin{figure}[!h]
\textbf{ \hspace{3.4 cm} (A) \hspace{3.2 cm} (B) \hspace{3.2 cm} (C) \hspace{3.4 cm} (D)} \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_ff/Sp_11_Spearman_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_ff/Sp_11_Kendall_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_ff/Sp_11_Corl_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_ff/Sp_11_Coru_vs_D.pdf}\\
\end{center}
\textbf{ \hspace{3.4 cm} (E) \hspace{3.2 cm} (F) \hspace{3.2 cm} (G) \hspace{3.4 cm} (H)} \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_ff/Sp_11_Pl_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_ff/Sp_11_Pu_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_ff/Sp_11_D2u_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_ff/Sp_11_D2l_vs_D.pdf}\\
\end{center}
\textbf{ \hspace{5.4 cm} (I) \hspace{3.4 cm} (J) \hspace{3.2 cm} (K) } \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_ff/Sp_11_Corl-Coru_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_ff/Sp_11_Pl-Pu_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_ff/Sp_11_D2u-D2l_vs_D.pdf}
\caption[Nonparametric tail statistics versus distance between locations for leaf-curling plum aphid first flight data]{Nonparametric tail statistics versus great-circle distance, $D$, between sampling locations for leaf-curling plum aphid first flight data. \label{fig_aphid_ff}}
\end{center}
\end{figure}

<!--Load the data-->
```{r read_plankton_north_sea_data, echo=F}
source("plankton_north_sea_data_calling.R")
d0_plankton_north_sea<-as.matrix(read.csv("./Data/Plankton_North_Sea_data/planktontimeseries201117_4.csv",header=F))
data_plankton_north_sea<-d_allsp_data_plankton_north_sea(d0_plankton_north_sea)
```

```{r modelsel_and_npa_pns_allsp, echo=F, cache=T, cache.extra=list(seed,data_plankton_north_sea,longs_pns,lats_pns,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"),mtime("NonParamStat.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}

set.seed(seed)
library(parallel)
source("FittingCopula_selective_loc.R")
RES_all_sp_pns<-mclapply(X=c(1:22),FUN=RES_single_sp,d_allsp=data_plankton_north_sea,
                              families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=45,
                              good_loc_given=F,good_loc=NA,mc.cores=3)
saveRDS(RES_all_sp_pns,"./Results/fitting_results/RES_all_sp_pns.RDS")

source("good_loclist.R")
source("NonParamStat.R")
resloc<-"./Results/stat_results/stat_plankton_north_sea/"

stat_plankton_north_sea_allsp<-vector("list",22)

for(sp in 1:22){
  nm<-paste("sp_",sp,sep="")
  resloc2<-paste(resloc,nm,sep="")
   if (!dir.exists(resloc2)){
    dir.create(resloc2)
  }
  good_loc<-good_loclist(d_allsp=data_plankton_north_sea,sp=sp,data_pt_thrs=45)

  temp<-multcall(d_allsp=data_plankton_north_sea,sp=sp,lats=lats_pns,longs=longs_pns,
                                      pfname=paste(resloc2,"/","Sp_",sp,sep=''),good_loc=good_loc)
  stat_plankton_north_sea_allsp[[sp]]<-temp
}
saveRDS(stat_plankton_north_sea_allsp,paste(resloc,file="stat_plankton_north_sea_allsp.RDS",sep=''))

summary_pns_allsp<-as.data.frame(matrix(NA,nrow=22,ncol=8))
colnames(summary_pns_allsp)<-c("species","#LTmUT>0","#LTmUT<0","val_LTmUT>0","val_LTmUT<0",
                                     "CorlmCoru_mean_CI","PlmPu_mean_CI","D2umD2l_mean_CI")

for(sp in c(1:22)){
  #cat("sp=",sp,"\n")
  summary_pns_allsp$species[sp]<-sp
  LTmUT<-(RES_all_sp_pns[[sp]]$LTdep_AICw - RES_all_sp_pns[[sp]]$UTdep_AICw) 
  summary_pns_allsp$`#LTmUT>0`[sp]<-sum(LTmUT>0,na.rm=T)
  summary_pns_allsp$`#LTmUT<0`[sp]<-sum(LTmUT<0,na.rm=T)
  ind_LTmUT_pos<-which(LTmUT>0,arr.ind = T)
  summary_pns_allsp$`val_LTmUT>0`[sp]<-sum(LTmUT[ind_LTmUT_pos])
  ind_LTmUT_neg<-which(LTmUT<0,arr.ind = T)
  summary_pns_allsp$`val_LTmUT<0`[sp]<-sum(LTmUT[ind_LTmUT_neg])

  summary_pns_allsp$CorlmCoru_mean_CI[sp]<-list(
    c(unname(round(stat_plankton_north_sea_allsp[[sp]]$numericdf[9,c(2:4)],3))))
  summary_pns_allsp$PlmPu_mean_CI[sp]<-list(
    c(unname(round(stat_plankton_north_sea_allsp[[sp]]$numericdf[10,c(2:4)],3))))
  summary_pns_allsp$D2umD2l_mean_CI[sp]<-list(
    c(unname(round(stat_plankton_north_sea_allsp[[sp]]$numericdf[11,c(2:4)],3))))
}
saveRDS(summary_pns_allsp,"./Results/summary_pns_allsp.RDS")
```

<!--Do the model selection for plankton data and then showing results in Tables-->
```{r RES_plankton_north_sea, echo=F, cache=T, results="hide", cache.extra=list(seed,data_plankton_north_sea,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}

source("FittingCopula_selective_loc.R")  # top most hierarchy source file

set.seed(seed)
#cat(paste("start-time: ",Sys.time(),"\n")) 
RES_plankton_north_sea<-RES_single_sp(d_allsp=data_plankton_north_sea,sp=16,families=c(1,3:10,13,14,16:20),level=0.05,
                                      data_pt_thrs=45,good_loc_given=F,good_loc=NA)
#cat(paste("stop-time: ",Sys.time(),"\n"))    
resloc<-"./Results/fitting_results/"
saveRDS(RES_plankton_north_sea,paste(resloc,file="Plankton_North_Sea_CopulaFit_selecloc_species_16.RDS",sep=""))
```

```{r tab_pns_bestcop, echo=F, results="markup"}
knitr::kable(RES_plankton_north_sea$info_ord_copname[,,1], #longtable=T,
             format='latex', linesep = "",
             caption.short = "Best-fitting copulas, \\textit{Ceratium furca} abundance data",
             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for \\textit{Ceratium furca} abundance data. See methods for Q1 for copula abbreviations. Indep.=Independent. NA occurs along the diagonal because diagonal entries represent a sampling location compared to itself.\\label{tab_pns_bestcop}",
             booktabs=TRUE
             )
```

```{r tab_pns_AICw,echo=F,results="markup"}
h<-RES_plankton_north_sea$info_ord_AICw[,,1]
h[is.infinite(h)]<-NA
knitr::kable(h,
             format="latex",
             digits=3,linesep = "",
             caption.short="AIC weights, best-fitting copulas, \\textit{Ceratium furca} abundance data",
             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for \\textit{Ceratium furca} abundance data.\\label{tab_pns_AICw}",
             booktabs=TRUE)
```

```{r tab_pns_AICw_normal,echo=F,results="markup"}
source("normal_cop_AICw.R")
h<-normal_cop_AICw_matrix(r=RES_plankton_north_sea)
h[is.infinite(h)]<-NA
knitr::kable(h,
             format="latex",linesep = "",
             digits=3,
             caption.short="AIC weights, N copulas, \\textit{Ceratium furca} abundance data",
             caption = "$\\text{AIC}_{\\text{w}}$ for the N copula family for each location pair for \\textit{Ceratium furca} abundance data, for comparison with Table \\ref{tab_pns_AICw}.\\label{tab_pns_AICw_normal}",
             booktabs=TRUE)
```

```{r tab_pns_p_CvM,echo=F,results="markup"}
h<-RES_plankton_north_sea$gfc_p_CvM
h[is.infinite(h)]<-NA
knitr::kable(h,
             format="latex",linesep = "",
             digits=3,
             caption.short="Goodness of fit of best-fitting copulas, \\textit{Ceratium furca} abundance data",
             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for \\textit{Ceratium furca} abundance data.\\label{tab_pns_p_CvM}",
             booktabs=TRUE)
```

```{r tab_pns_p_KS,echo=F,results="markup"}
h<-RES_plankton_north_sea$gfc_p_KS
h[is.infinite(h)]<-NA
knitr::kable(h,
             format="latex",linesep = "",
             digits=3,
             caption.short="Goodness of fit of best-fitting copulas, \\textit{Ceratium furca} abundance data",
             caption = "$P$-value results of a Kolmogorov-Smirnov-based test of
the goodness of fit of the lowest-AIC copula family for each pair of
locations for \\textit{Ceratium furca} abundance data.\\label{tab_pns_p_KS}",
             booktabs=TRUE)
```

```{r tab_pns_LTdep_AICw,echo=F,results="markup"}
h<-RES_plankton_north_sea$LTdep_AICw
h[is.infinite(h)]<-NA
knitr::kable(h,
             format="latex",linesep = "",
             digits=3,
             caption.short="Lower-tail dependence, \\textit{Ceratium furca} abundance data",
             caption = "Model-averaged lower-tail dependence statistics for each location pair for \\textit{Ceratium furca} abundance data.\\label{tab_pns_LTdep_AICw}",
             booktabs=TRUE)
```

```{r tab_pns_UTdep_AICw,echo=F,results="markup"}
h<-RES_plankton_north_sea$UTdep_AICw
h[is.infinite(h)]<-NA
knitr::kable(h,
             format="latex",linesep = "",
             digits=3,
             caption.short="Upper-tail dependence, \\textit{Ceratium furca} abundance data",
             caption = "Model-averaged upper-tail dependence statistics for each location pair for \\textit{Ceratium furca} abundance data.\\label{tab_pns_UTdep_AICw}",
             booktabs=TRUE)
```

```{r tab_pns_LTmUT,echo=F,results="markup"}
h<-RES_plankton_north_sea$LTdep_AICw-RES_plankton_north_sea$UTdep_AICw
h[is.nan(h)]<-NA
knitr::kable(h,
             format="latex",linesep = "",
             digits=3,
             caption.short="Lower- minus upper-tail dependence, \\textit{Ceratium furca} abundance data",
             caption = "Model-averaged lower-tail dependence statistics minus model-averaged upper-tail dependence statistics for each location pair for \\textit{Ceratium furca} abundance data.\\label{tab_pns_LTmUT}",
             booktabs=TRUE)
```

<!--Do the main computations for nonparametric analysis of plankton data-->
```{r stat_plankton_north_sea, echo=F, cache=T, cache=T,  cache.extra=list(seed,data_plankton_north_sea,longs_pns,lats_pns,mtime("good_loclist.R"),mtime("NonParamStat.R"),mtime("vivj_matrix.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}

source("good_loclist.R")
source("NonParamStat.R")
set.seed(seed)
resloc<-"./Results/stat_results/stat_plankton_north_sea/"

sp<-16
good_loc<-good_loclist(d_allsp=data_plankton_north_sea,sp=sp,data_pt_thrs=45)

stat_plankton_north_sea<-multcall(d_allsp=data_plankton_north_sea,sp=sp,lats=lats_pns,longs=longs_pns,
                                  pfname=paste(resloc,"Sp_",sp,sep=''),good_loc=good_loc) 
saveRDS(stat_plankton_north_sea,paste(resloc,file="stat_plankton_north_sea_sp_16.RDS",sep=''))
```

<!--display the main statistical results in a table-->
```{r tab_pns_resamp,echo=F,results="markup"}
s<-stat_plankton_north_sea$numericdf[,c(3,2,4)]
sc<-as.data.frame(c("Spearman","Kendall","cor$_{l}$","cor$_{u}$","P$_{l}$","P$_{u}$","D$_{l}^2$","D$_{u}^2$",
                    "cor$_{l}$ - cor$_{u}$","P$_{l}$ - P$_{u}$","D$_{u}^2$ - D$_{l}^2$"))
s<-cbind(sc,s)
knitr::kable(s, digits = 3,
             col.names=c("","2.5$^{th}$ quantile", "Mean", "97.5$^{th}$ quantile"),
             format = "latex",linesep = "", escape = F,align="c",
             caption.short="Results for nonparametric statistics, \\textit{Ceratium furca} abundance data",
             caption = "Average values of statistics across all locations pairs and confidence intervals based on spatial resampling for \\textit{Ceratium furca} abundance data.\\label{tab_pns_resamp}",
             booktabs=TRUE)
```

\begin{figure}[!h]
\textbf{ \hspace{3.4 cm} (A) \hspace{3.2 cm} (B) \hspace{3.2 cm} (C) \hspace{3.4 cm} (D)} \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Spearman_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Kendall_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Corl_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Coru_vs_D.pdf}\\
\end{center}
\textbf{ \hspace{3.4 cm} (E) \hspace{3.2 cm} (F) \hspace{3.2 cm} (G) \hspace{3.4 cm} (H)} \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Pl_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Pu_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_plankton_north_sea/Sp_16_D2u_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_plankton_north_sea/Sp_16_D2l_vs_D.pdf}\\
\end{center}
\textbf{ \hspace{5.4 cm} (I) \hspace{3.4 cm} (J) \hspace{3.2 cm} (K) } \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Corl-Coru_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Pl-Pu_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_plankton_north_sea/Sp_16_D2u-D2l_vs_D.pdf}
\caption[Nonparametric tail statistics versus distance between locations for \textit{Ceratium furca} abundance data]{Nonparametric tail statistics versus great-circle distance, $D$, between sampling locations for \textit{Ceratium furca} abundance data. \label{fig_pns}}
\end{center}
\end{figure}

<!--Load and format the data-->
```{r read_format_methane_data, echo=F}

methane_raw<-readRDS("./Data/Methane_data/methane_data_for_Shyamolina.rds")

myls <- vector("list", length = dim(methane_raw)[2]-1)
for(i in 2:dim(methane_raw)[2]){
  temp<-methane_raw[,c(1,i)]
  colnames(temp)<-c("Year","Dat")
  myls[[i-1]]<-temp
}
methane_data<-list(myls=myls)
```

<!--Do the main analysis-->
```{r RES_methane, echo=F, cache=T, results="hide", cache.extra=list(seed,methane_data,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}

source("FittingCopula_selective_loc.R")  # top most hierarchy source file

set.seed(seed)
#cat(paste("start-time: ",Sys.time(),"\n")) 
RES_methane<-RES_single_sp(d_allsp=methane_data,sp=1,families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=50,
                           good_loc_given=F,good_loc=NA)
#cat(paste("stop-time: ",Sys.time(),"\n"))    
resloc<-"./Results/fitting_results/"
saveRDS(RES_methane,paste(resloc,file="MethaneFluxFit_selecloc_data_pt_thrs_50.RDS",sep=""))
```

```{r tab_methane_bestcop, echo=F, results="markup"}
knitr::kable(RES_methane$info_ord_copname[,,1], #longtable=T,
             format='latex',linesep = "",
             caption.short = "Best-fitting copulas, methane flux data",
             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for methane flux data. See methods for Q1 for copula abbreviations. Indep.=Independent. NA occurs along the diagonal because diagonal entries represent a sampling location compared to itself.\\label{tab_methane_bestcop}",
             booktabs=TRUE
             )
```

```{r tab_methane_AICw,echo=F,results="markup"}
h<-RES_methane$info_ord_AICw[,,1]
h[is.infinite(h)]<-NA
knitr::kable(h,
             format="latex",linesep = "",
             digits=3,
             caption.short="AIC weights, best-fitting copulas, methane flux data",
             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for methane flux data.\\label{tab_methane_AICw}",
             booktabs=TRUE)
```

```{r tab_methane_AICw_normal,echo=F,results="markup"}
source("normal_cop_AICw.R")
h<-normal_cop_AICw_matrix(r=RES_methane)
h[is.infinite(h)]<-NA
knitr::kable(h,
             format="latex",linesep = "",
             digits=3,
             caption.short="AIC weights, N copulas, methane flux data",
             caption = "$\\text{AIC}_{\\text{w}}$ for the N copula family for each location pair for methane flux data, for comparison with Table \\ref{tab_methane_AICw}.\\label{tab_methane_AICw_normal}",
             booktabs=TRUE)
```

```{r tab_methane_p_CvM,echo=F,results="markup"}
h<-RES_methane$gfc_p_CvM
h[is.infinite(h)]<-NA
knitr::kable(h,
             format="latex",linesep = "",
             digits=3,
             caption.short="Goodness of fit of best-fitting copulas, methane flux data",
             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for methane flux data.\\label{tab_methane_p_CvM}",
             booktabs=TRUE)
```

```{r tab_methane_p_KS,echo=F,results="markup"}
h<-RES_methane$gfc_p_KS
h[is.infinite(h)]<-NA
knitr::kable(h,
             format="latex",linesep = "",
             digits=3,
             caption.short="Goodness of fit of best-fitting copulas, methane flux data",
             caption = "$P$-value results of a Kolmogorov-Smirnov-based test of
the goodness of fit of the lowest-AIC copula family for each pair of
locations for methane flux data.\\label{tab_methane_p_KS}",
             booktabs=TRUE)
```

```{r tab_methane_LTdep_AICw,echo=F,results="markup"}
h<-RES_methane$LTdep_AICw
h[is.infinite(h)]<-NA
knitr::kable(h,
             format="latex",linesep = "",
             digits=3,
             caption.short="Lower-tail dependence, methane flux data",
             caption = "Model-averaged lower-tail dependence statistics for each location pair for methane flux data.\\label{tab_methane_LTdep_AICw}",
             booktabs=TRUE)
```

```{r tab_methane_UTdep_AICw,echo=F,results="markup"}
h<-RES_methane$UTdep_AICw
h[is.infinite(h)]<-NA
knitr::kable(h,
             format="latex",linesep = "",
             digits=3,
             caption.short="Upper-tail dependence, methane flux data",
             caption = "Model-averaged upper-tail dependence statistics for each location pair for methane flux data.\\label{tab_methane_UTdep_AICw}",
             booktabs=TRUE)
```

```{r tab_methane_LTmUT,echo=F,results="markup"}
h<-RES_methane$LTdep_AICw-RES_methane$UTdep_AICw
h[is.nan(h)]<-NA
knitr::kable(h,
             format="latex",linesep = "",
             digits=3,
             caption.short="Lower- minus upper-tail dependence, methane flux data",
             caption = "Model-averaged lower-tail dependence statistics minus model-averaged upper-tail dependence statistics for each location pair for methane flux data.\\label{tab_methane_LTmUT}",
             booktabs=TRUE)
```

<!--Do the main computations for nonparametric analysis of methane-flux data-->
```{r stat_methane, echo=F, cache=T, cache=T,  cache.extra=list(seed,methane_data,mtime("good_loclist.R"),mtime("NonParamStat.R"),mtime("vivj_matrix.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}

source("good_loclist.R")
source("NonParamStat.R")
set.seed(seed)
resloc<-"./Results/stat_results/stat_methane/"

data_pt_thrs<-50
good_loc <- good_loclist(d_allsp=methane_data,sp=1,data_pt_thrs=data_pt_thrs)

latlon<-readRDS("./Data/Methane_data/chamberLocations.rds")
chamber_name<-colnames(methane_raw[,c(2:dim(methane_raw)[2])])
ndim<-length(chamber_name)

latlon_arranged<-data.frame(chamber_name=chamber_name,lat=NA*numeric(ndim),lon=NA*numeric(ndim))
for(i in 1:ndim){
  if (chamber_name[i] %in% latlon$chamber){
    ind<-which(latlon$chamber==chamber_name[i])
    latlon_arranged[i,2:3]<-unname(latlon[ind,2:3])
  }
}

stat_methane<-multcall(d_allsp=methane_data,sp=1,lats=latlon_arranged$lat,longs=latlon_arranged$lon,
                           pfname=paste(resloc,"data_pt_thrs",data_pt_thrs,sep=''),good_loc=good_loc)

saveRDS(stat_methane,paste(resloc,file="stat_methane.RDS",sep='')) 
```

```{r read_stat_methane,echo=F}
stat_methane<-readRDS("./Results/stat_results/stat_methane/stat_methane.RDS")
```

<!--display the main statistical results in a table-->
```{r tab_methane_resamp,echo=F,results="markup"}
s<-stat_methane$numericdf[,c(3,2,4)]

#rownames(s)<-c("Spearman","Kendall","cor_l","cor_u","P_l","P_u","D^2_l","D^2_u","cor_l - cor_u","P_l - P_u","D^2_u - D^2_l")

sc<-as.data.frame(c("Spearman","Kendall","cor$_{l}$","cor$_{u}$","P$_{l}$","P$_{u}$","D$_{l}^2$","D$_{u}^2$",
                    "cor$_{l}$ - cor$_{u}$","P$_{l}$ - P$_{u}$","D$_{u}^2$ - D$_{l}^2$"))
s<-cbind(sc,s)
knitr::kable(s,digits = 3,
             format = "latex",linesep = "", escape = F,align="c",
             col.names=c("","2.5$^{th}$ quantile", "Mean", "97.5$^{th}$ quantile"),
             caption.short="Results for nonparametric statistics, methane flux data",
             caption = "Average values of statistics across all locations pairs and confidence intervals based on spatial resampling for methane flux data.\\label{tab_methane_resamp}",
             booktabs=TRUE)
```

\begin{figure}[!h]
\textbf{ \hspace{3.4 cm} (A) \hspace{3.2 cm} (B) \hspace{3.2 cm} (C) \hspace{3.4 cm} (D)} \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/stat_results/stat_methane/data_pt_thrs50_Spearman_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_methane/data_pt_thrs50_Kendall_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_methane/data_pt_thrs50_Corl_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_methane/data_pt_thrs50_Coru_vs_D.pdf}\\
\end{center}
\textbf{ \hspace{3.4 cm} (E) \hspace{3.2 cm} (F) \hspace{3.2 cm} (G) \hspace{3.4 cm} (H)} \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/stat_results/stat_methane/data_pt_thrs50_Pl_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_methane/data_pt_thrs50_Pu_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_methane/data_pt_thrs50_D2u_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_methane/data_pt_thrs50_D2l_vs_D.pdf}\\
\end{center}
\textbf{ \hspace{5.4 cm} (I) \hspace{3.4 cm} (J) \hspace{3.2 cm} (K) } \\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=4 cm]{./Results/stat_results/stat_methane/data_pt_thrs50_Corl-Coru_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_methane/data_pt_thrs50_Pl-Pu_vs_D.pdf}
\includegraphics[width=4 cm]{./Results/stat_results/stat_methane/data_pt_thrs50_D2u-D2l_vs_D.pdf}
\caption[Nonparametric tail statistics versus distance between locations for methane-flux data]{Nonparametric tail statistics versus great-circle distance, $D$, between sampling locations for methane-flux data. \label{fig_methane}}
\end{center}
\end{figure}

```{r bivar_tab_summary_fn,echo=F}
bivar_tab_summary_fn<-function(r,rs){

  m<-matrix(NA,1,10)
  colnames(m)<-c("p.Independence.test","minimumAIC.bestfit.copula","Bestfit.copula.AICw",
                 "Normal.copula.AICw","p.CvM","p.KS",
                 "LT.model.avg","UT.model.avg","(LT-UT).model.avg",
                 "Rank.(Pl-Pu)")
  m<-as.data.frame(m)

  m$p.Independence.test<-round(r$IndepTestRes,3)
  m$minimumAIC.bestfit.copula<-r$InfCritRes$copname[which.min(r$InfCritRes$AIC)]
  m$Bestfit.copula.AICw<-round(max(r$InfCritRes$AICw),3)
  m$Normal.copula.AICw<-round(r$InfCritRes$AICw[1],3)
  m$p.CvM<-r$GofRes_CvM
  m$p.KS<-r$GofRes_KS
  m$LT.model.avg<-round(r$relLTdep_AICw,3)
  m$UT.model.avg<-round(r$relUTdep_AICw,3)
  m$`(LT-UT).model.avg`<-m$LT.model.avg-m$UT.model.avg
  m$`Rank.(Pl-Pu)`<-rs$Rank_PlmPu_K
  mt<-as.data.frame(t(m))
  return(mt)

}
#----------------------------------
#calling the above function on bivariable datasets
m1<-bivar_tab_summary_fn(r=BivMS_res_Loecke,rs=stat_soilCN_ga)
m2<-bivar_tab_summary_fn(r=BivMS_birds,rs=stat_birds_ga)
m3<-bivar_tab_summary_fn(r=BivMS_mammals,rs=stat_mammals_ga)
m4<-bivar_tab_summary_fn(r=BivMS_res_cc$`2000`,rs=stat_cc_ga$`2000`)
tab_bivar_summary<-cbind(m1,m2,m3,m4)
colnames(tab_bivar_summary)<-c("SoilCN","Birds' BMR","Mammals' BMR","Ceder-creek")
saveRDS(tab_bivar_summary,"./Results/tab_bivar_summary.RDS")
```

```{r multivar_tab_summary_fn,echo=F}
source("normal_cop_AICw.R")

multivar_tab_summary_fn<-function(r,rs){

  m<-matrix(NA,1,17)
  colnames(m)<-c("num.pairs","num.non-indep","num.non-N",
                 "Avg.best.AICw","Avg.N.AICw","num.good.fit",
                 "0.025-CI.LT","0.975-CI.LT","0.025-CI.UT","0.975-CI.UT",
                 "percent.LT-UT>0",
                 "0.025-CI.Pl","0.975-CI-Pl","0.025-CI.Pu","0.975-CI.Pu","0.025-CI.Pl-Pu","0.975-CI.Pl-Pu")
  m<-as.data.frame(m)

  normal_AICw<-normal_cop_AICw_matrix(r=r)

  d<-dim(r$gfc_numBS)[1]
  m$num.pairs<-(d^2)-d 
  m$`num.non-indep`<- (d^2)-d-2*(r$num_indep_loc_pair+r$num_neg_cor_loc_pair)
  bcop<-r$info_ord_copcode[,,1] # matrix with best fit copula name
  bcop[bcop==Inf]<-NaN # replacing Independent pairs with NaN
  m$`num.non-N`<-sum(bcop!=1,na.rm = T)
  r$info_ord_AICw[r$info_ord_AICw==Inf] <- NaN
  m$Avg.best.AICw<-round(mean(r$info_ord_AICw[,,1],na.rm = T),3)
  normal_AICw[normal_AICw==Inf]<-NaN
  m$Avg.N.AICw<-round(mean(normal_AICw,na.rm=T),3)
  pcvm<-r$gfc_p_CvM
  pcvm[pcvm==Inf]<-NaN
  pks<-r$gfc_p_KS
  pks[pks==Inf]<-NaN
  temp<-which((pcvm>0.01) & (pks>0.01),arr.ind=T)
  m$num.good.fit<-dim(temp)[1]
  t<-r$LTdep_AICw[is.finite(r$LTdep_AICw)]
  CI_LT<-unname(quantile(t,probs=c(0.025,0.975),na.rm = T))
  m$`0.025-CI.LT`<-round(CI_LT[1],3)
  m$`0.975-CI.LT`<-round(CI_LT[2],3)
  t<-r$UTdep_AICw[is.finite(r$UTdep_AICw)]
  CI_UT<-unname(quantile(t,probs=c(0.025,0.975),na.rm = T))
  m$`0.025-CI.UT`<-round(CI_UT[1],3)
  m$`0.975-CI.UT`<-round(CI_UT[2],3)
  m$`percent.LT-UT>0`<-round((sum(r$LTdep_AICw-r$UTdep_AICw>0,na.rm=T)/((d^2)-d-2*(r$num_indep_loc_pair+r$num_neg_cor_loc_pair)))*100,2)
  m$`0.025-CI.Pl`<-round(rs$numericdf[5,3],3)
  m$`0.975-CI-Pl`<-round(rs$numericdf[5,4],3)
  m$`0.025-CI.Pu`<-round(rs$numericdf[6,3],3)
  m$`0.975-CI.Pu`<-round(rs$numericdf[6,4],3)
  m$`0.025-CI.Pl-Pu`<-round(rs$numericdf[10,3],3)
  m$`0.975-CI.Pl-Pu`<-list(round(rs$numericdf[10,4],3))
  #m$CI_Pl<-list(c(round(rs$numericdf[5,3],3),round(rs$numericdf[5,4],3)))
  #m$CI_Pu<-list(c(round(rs$numericdf[6,3],3),round(rs$numericdf[6,4],3)))
  #m$`CI_Pl-Pu`<-list(c(round(rs$numericdf[10,3],3),round(rs$numericdf[10,4],3)))

  mt<-as.data.frame(t(m))
  return(mt)

}
#----------------------------------
#calling the above function on 4 multivariable datasets
m1<-multivar_tab_summary_fn(r=RES_aphid_count,rs=stat_aphid_count)
m2<-multivar_tab_summary_fn(r=RES_aphid_ff,rs=stat_aphid_ff)
m3<-multivar_tab_summary_fn(r=RES_plankton_north_sea,rs=stat_plankton_north_sea)
m4<-multivar_tab_summary_fn(r=RES_methane,rs=stat_methane)
tab_multivar_summary<-cbind(m1,m2,m3,m4)
colnames(tab_multivar_summary)<-c("Green-spruce aphid_count","Leaf-plum curling aphid_firstflight","C. furca_count","Methane-flux")
saveRDS(tab_multivar_summary,"./Results/tab_multivar_summary.RDS")
```

```{r sample_plot_aphid_count_ff,echo=F,results="hide"}
#-------------------------------------------------------------------------------
# code to plot raw data and copula plot for aphid count @location pair (4,11)
# bestfit cop : SJ

d1<-data_aphid_count$`Green spruce aphid`[[4]]
d2<-data_aphid_count$`Green spruce aphid`[[11]]
pdf("./Results/Aphid_count_raw_loc_4_11.pdf",height = 8,width=8)
op<-par(mar=c(10,10, 0.2, 2.1),mgp=c(3,1,0))
plot(d1$Dat,d2$Dat,type="p",xlab="",
     ylab="Green-spruce aphid \ncount (loc. 11)",cex.lab=3.5,cex.axis=2.5)
mtext(paste("Green-spruce aphid \ncount (loc. 4)"),line=-38,cex=3.5)
par(op)
dev.off()

source("./vivj_matrix.R")
m<-vivj_matrix(d_allsp = data_aphid_count,sp=10,i=4,j=11)
pdf("./Results/Aphid_count_cop_loc_4_11.pdf",height = 8,width=8)
op<-par(mar=c(6,6, 0.2, 2.1),mgp=c(3,1,0))
plot(m[,1],m[,2],type="p",xlab="u",ylab="v",
     cex.lab=4,cex.axis=3,
     xlim=c(0,1),ylim=c(0,1),asp=1)
rect(0,0,1,1)
lines(c(0,1),c(0,1),type='l',col="grey")
par(op)
dev.off()

# code to plot raw data and copula plot for aphid ff @location pair (2,5)
# bestfit cop : J
d1<-data_aphid_ff$`Leaf-curling plum aphid`[[2]]
d2<-data_aphid_ff$`Leaf-curling plum aphid`[[5]]
pdf("./Results/Aphid_ff_raw_loc_2_5.pdf",height = 8,width=8)
op<-par(mar=c(10,10, 0.2, 2.1),mgp=c(3,1,0))
plot(d1$Dat,d2$Dat,type="p",xlab="",ylab="Leaf-plum curling's \nfirst-flight (loc. 5)",cex.lab=3.5,cex.axis=2.5)
mtext(paste("Leaf-plum curling's \nfirst-flight (loc. 2)"),line=-38,cex=3.5)
par(op)
dev.off()

m<-vivj_matrix(d_allsp = data_aphid_ff,sp=11,i=2,j=5)
pdf("./Results/Aphid_ff_cop_loc_2_5.pdf",height = 8,width=8)
op<-par(mar=c(6,6, 0.2, 2.1),mgp=c(3,1,0))
plot(m[,1],m[,2],type="p",xlab="u",ylab="v",
     cex.lab=4,cex.axis=3,
     xlim=c(0,1),ylim=c(0,1),asp=1)
rect(0,0,1,1)
lines(c(0,1),c(0,1),type='l',col="grey")
par(op)
dev.off()
```

```{r cause4copula_pGOF_500, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
#source("Cause4copula.R")
#-----------------
set.seed(seed)
#---------------------------------------
#pdf("./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_500.pdf", width=8, height=3)
#op<-par(mfrow=c(1,2))

#Plotter_Cause4copula_GOF(N=5000,fcode=3,method = "spearman",num_keep_last=500,BS=1000)
#Plotter_Cause4copula_GOF(N=5000,fcode=13,method="spearman",num_keep_last=500,BS=1000)
#Plotter_Cause4copula_GOF(N=5000,fcode=4,method = "spearman",num_keep_last=500,BS=1000)
#Plotter_Cause4copula_GOF(N=5000,fcode=14,method="spearman",num_keep_last=500,BS=1000)

#Plotter_Cause4copula_GOF(N=5000,fcode=3,method = "kendall",num_keep_last=500,BS=1000)
#Plotter_Cause4copula_GOF(N=5000,fcode=13,method="kendall",num_keep_last=500,BS=1000)
#Plotter_Cause4copula_GOF(N=5000,fcode=4,method = "kendall",num_keep_last=500,BS=1000)
#Plotter_Cause4copula_GOF(N=5000,fcode=14,method="kendall",num_keep_last=500,BS=1000)

#par(op)

#op<-par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 1, 0, 0), new = TRUE)
#plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
#legend("top", c("noise copula,", "population copula,",  "Gof test : p-value(CvM),", "Gof test : p-value(KS)"), col = c("red","blue", #"magenta", "green2"),
#       cex = 0.9, pch = c(2, 6, 16, 16), xpd = TRUE, horiz = TRUE, inset = c(0,0),
#       bty = "n") 
#par(op)
#dev.off()
```

```{r cause4copula_pGOF_2500, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
#source("Cause4copula.R")
#-----------------
#set.seed(seed)
#---------------------------------------

#pdf("./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_2500.pdf", width=8, height=3)
#op<-par(mfrow=c(1,2))

#Plotter_Cause4copula_GOF(N=5000,fcode=3,method = "spearman",num_keep_last=2500,BS=1000)
#Plotter_Cause4copula_GOF(N=5000,fcode=13,method="spearman",num_keep_last=2500,BS=1000)
#Plotter_Cause4copula_GOF(N=5000,fcode=4,method = "spearman",num_keep_last=2500,BS=1000)
#Plotter_Cause4copula_GOF(N=5000,fcode=14,method="spearman",num_keep_last=2500,BS=1000)

#Plotter_Cause4copula_GOF(N=5000,fcode=3,method = "kendall",num_keep_last=2500,BS=1000)
#Plotter_Cause4copula_GOF(N=5000,fcode=13,method="kendall",num_keep_last=2500,BS=1000)
#Plotter_Cause4copula_GOF(N=5000,fcode=4,method = "kendall",num_keep_last=2500,BS=1000)
#Plotter_Cause4copula_GOF(N=5000,fcode=14,method="kendall",num_keep_last=2500,BS=1000)

#op<-par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 1, 0, 0), new = TRUE)
#plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
#legend("top", c("noise copula,", "population copula,",  "Gof test : p-value(CvM),", "Gof test : p-value(KS)"), col = c("red","blue", #"magenta", "green2"),
#       cex = 0.9, pch = c(2, 6, 16, 16), xpd = TRUE, horiz = TRUE, inset = c(0,0),
#       bty = "n") 
#par(op)
#dev.off()
```

```{r cause4copula_stat_ar1_clayton, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("Cause4copula.R")
#-----------------
set.seed(seed)
#---------------------------------------
# This chunk stores results for ar1 model : clayton

model<-"ar1"
fcode<-3
nsd<-1

# for positive beta
params<-c(0.5) # this parameter is beta in ar1 model
resloc<-"./Results/Cause4copula_results/"

tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")

if (!dir.exists(tempo)){
  dir.create(tempo)
}
resloc<-paste(tempo,"/",sep="")
# for clayton with spearman and kendall method
ar1_C_spear_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
                                           num_keep_last=500,resloc=resloc,
                                           params=params,p0=c(0,0),model=model)

saveRDS(ar1_C_spear_beta0.5,paste(resloc,"ar1_C_spear_beta0.5.RDS",sep=""))

ar1_C_kend_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
                                          num_keep_last=500,resloc=resloc,
                                          params=params,p0=c(0,0),model=model)

saveRDS(ar1_C_kend_beta0.5,paste(resloc,"ar1_C_kend_beta0.5.RDS",sep=""))

# for negative beta
resloc<-"./Results/Cause4copula_results/"
params<-c(-0.5) # this parameter is beta in ar1 model
tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")

if (!dir.exists(tempo)){
  dir.create(tempo)
}

resloc<-paste(tempo,"/",sep="")
# for clayton with spearman and kendall method
ar1_C_spear_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
                                              num_keep_last=500,resloc=resloc,
                                              params=params,p0=c(0,0),model=model)

saveRDS(ar1_C_spear_negbeta0.5,paste(resloc,"ar1_C_spear_negbeta0.5.RDS",sep=""))

ar1_C_kend_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
                                               num_keep_last=500,resloc=resloc,
                                               params=params,p0=c(0,0),model=model)

saveRDS(ar1_C_kend_negbeta0.5,paste(resloc,"ar1_C_kend_negbeta0.5.RDS",sep=""))
```

```{r plot_cause4copula_stat_ar1_clayton, echo=F, results="hide", cache=T, cache.extra=list(seed,ar1_C_spear_beta0.5,ar1_C_spear_negbeta0.5,ar1_C_kend_beta0.5,ar1_C_kend_negbeta0.5,mtime("Plotter_Cause4copula_stat.R"))}

set.seed(seed)
source("Plotter_Cause4copula_stat.R")
#---------------------------------------
# This chunk plots results for ar1 model : clayton

model<-"ar1"
fcode<-3
resloc0<-"./Results/Cause4copula_results/"

# for positive beta
params<-c(0.5) # this parameter is beta in ar1 model

tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
resloc<-paste(tempo,"/",sep="")

readres<-readRDS(paste(resloc,"ar1_C_spear_beta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)
readres<-readRDS(paste(resloc,"ar1_C_kend_beta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)

# for negative beta
params<-c(-0.5) # this parameter is beta in ar1 model
tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
resloc<-paste(tempo,"/",sep="")

readres<-readRDS(paste(resloc,"ar1_C_spear_negbeta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)
readres<-readRDS(paste(resloc,"ar1_C_kend_negbeta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)
```

```{r cause4copula_stat_ar1_survivalclayton, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("Cause4copula.R")
#-----------------
set.seed(seed)
#---------------------------------------
# This chunk stores results for ar1 model : survival clayton
model<-"ar1"
fcode<-13
nsd<-1

# for positive beta
params<-c(0.5) # this parameter is beta in ar1 model

resloc<-"./Results/Cause4copula_results/"
tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")

if (!dir.exists(tempo)){
  dir.create(tempo)
}
resloc<-paste(tempo,"/",sep="")
# for survival clayton with spearman and kendall method
ar1_SC_spear_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
                                            num_keep_last=500,resloc=resloc,
                                            params=params,p0=c(0,0),model=model)

saveRDS(ar1_SC_spear_beta0.5,paste(resloc,"ar1_SC_spear_beta0.5.RDS",sep=""))

ar1_SC_kend_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
                                           num_keep_last=500,resloc=resloc,
                                           params=params,p0=c(0,0),model=model)

saveRDS(ar1_SC_kend_beta0.5,paste(resloc,"ar1_SC_kend_beta0.5.RDS",sep=""))
# for negative beta
params<-c(-0.5) # this parameter is beta in ar1 model

resloc<-"./Results/Cause4copula_results/"
tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")

if (!dir.exists(tempo)){
  dir.create(tempo)
}

resloc<-paste(tempo,"/",sep="")
# for survival clayton with spearman and kendall method
ar1_SC_spear_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
                                               num_keep_last=500,resloc=resloc,
                                               params=params,p0=c(0,0),model=model)
saveRDS(ar1_SC_spear_negbeta0.5,paste(resloc,"ar1_SC_spear_negbeta0.5.RDS",sep=""))

ar1_SC_kend_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
                                              num_keep_last=500,resloc=resloc,
                                              params=params,p0=c(0,0),model=model)
saveRDS(ar1_SC_kend_negbeta0.5,paste(resloc,"ar1_SC_kend_negbeta0.5.RDS",sep=""))
```

```{r plot_cause4copula_stat_ar1_survivalclayton, echo=F, results="hide", cache=T, cache.extra=list(seed,ar1_SC_spear_beta0.5,ar1_SC_spear_negbeta0.5,ar1_SC_kend_beta0.5,ar1_SC_kend_negbeta0.5,mtime("Plotter_Cause4copula_stat.R"))}

set.seed(seed)
source("Plotter_Cause4copula_stat.R")
#---------------------------------------
# This chunk plots results for ar1 model : clayton

model<-"ar1"
fcode<-13
resloc0<-"./Results/Cause4copula_results/"

# for positive beta
params<-c(0.5) # this parameter is beta in ar1 model

tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
resloc<-paste(tempo,"/",sep="")

readres<-readRDS(paste(resloc,"ar1_SC_spear_beta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)
readres<-readRDS(paste(resloc,"ar1_SC_kend_beta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)

# for negative beta
params<-c(-0.5) # this parameter is beta in ar1 model
tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
resloc<-paste(tempo,"/",sep="")

readres<-readRDS(paste(resloc,"ar1_SC_spear_negbeta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)
readres<-readRDS(paste(resloc,"ar1_SC_kend_negbeta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)
```

```{r cause4copula_stat_ar1_gumbel, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("Cause4copula.R")
#-----------------
set.seed(seed)
#---------------------------------------
# This chunk stores results for ar1 model : clayton

model<-"ar1"
fcode<-4
nsd<-1

# for positive beta
params<-c(0.5) # this parameter is beta in ar1 model
resloc<-"./Results/Cause4copula_results/"

tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")

if (!dir.exists(tempo)){
  dir.create(tempo)
}
resloc<-paste(tempo,"/",sep="")
# for gumbel with spearman and kendall method
ar1_G_spear_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
                                              num_keep_last=500,resloc=resloc,
                                              params=params,p0=c(0,0),model=model)
saveRDS(ar1_G_spear_beta0.5,paste(resloc,"ar1_G_spear_beta0.5.RDS",sep=""))

ar1_G_kend_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
                                             num_keep_last=500,resloc=resloc,
                                             params=params,p0=c(0,0),model=model)

saveRDS(ar1_G_kend_beta0.5,paste(resloc,"ar1_G_kend_beta0.5.RDS",sep=""))

# for negative beta
resloc<-"./Results/Cause4copula_results/"
params<-c(-0.5) # this parameter is beta in ar1 model
tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")

if (!dir.exists(tempo)){
  dir.create(tempo)
}

resloc<-paste(tempo,"/",sep="")
# for gumbel with spearman and kendall method
ar1_G_spear_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
                                              num_keep_last=500,resloc=resloc,
                                              params=params,p0=c(0,0),model=model)
saveRDS(ar1_G_spear_negbeta0.5,paste(resloc,"ar1_G_spear_negbeta0.5.RDS",sep=""))

ar1_G_kend_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
                                             num_keep_last=500,resloc=resloc,
                                             params=params,p0=c(0,0),model=model)

saveRDS(ar1_G_kend_negbeta0.5,paste(resloc,"ar1_G_kend_negbeta0.5.RDS",sep=""))
```

```{r plot_cause4copula_stat_ar1_gumbel, echo=F, results="hide", cache=T, cache.extra=list(seed,ar1_G_spear_beta0.5,ar1_G_spear_negbeta0.5,ar1_G_kend_beta0.5,ar1_G_kend_negbeta0.5,mtime("Plotter_Cause4copula_stat.R"))}

set.seed(seed)
source("Plotter_Cause4copula_stat.R")
#---------------------------------------
# This chunk plots results for ar1 model : clayton

model<-"ar1"
fcode<-4
resloc0<-"./Results/Cause4copula_results/"

# for positive beta
params<-c(0.5) # this parameter is beta in ar1 model

tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
resloc<-paste(tempo,"/",sep="")

readres<-readRDS(paste(resloc,"ar1_G_spear_beta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)
readres<-readRDS(paste(resloc,"ar1_G_kend_beta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)

# for negative beta
params<-c(-0.5) # this parameter is beta in ar1 model
tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
resloc<-paste(tempo,"/",sep="")

readres<-readRDS(paste(resloc,"ar1_G_spear_negbeta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)
readres<-readRDS(paste(resloc,"ar1_G_kend_negbeta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)
```

```{r cause4copula_stat_ar1_survivalgumbel, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("Cause4copula.R")
#-----------------
set.seed(seed)
#---------------------------------------
# This chunk stores results for ar1 model : survival clayton
model<-"ar1"
fcode<-14
nsd<-1

# for positive beta
params<-c(0.5) # this parameter is beta in ar1 model

resloc<-"./Results/Cause4copula_results/"
tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")

if (!dir.exists(tempo)){
  dir.create(tempo)
}
resloc<-paste(tempo,"/",sep="")
# for survival gumbel with spearman and kendall method
ar1_SG_spear_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
                                              num_keep_last=500,resloc=resloc,
                                              params=params,p0=c(0,0),model=model)
saveRDS(ar1_SG_spear_beta0.5,paste(resloc,"ar1_SG_spear_beta0.5.RDS",sep=""))

ar1_SG_kend_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
                                             num_keep_last=500,resloc=resloc,
                                             params=params,p0=c(0,0),model=model)

saveRDS(ar1_SG_kend_beta0.5,paste(resloc,"ar1_SG_kend_beta0.5.RDS",sep=""))

# for negative beta
params<-c(-0.5) # this parameter is beta in ar1 model

resloc<-"./Results/Cause4copula_results/"
tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")

if (!dir.exists(tempo)){
  dir.create(tempo)
}

resloc<-paste(tempo,"/",sep="")
# for survival gumbel with spearman and kendall method
ar1_SG_spear_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
                                              num_keep_last=500,resloc=resloc,
                                              params=params,p0=c(0,0),model=model)
saveRDS(ar1_SG_spear_negbeta0.5,paste(resloc,"ar1_SG_spear_negbeta0.5.RDS",sep=""))

ar1_SG_kend_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
                                             num_keep_last=500,resloc=resloc,
                                             params=params,p0=c(0,0),model=model)

saveRDS(ar1_SG_kend_negbeta0.5,paste(resloc,"ar1_SG_kend_negbeta0.5.RDS",sep=""))
```

```{r plot_cause4copula_stat_ar1_survivalgumbel, echo=F, results="hide", cache=T, cache.extra=list(seed,ar1_SG_spear_beta0.5,ar1_SG_spear_negbeta0.5,ar1_SG_kend_beta0.5,ar1_SG_kend_negbeta0.5,mtime("Plotter_Cause4copula_stat.R"))}

set.seed(seed)
source("Plotter_Cause4copula_stat.R")
#---------------------------------------
# This chunk plots results for ar1 model : clayton

model<-"ar1"
fcode<-14
resloc0<-"./Results/Cause4copula_results/"

# for positive beta
params<-c(0.5) # this parameter is beta in ar1 model

tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
resloc<-paste(tempo,"/",sep="")

readres<-readRDS(paste(resloc,"ar1_SG_spear_beta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)
readres<-readRDS(paste(resloc,"ar1_SG_kend_beta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)

# for negative beta
params<-c(-0.5) # this parameter is beta in ar1 model
tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
resloc<-paste(tempo,"/",sep="")

readres<-readRDS(paste(resloc,"ar1_SG_spear_negbeta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)
readres<-readRDS(paste(resloc,"ar1_SG_kend_negbeta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)
```

```{r cause4copula_stat_ar1_joe, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("Cause4copula.R")
#-----------------
set.seed(seed)
#---------------------------------------
# This chunk stores results for ar1 model : clayton

model<-"ar1"
fcode<-6
nsd<-1

# for positive beta
params<-c(0.5) # this parameter is beta in ar1 model
resloc<-"./Results/Cause4copula_results/"

tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")

if (!dir.exists(tempo)){
  dir.create(tempo)
}
resloc<-paste(tempo,"/",sep="")
# for joe with kendall method only

ar1_J_kend_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
                                             num_keep_last=500,resloc=resloc,
                                             params=params,p0=c(0,0),model=model)

saveRDS(ar1_J_kend_beta0.5,paste(resloc,"ar1_J_kend_beta0.5.RDS",sep=""))

# for negative beta
resloc<-"./Results/Cause4copula_results/"
params<-c(-0.5) # this parameter is beta in ar1 model
tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")

if (!dir.exists(tempo)){
  dir.create(tempo)
}

resloc<-paste(tempo,"/",sep="")
# for joe with kendall method only

ar1_J_kend_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
                                             num_keep_last=500,resloc=resloc,
                                             params=params,p0=c(0,0),model=model)

saveRDS(ar1_J_kend_negbeta0.5,paste(resloc,"ar1_J_kend_negbeta0.5.RDS",sep=""))
```

```{r plot_cause4copula_stat_ar1_joe, echo=F, results="hide", cache=T, cache.extra=list(seed,ar1_J_kend_beta0.5,ar1_J_kend_negbeta0.5,mtime("Plotter_Cause4copula_stat.R"))}

set.seed(seed)
source("Plotter_Cause4copula_stat.R")
#---------------------------------------
# This chunk plots results for ar1 model : clayton

model<-"ar1"
fcode<-6
resloc0<-"./Results/Cause4copula_results/"

# for positive beta
params<-c(0.5) # this parameter is beta in ar1 model

tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
resloc<-paste(tempo,"/",sep="")

readres<-readRDS(paste(resloc,"ar1_J_kend_beta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)

# for negative beta
params<-c(-0.5) # this parameter is beta in ar1 model
tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
resloc<-paste(tempo,"/",sep="")

readres<-readRDS(paste(resloc,"ar1_J_kend_negbeta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)
```

```{r cause4copula_stat_ar1_survivaljoe, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("Cause4copula.R")
#-----------------
set.seed(seed)
#---------------------------------------
# This chunk stores results for ar1 model : survival clayton
model<-"ar1"
fcode<-16
nsd<-1

# for positive beta
params<-c(0.5) # this parameter is beta in ar1 model

resloc<-"./Results/Cause4copula_results/"
tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")

if (!dir.exists(tempo)){
  dir.create(tempo)
}
resloc<-paste(tempo,"/",sep="")
# for survival joe with kendall method only

ar1_SJ_kend_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
                                             num_keep_last=500,resloc=resloc,
                                             params=params,p0=c(0,0),model=model)

saveRDS(ar1_SJ_kend_beta0.5,paste(resloc,"ar1_SJ_kend_beta0.5.RDS",sep=""))

# for negative beta
params<-c(-0.5) # this parameter is beta in ar1 model

resloc<-"./Results/Cause4copula_results/"
tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")

if (!dir.exists(tempo)){
  dir.create(tempo)
}

resloc<-paste(tempo,"/",sep="")
# for survival joe with kendall method only

ar1_SJ_kend_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
                                             num_keep_last=500,resloc=resloc,
                                             params=params,p0=c(0,0),model=model)

saveRDS(ar1_SJ_kend_negbeta0.5,paste(resloc,"ar1_SJ_kend_negbeta0.5.RDS",sep=""))
```

```{r plot_cause4copula_stat_ar1_survivaljoe, echo=F, results="hide", cache=T, cache.extra=list(seed,ar1_SJ_kend_beta0.5,ar1_SJ_kend_negbeta0.5,mtime("Plotter_Cause4copula_stat.R"))}

set.seed(seed)
source("Plotter_Cause4copula_stat.R")
#---------------------------------------
# This chunk plots results for ar1 model : clayton

model<-"ar1"
fcode<-16
resloc0<-"./Results/Cause4copula_results/"

# for positive beta
params<-c(0.5) # this parameter is beta in ar1 model

tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
resloc<-paste(tempo,"/",sep="")

readres<-readRDS(paste(resloc,"ar1_SJ_kend_beta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)

# for negative beta
params<-c(-0.5) # this parameter is beta in ar1 model
tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
resloc<-paste(tempo,"/",sep="")

readres<-readRDS(paste(resloc,"ar1_SJ_kend_negbeta0.5.RDS",sep=""))
Plotter_Cause4copula_stat(res=readres)
```

```{r cause4copula_stat_ricker_clayton, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("Cause4copula.R")
#-----------------
set.seed(seed)
#---------------------------------------
# This chunk stores results for ar1 model : clayton
tag_ricker_C<-"999C" #this is a caching tag used later to cache output from present chunk

model<-"ricker"
fcode<-3

nsdlist<-c(0.1,1)
Klist<-c(100)

# for varying r (@r=1 transition from under- to over-compensatory dynamics )
rlist<-seq(from=0.1,to=1.3,by=0.2)

for(insd in nsdlist){
  for(iK in Klist){
   for(i in rlist){
    params<-c(i,iK) # this parameter is a vector with r and K
    p0<-c(params[2],params[2])

    resloc<-"./Results/Cause4copula_results/"

    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")

    if (!dir.exists(tempo)){
      dir.create(tempo)
    }
    resloc<-paste(tempo,"/",sep="")

  # for clayton with spearman and kendall method
    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="spearman",lb=0,ub=0.1,
                                           num_keep_last=500,resloc=resloc,
                                           params=params,p0=p0,model=model)

    saveRDS(myres,paste(resloc,"ricker_C_spear.RDS",sep=""))

    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,
                                          num_keep_last=500,resloc=resloc,
                                          params=params,p0=p0,model=model)

    saveRDS(myres,paste(resloc,"ricker_C_kend.RDS",sep=""))

  }
 }
}
```

```{r plot_cause4copula_stat_ricker_C, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_C,mtime("Plotter_Cause4copula.R"))}

set.seed(seed)
source("Plotter_Cause4copula_stat.R")
#---------------------------------------
# This chunk plots results for ar1 model : clayton

model<-"ricker"
fcode<-3
resloc0<-"./Results/Cause4copula_results/"

nsdlist<-c(0.1,1)
Klist<-c(100)

# for varying r (@r=1 transition from under- to over-compensatory dynamics )
rlist<-seq(from=0.1,to=1.3,by=0.2)

for(insd in nsdlist){
  for(iK in Klist){
   for(i in rlist){
    params<-c(i,iK) # this parameter is a vector with r and K
    p0<-c(params[2],params[2])

    resloc<-"./Results/Cause4copula_results/"

    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")

    resloc<-paste(tempo,"/",sep="")

    readres<-readRDS(paste(resloc,"ricker_C_spear.RDS",sep=""))
    Plotter_Cause4copula_stat(res=readres)
    readres<-readRDS(paste(resloc,"ricker_C_kend.RDS",sep=""))
    Plotter_Cause4copula_stat(res=readres)

  }
 }
}
```

```{r cause4copula_stat_ricker_survivalclayton, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("Cause4copula.R")
#-----------------
set.seed(seed)
#---------------------------------------
tag_ricker_SC<-"999SC" #this is a caching tag used later to cache output from present chunk

# This chunk stores results for ar1 model : survival clayton

model<-"ricker"
fcode<-13
nsdlist<-c(0.1,1)
Klist<-c(100)
# for varying r (@r=1 transition from under- to over-compensatory dynamics )
rlist<-seq(from=0.1,to=1.3,by=0.2)

for(insd in nsdlist){
  for(iK in Klist){
   for(i in rlist){
    params<-c(i,iK) # this parameter is a vector with r and K
    p0<-c(params[2],params[2])

    resloc<-"./Results/Cause4copula_results/"

    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")

    if (!dir.exists(tempo)){
      dir.create(tempo)
    }
    resloc<-paste(tempo,"/",sep="")

  # for survival clayton with spearman and kendall method
    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="spearman",lb=0,ub=0.1,
                                           num_keep_last=500,resloc=resloc,
                                           params=params,p0=p0,model=model)

    saveRDS(myres,paste(resloc,"ricker_SC_spear.RDS",sep=""))

    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,
                                          num_keep_last=500,resloc=resloc,
                                          params=params,p0=p0,model=model)

    saveRDS(myres,paste(resloc,"ricker_SC_kend.RDS",sep=""))
  }
 }
}
```

```{r plot_cause4copula_stat_ricker_SC, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_SC,mtime("Plotter_Cause4copula.R"))}

set.seed(seed)
source("Plotter_Cause4copula_stat.R")
#---------------------------------------

model<-"ricker"
fcode<-13
resloc0<-"./Results/Cause4copula_results/"

nsdlist<-c(0.1,1)
Klist<-c(100)

# for varying r (@r=1 transition from under- to over-compensatory dynamics )
rlist<-seq(from=0.1,to=1.3,by=0.2)

for(insd in nsdlist){
  for(iK in Klist){
   for(i in rlist){
    params<-c(i,iK) # this parameter is a vector with r and K
    p0<-c(params[2],params[2])

    resloc<-"./Results/Cause4copula_results/"

    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")

    resloc<-paste(tempo,"/",sep="")

    readres<-readRDS(paste(resloc,"ricker_SC_spear.RDS",sep=""))
    Plotter_Cause4copula_stat(res=readres)
    readres<-readRDS(paste(resloc,"ricker_SC_kend.RDS",sep=""))
    Plotter_Cause4copula_stat(res=readres)

  }
 }
}
```

```{r cause4copula_stat_ricker_gumbel, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("Cause4copula.R")
#-----------------
set.seed(seed)
#---------------------------------------
tag_ricker_G<-"999G" #this is a caching tag used later to cache output from present chunk

# This chunk stores results for ar1 model : gumbel

model<-"ricker"
fcode<-4
nsdlist<-c(0.1,1)
Klist<-c(100)
# for varying r (@r=1 transition from under- to over-compensatory dynamics )
rlist<-seq(from=0.1,to=1.3,by=0.2)

for(insd in nsdlist){
  for(iK in Klist){
   for(i in rlist){
    params<-c(i,iK) # this parameter is a vector with r and K
    p0<-c(params[2],params[2])

    resloc<-"./Results/Cause4copula_results/"

    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")

    if (!dir.exists(tempo)){
      dir.create(tempo)
    }
    resloc<-paste(tempo,"/",sep="")

  # for gumbel with spearman and kendall method
    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="spearman",lb=0,ub=0.1,
                                           num_keep_last=500,resloc=resloc,
                                           params=params,p0=p0,model=model)

    saveRDS(myres,paste(resloc,"ricker_G_spear.RDS",sep=""))

    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,
                                          num_keep_last=500,resloc=resloc,
                                          params=params,p0=p0,model=model)

    saveRDS(myres,paste(resloc,"ricker_G_kend.RDS",sep=""))
  }
 }
}
```

```{r plot_cause4copula_stat_ricker_G, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_G,mtime("Plotter_Cause4copula.R"))}

set.seed(seed)
source("Plotter_Cause4copula_stat.R")
#---------------------------------------

model<-"ricker"
fcode<-4
resloc0<-"./Results/Cause4copula_results/"

nsdlist<-c(0.1,1)
Klist<-c(100)

# for varying r (@r=1 transition from under- to over-compensatory dynamics )
rlist<-seq(from=0.1,to=1.3,by=0.2)

for(insd in nsdlist){
  for(iK in Klist){
   for(i in rlist){
    params<-c(i,iK) # this parameter is a vector with r and K
    p0<-c(params[2],params[2])

    resloc<-"./Results/Cause4copula_results/"

    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")

    resloc<-paste(tempo,"/",sep="")

    readres<-readRDS(paste(resloc,"ricker_G_spear.RDS",sep=""))
    Plotter_Cause4copula_stat(res=readres)
    readres<-readRDS(paste(resloc,"ricker_G_kend.RDS",sep=""))
    Plotter_Cause4copula_stat(res=readres)

  }
 }
}
```

```{r cause4copula_stat_ricker_survivalgumbel, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("Cause4copula.R")
#-----------------
set.seed(seed)
#---------------------------------------
tag_ricker_SG<-"999SG" #this is a caching tag used later to cache output from present chunk

# This chunk stores results for ar1 model : survival gumbel

model<-"ricker"
fcode<-14
nsdlist<-c(0.1,1)
Klist<-c(100)
# for varying r (@r=1 transition from under- to over-compensatory dynamics )
rlist<-seq(from=0.1,to=1.3,by=0.2)

for(insd in nsdlist){
  for(iK in Klist){
   for(i in rlist){
    params<-c(i,iK) # this parameter is a vector with r and K
    p0<-c(params[2],params[2])

    resloc<-"./Results/Cause4copula_results/"

    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")

    if (!dir.exists(tempo)){
      dir.create(tempo)
    }
    resloc<-paste(tempo,"/",sep="")

  # for survival gumbel with spearman and kendall method
    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="spearman",lb=0,ub=0.1,
                                           num_keep_last=500,resloc=resloc,
                                           params=params,p0=p0,model=model)

    saveRDS(myres,paste(resloc,"ricker_SG_spear.RDS",sep=""))

    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,
                                          num_keep_last=500,resloc=resloc,
                                          params=params,p0=p0,model=model)

    saveRDS(myres,paste(resloc,"ricker_SG_kend.RDS",sep=""))
  }
 }
}
```

```{r plot_cause4copula_stat_ricker_SG, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_SG,mtime("Plotter_Cause4copula.R"))}

set.seed(seed)
source("Plotter_Cause4copula_stat.R")
#---------------------------------------

model<-"ricker"
fcode<-14
resloc0<-"./Results/Cause4copula_results/"

nsdlist<-c(0.1,1)
Klist<-c(100)

# for varying r (@r=1 transition from under- to over-compensatory dynamics )
rlist<-seq(from=0.1,to=1.3,by=0.2)

for(insd in nsdlist){
  for(iK in Klist){
   for(i in rlist){
    params<-c(i,iK) # this parameter is a vector with r and K
    p0<-c(params[2],params[2])

    resloc<-"./Results/Cause4copula_results/"

    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")

    resloc<-paste(tempo,"/",sep="")

    readres<-readRDS(paste(resloc,"ricker_SG_spear.RDS",sep=""))
    Plotter_Cause4copula_stat(res=readres)
    readres<-readRDS(paste(resloc,"ricker_SG_kend.RDS",sep=""))
    Plotter_Cause4copula_stat(res=readres)

  }
 }
}
```

```{r cause4copula_stat_ricker_joe, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("Cause4copula.R")
#-----------------
set.seed(seed)
#---------------------------------------
tag_ricker_J<-"999J" #this is a caching tag used later to cache output from present chunk

# This chunk stores results for ar1 model : joe

model<-"ricker"
fcode<-6
nsdlist<-c(0.1,1)
Klist<-c(100)
# for varying r (@r=1 transition from under- to over-compensatory dynamics )
rlist<-seq(from=0.1,to=1.3,by=0.2)

for(insd in nsdlist){
  for(iK in Klist){
   for(i in rlist){
    params<-c(i,iK) # this parameter is a vector with r and K
    p0<-c(params[2],params[2])

    resloc<-"./Results/Cause4copula_results/"

    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")

    if (!dir.exists(tempo)){
      dir.create(tempo)
    }
    resloc<-paste(tempo,"/",sep="")

  # for joe with spearman and kendall method

    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,
                                          num_keep_last=500,resloc=resloc,
                                          params=params,p0=p0,model=model)

    saveRDS(myres,paste(resloc,"ricker_J_kend.RDS",sep=""))
  }
 }
}
```

```{r plot_cause4copula_stat_ricker_J, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_J,mtime("Plotter_Cause4copula.R"))}

set.seed(seed)
source("Plotter_Cause4copula_stat.R")
#---------------------------------------

model<-"ricker"
fcode<-6
resloc0<-"./Results/Cause4copula_results/"

nsdlist<-c(0.1,1)
Klist<-c(100)

# for varying r (@r=1 transition from under- to over-compensatory dynamics )
rlist<-seq(from=0.1,to=1.3,by=0.2)

for(insd in nsdlist){
  for(iK in Klist){
   for(i in rlist){
    params<-c(i,iK) # this parameter is a vector with r and K
    p0<-c(params[2],params[2])

    resloc<-"./Results/Cause4copula_results/"

    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")

    resloc<-paste(tempo,"/",sep="")

    readres<-readRDS(paste(resloc,"ricker_J_kend.RDS",sep=""))
    Plotter_Cause4copula_stat(res=readres)

  }
 }
}
```

```{r cause4copula_stat_ricker_survivaljoe, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("Cause4copula.R")
#-----------------
set.seed(seed)
#---------------------------------------
tag_ricker_SJ<-"999SJ" #this is a caching tag used later to cache output from present chunk

# This chunk stores results for ar1 model : survival joe

model<-"ricker"
fcode<-16
nsdlist<-c(0.1,1)
Klist<-c(100)
# for varying r (@r=1 transition from under- to over-compensatory dynamics )
rlist<-seq(from=0.1,to=1.3,by=0.2)

for(insd in nsdlist){
  for(iK in Klist){
   for(i in rlist){
    params<-c(i,iK) # this parameter is a vector with r and K
    p0<-c(params[2],params[2])

    resloc<-"./Results/Cause4copula_results/"

    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")

    if (!dir.exists(tempo)){
      dir.create(tempo)
    }
    resloc<-paste(tempo,"/",sep="")

  # for survival joe with spearman and kendall method

    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,
                                          num_keep_last=500,resloc=resloc,
                                          params=params,p0=p0,model=model)

    saveRDS(myres,paste(resloc,"ricker_SJ_kend.RDS",sep=""))
  }
 }
}
```

```{r plot_cause4copula_stat_ricker_SJ, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_SJ,mtime("Plotter_Cause4copula.R"))}

set.seed(seed)
source("Plotter_Cause4copula_stat.R")
#---------------------------------------

model<-"ricker"
fcode<-16
resloc0<-"./Results/Cause4copula_results/"

nsdlist<-c(0.1,1)
Klist<-c(100)

# for varying r (@r=1 transition from under- to over-compensatory dynamics )
rlist<-seq(from=0.1,to=1.3,by=0.2)

for(insd in nsdlist){
  for(iK in Klist){
   for(i in rlist){
    params<-c(i,iK) # this parameter is a vector with r and K
    p0<-c(params[2],params[2])

    resloc<-"./Results/Cause4copula_results/"

    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")

    resloc<-paste(tempo,"/",sep="")

    readres<-readRDS(paste(resloc,"ricker_SJ_kend.RDS",sep=""))
    Plotter_Cause4copula_stat(res=readres)

  }
 }
}
```
<!--
\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_2500.pdf}
\caption[Comparing noise-copula parameters and estimated population copula parameters with $2500$ sampling points]{Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last \textbf{2500} points for both noise and population data for comparison). p-values $> 0.05$ (above horizontal dashed line) indicate good fit of population copula with same noise cpoula family. \label{fig_Cause4copula_params_GOF_2500}}
\end{center}
\end{figure}
-->

\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2u-D2l_vs_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, AR(1) model,
C copula.]{Comparison of copula structure of noise inputs and model 
outputs, AR(1) model, C copula. Red and blue points give 
means over 50 replicate simulations of the statistics listed on the left axes, 
red points are values for noise inputs and blue points are for
model outputs. Error bars are standard errors. 
$P$-values (purple points, right axis) are for a 
paired $t$-test of the null hypothesis that the distributions of 50 noise and 
output values have the same mean; $p$-values less than $0.05$ indicate that 
copula structure as measured by the listed statistic differs significant
between noise inputs and model outputs. See methods section for Q2 
for additional details.\label{fig_Cause4copula_C_kend_ar1}}
\end{center}
\end{figure}
<!--***DAN: Shya, can you please change the top legend of this and the similar-format
figures so instead of saying "noise copula" for the red points and "population copula" 
for the blue points it says "input noise copula" for the red points and "model output 
copula" for the blue ones? This is because one of our models is much more general than
just a population model, so I have changed the langauge in the text to reflect that.
Please be aware that when you make this change it will mean the next knit will take 
several hours, unless you go to the effort of reorganizing your code to put the sims
and the plots in separate chunks. So maybe make this change right before you go home,
so you can knit over night. Also please let me know, so I can knit overnight, too.

At the same time that you do that above, please see if you can fix the alignment
between the bottom three panels of Fig S32 (and subsequent analogous figures) so they
line up with the panels above that.
***Shya: Done.-->

\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_D2u-D2l_vs_Kendall's Tau.pdf}\\
\hspace{-0.8cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, AR(1) model, 
SC copula.]{Comparison of copula structure of noise inputs and model outputs, AR(1) 
model, SC copula. Same format as Figs \ref{fig_Cause4copula_C_kend_ar1} and 
\ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, see captions of those figures for 
details.\label{fig_Cause4copula_SC_kend_ar1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_D2u-D2l_vs_Kendall's Tau.pdf}\\
\hspace{-0.8cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, AR(1) model, 
G copula.]{Comparison of copula structure of noise inputs and model outputs, AR(1) 
model, G copula. Same format as Figs \ref{fig_Cause4copula_C_kend_ar1} and 
\ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, see captions of those figures for 
details.\label{fig_Cause4copula_G_kend_ar1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_D2u-D2l_vs_Kendall's Tau.pdf}\\
\hspace{-0.8cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, AR(1) model, 
SG copula.]{Comparison of copula structure of noise inputs and model outputs, AR(1) 
model, SG copula. Same format as Figs \ref{fig_Cause4copula_C_kend_ar1} and 
\ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, see captions of those figures for 
details.\label{fig_Cause4copula_SG_kend_ar1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_D2u-D2l_vs_Kendall's Tau.pdf}\\
\hspace{-0.8cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_scatter_D2umD2l_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, AR(1) model, 
J copula.]{Comparison of copula structure of noise inputs and model outputs, AR(1) 
model, J copula. Same format as Figs \ref{fig_Cause4copula_C_kend_ar1} and 
\ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, see captions of those figures for 
details.\label{fig_Cause4copula_J_kend_ar1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_D2u-D2l_vs_Kendall's Tau.pdf}\\
\hspace{-0.8cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_scatter_D2umD2l_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, AR(1) model, 
SJ copula.]{Comparison of copula structure of noise inputs and model outputs, AR(1) 
model, SJ copula. Same format as Figs \ref{fig_Cause4copula_C_kend_ar1} and 
\ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, see captions of those figures for 
details.\label{fig_Cause4copula_SJ_kend_ar1}}
\end{center}
\end{figure}

<!--adding results for stochastic ricker with weak noise : nsd=0.1-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_D2u-D2l_vs_Kendall's Tau.pdf}\\
\hspace{-0.8cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
model, weak noise, C copula.]{Comparison of copula structure of noise inputs and model 
outputs, stochastic Ricker model, weak noise, C copula. Same format as Figs 
\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
see captions of those figures for details.\label{fig_Cause4copula_C_kend_sRicker_nsd_0.1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_D2u-D2l_vs_Kendall's Tau.pdf}\\
\hspace{-0.8cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
model, weak noise, SC copula.]{Comparison of copula structure of noise inputs and model 
outputs, stochastic Ricker model, weak noise, SC copula. Same format as Figs 
\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
see captions of those figures for details. \label{fig_Cause4copula_SC_kend_sRicker_nsd_0.1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_D2u-D2l_vs_Kendall's Tau.pdf}\\
\hspace{-0.8cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
model, weak noise, G copula.]{Comparison of copula structure of noise inputs and model 
outputs, stochastic Ricker model, weak noise, G copula. Same format as Figs 
\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
see captions of those figures for details.\label{fig_Cause4copula_G_kend_sRicker_nsd_0.1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_D2u-D2l_vs_Kendall's Tau.pdf}\\
\hspace{-0.8cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
model, weak noise, SG copula.]{Comparison of copula structure of noise inputs and model 
outputs, stochastic Ricker model, weak noise, SG copula. Same format as Figs 
\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
see captions of those figures for details. \label{fig_Cause4copula_SG_kend_sRicker_nsd_0.1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_D2u-D2l_vs_Kendall's Tau.pdf}\\
\hspace{-0.8cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_scatter_D2umD2l_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
model, weak noise, J copula.]{Comparison of copula structure of noise inputs and model 
outputs, stochastic Ricker model, weak noise, J copula. Same format as Figs 
\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
see captions of those figures for details.\label{fig_Cause4copula_J_kend_sRicker_nsd_0.1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_D2u-D2l_vs_Kendall's Tau.pdf}\\
\hspace{-0.8cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_scatter_D2umD2l_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
model, weak noise, SJ copula.]{Comparison of copula structure of noise inputs and model 
outputs, stochastic Ricker model, weak noise, SJ copula. Same format as Figs 
\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
see captions of those figures for details.\label{fig_Cause4copula_SJ_kend_sRicker_nsd_0.1}}
\end{center}
\end{figure}

<!--adding results for stochastic ricker with strong noise : nsd=1-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_D2u-D2l_vs_Kendall's Tau.pdf}\\
\hspace{-0.8cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
model, strong noise, C copula.]{Comparison of copula structure of noise inputs and model 
outputs, stochastic Ricker model, strong noise, C copula. Same format as Figs 
\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
see captions of those figures for details.\label{fig_Cause4copula_C_kend_sRicker_nsd_1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_D2u-D2l_vs_Kendall's Tau.pdf}\\
\hspace{-0.8cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
model, strong noise, SC copula.]{Comparison of copula structure of noise inputs and model 
outputs, stochastic Ricker model, strong noise, SC copula. Same format as Figs 
\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
see captions of those figures for details. \label{fig_Cause4copula_SC_kend_sRicker_nsd_1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_D2u-D2l_vs_Kendall's Tau.pdf}\\
\hspace{-0.8cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
model, strong noise, G copula.]{Comparison of copula structure of noise inputs and model 
outputs, stochastic Ricker model, strong noise, G copula. Same format as Figs 
\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
see captions of those figures for details.\label{fig_Cause4copula_G_kend_sRicker_nsd_1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_D2u-D2l_vs_Kendall's Tau.pdf}\\
\hspace{-0.8cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
model, strong noise, SG copula.]{Comparison of copula structure of noise inputs and model 
outputs, stochastic Ricker model, strong noise, SG copula. Same format as Figs 
\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
see captions of those figures for details. \label{fig_Cause4copula_SG_kend_sRicker_nsd_1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_D2u-D2l_vs_Kendall's Tau.pdf}\\
\hspace{-0.8cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_scatter_D2umD2l_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
model, strong noise, J copula.]{Comparison of copula structure of noise inputs and model 
outputs, stochastic Ricker model, strong noise, J copula. Same format as Figs 
\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
see captions of those figures for details.\label{fig_Cause4copula_J_kend_sRicker_nsd_1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_D2u-D2l_vs_Kendall's Tau.pdf}\\
\hspace{-0.8cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_scatter_D2umD2l_Kendall's Tau.pdf}
\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
model, strong noise, SJ copula.]{Comparison of copula structure of noise inputs and model 
outputs, stochastic Ricker model, strong noise, SJ copula. Same format as Figs 
\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
see captions of those figures for details.\label{fig_Cause4copula_SJ_kend_sRicker_nsd_1}}
\end{center}
\end{figure}

```{r moran_effect_on_aphid_ff, echo=F, results="hide", cache=T, cache.extra=list(seed,temp_aphid,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
set.seed(seed)
source("FittingCopula_selective_loc.R")  # top most hierarchy source file

#cat(paste("start-time: ",Sys.time(),"\n"))
RES_temp_aphid<-RES_single_sp(d_allsp=temp_aphid,sp=1,families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=30,
                              good_loc_given=T,good_loc=c(2:11))
#cat(paste("stop-time: ",Sys.time(),"\n")) 
resloc<-"./Results/fitting_results/"
saveRDS(RES_temp_aphid,paste(resloc,file="AphidCopulaFit_selecloc_winter_neg_temp.RDS",sep=""))

LTmUT_temp<-(RES_temp_aphid$LTdep_AICw - RES_temp_aphid$UTdep_AICw)

sum(LTmUT_temp>0,na.rm = T)
sum(LTmUT_temp<0,na.rm = T) #66 out of 90 : suggests UT in leaf-plum curling aphids firstflight data (90 out of 90 UT)
```

<!--
```{r tab_temp_aphid_bestcop, echo=F, results="markup"}
knitr::kable(RES_temp_aphid$info_ord_copname[,,1], #longtable = T,
             format='latex',
             #caption.short = "Best-fitting copulas for each location pair from among those considered, according to AIC, for first flight data for the leaf-curling plum aphid.",
             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for **negative average winter temperature data (1958-2013)**. Copulas' abbreviations were given in \\nameref{MT-Methods}. NA occurs along the diagonal of the matrix because the diagonal represents a sampling location compared to itself.  \\label{tab_temp_aphid_bestcop}",
             booktabs=TRUE
             )
```

```{r tab_temp_aphid_AICw, echo=F, results='markup'}
knitr::kable(RES_temp_aphid$info_ord_AICw[,,1], 
             format='latex',
             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for **negative average winter temperature data**.\\label{tab_temp_aphid_AICw}",
             booktabs=TRUE
             )
```

```{r tab_temp_aphid_AICw_normal, echo=F, results='markup'}
source("normal_cop_AICw.R")
knitr::kable(normal_cop_AICw_matrix(r=RES_temp_aphid),
             format='latex',
             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for **negative average winter temperature data**, for comparison with Table \\ref{tab_ff_AICw}.\\label{tab_temp_aphid_AICw_normal}",
             booktabs=TRUE
             )
```

<!--
```{r tab_temp_aphid_p_CvM, echo=F, results='markup'}
knitr::kable(RES_temp_aphid$gfc_p_CvM,
             format='latex',
             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for **negative average winter temperature data**. Tests were based on bootstrapping as described in Methods.\\label{tab_temp_aphid_p_CvM}",
             booktabs=TRUE
             )
```
-->

<!--
```{r tab_temp_aphid_p_KS, echo=F, results='markup'}
knitr::kable(RES_temp_aphid$gfc_p_KS,
             format='latex',
             caption = "$P$-value results of a Kolmogorov-Smirnov-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for **negative average winter temperature data**. Tests were based on bootstrapping as described in Methods.\\label{tab_temp_aphid_p_KS}",
             booktabs=TRUE
             )
```
-->

<!--
```{r tab_temp_aphid_LTdep_AICw, echo=F, results='markup'}
knitr::kable(RES_temp_aphid$LTdep_AICw,
             format='latex',
             caption = "Model-averaged lower tail dependence statistic for each location pair for **negative average winter temperature data**.\\label{tab_temp_aphid_LTdep_AICw}",
             booktabs=TRUE
             )
```
-->

<!--
```{r tab_temp_aphid_UTdep_AICw, echo=F, results='markup'}
knitr::kable(RES_temp_aphid$UTdep_AICw,
             format='latex',
             caption = "Model-averaged upper tail dependence statistic for each location pair for **negative average winter temperature data**.\\label{tab_temp_aphid_UTdep_AICw}",
             booktabs=TRUE
             )
```
-->

<!--
```{r tab_temp_aphid_LTmUT, echo=F, results='markup'}
knitr::kable(RES_temp_aphid$LTdep_AICw - RES_temp_aphid$UTdep_AICw,
             digits = 5,
             format='latex',
             caption = "Model averaged lower tail dependence minus model averaged upper tail dependence statistic for each location pair for **negative average winter temperature data**.\\label{tab_temp_aphid_LTmUT}",
             booktabs=TRUE
             )
```
-->

```{r asym_sens_b_variation, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"),mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
set.seed(seed)
#-----------------------------------------------------------------------------------
source("OurBiCopSelect.R")
source("bivfunctionplot.R")
source("asym_sensitivity.R")
#-----------------------------------------------------------------------------------
# This function takes 
# Input :
#       mg : global noise matrix (dim = N by 2)
#       ml : local noise matrix (dim = N by 2)
#       a : a number : local noise coeeficient
#       b : growth factor of pop
#       decrease : logical (controls the nature of global noise)
#       numkeep_last : number of points you want to keep for the pop matrix from the end
#       ploton : logical for optional plot

# Output :
#     a list of 4 pop matrices 
#                         1. (for b=+ve, decrease=F)
#                         2. (for b=+ve, decrease=T)
#                         3. (for b=-ve, decrease=F)
#                         4. (for b=-ve, decrease=T)

plotter_asym_sens_b_varn<-function(mg,ml,a,b,numkeep_last,ploton,resloc){

  pop_g<-Simulator_asym_sens(mg=mg,ml=ml,a=a,b=b,p0=c(0,0),decrease=F,numkeep_last=numkeep_last)
  pop_d_g<-Simulator_asym_sens(mg=mg,ml=ml,a=a,b=b,p0=c(0,0),decrease=T,numkeep_last=numkeep_last)

  if(ploton==T){
    pdf(paste0(resloc,"a_",a,"_b_",b,"_r_",r,"_asym_sens.pdf",sep=""),height=3,width=6)
    op<-par(mfrow=c(1,2),mar=c(4,4.2,1.5,1),mgp=c(2.5,0.6,0))
    #plot(pop[,1],pop[,2],col="red",main = bquote(beta ==.(beta)))
    pop<-pop_g$pop
    geps<-pop_g$g_noise
    #plot(mg[,1],geps[,1],cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[1])),main="location=1")
    #plot(mg[,2],geps[,2],cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[1])),main="location=2")
    plot(pop[,1],pop[,2],xlab=expression(paste(P[1])), ylab=expression(paste(P[2])),cex.lab=2,cex.axis=2,col="grey",pch=20,cex=1.4)
    #title(expression(paste("g = ",g[1]," , ")), adj=0.06,cex.main=1.5)
    #mtext(paste("a = ",a,", b = ",b),side = 3, line=0.2, adj=0.9)

    pop_d<-pop_d_g$pop
    geps<-pop_d_g$g_noise
    #plot(mg[,1],geps[,1],cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[2])),main="location=1")
    #plot(mg[,2],geps[,2],cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[2])),main="location=2")
    plot(pop_d[,1],pop_d[,2],xlab=expression(paste(P[1])), ylab=expression(paste(P[2])),cex.lab=2,cex.axis=2,col="grey",pch=20,cex=1.4)
    #title(expression(paste("g = ",g[2]," , ")), adj=0.06,cex.main=1.5)
    #mtext(paste("a = ",a,", b = ",b),side = 3, line=0.2, adj=0.9)

    par(op)
    dev.off()
  }

  return(list(pop=pop,
              pop_d=pop_d))

}
#----------------------------------------------------------
N<-25000
r<-0.8
a<-0.2
numkeep_last<-2500
temp<-get_noise_glob_loc_mat(N=N,r=r)
mg<-temp$mg
ml<-temp$ml

resloc<-"./Results/asym_sens_results/"
ans_b_0.1<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=0.1,numkeep_last=numkeep_last,ploton=T,resloc=resloc)
ans_b_neg0.1<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=-0.1,numkeep_last=numkeep_last,ploton=T,resloc=resloc)
ans_b_0.5<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=0.5,numkeep_last=numkeep_last,ploton=T,resloc=resloc)
ans_b_neg0.5<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=-0.5,numkeep_last=numkeep_last,ploton=T,resloc=resloc)

vlist_b_0.1_b_0.5<-c(ans_b_0.1,ans_b_neg0.1,ans_b_0.5,ans_b_neg0.5) # a list of 8 matrices

summary_stat<-as.data.frame(matrix(NA,7,8),stringsAsFactors = F) # Table for model selection summary
rownames(summary_stat)<-c("bestcop_AICw","N_AICw","relLTdep_AICw","relUTdep_AICw","p_CvM","p_KS","BS_success%")
#bg <- intToUtf8(0x03B2) 
cnm<-c("b_0.1_g~1~","b_0.1_g~2~","b_-0.1_g~1~","b_-0.1_g~2~","b_0.5_g~1~","b_0.5_g~2~","b_-0.5_g~1~","b_-0.5_g~2~")
#cname<-paste0(bg,cnm)
colnames(summary_stat)<-cnm

ms_asym_b_0.1_b_0.5<-vector(mode = "list", length = 8) # initializing list to store model selection results for all possible b variation
names(ms_asym_b_0.1_b_0.5)<-cnm

stat_npa_asym_b_0.1_b_0.5<-vector(mode = "list", length = 8) # initializing list to store NPA results for all possible b variation
names(stat_npa_asym_b_0.1_b_0.5)<-cnm

for(ic in c(1:8)){
  v<-vlist_b_0.1_b_0.5[[ic]]
  res<-OurBiCopSelect(u1=v[,1],u2=v[,2],families=c(1,3:10,13:14,16:20),level=0.05,
                      AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,
                      gofnormal=F,status=F)

  ms_asym_b_0.1_b_0.5[[ic]]<-res

  summary_stat[1,ic]<-max(res$InfCritRes$AICw)
  summary_stat[2,ic]<-res$InfCritRes$AICw[1]
  summary_stat[3,ic]<-res$relLTdep_AICw
  summary_stat[4,ic]<-res$relUTdep_AICw
  summary_stat[5,ic]<-res$GofRes_CvM
  summary_stat[6,ic]<-res$GofRes_KS
  summary_stat[7,ic]<-(res$Numboot_success/res$Numboot)*100

  stat_npa_asym_b_0.1_b_0.5[[ic]]<-bivfunctionplot(v=v,resloc="./Results/asym_sens_results/",nametag=cnm[ic],numbin=10)

}

saveRDS(summary_stat,"./Results/asym_sens_results/b_variation_summary_stat.RDS")
saveRDS(ms_asym_b_0.1_b_0.5,"./Results/asym_sens_results/ms_asym_b_0.1_b_0.5.RDS")
saveRDS(stat_npa_asym_b_0.1_b_0.5,"./Results/asym_sens_results/stat_npa_asym_b_0.1_b_0.5.RDS")
```

```{r overwritting_plots_asym_cop,echo=F,results="hide"}
pdf("./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf",height=3,width=6)
op<-par(mfrow=c(1,2),mar=c(3.5,4,1,1),mgp=c(2.2,0.5,0))
pop<-ans_b_0.1$pop
plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
     cex.axis=1,col=rgb(0,0,0,0.5),pch=20,cex=0.6)
text(0.1,0.95,"A",cex=2)
pop<-ans_b_0.1$pop_d
plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
     cex.axis=1,pch=20,col=rgb(0,0,0,0.5),cex=0.6)
text(0.1,0.95,"B",cex=2)
par(op)
dev.off()

pdf("./Results/asym_sens_results/a_0.2_b_-0.1_r_0.8_asym_sens.pdf",height=3,width=6)
op<-par(mfrow=c(1,2),mar=c(3.5,4,1,1),mgp=c(2.2,0.5,0))
pop<-ans_b_neg0.1$pop
plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
     cex.axis=1,pch=20,col=rgb(0,0,0,0.5),cex=0.6)
text(0.1,0.95,"A",cex=2)
pop<-ans_b_neg0.1$pop_d
plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
     cex.axis=1,col=rgb(0,0,0,0.5),pch=20,cex=0.6)
text(0.1,0.95,"B",cex=2)
par(op)
dev.off()

pdf("./Results/asym_sens_results/a_0.2_b_0.5_r_0.8_asym_sens.pdf",height=3,width=6)
op<-par(mfrow=c(1,2),mar=c(3.5,4,1,1),mgp=c(2.2,0.5,0))
pop<-ans_b_0.5$pop
plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
     cex.axis=1,col=rgb(0,0,0,0.5),pch=20,cex=0.6)
text(0.1,0.95,"A",cex=2)
pop<-ans_b_0.5$pop_d
plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
     cex.axis=1,col=rgb(0,0,0,0.5),pch=20,cex=0.6)
text(0.1,0.95,"B",cex=2)
par(op)
dev.off()

pdf("./Results/asym_sens_results/a_0.2_b_-0.5_r_0.8_asym_sens.pdf",height=3,width=6)
op<-par(mfrow=c(1,2),mar=c(3.5,4,1,1),mgp=c(2.2,0.5,0))
pop<-ans_b_neg0.5$pop
plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
     cex.axis=1,col=rgb(0,0,0,0.5),pch=20,cex=0.6)
text(0.1,0.95,"A",cex=2)
pop<-ans_b_neg0.5$pop_d
plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
     cex.axis=1,col=rgb(0,0,0,0.5),pch=20,cex=0.6)
text(0.1,0.95,"B",cex=2)
par(op)
dev.off()
```

```{r tab_summary_stat_asym_sens,echo=F,results="markup"}
summary_stat<-summary_stat[c(1:6),]
rownames(summary_stat)<-c("Best-fit copula AICw","N-copula AICw","Model-avg. LT","Model-avg. UT","p, CvM","p, KS")
knitr::kable(summary_stat, booktabs = T, escape=F,
             format = "latex", linesep = "",
             digits=4, col.names=c("b=0.1, g$_{1}$","b=0.1, g$_{2}$",
                                   "b=-0.1, g$_{1}$","b=-0.1, g$_{2}$",
                                   "b=0.5, g$_{1}$","b=0.5, g$_{2}$",
                                   "b=-0.5, g$_{1}$","b=-0.5, g$_{2}$"),
      caption.short="Asymmetric sensitivity model, model selection results",
      caption="Asymmetric sensitivity model, model selection results. 
The output of simulations from the second model detailed in 
the methods section for Q2 (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
\\ref{fig_ms_asym_b_-0.1}-\\ref{fig_ms_asym_b_-0.5}) were subjected to the model 
selection procedures detailed in the 
methods section for Q1. Best-fit copula AICw: the AIC weight of the best-fitting copula family;
N-copula AICw: the AIC weight of the N copula family; Model-avg. LT: the lower-tail 
statistic, weighted-averaged across models using AICw; Model-avg. UT: same, but for the 
upper-tail statistic; p, CvM: p-value for goodness of fit test of the best-fitting copula
family using the test based on the Cramer-von Mises statistic; p, KS: p-value for goodness of fit 
test of the best-fitting copula family using the test based on the Kolmogorov-Smirnov 
statistic.\\label{tab_summary_stat_asym_sens}")
```

```{r read_tab_stat_npa_asym,echo=F,results="asis"}
#knitr::opts_knit$set(kable.force.latex = TRUE)
#tab_asym<-readRDS("./Results/asym_sens_results/stat_npa_asym_b_0.1_b_0.9.RDS")

#for(i in c(1:8)){
#  temp<-make_table_stat_fn(data=tab_asym[[i]])
#  temp2<-knitr::kable(temp, booktabs = T, 
#             format = "latex",
      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
#      caption =paste(cnm[i]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence."))
#  print(temp2)
#  cat("\\newpage")
#}
```

```{r tab_stat_npa_asym,echo=F,results="markup"}
library(kableExtra)

tab_asym<-readRDS("./Results/asym_sens_results/stat_npa_asym_b_0.1_b_0.5.RDS")
tab_stat_npa_asym<-c()
for(i in c(1:8)){
  temp<-make_table_stat_fn(data=tab_asym[[i]])
  tab_stat_npa_asym<-rbind(tab_stat_npa_asym,temp)
}
gtext<-c("b=0.1, g$_{1}$","b=0.1, g$_{2}$","b=-0.1, g$_{1}$","b=-0.1, g$_{2}$","b=0.5, g$_{1}$","b=0.5, g$_{2}$","b=-0.5, g$_{1}$","b=-0.5, g$_{2}$")
tab_stat_npa_asym<-cbind(b=rep(c(0.1,0.1,-0.1,-0.1,0.5,0.5,-0.5,-0.5),each=3),
  g=rep(c("g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$"),each=3),tab_stat_npa_asym)
knitr::kable(tab_stat_npa_asym, booktabs = T, longtable=T,
             format = "latex",linesep = "",escape=F,
             caption.short = "Asymmetric sensitivity model, nonparametric statistics results",
             caption ="Asymmetric sensitivity model, nonparametric statistics results. 
The output of simulations from the second model detailed in 
the methods section for Q2 (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
\\ref{fig_ms_asym_b_-0.1}-\\ref{fig_ms_asym_b_-0.5}) were subjected to nonparamtric statistical 
analyses decribed in the methods section for Q1. The listed statistic was computed for the
output of the model with the indicated b and g, and the value was compared to 1000 values
of the same statistic computed on surrogate time series randomized to have N-copula structure
(see methods for Q1). A table entry $<X$ (respectively, $>X$) indicates the value of the 
given statistic on the data was less (respectively, more) than its value on $X$ of the 
surrogates, so entries of the form $<X$ (respectively, $>X$) for $X$ equal to $975$ or above 
indicate that upper-tail (respectively, lower-tail) dependence was significantly stronger than 
lower-tail dependence compared to an N-copula null hypothesis. `FS' stands for `first slice'
and refers to the bounds 0 to 0.1; `LS' stands for `last slice' and refers to the bounds
0.9 to 1.\\label{tab_stat_npa_asym}")#%>%
 #            kable_styling()%>%
 #            group_rows(index=setNames(c(3,3,3,3,3,3,3,3), gtext))
 #            group_rows("b=0.1, g$_{1}$", 2,4, escape=F) %>%
 #            group_rows("b=0.1, g$_{2}$", 5,7, escape=F) %>%
 #            group_rows("b=-0.1, g$_{1}$", 8,10, escape=F) %>%
 #            group_rows("b=-0.1, g$_{2}$", 11,13, escape=F) %>%
 #            group_rows("b=0.9, g$_{1}$", 14,16, escape=F) %>%
 #            group_rows("b=0.9, g$_{2}$", 17,19, escape=F) %>%
 #            group_rows("b=-0.9, g$_{1}$", 20,22, escape=F) %>%
 #            group_rows("b=-0.9, g$_{2}$", 23,25, escape=F)
```

\begin{figure}[!h]
\begin{center}
\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_-0.1_r_0.8_asym_sens.pdf}
\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=-0.1$.\label{fig_ms_asym_b_-0.1}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_0.5_r_0.8_asym_sens.pdf}
\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=0.9$.\label{fig_ms_asym_b_0.5}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_-0.5_r_0.8_asym_sens.pdf}
\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=-0.9$.\label{fig_ms_asym_b_-0.5}}
\end{center}
\end{figure}

```{r skewness_aphidcount_allsp_plot, echo=F, results="hide", cache=T, cache.extra=list(seed,data_aphid_count,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}

# For sp =1, sp=4 and sp=6 : kendall preserving surrogates gives warning : Error in ncsurrog: ncov is not positive semidefinite
# also it can't handle ties for "kendall" and code crashes
# you can see those msg by deleting results="hide" and warning=F option in r chunk

source("good_loclist.R")
source("skewness_multivar_dataset.R")
set.seed(seed)

numsurrog<-10000

splist<-c(1:20)
skw_aphid_count_spearman<-as.data.frame(matrix(NA,nrow=length(splist),ncol=3))
colnames(skw_aphid_count_spearman)<-c("realskw","p_left","p_right")
rownames(skw_aphid_count_spearman)<-paste("sp",splist,sep="")
pdf("./Results/skewness_results/skewness_aphid_count/aphid_count_allsp_skewness_spearman.pdf",height=25,width=25)
op<-par(mfrow=c(4,5),mar=c(5,5,1,1),mgp=c(3,1,0))
cat("Now for spearman preserving surrogs : aphid count","\n")
for(i in splist){
  cat("sp=",i,"\n")
  loclist<-good_loclist(d_allsp=data_aphid_count,sp=i,data_pt_thrs=30)
  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_aphid_count),loclist=loclist,
                        numsurrog=numsurrog,
                        ploton=T,
                        corpres="spearman") 
  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
  p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
  skw_aphid_count_spearman$realskw[i]<-ans$realskw
  skw_aphid_count_spearman$p_left[i]<-p_left
  skw_aphid_count_spearman$p_right[i]<-p_right
  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
}
par(op)
dev.off()
saveRDS(skw_aphid_count_spearman,"./Results/skewness_results/skewness_aphid_count/skw_aphid_count_spearman.RDS")

#skw_aphid_count_kendall<-as.data.frame(matrix(NA,nrow=length(splist),ncol=3))
#colnames(skw_aphid_count_kendall)<-c("realskw","p_left","p_right")
#rownames(skw_aphid_count_kendall)<-paste("sp",splist,sep="")
#pdf("./Results/skewness_results/skewness_aphid_count/aphid_count_allsp_skewness_kendall.pdf",height=25,width=25)
#op<-par(mfrow=c(4,5),mar=c(5,5,1,1),mgp=c(3,1,0))
#cat("Now for kendall preserving surrogs : aphid count","\n")
#for(i in splist){
#  cat("sp=",i,"\n")
#  loclist<-good_loclist(d_allsp=data_aphid_count,sp=i,data_pt_thrs=30)
#  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_aphid_count),loclist=loclist,
#                        numsurrog=numsurrog,
#                        ploton=T,
#                        corpres="kendall") 
#  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
# p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
#  skw_aphid_count_kendall$realskw[i]<-ans$realskw
#  skw_aphid_count_kendall$p_left[i]<-p_left
#  skw_aphid_count_kendall$p_right[i]<-p_right
#  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
#}
#par(op)
#dev.off()
#saveRDS(skw_aphid_count_kendall,"./Results/skewness_results/skewness_aphid_count/skw_aphid_count_kendall.RDS")
```

```{r skewness_aphidff_allsp_plot, echo=F, results="hide", cache=T, cache.extra=list(seed,data_aphid_ff,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}

# For sp =12 and sp=16 : kendall preserving surrogates gives warning : Error in ncsurrog: ncov is not positive semidefinite,
# also it can't handle ties for "kendall" and code crashes
# you can see those msg by deleting results="hide" and warning=F option in r chunk

source("good_loclist.R")
source("skewness_multivar_dataset.R")
set.seed(seed)

numsurrog<-10000

splist<-c(1:20)
skw_aphid_ff_spearman<-as.data.frame(matrix(NA,nrow=length(splist),ncol=3))
colnames(skw_aphid_ff_spearman)<-c("realskw","p_left","p_right")
rownames(skw_aphid_ff_spearman)<-paste("sp",splist,sep="")
pdf("./Results/skewness_results/skewness_aphid_ff/aphid_ff_allsp_skewness_spearman.pdf",height=25,width=25)
op<-par(mfrow=c(4,5),mar=c(5,5,1,1),mgp=c(3,1,0))
cat("Now for spearman preserving surrogs : aphid ff","\n")
for(i in splist){
  cat("sp=",i,"\n")
  loclist<-good_loclist(d_allsp=data_aphid_ff,sp=i,data_pt_thrs=30)
  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_aphid_ff),loclist=loclist,
                        numsurrog=numsurrog,
                        ploton=T,
                        corpres="spearman") 
  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
  p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
  skw_aphid_ff_spearman$realskw[i]<-ans$realskw
  skw_aphid_ff_spearman$p_left[i]<-p_left
  skw_aphid_ff_spearman$p_right[i]<-p_right
  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
}
par(op)
dev.off()
saveRDS(skw_aphid_ff_spearman,"./Results/skewness_results/skewness_aphid_ff/skw_aphid_ff_spearman.RDS")

#skw_aphid_ff_kendall<-as.data.frame(matrix(NA,nrow=length(splist),ncol=3))
#colnames(skw_aphid_ff_kendall)<-c("realskw","p_left","p_right")
#rownames(skw_aphid_ff_kendall)<-paste("sp",splist,sep="")
#pdf("./Results/skewness_results/skewness_aphid_ff/aphid_ff_allsp_skewness_kendall.pdf",height=25,width=25)
#op<-par(mfrow=c(4,5),mar=c(5,5,1,1),mgp=c(3,1,0))
#cat("Now for kendall preserving surrogs : aphid ff","\n")
#for(i in splist){
#  cat("sp=",i,"\n")
#  loclist<-good_loclist(d_allsp=data_aphid_ff,sp=i,data_pt_thrs=30)
#  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_aphid_ff),loclist=loclist,
#                        numsurrog=numsurrog,
#                        ploton=T,
#                        corpres="kendall") 
#  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
#  p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
#  skw_aphid_ff_kendall$realskw[i]<-ans$realskw
#  skw_aphid_ff_kendall$p_left[i]<-p_left
#  skw_aphid_ff_kendall$p_right[i]<-p_right
#  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
#}
#par(op)
#dev.off()
#saveRDS(skw_aphid_ff_kendall,"./Results/skewness_results/skewness_aphid_ff/skw_aphid_ff_kendall.RDS")
```

```{r skewness_plankton_allsp_plot, echo=F, results="hide", cache=T, cache.extra=list(seed,data_plankton_north_sea,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}

# There is no error msg or warning coming from ncsurrog : ncov semidefinite matrix for any species with spearman or kendall preserving surrogates.
# But there may be error handling ties with "kendall"
# So,for safety, now we will only analyze with "spearman"

source("good_loclist.R")
source("skewness_multivar_dataset.R")
set.seed(seed)

splist<-c(1:22)
skw_plankton_spearman<-as.data.frame(matrix(NA,nrow=length(splist),ncol=3))
colnames(skw_plankton_spearman)<-c("realskw","p_left","p_right")
rownames(skw_plankton_spearman)<-paste("sp",splist,sep="")
#skw_plankton_kendall<-skw_plankton_spearman
pdf("./Results/skewness_results/skewness_plankton_north_sea/plankton_allsp_skewness_spearman.pdf",height=30,width=30)
op<-par(mfrow=c(4,6),mar=c(5,5,1,1),mgp=c(3,1,0))
#cat("Now for spearman preserving surrogs : pns","\n")
for(i in splist){
  #cat("sp=",i,"\n")
  loclist<-good_loclist(d_allsp=data_plankton_north_sea,sp=i,data_pt_thrs=45)
  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_plankton_north_sea),loclist=loclist,
                        numsurrog=numsurrog,
                        ploton=T,
                        corpres="spearman") 
  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
  p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
  skw_plankton_spearman$realskw[i]<-ans$realskw
  skw_plankton_spearman$p_left[i]<-p_left
  skw_plankton_spearman$p_right[i]<-p_right
  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
}
par(op)
dev.off()
saveRDS(skw_plankton_spearman,"./Results/skewness_results/skewness_plankton_north_sea/skw_plankton_spearman.RDS")

#pdf("./Results/skewness_results/skewness_plankton_north_sea/plankton_allsp_skewness_kendall.pdf",height=30,width=30)
#op<-par(mfrow=c(4,6),mar=c(5,5,1,1),mgp=c(3,1,0))
#cat("Now for kendall preserving surrogs : pns","\n")
#for(i in splist){
  #cat("sp=",i,"\n")
 # loclist<-good_loclist(d_allsp=data_plankton_north_sea,sp=i,data_pt_thrs=45)
#  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_plankton_north_sea),loclist=loclist,
#                        numsurrog=numsurrog,
#                        ploton=T,
#                        corpres="kendall") 
#  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
#  p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
#  skw_plankton_kendall$realskw[i]<-ans$realskw
#  skw_plankton_kendall$p_left[i]<-p_left
#  skw_plankton_kendall$p_right[i]<-p_right
#  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
#}
#par(op)
#dev.off()
#saveRDS(skw_plankton_kendall,"./Results/skewness_results/skewness_plankton_north_sea/skw_plankton_kendall.RDS")
```

```{r skewness_count_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}

# Only done for "Spearman" method

source("good_loclist.R")
source("skewness_multivar_dataset.R")
set.seed(seed)

numsurrog<-10000
loclist<-good_loclist(d_allsp=data_aphid_count,sp=10,data_pt_thrs=30)

#pdf("./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10_kendall.pdf",height=6,width=8)
#op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
#answer1k<-skewness_testing(ts_matrix=sp_data(sp=10,d_allsp=data_aphid_count),loclist=loclist,numsurrog=numsurrog,ploton=T,
#                          corpres="kendall") 
#p_left<-sum(answer1k$surrogskw < answer1k$realskw)/numsurrog
#mtext(paste0("p = ",p_left),side=3,line=-0.5,cex = 2)
#par(op)
#dev.off()

pdf("./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10_spearman.pdf",height=5,width=5)
op<-par(mar=c(4,5,1,1),mgp=c(3,1,0))
#op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
answer1s<-skewness_testing(ts_matrix=sp_data(sp=10,d_allsp=data_aphid_count),loclist=loclist,numsurrog=numsurrog,ploton=F,
                          corpres="spearman") 
hist(answer1s$surrogskw,breaks=100,main="",xlab="Skewness",ylab="Frequency",cex.lab=2,cex.axis=2,col="grey",border=F)
#abline(v=realskw,col="red")
points(x=answer1s$realskw,y=0,col="black",pch=15,cex=2)

p_left<-sum(answer1s$surrogskw < answer1s$realskw)/numsurrog
mtext(paste0("p = ",p_left),side=3,line=-0.5,cex = 2)
par(op)
dev.off()
```

```{r skewness_ff_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,data_aphid_ff,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}

# Only done for "Spearman" method

source("good_loclist.R")
source("skewness_multivar_dataset.R")
set.seed(seed)

numsurrog<-10000
loclist<-good_loclist(d_allsp=data_aphid_ff,sp=11,data_pt_thrs=30)

#pdf("./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11_kendall.pdf",height=6,width=8)
#op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
#answer2k<-skewness_testing(ts_matrix=sp_data(sp=11,d_allsp=data_aphid_ff),loclist=loclist,numsurrog=numsurrog,ploton=T,
#                          corpres="kendall") 
#p_right<-sum(answer2k$surrogskw > answer2k$realskw)/numsurrog
#mtext(paste0("p = ",p_right),side=3,line=-0.5,cex = 2)
#par(op)
#dev.off()

pdf("./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11_spearman.pdf",height=5,width=5)
op<-par(mar=c(4,5,1,1),mgp=c(3,1,0))
answer2s<-skewness_testing(ts_matrix=sp_data(sp=11,d_allsp=data_aphid_ff),loclist=loclist,numsurrog=numsurrog,ploton=F,
                          corpres="spearman")
hist(answer2s$surrogskw,breaks=100,main="",xlab="Skewness",ylab="Frequency",cex.lab=2,cex.axis=2,col="grey",border=F)
#abline(v=realskw,col="red")
points(x=answer2s$realskw,y=0,col="black",pch=15,cex=2)

p_right<-sum(answer2s$surrogskw > answer2s$realskw)/numsurrog
mtext(paste0("p = ",p_right),side=3,line=-0.5,cex = 2)
par(op)
dev.off()
```

```{r skewness_pns_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,data_plankton_north_sea,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}

# Only done for "Spearman" method

source("good_loclist.R")
source("skewness_multivar_dataset.R")
set.seed(seed)

numsurrog<-10000
loclist<-good_loclist(d_allsp=data_plankton_north_sea,sp=16,data_pt_thrs=45)

#pdf("./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16_kendall.pdf",height=6,width=8)
#op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
#answer3k<-skewness_testing(ts_matrix=sp_data(sp=16,d_allsp=data_plankton_north_sea),loclist=loclist,numsurrog=numsurrog,ploton=T,
#                          corpres="kendall") 
#p_left<-sum(answer3k$surrogskw < answer3k$realskw)/numsurrog
#mtext(paste0("p = ",p_left),side=3,line=-0.5,cex = 2)
#par(op)
#dev.off()

pdf("./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16_spearman.pdf",height=5,width=5)
op<-par(mar=c(4,5,1,1),mgp=c(3,1,0))
answer3s<-skewness_testing(ts_matrix=sp_data(sp=16,d_allsp=data_plankton_north_sea),loclist=loclist,numsurrog=numsurrog,ploton=F,
                          corpres="spearman") 
hist(answer3s$surrogskw,breaks=100,main="",xlab="Skewness",ylab="Frequency",cex.lab=2,cex.axis=2,col="grey",border=F)
#abline(v=realskw,col="red")
points(x=answer3s$realskw,y=0,col="black",pch=15,cex=2)

p_left<-sum(answer3s$surrogskw < answer3s$realskw)/numsurrog
mtext(paste0("p = ",p_left),side=3,line=-0.5,cex = 2)
par(op)
dev.off()
```

```{r skewness_methane_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,data_plankton_north_sea,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}

# Only done for "Spearman" method

source("good_loclist.R")
source("skewness_multivar_dataset.R")
set.seed(seed)

numsurrog<-10000
loclist<-good_loclist(d_allsp=methane_data,sp=1,data_pt_thrs=50)

#pdf("./Results/skewness_results/skewness_methane/skewness_methane_kendall.pdf",height=6,width=8)
#op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
#answer1k<-skewness_testing(ts_matrix=sp_data(sp=1,d_allsp=methane_data),loclist=loclist,numsurrog=numsurrog,ploton=T,
#                           corpres="kendall") 
#p_left<-sum(answer1k$surrogskw < answer1k$realskw)/numsurrog
#mtext(paste0("p = ",p_left),side=3,line=-0.5,cex = 2)
#par(op)
#dev.off()

pdf("./Results/skewness_results/skewness_methane/skewness_methane_spearman.pdf",height=5,width=5)
op<-par(mar=c(4,5,1,1),mgp=c(3,1,0))
answer1s<-skewness_testing(ts_matrix=sp_data(sp=1,d_allsp=methane_data),loclist=loclist,numsurrog=numsurrog,ploton=F,
                           corpres="spearman") 
hist(answer1s$surrogskw,breaks=100,main="",xlab="Skewness",ylab="Frequency",cex.lab=2,cex.axis=2,col="grey",border=F)
#abline(v=realskw,col="red")
points(x=answer1s$realskw,y=0,col="black",pch=15,cex=2)

p_left<-sum(answer1s$surrogskw < answer1s$realskw)/numsurrog
mtext(paste0("p = ",p_left),side=3,line=-0.5,cex = 2)
par(op)
dev.off()
```

<!--make surrogate datasets for TL analysis to follow-->
```{r taylors_law_make_surrogs, echo=F, cache=T, cache.extra=list(seed,mtime("copsurrognd.R"))}
source("copsurrognd.R")
set.seed(seed)

#***constants
N<-50 #length of time series
n<-25 #number of sampling locations
#numdat<-10 #the number of datasets to get from the normal copula
numdat<-1000 #the number of datasets to get from the normal copula
rho<-0.7 #normal-copula parameter for dependence between locations
          #this is a covariance, see below call to rmvnorm

#***generate the starting data, which has normal copula structure
#because it comes originally from a multivariate normal, before 
#transformation
sig<-matrix(rho,n,n)
diag(sig)<-1
m<-rmvnorm(N*numdat,mean=rep(0,n),sigma=sig) # each column of m has normal histogram
m<-array(m,c(N,numdat,n))
m<-aperm(m,c(1,3,2)) # 2nd dim is normally distributed e.g. hist(m[,1,])
gshape<-7.5
gscale<-1 #gamma parameters

cop_to_marg_trans<-function(x){
  return(qgamma(x,shape=gshape,scale=gscale)) #this is where we choose the marginals,
                                  #should be something skewed
}

marg_to_cop_trans<-function(x){ #transforms to copula space
  return(pgamma(x,shape=gshape,scale=gscale))
}

# pnorm(m) makes normal to uniform
# then qgamma(...) makes uniform to gamma distributed
m<-cop_to_marg_trans(pnorm(m))

#***surrogates using a Clayton copula, kendall preserving

#get the Clayton parameter to use
ncop<-normalCopula(rho,2)
ccop<-claytonCopula(2,2)
cparam<-iTau(ccop,tau(ncop))
ccop<-claytonCopula(cparam,n)

#generate the surrogates
csurrog<-array(NA,dim(m))
for (counter in 1:dim(m)[3]){
  csurrog[,,counter]<-copsurrognd(m[,,counter],targetcop=ccop,numsurrog=1)
}

# #test to see whether kendalls are similar between m and csurrog - 
#only works at all as a test if N is large (I used 1000)
# cor(m[,1,1],m[,2,1],method='kendall')
# cor(csurrog[,1,1],csurrog[,2,1],method='kendall')
# cor(m[,3,1],m[,4,1],method='kendall')
# cor(csurrog[,3,1],csurrog[,4,1],method='kendall')
# cor(m[,5,1],m[,6,1],method='kendall')
# cor(csurrog[,5,1],csurrog[,6,1],method='kendall')
# 
# cor(m[,1,2],m[,2,2],method='kendall')
# cor(csurrog[,1,2],csurrog[,2,2],method='kendall')
# cor(m[,3,2],m[,4,2],method='kendall')
# cor(csurrog[,3,2],csurrog[,4,2],method='kendall')
# cor(m[,5,2],m[,6,2],method='kendall')
# cor(csurrog[,5,2],csurrog[,6,2],method='kendall')

#***surrogates using a Frank copula, kendall preserving
#get the Frank parameter to use
fcop<-frankCopula(4,2)
fparam<-iTau(fcop,tau(ncop))
fcop<-frankCopula(fparam,n)

#generate the surrogates
fsurrog<-array(NA,dim(m))
for (counter in 1:dim(m)[3]){
  fsurrog[,,counter]<-copsurrognd(m[,,counter],targetcop=fcop,numsurrog=1)
}

#***surrogates using a survival Clayton, kendall preserving
#take the clayton surrogates and transform them to the 
#copula space, then flip, then transform back
scsurrog<-cop_to_marg_trans((1-marg_to_cop_trans(csurrog)))

# check : plot(csurrog[,1,1],csurrog[,2,1]) and plot(scsurrog[,1,1],scsurrog[,2,1])

# #test: should have identical taus as csurrog
# cor(csurrog[,1,1],csurrog[,2,1],meth='kendall')
# cor(scsurrog[,1,1],scsurrog[,2,1],meth='kendall')
# cor(csurrog[,1,2],csurrog[,2,2],meth='kendall')
# cor(scsurrog[,1,2],scsurrog[,2,2],meth='kendall')

kend.surrog<-list(cla=csurrog,nor=m,fra=fsurrog,scl=scsurrog)

#***surrogates using a Clayton copula, spearman preserving
#get the Clayton parameter to use
ccop<-claytonCopula(2,2)
cparam<-iRho(ccop,rho(ncop))
ccop<-claytonCopula(cparam,n)

#generate the surrogates
csurrog<-array(NA,dim(m))
for (counter in 1:dim(m)[3]){
  csurrog[,,counter]<-copsurrognd(m[,,counter],targetcop=ccop,numsurrog=1)
}

#***surrogates using a Frank copula, spearman preserving
#get the Frank parameter to use
fcop<-frankCopula(4,2)
fparam<-iRho(fcop,rho(ncop))
fcop<-frankCopula(fparam,n)

#generate the surrogates
fsurrog<-array(NA,dim(m))
for (counter in 1:dim(m)[3]){
  fsurrog[,,counter]<-copsurrognd(m[,,counter],targetcop=fcop,numsurrog=1)
}

#***surrogates using a survival Clayton, spearman preserving
#take the clayton surrogates and transform them to the 
#copula space, then flip, then transform back
scsurrog<-cop_to_marg_trans((1-marg_to_cop_trans(csurrog)))

spea.surrog<-list(cla=csurrog,nor=m,fra=fsurrog,scl=scsurrog)
```

<!--do a TL analysis on the data just created-->
```{r tl_analysis, echo=F, cache=T, cache.extra=list(kend.surrog,spea.surrog,mtime("mncurvat.R"),mtime("TaylorsLawTools.R"),seed,mtime("copsurrognd.R"))}
source("copsurrognd.R")
source("TaylorsLawTools.R")

#We are only doing spatial taylor's law because we know a priori temporal
#won't give anything interesting for the surrogates we use

kend.stats.res<-list()
kend.points.res<-list()
spea.stats.res<-list()
spea.points.res<-list()
for (counter in 1:length(kend.surrog)){
 ks<-kend.surrog[[counter]]
 ss<-spea.surrog[[counter]]
 kend.stats.res[[counter]]<-matrix(NA,7,numdat) # to keep spatial TL stats
 spea.stats.res[[counter]]<-matrix(NA,7,numdat)
 kend.points.res[[counter]]<-array(NA,c(N,2,numdat)) # to keep log(mu), log(var) of spatial TL as a dataframe
 spea.points.res[[counter]]<-array(NA,c(N,2,numdat))
 for (sc in 1:numdat){
   khold<-tl_stats(ks[,,sc])
   kend.stats.res[[counter]][,sc]<-khold$s.stats #spatial TL only
   rownames(kend.stats.res[[counter]])<-names(khold$s.stats)
   kend.points.res[[counter]][,,sc]<-as.matrix(khold$s.points)

   shold<-tl_stats(ss[,,sc])
   spea.stats.res[[counter]][,sc]<-shold$s.stats #spatial TL only
   rownames(spea.stats.res[[counter]])<-names(shold$s.stats)
   spea.points.res[[counter]][,,sc]<-as.matrix(shold$s.points)
 }
}
names(kend.stats.res)<-names(kend.surrog)
names(kend.points.res)<-names(kend.surrog)
names(spea.stats.res)<-names(spea.surrog)
names(spea.points.res)<-names(spea.surrog)
```

<!--Now make some plots, these one for kendall surrogates-->
```{r taylor_plots_kendall, echo=F, results="hide"}
xlimits<-range(kend.points.res[[1]][,1,],kend.points.res[[2]][,1,],kend.points.res[[3]][,1,],kend.points.res[[4]][,1,])
ylimits<-range(kend.points.res[[1]][,2,],kend.points.res[[2]][,2,],kend.points.res[[3]][,2,],kend.points.res[[4]][,2,])

#pdf("./Results/taylorslaw_results/SpatialTL_CNFSC_kendall.pdf",height=4,width=16)
png("./Results/taylorslaw_results/SpatialTL_CNFSC_kendall.png",width=16,height=4,units="in",res=400)
op<-op<-par(mfrow=c(1,4),mar=c(4,5,1,1),mgp=c(3,1,0))
##plot TL for clayton
plot(kend.points.res[[1]][,1,],kend.points.res[[1]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
     cex.axis=1.5,cex.lab=1.5)  
points(kend.points.res[[1]][,1,1],kend.points.res[[1]][,2,1],type='p',pch=20,cex=.75,col='red')
points(kend.points.res[[1]][,1,2],kend.points.res[[1]][,2,2],type='p',pch=20,cex=.75,col='green')
mtext(paste("Clayton"),side = 3, line=-2, adj=0.1, col="black")

##plot TL for normal copula
plot(kend.points.res[[2]][,1,],kend.points.res[[2]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
     cex.axis=1.5,cex.lab=1.5)  
points(kend.points.res[[2]][,1,1],kend.points.res[[2]][,2,1],type='p',pch=20,cex=.75,col='red')
points(kend.points.res[[2]][,1,2],kend.points.res[[2]][,2,2],type='p',pch=20,cex=.75,col='green')
mtext(paste("Normal"),side = 3, line=-2, adj=0.1, col="black")

#plot TL for frank
plot(kend.points.res[[3]][,1,],kend.points.res[[3]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
     cex.axis=1.5,cex.lab=1.5)  
points(kend.points.res[[3]][,1,1],kend.points.res[[3]][,2,1],type='p',pch=20,cex=.75,col='red')
points(kend.points.res[[3]][,1,2],kend.points.res[[3]][,2,2],type='p',pch=20,cex=.75,col='green')
mtext(paste("Frank"),side = 3, line=-2, adj=0.1, col="black")

#plot TL for survival clayton
plot(kend.points.res[[4]][,1,],kend.points.res[[4]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
     cex.axis=1.5,cex.lab=1.5)  
points(kend.points.res[[4]][,1,1],kend.points.res[[4]][,2,1],type='p',pch=20,cex=.75,col='red')
points(kend.points.res[[4]][,1,2],kend.points.res[[4]][,2,2],type='p',pch=20,cex=.75,col='green')
mtext(paste("Survival Clayton"),side=3, line=-2, adj=0.1, col="black")

par(op)
dev.off()
#---------------------------------------------------
#prepare for plots coming below
s.lin<-cbind(kend.stats.res[[1]][1,],kend.stats.res[[2]][1,],kend.stats.res[[3]][1,],kend.stats.res[[4]][1,])
s.hom<-cbind(kend.stats.res[[1]][2,],kend.stats.res[[2]][2,],kend.stats.res[[3]][2,],kend.stats.res[[4]][2,])
s.rmse<-cbind(kend.stats.res[[1]][3,],kend.stats.res[[2]][3,],kend.stats.res[[3]][3,],kend.stats.res[[4]][3,])
s.int<-cbind(kend.stats.res[[1]][4,],kend.stats.res[[2]][4,],kend.stats.res[[3]][4,],kend.stats.res[[4]][4,])
s.slope<-cbind(kend.stats.res[[1]][5,],kend.stats.res[[2]][5,],kend.stats.res[[3]][5,],kend.stats.res[[4]][5,])
s.quadcoeff<-cbind(kend.stats.res[[1]][6,],kend.stats.res[[2]][6,],kend.stats.res[[3]][6,],kend.stats.res[[4]][6,])
s.mncurv<-cbind(kend.stats.res[[1]][7,],kend.stats.res[[2]][7,],kend.stats.res[[3]][7,],kend.stats.res[[4]][7,])

pdf("./Results/taylorslaw_results/SpatialTL_7stats_kendall.pdf",height=3,width=21)
op<-par(mfrow=c(1,7),mar=c(4.2,5,1,1),mgp=c(3,1,0))
#***DAN: x axis needs to be redone to show the copula names used in the surrogates

#x<-1:4
x2<-c("C", "N", "F", "SC")
#plot(x,s.lin[1,],type="l",col="grey",ylim=range(s.lin),xlab="Set of surrogates",ylab="Linearity test p-vals",
#     cex.axis=1.5,cex.lab=1.5,xaxt="n")
#points(x,s.lin[1,],col="red")
#axis(1, at = c(1:4), x2, cex.axis=1.5)
#for (counter in 2:dim(s.lin)[1]){
#  lines(x,s.lin[counter,],col="grey")
#  points(x,s.lin[counter,],col="red")
#}
boxplot(s.lin , col=rainbow(4), xlab="Copula family", ylab="Linearity test p-vals", cex.axis=1.5,cex.lab=1.5,xaxt="n")
axis(1, at = c(1:4), x2, cex.axis=1.5)

#plot s.hom distr for each set of surrogates
boxplot(s.hom , col=rainbow(4), xlab="Copula family", ylab="Homoskedasticity test p-vals", cex.axis=1.5,cex.lab=1.5,xaxt="n")
axis(1, at = c(1:4), x2, cex.axis=1.5)

#plot s.rmse distr for each set of surrogates
boxplot(s.rmse , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL rmse", cex.axis=1.5,cex.lab=1.5,xaxt="n")
axis(1, at = c(1:4), x2, cex.axis=1.5)

#plot s.int distr for each set of surrogates
boxplot(s.int , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL intercept", cex.axis=1.5,cex.lab=1.5,xaxt="n")
axis(1, at = c(1:4), x2, cex.axis=1.5)

#plot s.slope distr for each set of surrogates
boxplot(s.slope , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL slope", cex.axis=1.5,cex.lab=1.5,xaxt="n")
axis(1, at = c(1:4), x2, cex.axis=1.5)

#plot s.quadcoeff distr for each set of surrogates
boxplot(s.quadcoeff , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL quadratic coefficient", cex.axis=1.5,cex.lab=1.5,xaxt="n")
axis(1, at = c(1:4), x2, cex.axis=1.5)

#plot s.mncurv distr for each set of surrogates
boxplot(s.mncurv , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL mean curvature", cex.axis=1.5,cex.lab=1.5,xaxt="n")
axis(1, at = c(1:4), x2, cex.axis=1.5)

par(op)
dev.off()
```

<!--Now make some plots, these one for spearman surrogates-->
```{r taylor_plots_spearman, echo=F, results="hide"}
xlimits<-range(spea.points.res[[1]][,1,],spea.points.res[[2]][,1,],spea.points.res[[3]][,1,],spea.points.res[[4]][,1,])
ylimits<-range(spea.points.res[[1]][,2,],spea.points.res[[2]][,2,],spea.points.res[[3]][,2,],spea.points.res[[4]][,2,])

#pdf("./Results/taylorslaw_results/SpatialTL_CNFSC_spearman.pdf",height=4,width=16)
png("./Results/taylorslaw_results/SpatialTL_CNFSC_spearman.png",height=4,width=16,units="in",res=400)
op<-op<-par(mfrow=c(1,4),mar=c(4,5,1,1),mgp=c(3,1,0))
##plot TL for clayton
plot(spea.points.res[[1]][,1,],spea.points.res[[1]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
     cex.axis=1.5,cex.lab=1.5)  
points(spea.points.res[[1]][,1,1],spea.points.res[[1]][,2,1],type='p',pch=20,cex=.75,col='red')
points(spea.points.res[[1]][,1,2],spea.points.res[[1]][,2,2],type='p',pch=20,cex=.75,col='green')
mtext(paste("Clayton"),side = 3, line=-2, adj=0.1, col="black")

##plot TL for normal copula
plot(spea.points.res[[2]][,1,],spea.points.res[[2]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
     cex.axis=1.5,cex.lab=1.5)  
points(spea.points.res[[2]][,1,1],spea.points.res[[2]][,2,1],type='p',pch=20,cex=.75,col='red')
points(spea.points.res[[2]][,1,2],spea.points.res[[2]][,2,2],type='p',pch=20,cex=.75,col='green')
mtext(paste("Normal"),side = 3, line=-2, adj=0.1, col="black")

#plot TL for frank
plot(spea.points.res[[3]][,1,],spea.points.res[[3]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
     cex.axis=1.5,cex.lab=1.5)  
points(spea.points.res[[3]][,1,1],spea.points.res[[3]][,2,1],type='p',pch=20,cex=.75,col='red')
points(spea.points.res[[3]][,1,2],spea.points.res[[3]][,2,2],type='p',pch=20,cex=.75,col='green')
mtext(paste("Frank"),side = 3, line=-2, adj=0.1, col="black")

#plot TL for survival clayton
plot(spea.points.res[[4]][,1,],spea.points.res[[4]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
     cex.axis=1.5,cex.lab=1.5)  
points(spea.points.res[[4]][,1,1],spea.points.res[[4]][,2,1],type='p',pch=20,cex=.75,col='red')
points(spea.points.res[[4]][,1,2],spea.points.res[[4]][,2,2],type='p',pch=20,cex=.75,col='green')
mtext(paste("Survival Clayton"),side=3, line=-2, adj=0.1, col="black")

par(op)
dev.off()
#---------------------------------------------------
#prepare for plots coming below
s.lin<-cbind(spea.stats.res[[1]][1,],spea.stats.res[[2]][1,],spea.stats.res[[3]][1,],spea.stats.res[[4]][1,])
s.hom<-cbind(spea.stats.res[[1]][2,],spea.stats.res[[2]][2,],spea.stats.res[[3]][2,],spea.stats.res[[4]][2,])
s.rmse<-cbind(spea.stats.res[[1]][3,],spea.stats.res[[2]][3,],spea.stats.res[[3]][3,],spea.stats.res[[4]][3,])
s.int<-cbind(spea.stats.res[[1]][4,],spea.stats.res[[2]][4,],spea.stats.res[[3]][4,],spea.stats.res[[4]][4,])
s.slope<-cbind(spea.stats.res[[1]][5,],spea.stats.res[[2]][5,],spea.stats.res[[3]][5,],spea.stats.res[[4]][5,])
s.quadcoeff<-cbind(spea.stats.res[[1]][6,],spea.stats.res[[2]][6,],spea.stats.res[[3]][6,],spea.stats.res[[4]][6,])
s.mncurv<-cbind(spea.stats.res[[1]][7,],spea.stats.res[[2]][7,],spea.stats.res[[3]][7,],spea.stats.res[[4]][7,])

pdf("./Results/taylorslaw_results/SpatialTL_7stats_spearman.pdf",height=3,width=21)
op<-par(mfrow=c(1,7),mar=c(4.2,5,1,1),mgp=c(3,1,0))
x2<-c("C", "N", "F", "SC")

#plot s.lin distr for each set of surrogates
boxplot(s.lin , col=rainbow(4), xlab="Copula family", ylab="Linearity test p-vals", cex.axis=1.5,cex.lab=1.5,xaxt="n")
axis(1, at = c(1:4), x2, cex.axis=1.5)

#plot s.hom distr for each set of surrogates
boxplot(s.hom , col=rainbow(4), xlab="Copula family", ylab="Homoskedasticity test p-vals", cex.axis=1.5,cex.lab=1.5,xaxt="n")
axis(1, at = c(1:4), x2, cex.axis=1.5)

#plot s.rmse distr for each set of surrogates
boxplot(s.rmse , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL rmse", cex.axis=1.5,cex.lab=1.5,xaxt="n")
axis(1, at = c(1:4), x2, cex.axis=1.5)

#plot s.int distr for each set of surrogates
boxplot(s.int , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL intercept", cex.axis=1.5,cex.lab=1.5,xaxt="n")
axis(1, at = c(1:4), x2, cex.axis=1.5)

#plot s.slope distr for each set of surrogates
boxplot(s.slope , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL slope", cex.axis=1.5,cex.lab=1.5,xaxt="n")
axis(1, at = c(1:4), x2, cex.axis=1.5)

#plot s.quadcoeff distr for each set of surrogates
boxplot(s.quadcoeff , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL quadratic coefficient", cex.axis=1.5,cex.lab=1.5,xaxt="n")
axis(1, at = c(1:4), x2, cex.axis=1.5)

#plot s.mncurv distr for each set of surrogates
boxplot(s.mncurv , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL mean curvature", cex.axis=1.5,cex.lab=1.5,xaxt="n")
axis(1, at = c(1:4), x2, cex.axis=1.5)

par(op)
dev.off()
```

\begin{figure}[!h]
\textbf{ \hspace{2.8 cm} (A) \hspace{3.6 cm} (B) \hspace{3.8 cm} (C) \hspace{3.6 cm} (D)} \\
\vspace{-0.5 cm} 
\begin{center}
\includegraphics[width=18 cm]{./Results/taylorslaw_results/SpatialTL_CNFSC_spearman.png}\\
\end{center}
\textbf{ \hspace{1.8 cm} (E) \hspace{1.6 cm} (F) \hspace{1.8 cm} (G) \hspace{1.8 cm} (H) \hspace{1.8 cm} (I) \hspace{1.8 cm} (J) \hspace{1.8 cm} (K)}\\
\vspace{-0.5 cm}
\begin{center}
\includegraphics[width=18 cm]{./Results/taylorslaw_results/SpatialTL_7stats_spearman.pdf}
\caption[Spatial Taylor's law results]{Spatial Taylor's law results. Log(spatial variance) vs. 
log(spatial mean) relationships for $1000$ simulations over $50$ time steps for $25$ locations 
using Clayton (A), normal (B), Frank (C), and survival Clayton (D) copula structures between
locations. Results from all $1000$ 
simulations were plotted on the same axes for each panel, but two example simulations are 
shown in red and green to help assess to what extent variation on the plots was between 
simulations or within simulations. For each simulation, we tested the linearity 
(E) and homoskedasticity (F) of the log(variance) vs. log(mean) plot for that simulation, 
and quantified the root mean squared error of points from the linear regression line (G).
We also calculated the linear regression intercept (H) and slope (I), the quadratic
term of a quadratic regression through the points on the log(variance) vs. log(mean) plot
(J), and the curvature of that quadratic regression (K). Distributions of values 
across all 1000 simulations are displayed. Spearman-preserving surrogates were used, 
though results using Kendall-preserving surrogates were very similar. See 
Appendix \ref{Consequence_TL_simulation} for 
simulation and plotting details.\label{fig_spTL}}
\end{center}
\end{figure}

```{r spatialTL_on_realdata_aphid_count, results="hide", echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("good_loclist.R"),mtime("TaylorsLawTools.R"),mtime("mncurvat.R"),mtime("ncsurrog.R"))}

set.seed(seed)
source("ncsurrog.R")
source("TaylorsLawTools.R")
source("good_loclist.R")

TL_on_realdata<-function(data,numsurrog=1000,sp,data_pt_thrs,ploton,tag,labelon=T){

  loclist<-good_loclist(d_allsp=data,sp=sp,data_pt_thrs=data_pt_thrs)
  data<-data[[sp]]
  df<-c()
  for(i in loclist){
    tem<-data[[i]]$Dat
    df<-cbind(df,tem)
  }
  colnames(df)<-paste("loc",loclist,sep="")
  df<-na.omit(df)

  mat_real<-as.matrix(df)
  check_rep<-function(y){all( abs(y - mean(y)) < 1e-8 )} #This fn checks for repeatation of same values in a vector
  call_check_rep<-apply(mat_real,1,check_rep) # checks the same for mat_real matrix along all row

  if(all(call_check_rep==F)==T){

    realTL_stats<-tl_stats(m = mat_real)
    real_spatialTLslope<-unname(realTL_stats$s.stats[5])
    real_sTL_quadcoeff<-unname(realTL_stats$s.stats[6])
    real_sTL_mncurv<-unname(realTL_stats$s.stats[7])

    #generate the normal surrogates with spearman correlation preserving
    array_surrog<-ncsurrog(m=mat_real,corpres="spearman",numsurrog=numsurrog)

    bad_id<-c()
    for(i in 1:numsurrog){
      m<-array_surrog[,,i]
      mean_m<-apply(X=m, MARGIN = 1, FUN = mean)
      var_m<-apply(X=m, MARGIN = 1, FUN = var)
      if(any(mean_m==0) | any(var_m==0)){
        bad_id<-c(bad_id,i)
      }
    }

    m<-array_surrog
    if(length(bad_id)!=0){
      m<-array_surrog[,,-bad_id]
    }
    ts<-apply(m,3,tl_stats)

    good_surrog_slope<-c()
    good_surrog_quadcoeff<-c()
    good_surrog_mncurv<-c()
    for(i in c(1:length(ts))){
      good_surrog_slope<-c(good_surrog_slope,unname(ts[[i]]$s.stats[5]))
      good_surrog_quadcoeff<-c(good_surrog_quadcoeff,unname(ts[[i]]$s.stats[6]))
      good_surrog_mncurv<-c(good_surrog_mncurv,unname(ts[[i]]$s.stats[7]))
    }

    numsurrog_good<- numsurrog - length(bad_id)

    p_left<-sum(good_surrog_slope < real_spatialTLslope)/numsurrog_good
    p_right<-sum(good_surrog_slope > real_spatialTLslope)/numsurrog_good

    p_left_quadcoeff<-sum(good_surrog_quadcoeff < real_sTL_quadcoeff)/numsurrog_good
    p_right_quadcoeff<-sum(good_surrog_quadcoeff > real_sTL_quadcoeff)/numsurrog_good

    p_left_mncurv<-sum(good_surrog_mncurv < real_sTL_mncurv)/numsurrog_good
    p_right_mncurv<-sum(good_surrog_mncurv > real_sTL_mncurv)/numsurrog_good

    if(ploton==T & tag=="slope"){
      hist(good_surrog_slope,breaks=100,main="",xlab="surrogate_TL_slope",cex.axis=1.5,cex.lab=1.5)
      points(x=real_spatialTLslope,y=0,col="red",pch=15)
      if(labelon==T){
        mtext(paste0("sp= ",sp,", p_left= ",round(p_left,3),", p_right=",round(p_right,3),sep=""),
              side = 3, line=-2, adj=0, col="blue")
      }
    }

    if(ploton==T & tag=="quadcoeff"){
      hist(good_surrog_quadcoeff,breaks=100,main="",xlab="surrogate_TL_quadcoeff",cex.axis=1.5,cex.lab=1.5)
      points(x=real_sTL_quadcoeff,y=0,col="red",pch=15)
      if(labelon==T){
        mtext(paste0("sp= ",sp,", p_left= ",round(p_left_quadcoeff,3),", p_right=",round(p_right_quadcoeff,3),sep=""),
              side = 3, line=-2, adj=0, col="blue")
      }
    }

    if(ploton==T & tag=="mncurv"){
      hist(good_surrog_mncurv,breaks=100,main="",xlab="surrogate_TL_meanCurvature",cex.axis=1.5,cex.lab=1.5)
      points(x=real_sTL_mncurv,y=0,col="red",pch=15)
      if(labelon==T){
        mtext(paste0("sp= ",sp,", p_left= ",round(p_left_mncurv,3),", p_right=",round(p_right_mncurv,3),sep=""),
              side = 3, line=-2, adj=0, col="blue")
      }
    }

  }else{
    real_spatialTLslope<-NA
    p_left<-NA
    p_right<-NA
    numsurrog_good<-NA
    real_sTL_quadcoeff<-NA
    p_left_quadcoeff<-NA
    p_right_quadcoeff<-NA
    real_sTL_mncurv<-NA
    p_left_mncurv<-NA
    p_right_mncurv<-NA
  }
  return(list(realTL_stats=realTL_stats,
              real_spatialTLslope=real_spatialTLslope,
              p_left=p_left,
              p_right=p_right,
              numsurrog_good=numsurrog_good,
              real_sTL_quadcoeff=real_sTL_quadcoeff,
              p_left_quadcoeff=p_left_quadcoeff,
              p_right_quadcoeff=p_right_quadcoeff,
              real_sTL_mncurv=real_sTL_mncurv,
              p_left_mncurv=p_left_mncurv,
              p_right_mncurv=p_right_mncurv,
              surrogTL_stats=ts,
              good_surrog_mncurv=good_surrog_mncurv))
}

#call on real data
# For all sp
splist<-c(1:20)
tl_aphid_count<-as.data.frame(matrix(NA,nrow=length(splist),ncol=10))
colnames(tl_aphid_count)<-c("real_sTL_slope","p_left","p_right","numsurrog_good",
                            "real_sTL_quadcoeff","p_left_quadcoeff","p_right_quadcoeff",
                            "real_sTL_mncurv","p_left_mncurv","p_right_mncurv")
rownames(tl_aphid_count)<-paste("sp",splist,sep="")
#pdf("aphid_count_allsp_spTLslope.pdf",height=20,width=20)
#pdf("./Results/taylorslaw_results/aphid_count_allsp_sTL_quadcoeff.pdf",height=20,width=20)
pdf("./Results/taylorslaw_results/aphid_count_allsp_sTL_mncurv.pdf",height=20,width=20)
op<-par(mfrow=c(4,5),mar=c(4.2,5,1,1),mgp=c(3,1,0))
for(i in splist){
  #cat("sp=",i,"\n")
  #ans<-TL_on_realdata(data=data_aphid_count,numsurrog=10000,sp=i,data_pt_thrs = 30,ploton=T,tag="slope",labelon=T)
  #ans<-TL_on_realdata(data=data_aphid_count,numsurrog=10000,sp=i,data_pt_thrs = 30,ploton=T,tag="quadcoeff",labelon=T)
  ans<-TL_on_realdata(data=data_aphid_count,numsurrog=10000,sp=i,data_pt_thrs = 30,ploton=T,tag="mncurv",labelon=T)
  if(i==6){
    saveRDS(ans,"./Results/taylorslaw_results/tl_aphid_count_sp6.RDS") #saving sp=6 results
  }
  tl_aphid_count$real_sTL_slope[i]<-ans$real_spatialTLslope
  tl_aphid_count$p_left[i]<-ans$p_left
  tl_aphid_count$p_right[i]<-ans$p_right
  tl_aphid_count$numsurrog_good[i]<-ans$numsurrog_good
  tl_aphid_count$real_sTL_quadcoeff[i]<-ans$real_sTL_quadcoeff
  tl_aphid_count$p_left_quadcoeff[i]<-ans$p_left_quadcoeff
  tl_aphid_count$p_right_quadcoeff[i]<-ans$p_right_quadcoeff
  tl_aphid_count$real_sTL_mncurv[i]<-ans$real_sTL_mncurv
  tl_aphid_count$p_left_mncurv[i]<-ans$p_left_mncurv
  tl_aphid_count$p_right_mncurv[i]<-ans$p_right_mncurv
}
par(op)
dev.off()
saveRDS(tl_aphid_count,"./Results/taylorslaw_results/tl_aphid_count.RDS")
```

```{r spatialTL_on_realdata_aphid_count_sp6, results="hide", echo=F}
#Now for sp=6

ans<-readRDS("./Results/taylorslaw_results/tl_aphid_count_sp6.RDS")

pdf("./Results/taylorslaw_results/aphid_count_sp6_sTL_mncurv.pdf",height=5,width=5)
op<-par(mar=c(4,5,1,1),mgp=c(3,1,0))
hist(ans$good_surrog_mncurv,breaks=100,main="",xlab="Mean curvature",cex.axis=2,cex.lab=2,cex=2,col="grey",border=F)
points(x=ans$real_sTL_mncurv,y=0,col="black",pch=15,cex=2)
mtext(paste0("p= ",ans$p_left_mncurv),side = 3, line=-2, adj=0.5, col="black",cex=2)
par(op)
dev.off()

pdf("./Results/taylorslaw_results/aphid_count_sp6_logv_vs_logm.pdf",height=5,width=10)
op<-par(mfrow=c(1,2),mar=c(4,5,1,1),mgp=c(3,1,0))
plot(ans$realTL_stats$s.points,xlab=expression(log[10]("spatial mean")),
     ylab=expression(log[10]("spatial variance")),col="grey",pch=20,cex=1.5,
     cex.axis=2,cex.lab=2)
mtext(paste0("mean curvature= ",round(unname(ans$realTL_stats$s.stats[7]),3)),side = 3, line=-2, adj=0, col="black",cex=2)

target_mncurv<-0.1
id<-which(abs(ans$good_surrog_mncurv-target_mncurv)<1e-4)[1]
ans_s<-ans$surrogTL_stats
plot(ans_s[[id]]$s.points,xlab=expression(log[10]("spatial mean")),
     ylab=expression(log[10]("spatial variance")),col="grey",pch=20,cex=1.5,
     cex.axis=2,cex.lab=2)
mtext(paste0("mean curvature= ",round(ans$good_surrog_mncurv[id],3)),side = 3, line=-2, adj=0, col="black",cex=2)

par(op)
dev.off()
```

```{r sim_ext_risk, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}

source("mydispersal.R")
set.seed(seed)

model<-"lcohen"
numsims<-10000
numsteps<-25
params<-0
ext_thrs<-1
scl<-1
p0<-50

sl5<-varying_d(numsims=numsims,numsteps =numsteps,numlocs=5,p0=p0,params=params,
                 ext_thrs=ext_thrs,scl=scl,model=model,disp_everywhere=F,ploton=F)
saveRDS(sl5,"./Results/ext_risk_copula/ext_risk_local_lcohen_numlocs5.RDS")

sl25<-varying_d(numsims=numsims,numsteps =numsteps,numlocs=25,p0=p0,params=params,
                 ext_thrs=ext_thrs,scl=scl,model=model,disp_everywhere=F,ploton=F)
saveRDS(sl25,"./Results/ext_risk_copula/ext_risk_local_lcohen_numlocs25.RDS")

sg5<-varying_d(numsims=numsims,numsteps =numsteps,numlocs=5,p0=p0,params=params,
                 ext_thrs=ext_thrs,scl=scl,model=model,disp_everywhere=T,ploton=F)
saveRDS(sg5,"./Results/ext_risk_copula/ext_risk_global_lcohen_numlocs5.RDS")

sg25<-varying_d(numsims=numsims,numsteps =numsteps,numlocs=25,p0=p0,params=params,
                 ext_thrs=ext_thrs,scl=scl,model=model,disp_everywhere=T,ploton=F)
saveRDS(sg25,"./Results/ext_risk_copula/ext_risk_global_lcohen_numlocs25.RDS")
```

```{r fig_ext_risk, echo=F,  results="hide", cache=T, cache.extra=list(seed,sl5,sl25,sg5,sg25)}
sl5<-readRDS("./Results/ext_risk_copula/ext_risk_local_lcohen_numlocs5.RDS")
sl25<-readRDS("./Results/ext_risk_copula/ext_risk_local_lcohen_numlocs25.RDS")
sg5<-readRDS("./Results/ext_risk_copula/ext_risk_global_lcohen_numlocs5.RDS")
sg25<-readRDS("./Results/ext_risk_copula/ext_risk_global_lcohen_numlocs25.RDS")

# plot for local dispersal in D matrix 
pdf("./Results/ext_risk_copula/lcohen_ext_risk_local_disp.pdf",heigh=5,width=10)
op<-par(mfrow=c(1,2),mar=c(5,5,1.8,1),mgp=c(3,1,0))
plot(sl5$d_seq[1:6],sl5$risk_left[1:6],xlim=c(0,0.5),ylim=c(0,1),type="b",col="black",
     xlab="d",ylab="Extinction risk",
     cex.lab=2,cex.axis=1.8,cex=1.5,pch=16,lwd=1.5)
lines(sl5$d_seq[1:6],sl5$risk_right[1:6],type="b",lwd=1.5,cex=1.5,pch=16,col="grey")
legend(x=0.05,y=0.9, title=paste0("r = 0 , N = 5"), c("Left tail-dep. in noise","Right tail-dep. in noise"), 
       pch=c(16,16), lty=c(1,1), cex=1.4, pt.cex=c(1.5,1.5), col=c("black","grey"),
       horiz=F, bty='n')

plot(sl25$d_seq[1:6],sl25$risk_left[1:6],xlim=c(0,0.5),ylim=c(0,1),type="b",col="black",
     xlab="d",ylab="Extinction risk",
     cex.lab=2,cex.axis=1.8,cex=1.5,pch=16,lwd=1.5)
lines(sl25$d_seq[1:6],sl25$risk_right[1:6],type="b",lwd=1.5,cex=1.5,pch=16,col="grey")
legend(x=0.05,y=0.9, title=paste0("r = 0 , N = 25"), c("Left tail-dep. in noise","Right tail-dep. in noise"), 
       pch=c(16,16), lty=c(1,1), cex=1.4, pt.cex=c(1.5,1.5), col=c("black","grey"),
       horiz=F, bty='n')

par(op)
dev.off()

# plot for global dispersal in D matrix 
pdf("./Results/ext_risk_copula/lcohen_ext_risk_global_disp.pdf",heigh=5,width=10)
op<-par(mfrow=c(1,2),mar=c(5,5,1.8,1),mgp=c(3,1,0))
plot(sg5$d_seq[1:6],sg5$risk_left[1:6],xlim=c(0,0.5),ylim=c(0,1),type="b",col="black",
     xlab="d",ylab="Extinction risk",
     cex.lab=2,cex.axis=1.8,cex=1.5,pch=16,lwd=1.5)
lines(sg5$d_seq[1:6],sg5$risk_right[1:6],type="b",lwd=1.5,cex=1.5,pch=16,col="grey")
legend(x=0.05,y=0.9, title=paste0("r = 0 , N = 5"), c("Left tail-dep. in noise","Right tail-dep. in noise"), 
       pch=c(16,16), lty=c(1,1), cex=1.4, pt.cex=c(1.5,1.5), col=c("black","grey"),
       horiz=F, bty='n')

plot(sg25$d_seq[1:6],sg25$risk_left[1:6],xlim=c(0,0.5),ylim=c(0,1),type="b",col="black",
     xlab="d",ylab="Extinction risk",
     cex.lab=2,cex.axis=1.8,cex=1.5,pch=16,lwd=1.5)
lines(sg25$d_seq[1:6],sg25$risk_right[1:6],type="b",lwd=1.5,cex=1.5,pch=16,col="grey")
legend(x=0.05,y=0.9, title=paste0("r = 0 , N = 25"), c("Left tail-dep. in noise","Right tail-dep. in noise"), 
       pch=c(16,16), lty=c(1,1), cex=1.4, pt.cex=c(1.5,1.5), col=c("black","grey"),
       horiz=F, bty='n')
par(op)
dev.off()
```

\begin{figure}[!h]
\textbf{ \hspace{7.3 cm} (A) \hspace{4 cm} (B)} \\
\vspace{-0.8 cm}
\begin{center}
\includegraphics[width=10 cm]{./Results/ext_risk_copula/lcohen_ext_risk_global_disp.pdf}
\caption{Analogous to Fig. \ref{MT-fig_extrisk_lcohen_local} in the main text, but with global dispersal.\label{fig_extrisk_lcohen_global}}
\end{center}
\end{figure}

# Surrogate testing with multivariate datasets\label{Surrog_multi}

For multivariate datasets, surrogates were produced as follows:

- Given data $x_i(t)$ for $i=1,\ldots,N$ and $t=1,\ldots,T$, let $\tau_{ij}$ be the Spearman (or Kendall) 
correlation of $x_i(t)$ and $x_j(t)$.
- Calculate the covariance matrix of a multivariate normal distribution with standard-normal
marginals and with pairwise Spearman (or Kendall) correlations equal to the $\tau_{ij}$. This is possible.
using the `iRho` function (or the `iTau` function) of the `copula` package
- Generate data $a_i(t)$ for $i=1,\ldots,N$ and $t=1,\ldots,T$ by taking $T$ independent draws
from the multivariate normal distribution of the previous step.
- For any $i,t$ for which the datum $x_i(t)$ was missing, replace $a_i(t)$ by `NA`.
- For each $k$ and $i$ for which $a_i(t)$ is not `NA`, replace the 
$k$th-smallest (non-`NA`) element of the time series 
$a_i(t)$, $t=1,\ldots,T$ by the $k$th smallest (non-`NA`) 
element of $x_i(t)$, $t=1,\ldots,T$.
Call the result $b_i(t)$. This is a surrogate dataset.

\noindent Our code for this algorithm is `ncsurrog` in the `BIVAN` repository 
(\url{https://github.com/sghosh89/BIVAN}).
The code for the Kendall version of the
algorithm throws an error if there are ties in any of the time series. This is
because the Kendall correlation in the very commonly used `cor` function in R 
is the Kendall tau-b coefficient in the event of ties, and index which 
differs in its definition from the commonly used tau-a that works when there are
no ties; and `tau` and `iTau` are intended for continuous copulas, which cannot 
produce data with ties. Our code also throws an error in the event that
the matrix generated in the second step is not positive semidefinite. Because
ties are common in our data, we only used the Spearman case in analyses using `ncsurrog`,
though, for expansibility, the `ncsurrog` code itself is written to handle 
the Kendall case as well when there are no ties.

Because, for each $i$, the final time series 
$b_i(t)$ is a permuted version of $x_i(t)$ (and with `NA`s for the same $t$), 
the marginal distributions of the surrogate dataset (i.e., the distributions of 
values for each sampling location, $i$) are exactly the same as for the original data.
Because ranks are preserved in passing from $a_i(t)$ to $b_i(t)$, pairwise Spearman 
(or Kendall) 
correlations for the $b_i(t)$ will be the same as those of the $a_i(t)$. 
Discrepancies between the $\tau_{ij}$ and the pairwise Spearman (or Kendall) 
correlations of the
$b_i(t)$ will therefore come about only from the presence of NAs and from sampling variation. 
These discrepancies should be minor if the 
numbers of NAs are not too large and $T$ is not too small.

# Extinction risk noise and dispersal matrix \label{ext_risk_model}

<!--How noise was generated ? code : ExtremeTailDep.R-->
\noindent Noise was generated by first producing a 25 by $N$ matrix, $m$, of 
independent standard-normal random draws. Then, each row of $m$ 
was identified for a first set of modifications randomly 
and independently with a $50\%$ chance; 
for those rows identified, every element of the row was replaced with the 
absolute value of the first element of the row (including the first element itself).
For other rows, every element in the row was replaced by the negative of its
own absolute value. This produces right-tail dependent noise, normally distributed
for each location. If left-tail dependent noise is desired, take the negative of the
entire matrix.

<!--How dispersal matrix was generated ? code : mydispersal.R-->
For local dispersal, the $N \times N$ dispersal matrix $D$ was generated by assuming, 
for simplicity, that the $N$ habitat patches
were arranged, evenly spaced, in a line. A fraction $d$ of each population dispersed during
any given time step, equally distributed to the two or one nearest neighbors. 
For global dispersal, the fraction $d$ that dispersed from each patch was divided up equally
among all other patches.

# Effects of non-N copula structure on Taylor's law \label{Consequence_TL_simulation}

\noindent Pursuing ideas that sprang from earlier theoretical work [@Reuman2017], we 
tested whether the copula structure of the dependences between population variables 
measured in different locations influences the spatial version of Taylor's
law. Taylor's law is an extremely commonly observed and widely studied 
[@Taylor1961; @Taylor1988; @Cohen2013; @Xu2016] empirical pattern that relates the 
variances of groups, $g$, of population measurements to the means of the 
groups via a power law, $v_g = a \times m_g^b$, or equivalently, 
$\log(v_g)=\log(a)+b \times \log(m_g)$. Here $b$ is called 
the Taylor's law *exponent* or 
*slope*, and we refer to $\log(a)$ as the Taylor's law intercept. 
There are several versions of Taylor's law. For spatial Taylor's law,
given population density or abundance data $x_i(t)$ measured in locations
$i=1,\ldots,N$ and times $t=1,\ldots,T$, a group $g$ consists of 
all the measurements $x_i(t)$, $i=1,\ldots,N$ made in different locations at the same time.
So means $m_g$ and variances $v_g$ are computed across space. We view the assumptions
of Taylor's law to be validated for a dataset if the 
$\log(v_g)$ versus $\log(m_g)$ scatter plot 
for the dataset shows a linear, homoskedastic pattern.
Taylor's law has been 
verified empirically for hundreds of taxa and has been applied in numerous fields
including fisheries management, estimation of species persistence times, and
agriculture [@Cohen2015; @Reuman2017].

We examined using simulations whether copula structure influences spatial Taylor's law, as 
follows. First, an $n \times n$ covariance matrix $\Sigma$ was created
with unit diagonal entries 
and off-diagonal entries $\rho=0.7$; $n=25$ was the number of locations at which populations
were sampled. Then $N \times 1000$ random vectors were generated from a multivariate normal 
distribution with mean vector $(0,\ldots,0)$ and covariance matrix $\Sigma$. These 
were organized into an $N \times n \times 1000$ array representing $1000$ replicate 
simulations of (temporally independent, in this example) 
spatio-temporal population dynamics measured in 
$n$ locations over $N$ time steps. We used $N=50$.
The cumulative distribution function
(cdf) of the standard normal distribution was applied to these numbers, followed by the 
inverse of the cdf of a gamma distribution with shape and scale parameters $7.5$ and $1$. 
These transformations made gamma distributed the marginal distributions of populations in 
each sampling location for each simulation. 
A skewed population marginal distribution is an 
important factor in Taylor's law [@Cohen2015; @Reuman2017]. We refer henceforth to the 
resulting $N \times n \times 1000$ array as our *N-copula population array* 
for this simulation exercise;
copula structure of dependences between populations in different locations
was normal (N) for this array.

We then generated an series of arrays that were the same dimensions ($N \times n \times 1000$)
as the N-copula population array, and that were statistically similar to 
the N-copula population array in all ways except the new
arrays were constructed to have C (respectively, F, SC) copula structure for 
the dependences between 
populations in different sampling locations within a simulation. We 
refer to the new arrays as our
*C-copula*, *F-copula*, and *SC-copula population arrays*, respectively. The C-copula array was
constructed as follows; the others were constructed analogously. First, we constructed an
$n$-dimensional C copula with parameter chosen so that the 
Spearman correlations between each pair of the $n$ components were the same as those
of the N distribution constructed above. This was done using the `rho` and `iRho` functions 
in the `copula` package in R. Because the C copula is a one-paremeter family, this 
step relies on the fact that all pairs of locations were 
equally correlated with each other in the original multivariate normal distribution
constructed above. We then randomized each simulation in 
the N-copula population array in a special way 
(see Appendix \ref{sect:copsurrognd}) so that the resulting 
surrogate/randomized dataset had the copula structure
of the C copula just constructed. The randomization procedure worked 
by permuting time series of values in the N-copula array, separately for each 
sampling location and simulation. So population marginal distrbutions for each location
and simulation were *exactly* the same as those in the N-copula population array. 
Permutations were done in such a way
that resulting pairwise Spearman correlations were the same as the C-copula 
constructed above, and hence were the same as the N-copula array, 
except for sampling variation. Thus, to summarize using mathematical notation,
if $m^{(\text{N})}$ is our N-copula population array and $m^{(\text{C})}$ is our 
C-copula population array, then,
for any given $i$ and $j$, the sets $\{m^{(\text{N})}_{tij}: t=1,\ldots,N\}$ and 
$\{m^{(\text{C})}_{tij}: t=1,\ldots,N\}$ are exactly the same (when considered 
as unordered sets); 
and for any given $i_1$, $i_2$ and $j$, the Spearman correlations $\cor_t(m^{(\text{C})}_{ti_1j},m^{(\text{C})}_{ti_2j})$
and $\cor_t(m^{(\text{N})}_{ti_1j},m^{(\text{N})}_{ti_2j})$ are the same, to within 
sampling variation. The whole exercise was repeated preserving Kendall instead of 
Spearman correlations.

A variety of Taylor's law statistics were computed for each simulation $m^{(X)}_{tij}$ 
($i=1,\ldots,n$, $t=1,\ldots,N$) for each copula structure $X$ used. First, the $N=50$ 
spatial means,
$m_t$, and variances, $v_t$, were computed, and $\log(v_t)$ vs. $\log(m_t)$ plots 
were considered. As a shorthand, denote $y=\log(v_t)$ and $x=\log(m_t)$ for a 
single simulation. Linearity of the $y$ vs. $x$ relationship was 
tested for each simulation by comparing the linear regression
through these $50$ points to a quadratic alternative via an $F$-test, 
producing a $p$-value result. If 
these $p$-values were uniformly distributed across the unit interval for the 1000 replicate 
simulations, it supported the linearity assumption of Taylor's law; whereas if 
they were clustered toward smaller values
the test tended to reject that assumption. We also tested, for each simulation, 
the assumption of Taylor's 
law that the $y$ vs. $x$ plot was homoskedastic: we regressed the absolute residuals of the 
linear regression of $y$ versus $x$ against the predictions of that regression. 
A significant $p$-value
result of this test indicates heteroskedasticity. For each simulation we also computed the
root mean squared error of $y$ vs. $x$ data from the linear regression
of $y$ versus $x$, as well as the intercept and slope of the linear regression.
Finally we recorded the quadratic coefficient of the regression of $y$ against
$x$ and $x^2$, and the mean curvature of the quadratic regression equation itself
across the values $x$. The distrubutions of all these statistics across 1000 replicate 
simulations were compared for the C, N, F, and SC simulations to determine if copula structure 
influences Taylor's law.

Taylor's law was strongly influenced, and was often even invalidated in its assumptions of 
linearity and homoskedasticity, by non-N copula structure. For N copula structure 
(i.e., for simulations in our N-copula population array), $p$-values for the linearty
and homoskedasticity tests were roughly uniformly distributed across replicate 
simulations and Taylor's law appeared visually to be a reasonable approximation
of the $\log(v_t)$ versus $\log(m_t)$ relationship (Fig. \ref{fig_spTL}B,E,F). 
Furthermore, quadratic coefficients
and curvature values were close to $0$ (Fig. \ref{fig_spTL}J,K), and root mean squared
errors from the linear regression were relatively small (Fig. \ref{fig_spTL}G).
But linearity or homoskedasticity were violated more frequently for non-N copula structure
(Fig. \ref{fig_spTL}A,C,D,E,F); quadratic coefficients and curvatures were frequently 
non-zero (Fig. \ref{fig_spTL}J,K); and root mean squared
errors from the linear regression were much higher (Fig. \ref{fig_spTL}G). Slopes 
and intercepts of the linear regression were also strongly affected by copula
structure (Fig. \ref{fig_spTL}H,I), though some of the effect here was because
linear regressions do not always adequately represent the $\log(v_t)$ versus $\log(m_t)$
relationship when copula structure was not N.

# Another tool for producing surrogates for multivariate datasets\label{sect:copsurrognd}

\noindent Suppose one has data $x_i(t)$ for $i=1,\ldots,N$ and $t=1,\ldots,T$, with
no missing values and no ties, and an
$N$-dimensional copula $C$. The following algorithm produces a surrogate dataset
$y_i(t)$ ($i=1,\ldots,N$, $t=1,\ldots,T$) such that: 1) the Kendall and Spearman 
correlations $\cor_t(y_i(t),y_j(t))$ are the same,
up to sampling variation, as the Kendall and Spearman correlations, respectively, of the corresponding
components of $C$, for all $i\neq j$; and 2) for any $i$, the values of the unordered set 
$\{y_i(t):t=1,\ldots,T\}$ are the same as the values of the unordered set
$\{x_i(t):t=1,\ldots,T\}$.

- Generate data $a_i(t)$ for $i=1,\ldots,N$ and $t=1,\ldots,T$ by taking $T$ independent 
draws from $C$.
- For each $k$ and $i$, replace the $k$th-smallest element of the time series $a_i(t)$, 
$t=1,\ldots,T$ by the $k$th smallest element of $x_i(t)$, $t=1,\ldots,T$. Call the 
result $y_i(t)$. This is the surrogate dataset.

\clearpage

#  References

<!--
<div id="refs"></div>
-->