### 2019-10-01 Paper.Rmd
table and figure simplified as suggested by DCR

---
title: "Copulas and their potential for ecology"
author: "Shyamolina Ghosh, Lawrence W. Sheppard, Mark T. Holder, Terrance E. Loecke, Philip C. Reid, James D. Bever, Daniel C. Reuman"
date: ""
fontsize: 12pt
geometry: "left=2.54cm,right=2.54cm,top=2.54cm,bottom=2.54cm"

output: 
  pdf_document:
    number_sections: no
    keep_tex: yes
    fig_caption: yes
    includes:
      in_header: head_maintext.sty

tables: True
link-citations: True
urlcolor : blue
indent : True

csl: ecology-letters.csl
bibliography: REF_ALL.bib
---

```{r setup_Paper, echo=F}
knitr::opts_chunk$set(echo = TRUE, fig.pos = "H")
options(scipen = 1, digits = 3) #This option round all numbers appeared in the inline r code upto specified digit
#seed<-101
#source("mtime.R") #A function needed for caching
```

\raggedright
\setlength\parindent{2em}
\setlength{\parskip}{6pt}

\noindent \emph{Affiliations:}

\noindent Ghosh, Sheppard, Bever: Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66045, USA

\noindent Holder: Department of Ecology and Evolutionary Biology and Biodiversity Institute, University of Kansas, Lawrence, KS, 66045, USA

\noindent Loecke: Environmental Studies Program and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66047, USA

\noindent Reid: School of Biological & Marine Sciences, University of Plymouth, Drake Circus, Plymouth PL4 8AA, UK; Continuous Plankton Recorder Survey, The Marine Biological Association, The Laboratory, Citadel Hill, Plymouth PL1 2PB, UK

\noindent Reuman: Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66047, USA; Laboratory of Populations, Rockefeller University, 1230 York Ave, New York, NY, 10065, USA

<!--\noindent \emph{Possible alternative titles:}

Complex and informative dependencies propagate throughout ecology and can be revealed using copulas

Copulas reveal complex and informative dependencies propagating throughout ecology

Copulas and their importance for ecology

Copulas and their importance in ecology

Copulas and their importance throughout ecology

Copulas and their potential throughout ecology

Copulas and their importance

Copulas in ecology

Copulas

The importance of copulas

The importance of copulas in ecology-->

\noindent \emph{Correspondence:} Daniel Reuman, 2101 Constant Ave, Lawrence, KS, 66047, reuman@ku.edu, 626 560 7084.

\noindent \emph{The number of words in the abstract:} 266

\newpage

# Abstract
<!--***DANr: I have not edited this abstract at all since Ecol Monogr, it needs to be redone-->
All branches of ecology study relationships among environmental and biological variables. However, 
standard approaches to studying such relationships, based on correlation and regression, 
provide only a small slice of the complex information contained in the relationships. Other 
statistical approaches exist that provide a complete description of relationships between variables, 
based on the concept of the *copula*; they are applied in finance, neuroscience and other fields, 
but not in ecology. We here explore the concepts that underpin copulas and examine the potential 
for those concepts to improve our understanding of ecology. We find that informative copula 
structure in dependencies between variables is exceedingly common across all the environmental, 
species-trait, phenological, population, community, and ecosystem functioning datasets we 
considered. Many datasets exhibited asymmetric tail dependence, whereby two variables are more 
strongly related in their left compared to right tails, or *vice versa*. We describe mechanisms by which 
observed copula structure and tail dependence can arise in ecological data, including a Moran-like 
effect whereby dependence structures between environmental variables are inherited by ecological 
variables; and asymmetric or nonlinear influences of environments on ecological variables, such as 
under Liebig's law of the minimum. Copula structure and tail 
dependence can also help reveal causal relationships between variables. We also describe consequences of copula structure for ecological phenomena, including impacts 
on extinction risk and the stability through time of ecosystem services. By documenting the 
importance of a complete description of dependence, advancing conceptual frameworks, and 
demonstrating a powerful approach, we aim to elevate copula reasoning to widespread use in ecology, benefiting the discipline.

\noindent \emph{Keywords:} dependence, regression, correlation, copula, population, community, ecosystem functioning

\newpage

# Introduction\label{Intro}

All branches of ecology 
study relationships among environmental and biological 
variables. However, commonly used correlation and regression 
approaches to studying 
such relationships are limited, and 
provide only a partial description of a relationship between variables. 
For instance, datasets showing markedly different 
relationships may have the same correlation coefficient
(Fig. \ref{fig_cop_pedag1}A-B). The variables of Fig. \ref{fig_cop_pedag1}A 
(respectively, Fig. \ref{fig_cop_pedag1}B) are principally 
related in the left (respectively, right) portions of their distributions, 
an asymmetric pattern of dependence that can
have ecological significance, as discussed below, but that is not captured by 
correlation. Although correlation is only one way to 
study relationships between variables, other common metrics also 
provide only partial information.

\begin{figure}[!h]
\begin{center}
\includegraphics[width=3.5in]{./Results/PedagogFig1.pdf}
\caption{(A-D) Bivariate datasets showing diverse relationships between 
variables; some of these datasets nevertheless have the same Pearson (P)
or Spearman (S) correlation coefficients. C and D show normalized ranks of A, B, 
respectively. (E,F) Reflections
of C, D, respectively, about a vertical axis. Subscripts h and v 
in the axis labels stand for "horizontal" and "vertical".\label{fig_cop_pedag1}}
\end{center}
\end{figure}

Well-developed approaches do exist, however, and are 
applied widely in other fields, 
that provide a description of the dependence 
between two or more variables that can be mathematically proven 
to be complete [@nelsen2006_copula; @joe2014_dependence; @MaiScherer2017]. 
The approaches are based on 
the concept of the *copula*; copula approaches separate the information of a bivariate 
random variable into two distinct parts: the information in the 
marginal distributions (which says nothing about the dependence between the 
variables), and the remaining information (which is solely about
dependence). We here introduce some simple concepts, based on taking ranks,
that relate to copulas and give a conceptual 
flavor of copula ideas to be introduced 
more formally in the section on Background below. Given 
a sample $(x_i,y_i)$, $i=1,2,\ldots,n$ (e.g., 
Fig. \ref{fig_cop_pedag1}A-B), information about the structure of the dependence
between $x$ and $y$ can be separated from information contained 
in the marginal distributions by 
considering the bivariate plot of $u_i$ versus $v_i$, where
$u_i$ is the rank of $x_i$ in the set $x_j$ ($j=1,\ldots,n$), divided by
$n+1$; and $v_i$ is defined in the same way but using the $y_i$. The rank
of the smallest element in a set is $1$. The $u_i$ and 
$v_i$ are called *normalized ranks* of the $x_i$ and $y_i$; they relate to 
the empirical cummulative distribution function (cdf) of the $x_i$ and $y_i$,
respectively. See, for instance,
Fig. \ref{fig_cop_pedag1}C-D, which show the normalized ranks of 
A-B, respectively. Ranking makes the marginal distributions of 
the component datasets uniform, isolating the 
dependence structure. Dependence structure and marginals can then be studied 
separately. Normalized rank plots such as Fig. \ref{fig_cop_pedag1}C-D
relate to copulas in a way to be described in the section on Background below, 
after a definition of copulas
is provided. Note that Pearson correlation, even though it is the most commonly
used measure of relationships between variables, is modified by monotonic
tranformation of the component variables (Appendix 
\ref{SM-pedagog1_details}, Fig. \ref{SM-fig_transf_pearson_spearman}),
and therefore reflects not only dependence information, but also
information on the marginals [@Genest2007]. 
Spearman and Kendall correlations, being rank-based, 
are not influenced by monotonic transformations of the component variables; 
they provide information solely on the dependence between those variables.

A main benefit of a copula approach is 
that it can detect associations in the tails of distributions, and asymmetries of such associations. 
Tail association (introduced formally in the section on Background, below) 
is association between extreme values of variables. 
If smaller values of two positively associated variables are more 
strongly associated than are larger values, the variables are said to
have stronger lower- or left-tail association than right- or upper-tail association
(Fig. \ref{fig_cop_pedag1}C);
and *vice versa* if larger values are more strongly associated than smaller
ones (Fig. \ref{fig_cop_pedag1}D). Datasets of the same 
Spearman or Kendall correlation can have a range of tail association 
(Fig. \ref{SM-fig_cop_pedag}), which can be quantified (see sections on Background and
methods for Q1 below). Although copulas can be used to model the entire
complex dependence structure of data, we have found tail associations to be
a useful aspect of the full structure of dependence, so tail associations 
is one of the main foci of this paper.

The goal of this paper is to explore the potential for 
applications of copulas in ecology. We provide a basic introduction
to copula concepts in the Background section below, but this is a sketch
only - readers can find abbreviated [@Anderson2018; @Genest2007] 
and comprehensive introductions elsewhere
[@nelsen2006_copula; @joe2014_dependence; @MaiScherer2017]. Instead our goal is to 
help determine to what extent ecological understanding may benefit generally from
using copulas. Copulas have already been used very effectively 
in a few studies to illuminate ecological phenomena 
[@Valpine2014; @Anderson2018; @Popovic2018], but such usage is
still rare. Our results suggest that environmental, 
ecological and evolutionary processes 
may commonly generate and transmit complex dependence structures, including tail 
dependence, and that greater use of copulas can illuminate the underlying processes.

\begin{figure}[!h]
\begin{center}
\includegraphics[width=7in]{./MasterFigure.pdf}
\caption{Summary and guide for analyses presented in the text. Middle boxes
correspond to ecological datasets for which Q1 was examined (in the 
third through
fifth sections of the text, see also Table \ref{tab_data_info}). Upper boxes 
correspond to potential causes of non-normal (non-N) copula structure that
were examined (in the sixth and seventh sections of the text). Lower boxes 
correspond to potential consequences we examined of non-normal copula structure for
ecological understanding and for applications (in the eighth and ninth
sections of the text). Arrow labels (A-D for analyses pertaining 
to causes, Q2; and X-Z for analyses pertaining to consequences, Q3) 
link to locations in the text.\label{MasterFigure}}
\end{center}
\end{figure}

One way we may expect, *a priori*, a copula 
study of complex dependence structures between
ecological variables
to better reflect ecological relationships relates to Liebig's law of the
minimum. Liebig's law is 
the idea that growth is controlled not by total resources but by the resources which
are scarcest relative to organism needs. 
If, for instance, the growth of a plant depends on soil 
nitrogen, N, and other factors, a plot of plant growth rates versus soil N may 
look more like Fig. \ref{fig_cop_pedag1}A or C than like 
Fig. \ref{fig_cop_pedag1}B or D. Nitrogen 
will control plant growth, producing a clear
relationship, only when it is limiting, in the left portions of the distributions. 
The portions of the distributions that dominate the association are visible in a rank-based
plot, but may not be revealed by standard correlation techniques.

A second reason we may expect understanding of complex dependence structures
to benefit ecological
research is that prior work has demonstrated (using copulas) complex 
stucture in the spatial dependence of environmental 
variables [@Serinaldi2008; @She2018; @Goswami2018;
@li2013]. One may expect an environmental variable 
measured through time in two locations to show strong tail associations between
the locations if
intense meteorological events are also widespread in their effects, as seems 
frequently to be the case: extreme values are associated with intense events,
so happen in both places at the same time, whereas moderate values of the 
environmental variable may be associated with local phenomena, which differ
between the locations.
Spatial dependence in an environmental variable 
tends to beget spatial dependence 
in fluctuations of populations or other ecosystem variables 
(this is called \emph{spatial synchrony}) which are influenced by the environmental 
variable. This influence is called the \emph{Moran effect}. 
If complex dependence structure is transmitted, 
possibly in modified form, 
from environmental to ecological variables, then we may expect complex dependence 
structure to be a common feature of the 
spatial synchrony of population, community, 
biogeochemical and other environmentally influenced ecological variables.
Synchrony is a phenomenon of substantial interest in 
ecology [@Liebhold2004; @Sheppard2016; @walteretal2017].

A third reason we may expect an understanding of complex 
dependence structures to benefit 
ecological research is that such an approach may help illuminate 
the causal mechanisms between variables. For instance, if two species
Sp1 and Sp2 are strong competitors, abundances of the two species
across quadrats should be negatively related, as in 
Fig. \ref{fig_cop_pedag1}E-F. If Sp1 is plotted on the horizontal axis
and Sp2 on the vertical, then Fig. \ref{fig_cop_pedag1}E 
may suggest Sp1 is the dominant 
competitor: when Sp1 is abundant, Sp2 is necessarily rare because it is suppressed; 
whereas when Sp1 is rare, Sp2
may be abundant, or may also be relatively rare due to 
limiting factors other than Sp1. 
Alternatively, Fig. \ref{fig_cop_pedag1}F may suggest that Sp2
is the dominant competitor. Although other causal hypotheses may produce 
similar dependence structure, and it is usually impossible to obtain
complete information on causal pathways from analyses which are
fundamentally based on associations, more detailed 
examination of the patterns of dependence of variables may
well be combined with biological and other reasoning to rule out
some causal hypotheses which could not be eliminated solely 
with standard correlation approaches. @Popovic2018 has previously
used copulas to help illuminate causal aspects of relationships
between species.

A fourth reason we may expect a copula understanding of complex 
dependence structures to be useful
for ecology has to do with spatially aggregated or averaged quantities. 
Many ecological variables of applied importance 
depend on the spatial average or sum of local 
quantities. For instance, regional methane and CO$_2$ 
fluxes are the sum of local fluxes; 
and the total economic value of a fishery is the sum of local catches. 
We will explore how details, which can be captured using copulas, 
of the dependence between fluctuations of local
quantities can influence fluctuations of 
the spatial mean or sum, and how this may influence 
higher organizational levels in ecology and human concerns.
Illustrating the idea here, prior work demonstrated that the overall 
reliability of wind-generated electricity depends sensitively on details
of the dependence between wind speeds at multiple generator sites [@li2013]. 
Spatially aggregated ecological variables may be subject to a similar 
effect, with
possible consequences. For instance, if populations of a 
pest species in different locations are all positively associated and also more strongly
related to each other in their right tails, then local outbreaks will tend 
to occur together, creating regional epidemics. 
On the other hand, stronger left-tail associations in a pest species, even if overall 
correlation were unchanged, would have more benign effects.

We approached our overall goal of exploring the potential of copula approaches 
for ecology by addressing the following specific questions. (Q1) Do datasets
in ecology have dependence structure distinct from that of a multivariate 
Gaussian/normal
distribution (here called non-normal copula structure)? Do they show tail 
associations distinct from those of a normal distribution, and in particular
do they show asymmetric tail associations? Normal copula structure is assumed
by standard approaches that use multivariate 
normal distributions or distributions obtained by transforming
the marginals of a normal distribution. So Q1 asks whether ecological
data contain dependency information distinct from the standard or default case. 
(Q2) If the answer to
Q1 is "yes," then what are some possible causes/mechanisms of non-normal
copula structure and asymmetric tail associations in ecology? And (Q3) what are the 
consequences of non-normal copula
structure and asymmetric tail associations for ecological understanding and
applications? After a Background section (the second section), we address Q1 
(third through fifth sections of the paper) 
by analyzing several datasets, including 
environmental, species-trait, phenological, population, community, and 
ecosystem functioning data, selected to span multiple 
sub-fields and organizational levels of ecology. 
We address Q2 (sixth and seventh sections of the paper), using 
simple models. We address Q3 (eighth and ninth sections of the paper) using 
both data and models. The final section is the Discussion.
Multiple analyses are brought to bear on each question, and these
are summarized diagramatically in Fig. \ref{MasterFigure}, which can 
serve as a *post hoc* guide to the paper. 
Copula approaches have been used to great effect in neuroscience [@onken2009], 
bioinformatics [@Kim2008], medical research [@Emura2016], direct study of environmental 
variables [@Serinaldi2008; @She2018; @Goswami2018; @li2013], and finance [@Li2000],
and have been used very effectively, but rarely so far in ecology [@Valpine2014; 
@Anderson2018; @Popovic2018]. 
We argue that benefits of wider usage of complete 
copula descriptions of dependence in ecology will be substantial.

# Background on copulas

We give a brief introduction to copulas, focusing, for simplicity, on bivariate copulas
and on concepts needed
for the rest of the paper. See @nelsen2006_copula, @Genest2007, @joe2014_dependence, and
@MaiScherer2017 for general background, and see also @Anderson2018, part of 
which is a short introduction to 
copulas for ecologists. A bivariate *copula* can be defined as a bivariate distribution
function with both margins uniform on $(0,1)$ (@joe2014_dependence, p. 7). 
This will be the
distribution function of a bivariate random variable with uniform marginals, and 
the terminology *copula* is sometimes also applied to the random variable. 
A foundational theorem of
@sklar1959_theorem (see also @MaiScherer2017, p. 16) 
says that if $F$ is the bivariate distribution function
of a random vector $(X,Y)$, with margins $F_X$ and $F_Y$, then there 
exists a copula $C$
such that for all $(x,y)$ in the Euclidean plane, $F(x,y)=C(F_X(x),F_Y(y))$.
The theorem also says $C$ is unique if $F_X$ and $F_Y$ are continuous.
Thus $C$ couples the bivariate distribution, $F$, with the two marginal distributions. 
Finally, Sklar's theorem says that if $D$ is any bivariate copula and 
$G_A$ and $G_B$ are univariate distribution functions, then 
$D(G_A(x),G_B(y))$ is a distribution function.
The applications of this study fall within the case in which univariate marginal 
distributions have continuous, strictly 
monotonic distribution functions, and this case is simpler, so we 
henceforth make such an assumption. 
Then, "the" copula (now guaranteed unique) 
associated with $(X,Y)$ is the distribution
function of the random variable $(F_X(X),F_Y(Y))$ 
(Appendix \ref{SM-ContStrictMonoton}). And 
$D(G_A(x),G_B(y))$ is the distribution function of the random variable 
$(G_A^{-1}(U),G_B^{-1}(V))$, where $(U,V)$ is the random
vector with distribution function $D$ and
$G_A^{-1}$ and $G_B^{-1}$ are the inverses of $G_A$ and $G_B$
(Appendix \ref{SM-ContStrictMonoton}). Application of the 
distribution functions of the component random variables translates
between the uniform-marginal context, which isolates the 
dependence information, and the context of the original random variable.

Thus Sklar's theorem implies that
any random vector can be constructed from a copula and marginal 
distributions. And conversely,
any copula and any univariate distribution functions give rise to a random vector. 
Marginals contain no information about the
dependence structure of a random vector, so the copula contains
all such information - it is a complete description of the dependence
between variables. Sklar's theorem makes it possible to construct
bivariate distributions in two separate steps: by specifying the marginal 
distributions and by specifying the dependence structure. For instance,
construction of several bivariate distributions is carried out in this way in
Fig. \ref{PedagogFigCopThry}. The contrast between Fig. \ref{PedagogFigCopThry}C and D
illustrates in particular how a bivariate distribution can be changed, while retaining 
the same marginals, by changing the copula. Fig. \ref{PedagogFigCopThry}C is a 
bivariate normal distribution, but Fig. \ref{PedagogFigCopThry}D clearly 
is not, although its marginals
are the same. Fig. \ref{PedagogFigCopThry}D shows stronger association of the two variables
in the left than in the right tails. Sklar's theorem
also makes it possible to study and model dependence separately 
from marginal distributions. For instance, in a pedagogical 
example, @Anderson2018 modelled the dependence between abundance
data for two fish species, red moki and black angelfish, by first 
modelling the marginal distributions and then modelling 
the dependence using Gaussian copulas (their section 2).
See the Discussion for a few words on multivariate copulas, which can
be developed in the same way as above [@nelsen2006_copula; @joe2014_dependence;
@MaiScherer2017; @Anderson2018].
<!--***DAN: dont forget to put these few words in!-->

\begin{figure}[!h]
\begin{center}
\includegraphics[width=6.5in]{./Results/PedagogFigCopThry.pdf}
\caption{Normal (A) and Clayton (B) copulas were combined with standard normal 
marginals (C, D) and gamma marginals (E, F) via Sklar's theorem to produce 
bivariate distributions. Note that a normal copula is distinct from a normal marginal 
distribution: a normal copula is the copula of a bivariate normal distribution.
See the Background section for more information on normal, Clayton,
and other copulas. Each copula was used with both sets of marginals and each
set of marginals was used with both copulas, to demonstrate that: both copula and 
marginals contribute fundamentally to the resulting distribution, the copula 
contributing the information on the dependence between the variables; and that 
one can select the copula and marginals independently. Bivariate
distributions, including the copulas, are depicted via their log-scale probability density functions (pdfs), 
and marginals are depicted via their linear-scale pdfs. Grey dots are
50 random samples from each distribution. The parameter $0.7$ was used for the normal 
copula, the parameter $2$ for the Clayton 
copula, and shape and scale parameters $5$ and $1$ for the gamma marginals. 
Variables $u$ and $v$ were used for copulas
and $x$ and $y$ for distributions created by pairing a copula with
(non-uniform) marginals.
\label{PedagogFigCopThry}}
\end{center}
\end{figure}

<!--***DAN: In the Discussion, you need to point out not only that we
use only bivariate copulas (and there are multivariate ones), but also
that we use only continuous marginals. Probably an adequate approximation
for our data, but see the other papers I've seen recently which use count data,
and discrete marginals. Approaches such as that will be necessary for some 
datasets-->

We introduce some standard copula families, as examples of copulas and also
because we will fit these families to data later. The bivariate normal
copula family, which consists of copulas of bivariate normal
distributions, is a one-parameter family. The parameter, which we here call $p$,
corresponds to the correlation
of the related bivariate normal distribution, and controls the degree of dependence
between the variables. A normal copula with $p=0.7$ was already introduced (Fig.
\ref{PedagogFigCopThry}A). The formula for the
copula is $\Phi_{2,p}(\Phi_{1}^{-1}(u),\Phi_{1}^{-1}(v))$, where $\Phi_1^{-1}$ is the inverse of the
cdf of a (univariate) standard normal and $\Phi_{2,p}$ is the joint cdf of a
bivariate normal distribution with mean $(0,0)$ and covariance matrix having $1$s on the diagonal
and $p$ in the off diagonal positions. However, formulas for copulas are often not as 
directly informative about copula properties as pdf diagrams; Fig. \ref{figped_N} has 
example pdfs corresponding to bivariate normal copulas for various values of $p$.
Note that the pdfs are symmetric
around the diagonal line $v=-u+1$ (Fig. \ref{figped_N}), so normal copulas have symmetric 
associations between the two variables in the left and right tails (see
below for a more precise defintion of tail associations). 
The Clayton copula family,
of which we also already pictured an example (Fig.
\ref{PedagogFigCopThry}B), is another one-parameter family. In contrast to normal copulas, Clayton copulas
have stronger left- than right-tail association. The formula is 
$\left[  \reumax(u^{-p}+v^{-p}-1,0)\right]^{-1/p}$, for parameter $p$; though this again provides 
less direct intuition than do example pdfs (Fig. \ref{figped_C}). The survival
Clayton family is the symmetric opposite of the Clayton family, showing
stronger right- than left tail association (Fig. \ref{SM-figped_SC}). The BB1 copula 
family is a two-parameter family, thereby providing more flexibility: for some parameters
BB1 copulas have stronger left- than right-tail association, and for others the reverse
(Fig. \ref{figped_BB1}). See @nelsen2006_copula and @joe2014_dependence for further
information on these and other copula families.

<!--***DAN: Note: I am using tail dependence for the specific thing defined on p 62 of Joe,
and tail associations as a more general term. I have changed this in parts of the ms
upsteam from this point, but need to remember to change it as I move downstream.-->

Having introduced tail association conceptually and referred to it
in examples, we now need a more precise definition of a measure of tail association.
We use the term *tail association* to describe the general idea of the strength of
associations between two variables in the tails of their distributions. One of the
ways this is measured is called *tail dependence* (@joe2014_dependence, section 2.13; 
@nelsen2006_copula, section 5.4), which we here define. Given a random vector $(X,Y)$
with margins $F_X$, $F_Y$ and defining $U=F_X(X)$, $V=F_Y(Y)$, the upper-tail
dependence of $X$ and $Y$ is defined as 
$\lambda_U = \lim_{u \rightarrow 1^-} P[Y>F_Y^{-1}(u) | X>F_X^{-1}(u)]$. This equals
$\lim_{u \rightarrow 1^-} P[U>u | V>u]$, which in turn equals 
$\lim_{u \rightarrow 1^-} P[U>u, V>u]/P[V>u]=\lim_{u \rightarrow 1^-} P[U>u, V>u]/(1-u)$,
which shows upper-tail dependence is defined symmetrically in the two random variables.
Lower tail dependence is defined analogously, as 
$\lambda_L = \lim_{u \rightarrow 1^-} P[Y \leq F_Y^{-1}(u) | X \leq F_X^{-1}(u)]$.
Tail dependence is a property of the random variable $(U,V)$, so depends only
on the copula and not on the marginals combined with it. It can be shown 
(@nelsen2006_copula, section 5.4) that $\lambda_L=\lambda_U=0$ for the normal copula,
$\lambda_L=2^{-1/p}$ and $\lambda_U=0$ for a Clayton copula with parameter $p>0$,
and $\lambda_L=0$ and $\lambda_U=2^{-1/p}$ for a survival Clayton  copula with parameter $p>0$.
In the section on methods for Q1 we will introduce
ways tail dependence can be applied to data, by first fitting copulas to the data. We will also 
introduce nonparametric measures of tail association that apply directly to data.

Plots using normalized ranks were proposed in the Introduction as being conceptually
similar to copulas. We here explain the connection. As stated above, the copula associated
with a random variable $(X,Y)$ is the joint distribution function of the random variable
$(F_X(X),F_Y(Y))$. Given a sample $(x_i,y_i)$, $i=1,\ldots,n$, from $(X,Y)$, let $\hat{F}_X$ and
$\hat{F}_Y$ be the empirical cdfs associated with the $x_i$ and $y_i$, respectively.
These are step functions of $x$ and $y$, respectively, that start at $0$ for low values of 
$x$ and $y$ and jump by $1/n$
at each of the data points; for large $n$ they approximate the cdfs $F_X$ and $F_Y$. 
But $\hat{F}_X(x_i)$ equals the normalized rank of $x_i$ times
$n/(n+1)$, and likewise $\hat{F}_Y(y_i)$ equals the normalized rank of $y_i$ times
$n/(n+1)$. So the normalized rank in turn approximates the empirical cdf and therefore 
the cdf. Thus the normalized rank pairs $(u_i,v_i)$ for $i=1,\ldots,n$ can be 
regarded as approximate samples from the bivariate pdf 
associated with the copula of $(X,Y)$,
and the scatterplot of $v_i$ versus $u_i$ can be used to infer aspects of copula structure
(see section on methods for Q1 below).

\begin{figure}[!h]
\begin{center}
\includegraphics[width=\textwidth]{./Results/PedagogSuppMat_Normal.pdf}
\caption[Example normal copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example normal copulas. 
K is Kendall correlation; $p$ is the value of the parameter for the normal
family (it is a 
one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, respectively (see Background section for definition). The 
parameter range for the family is $p \in [-1,1]$, and lower- and upper-tail
dependence are always $0$.\label{figped_N}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=\textwidth]{./Results/PedagogSuppMat_Clayton.pdf}
\caption[Example Clayton copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example Clayton  
copulas. K is Kendall correlation; $p$ is the value of the parameter for the Clayton family (it 
is a one-parameter family); and LT and UT are the measures of lower- and 
upper-tail dependence, 
respectively (see Background section for definition). The parameter range for the family is $p \in (0,\infty)$, lower-tail dependence is $2^{-1/p}$ and upper-tail
dependence is $0$.\label{figped_C}}
\end{center}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=\textwidth]{./Results/PedagogSuppMat_BB1.pdf}
\caption[Example BB1 copulas]{Log-transformed pdfs for example BB1 copulas. K is Kendall correlation; $p_1$ and $p_2$
denote the two parameters of the family (it is a 
two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
respectively (see Background section for definition). The parameter ranges for 
the family are $p_1 \in (0,\infty)$ and $p_2 \in [1,\infty)$, 
lower-tail dependence is $2^{-1/(p_1 p_2)}$ and upper-tail
dependence is $2-2^{1/p_2}$.\label{figped_BB1}}
\end{center}
\end{figure}

# Data \label{Data}

```{r read_somedata, echo=F, results='hide'}
d_soilCN<-read.csv("./Data/SoilCN/SoilCNdata.csv")
rownames(d_soilCN)<-d_soilCN$X
d_soilCN<-d_soilCN[,-1]
d_soilCN<-na.omit(d_soilCN)
d_birdsBMR<-readRDS("./Data/BMR/BirdBodyMassesMetabolicRates/my_birdsBMR_data.RDS")
d_mammalsBMR<-readRDS("./Data/BMR/MammalianBodyMassesMetabolicRates/my_mammalsBMR_data.RDS")
d_cc<-readRDS("./Results/Ceder_creek_results/cop_HB_all_yr.RDS")
```

Datasets included environmental, species-trait, phenological, 
population, community, and ecosystem functioning data (Table \ref{tab_data_info}), selected to span multiple fields and levels
of organization within ecology. Copula structure describing 
the relationship between atmospheric weather variables such as rainfall or wind speed, 
as measured in multiple locations through time, has been 
examined previously in the meteorological literature [e.g., @Serinaldi2008; @li2013]. We therefore 
examined environmental variables from the soil instead, using the Rapid Carbon 
Assessment database [RaCA; @Wills2014].
<!--***DAN: Add Wills and Loecke pending reference if published in time--> 
The database comprises measurements of soil organic carbon and total soil 
nitrogen (Mg C or N per hectare of soil surface) from `r nrow(d_soilCN)` 
locations across the coterminous United States (Fig. \ref{SM-fig_SoilCNmapUSA} for sampling locations).
Species-trait data were average species basal metabolic rate (BMR, KJ per hour) and 
body mass (grams) for `r nrow(d_birdsBMR)`  species of birds [@Mcnab_2009] 
and `r nrow(d_mammalsBMR)`  species of mammals [@Mcnab_2008].
These data have been much studied, but to our knowledge the copula structure 
of the dependence between metabolic rate and body mass has not been examined. 
Species-trait data such as these reflect the coevolution of the two traits.
Phenological data were aphid first-flight dates from 10 locations (Fig. \ref{SM-fig_aphid_map}) 
across the United Kingdom (UK) for 20 aphid species (Table \ref{SM-tab_aphid_info}), for the 
35 years 1976 to 2010. These time series were computed from the Rothamsted Insect 
Survey (RIS) suction-trap dataset [@Harrington2014; @Bell2015]. The first of our two 
population-level datasets was also derived from the RIS suction-trap data, and comprised 
total counts of aphids trapped for the same locations, species, and years. 
Our second population dataset comprised average annual plankton abundance estimates 
for 14 locations (Fig. \ref{SM-fig_plankton_map})
in the North Sea and British seas for 22 taxa (Table \ref{SM-tab_plankton_info}) for the 
56 years 1958 to 2013. The locations are $2^\circ$ by $2^\circ$ grid cells. 
Community-level data, obtained from the Cedar Creek Ecosystem Science Reserve, were 
plant aboveground biomass [@ccdataple120] and Shannon's diversity index (computed from plant 
species percent cover data [@ccdatapce120]) for `r nrow(d_cc[[5]])` plots [@ccdataplotse120], 
each 9m by 9m, as sampled in the years 1996-2000 and 2007, each year 
analyzed separately [@Tilman2001; @Tilman2006]. 
Finally, ecosystem functioning data were methane (CH$_{4}$) fluxes between the soil or 
water surface and the troposphere, measured at 13 locations at daily to weekly intervals 
from September 2015 to September 2016 at the Great Miami Wetland mitigation bank in Trotwood, 
Ohio [@Holland1999; @Jarecke2016; @smyth2019using]. Each included location was measured on at least 50 dates. See Appendix \ref{SM-Data_details} for additional data details.

Our environmental, species-trait, and community datasets happen to be 
"bivariate" datasets in the sense that they comprise two quantities measured at 
different locations or for different species (Table \ref{tab_data_info}). 
Our phenology, population, and ecosystem functioning datasets are 
"multivariate" in that they comprise, for each taxon, measurements through time at multiple locations; 
copula structure can be studied for each pair of locations.

<!--summary table with info of all data used in Paper-->
```{r tab_summary_data_info, echo=F, results="asis",message=F}
library(kableExtra)
library(dplyr)
tab_data_info<-as.data.frame(matrix(NA,nrow=8,ncol=4))
colnames(tab_data_info)<-c("Data","Category","No. of measurements we used", "References")
tab_data_info$Data<-c("Soil C and N","Bird body masses and BMR", "Mammal body masses and BMR", 
                      "Green spruce and other aphid species abundances", "Leaf-curling plum and other aphid first flight dates","\\textit{Ceratium furca} and other plankton taxa abundances", "Plant diversity and aboveground biomass","Methane-flux")
tab_data_info$Category<-c("Environmental","Species-trait","Species-trait",
                          "Population","Phenological","Population","Community level","Ecosystem functioning")
tab_data_info$`No. of measurements we used`<-c(paste(nrow(d_soilCN)," locations",sep=""),
                                               paste(nrow(d_birdsBMR)," birds",sep=""),
                                               paste(nrow(d_mammalsBMR)," mammals",sep=""),
                                               "10 locations with at least 30-yr. timeseries for each of 20 sp.",
                                               "10 locations with at least 30-yr. timeseries for each of 20 sp.",
                                               "14 locations with at least 45-yr. timeseries for each of 22 taxa",
                                               paste(nrow(d_cc[[5]])," plots",sep=""),
                                               "13 locations with at least 50 dates of data")
tab_data_info$References<-c("US Rapid Carbon Assessment database (RaCA), Wills et al. (2014)", "McNab (2009)","McNab (2008)",
                            "Rothamsted Insect Survey","Rothamsted Insect Survey","Continuous Plankton Recorder Survey",
                            "Cedar Creek Ecosystem Science Reserve, biodiversity experiment e120",
                            "Great Miami Wetland Mitigation Bank, Smyth et al. (2019)")

knitr::kable(tab_data_info, 
             #format="latex", align="l",
             format="latex", align="l",escape=F,
             caption = "Summary table for the data we used. Bold entries are the multivariate data, the rest are bivariate datasets (see Data section). Basal metabolic rate=BMR. \\label{tab_data_info}", 
             booktabs = T,linesep = "\\addlinespace") %>%
             column_spec(1:4,width="3.4 cm")%>%
             row_spec(c(4,5,6,8), bold=T)
```

# Concepts and methods for Q1 \label{Methods}

We addressed Q1 first via a model selection procedure in which several
families of copulas were fitted to 
our ecological datasets and fits were compared via the Akaike and Bayesian 
Information Criteria (AIC and BIC). One of the fitted copulas was the normal
copula, making possible comparisons of the degree to which the normal copula versus
other copula families were good descriptions of data.

For bivariate datasets 
$(x_i,y_i)$ for $i=1,...,n$, model selection involved several steps. 
First, we produced normalized ranks 
$u_i$ and $v_i$ as in the Introduction. 
Second, we tested the independence of the $u_i$ and $v_i$
using the statistic $T=\sqrt{(9N(N-1))/(4N+10)}|\tau|$, where $\tau$
is Kendall's tau for the data [@Genest2007]. We used the 
implementation of this test given in the `BiCopIndTest` in the `VineCopula` package
in the R programming language. 
We tested for independence because our model selection 
algorithms were ineffective if data could not 
be distinguished from independent data 
since many of the copula families we considered include the independent copula. 
If independence could be rejected ($0.05$ significance level), model 
selection proceeded. Third, we fit 16 bivariate copula families (see below for 
information on the families used) to the 
normalized ranks via maximum likelihood. The approach of fitting
copula families to normalized ranks was studied in depth
and recommended by @Genest1995 and @Shih1995. Their estimator of copula
parameters, which we use, was shown to be consistent, asymptotically normal,
and fully efficient at indepencence [@Genest1995]. @Genest2007
recommend carrying out inferences of dependence structures (which was our
goal here) using normalized ranks.
We used the fitting implementation given in `BiCopEst` in `VineCopula`.
Fourth, we obtained AIC and BIC values and accompanying model weights 
$\text{AIC}_{\text{w}}$ and 
$\text{BIC}_{\text{w}}$ [@burnham2003_modelselection] for each fitted copula. 
`BiCopEst` also 
provided lower- and 
upper-tail dependence of the best-fitting member of each family. 
$\text{AIC}_{\text{w}}$ values were used to get
model-averaged lower- and upper-tail dependence values using standard model averaging 
formulas [@burnham2003_modelselection]; likewise for BIC.

Thus for each bivariate dataset, 
the end products of the procedure just described were threefold, corresponding to the
three parts of Q1 listed in the Introduction: A) the AIC, BIC, $\text{AIC}_{\text{w}}$ and 
$\text{BIC}_{\text{w}}$ values for each of our 16 copula families (see below for list),
including the normal family, providing an inference as to whether the normal
copula or an alternative was better supported by 
data; B) lower- and upper-tail dependence measures for each fitted copula and 
model-averaged tail dependence measures, providing information on whether,
and to what extent, tail dependence differed from the tail dependence of 
a normal copula (i.e., 0); and C) the difference of lower- and upper-tail
dependence measures for each fitted copula and model averages of those 
quantities, providing information on whether, and to what extent, tail
dependence was asymmetric.

Model selection methods give the *relative* support of several models, but do not 
indicate whether any of the models are an objectively good fit. To test that, 
we tested the goodness of fit of our AIC-best copula family using a 
bootstrapping procedure developed and studied by @Wang2000 and @Genest2006
and implemented as `BiCopGofTest` of `VineCopula`. The procedure performed 
one test based on a Cramer-von Mises statistic and another based on a 
Kolmogorov-Smirnov statistic. To keep computation times reasonable, 
a run using $100$ bootstraps was performed; if
the $p$-value from either test was less than $0.2$, tests were re-run with 
$1000$ boostraps.

We fit 16 bivariate copula families, exhibiting a variety of 
lower- and upper-tail dependence characteristics, with bivariate datasets. The purpose
of using a large collection of families was solely to include a 
variety of alternative dependence structures to have a robust model
selection and multi-model inference procedure. For that purpose, it 
is not important for the reader to
understand the details of these copulas,
and, additonally, these copulas have been 
described formally elsewhere [@Joe1997]; so 
we only briefly identify each family, say a few words about tail associations 
of the families, and provide pictorial descriptions in the Appendices.
We used several families that can exhibit positive lower-tail dependence 
(of strength depending on parameters) and 
zero upper-tail dependence: the Clayton, survival Gumbel, survival Joe
and survival BB6 familes. We used several families that can exhibit positive
upper-tail dependence (strength depends on parameters) and zero lower-tail
dependence: the survival Clayton, Gumbel, Joe and BB6 families. We used
families that show zero upper- and lower-tail dependence: the normal
and Frank families. These families have pdfs symmetric about the line
$v=-u+1$. We also used several families that can show both upper- and
lower-tail dependence, in relative amounts that depend on parameter
values: the BB1, survival BB1, BB7 and survival BB7 families. 
We also used the BB8 family, which shows zero lower-tail dependence,
and zero upper-tail dependence except for a boundary case for the parameters.
And we used the survival BB8 copula, which shows zero upper-tail dependence, 
and zero lower-tail dependence except for a boundary case.
"Survival" families are rotations of the copula with the similar name by
180 degrees. 
See @Joe1997 for details on all these families. We used the implementations
of these families provided in the `VineCopula` package for the R programming
language. See Figs \ref{figped_N}-\ref{figped_BB1} and \ref{SM-figped_SC} and \ref{SM-figped_G}-\ref{SM-figped_SBB8} 
for visual depictions of the pdfs of these copulas and how
pdfs and tail dependence are influenced by the parameters.

<!--***DAN: Info on families kept here for convenience, see also the file
CopulaFamilyParamRangesAndTailDependence.jpg, stored in the BIVAN folder:
lower not upper dependence:
Clayton (3), survival Gumbel (14), survival Joe (16),
SBB6 (18)
zero both, and symmetric:
  normal (1), Frank (5)
upper not lower dependence:
  SC (13), Gumbel (4), Joe (6)
  BB6 (8)
both, in relative amounts that depend on parameters:
  BB1 (7) - maybe need to look at formulas for this one to understand what the   
    limits are, since I only chose some random values of parameters in my pdf 
    pictures
  SBB1 (17)
  BB7 (9), SBB7 (19)
BB8 (10) - 0 LT dep, UT dep 0 except in a boundary case for the parameters
SBB8 (20) - reverse of BB8-->

For multivariate datasets, we performed the bivariate analysis
for all possible pairwise combinations of distinct locations. We carried out
pairwise bivariate analyses instead of trying to fit a multivariate copula,
for simplicity and because that approach was
sufficient to answer our research questions; but see the Discussion
for a few words on multivariate copulas. 
<!--***DAN: dont forget to put these in!-->
We present
the number of pairs for which a non-normal copula
was the AIC-best copula, and we characterize AIC differences between
the normal and AIC-best copulas across location pairs. 
We also computed the model-averaged lower- and
upper-tail statistics, and differences
between these, for each pair of locations, and we characterize
the distribution of these values across location pairs.

In addition to our model selection approach, we also used a
nonparametric approach, to provide greater 
confidence in our answers to Q1. We used three 
statistics which quantify the extent to which 
the normalized ranks $u_i$ and $v_i$ are 
related in any part of their distributions; we here describe the 
statistics, with some additional details in Appendix \ref{SM-nonparam_stats}.
The statistics are defined with positively associated
variables in mind. All our variables were positively associated
when they were significantly associated.
Given two bounds 
$0 \leq l_b < u_b \leq 1$, define the lines $u+v=2l_b$ and 
$u+v=2u_b$, which intersect the unit square (Fig. \ref{NonparamStatsFig}). 
Our statistics quantify the association
between $u_i$ and $v_i$ in the region bounded by these lines.
Using $l_b=0$ and $u_b\leq 0.5$ gives information about 
association in the left parts of the distributions of $u$ and $v$, and 
using $u_b=1$ and $l_b \geq 0.5$ gives information about association in 
the right parts. The first statistic, a partial Spearman correlation
\begin{equation}
\cor_{l_b,u_b}(u,v)=\frac{\sum 
   (u_i-\mean(u)) (v_i-\mean(v))}{(n-1)\sqrt{\var(u)\var(v)}},
\end{equation}
\noindent is the portion of the Spearman 
correlation of $u_i$ and $v_i$ that is attributable to the points 
between the bounds. 
Here sample means and variances are computed using all $n$ data 
points, but the
sum is over only the indices $i$ for which $u_i+v_i > 2l_b$ and 
$u_i+v_i < 2u_b$.
Larger values of the partial Spearman correlation indicate stronger positive association.
The sum of 
$\cor_{0,0.5}(u,v)$ and $\cor_{0.5,1}(u,v)$ (or some other choice of 
$\cor_{l_{b_k},u_{b_k}}(u,v)$ for bounds $l_{b_k},u_{b_k}$ that partition the
interval $(0,1)$) equals the standard Spearman correlation
as long as no points fall directly on the bounds.
We also defined a statistic $\Ps_{l_b,u_b}$ (Appendix \ref{SM-nonparam_stats}), 
which has a similar qualitative
interpretation to $\cor_{l_{b},u_{b}}$.
Our third statistic, $\Dsq_{l_b,u_b}$, is the average squared 
distance between points satisfying $u_i+v_i>2l_b$ and 
$u_i+v_i<2u_b$ and the line $v=u$.
Unlike $\cor_{l_b,u_b}$ and 
$\Ps_{l_b,u_b}$, for which large values inidicate strong association
between the bounds, small values of 
$\Dsq_{l_b,u_b}$ indicate strong association. These statistics are not estimators of the 
tail dependence quantities
defined previously, but rather are conceptually similar measures of associations in
the tail portions of the distributions when appropriate values of $l_b$ and $u_b$ are
selected.

\begin{figure}[!h]
\begin{center}
\includegraphics[width=3.5in]{./Results/NonparamStatsFig.pdf}
\caption{The partial Spearman correlation, $\cor_{l_{b},u_{b}}(u,v)$, 
within a band can be computed
for any band (methods for Q1) to describe how the strength of association
between $u$ and $v$
varies from one part of the two distributions to another, 
as can the statistics $\Ps_{l_b,u_b}(u,v)$ and $\Dsq_{l_b,u_b}(u,v)$. Diagonal lines
show two bands, the data in the left/lower band showing stronger
association than those in the right/upper band.\label{NonparamStatsFig}}
\end{center}
\end{figure}

For large datasets (large $n$),
we used $l_b$ and $u_b$ close together without incurring undue
sampling variation in our statistics, and we considered 
multiple bands $(l_b,u_b)$ to understand how dependence
varies in different parts of the distributions. But for datasets with 
smaller $n$ we considered only $l_b=0$, $u_b=0.5$ and $l_b=0.5$, $u_b=1$,
abbreviating $\cor_l=\cor_{0,0.5}$ ($l$ is for "lower") 
and $\cor_u=\cor_{0.5,1}$ ($u$ is for "upper"). Likewise $\Ps_l=\Ps_{0,.5}$,
$\Ps_u=\Ps_{0.5,1}$, $\Dsq_l=\Dsq_{0,0.5}$, $\Dsq_u=\Dsq_{0.5,1}$.
To test for asymmetry of dependence in upper and lower portions of distributions,
we used differences $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$
for smaller datasets (note the opposite order
in the last of these); and $\cor_{0,u_b}-\cor_{1-u_b,1}$, 
$\Ps_{0,u_b}-\Ps_{1-u_b,1}$ and
$\Dsq_{1-u_b,1}-\Dsq_{0,u_b}$ with $u_b$ close 
to $0$ for large datasets.
We tested our statistics in Appendix \ref{SM-Test_npa_stat} (see also 
Figs \ref{SM-fig_stat_testing35}-\ref{SM-fig_stat_testing1000}).

For bivariate datasets, we compared values 
of the statistics $\cor_{l_b,u_b}$,
$\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, $\cor_{0,u_b}-\cor_{1-u_b,1}$, 
$\Ps_{0,u_b}-\Ps_{1-u_b,1}$ and
$\Dsq_{1-u_b,1}-\Dsq_{0,u_b}$ to 
distributions of values of the same statistics computed on surrogate datasets that were produced from 
the empirical data by randomizing it in a special way to have no 
tail dependence (Appendix \ref{SM-surrog_test}). 
The surrogate/randomized datasets had exactly 
the same marginal distributions as the empirical data and had very similar Kendall or Spearman 
correlation (the surrogate algorithm had two versions, one for preserving each correlation coefficient), 
but had normal copula structure. Thus our comparisons tested the null hypothesis 
that our statistics took values on the empirical data no different from what would have been expected 
if the copula structure of the data were normal, but the data were otherwise 
statistically unchanged. The comparison of one of the
statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, as computed on the 
empirical data, to the distribution
of its values computed on surrogate datasets 
provides a test of whether association between the variables in the part of the distributions specified 
by $l_b$ and $u_b$ is different from
what would be expected from a null hypothesis of normal copula structure. 
Thus the comparison addresses the first two parts of Q1
for bivariate datasets. 
Significant deviations correspond to deviations from normal 
copula structure. In particular, deviations 
using $l_b=0$ and $u_b$ small (say, $0.1$) correspond to lower-tail associations 
different from that of a normal copula;
likewise, using $l_b=0.9$ and $u_b=1$ tests for upper-tail associations different 
from that of a normal copula.
The comparison of one of the statistics 
$\cor_{0,u_b}-\cor_{1-u_b,1}$, 
$\Ps_{0,u_b}-\Ps_{1-u_b,1}$ and
$\Dsq_{1-u_b,1}-\Dsq_{0,u_b}$,
as computed on the empirical data, to the distribution of its values on many surrogate datasets 
provides a test of asymmetry of tail associations. Thus this comparison helps address the
third part of Q1 for bivariate 
datasets.

For multivariate datasets, we calculated Spearman and Kendall correlations
and the statistics 
$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$,
$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$
for all pairs of sampling locations, and we then computed means across the pairs
for each statistic. We used a spatial resampling scheme 
(Appendix \ref{SM-spatial_resamp}) to calculate confidence intervals of these means.
The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
widely used spline correlogram method; the same scheme was also used by @walteretal2017 
as part of a method for 
performing model selection among matrix regression models.
Code and data for this
project are archived at \url{www.github.com/sghosh89/BIVAN}.

# Results for Q1: Ecological datasets have non-normal copula structure and asymmetric tail associations \label{Results}

```{r read_BivMS_soilCN_data,echo=F}
d<-read.csv("./Data/SoilCN/SoilCNdata.csv")
rownames(d)<-d$X
d<-d[,-1]
d<-d[,c("SOCstock100","TSNstock100")] #raw soilCN data
BivMS_res_Loecke<-readRDS("./Results/fitting_results/BivMS_soilCN.RDS")
```

```{r read_RES_aphid_count, echo=F}
RES_aphid_count<-readRDS("./Results/fitting_results/AphidCopulaFit_selecloc_count_species_10.RDS")
#------------ also reading some extra stuffs -------
stat_aphid_count<-readRDS("./Results/stat_results/stat_aphid_count/stat_aphid_count_sp_10.RDS")
stat_aphid_ff<-readRDS("./Results/stat_results/stat_aphid_ff/stat_aphid_ff_sp_11.RDS")
stat_plankton_north_sea<-readRDS("./Results/stat_results/stat_plankton_north_sea/stat_plankton_north_sea_sp_16.RDS")
stat_methane<-readRDS("./Results/stat_results/stat_methane/stat_methane.RDS")
tab_multivar_summary<-readRDS("./Results/tab_multivar_summary.RDS")
```

\begin{figure}[!h]
\begin{center}
\includegraphics[width=6.5in]{./Results/BivarDataPlot.pdf}
\caption{Upper panels show raw data plots for (A) $\log_{10}$(soil C) and $\log_{10}$(soil N) data, (B) $\log_{10}$(basal metabolic rate (BMR)) vs. $\log_{10}$(body mass) 
for birds, (C) the same for mammals, and (D) above-ground plant biomass vs. Shannon's index (\textit{H}) from Cedar Creek. Bottom panels show the corresponding normalized rank plots. See Data section for units used in upper panels. \label{fig_biv_multi_raw_cop}}
\end{center}
\end{figure}

Copula structures were non-normal and 
showed asymmetric tail 
associations for most, but not all datasets we examined, 
answering Q1 in the affirmative. 
To make results easier to absorb, we first 
present results for an example bivariate dataset, then we summarize results
for all bivariate datasets, then we present results for an example multivariate dataset, then we 
summarize for all multivariate datasets.

<!--bivariate example, model selection-->
For the soil C and N data (Data section, Table \ref{tab_data_info}, 
Fig. \ref{fig_biv_multi_raw_cop}A), variables were 
non-independent ($p=$ `r BivMS_res_Loecke$IndepTestRes`, to within the precision available 
from `BiCopIndTest` in the `VineCopula` package); non-independence is also visually apparent.
The Kendall correlation was `r round(BivMS_res_Loecke$TauVal,2)`. We fitted our 16 copulas
to the normalized ranks (Fig. \ref{fig_biv_multi_raw_cop}E) 
and computed AIC and BIC weights and corresponding
lower- and upper-tail statistics for the fitted distributions (Table \ref{tab_soilCNfit}).
`r BivMS_res_Loecke$InfCritRes$copname[which.min(BivMS_res_Loecke$InfCritRes$AIC)]` was the 
minimum-AIC copula, with AIC `r round(min(BivMS_res_Loecke$InfCritRes$AIC),2)`, whereas the 
normal copula had a much higher AIC, `r round(BivMS_res_Loecke$InfCritRes$AIC[1],2)`, answering the 
first part of Q1 for these data. Recall that AIC differences of $2$
or $3$ are considered meaningful and differences of $8$ or greater
are considered diffinitive, so that data provide essentially no support
for the higher-AIC model in that case [@burnham2003_modelselection].
Model-averaged lower- and upper-tail dependence statistics were `r round(BivMS_res_Loecke$relLTdep_AICw,3)` 
and `r round(BivMS_res_Loecke$relUTdep_AICw,3)`, respectively (AIC weights were used for averaging). 
These values are distinct from what a normal copula would give, namely, 0, helping to answer 
the second part 
Q1 for these data. The values also differed substantially 
from each other, helping answer the third part of Q1. 
The numbers reflect stronger upper- than lower-tail
dependence, and this is also visible in the extreme upper-right
and lower-left corners of the copula plot (Fig. \ref{fig_biv_multi_raw_cop}E).

Our model selection procedures do not reveal whether tail dependence parameters are 
*significantly* different from $0$ and from each other. 
Furthermore, we caution that, while our model selection results do convincingly show 
non-normal copula structure, model-averaged tail-dependence statistics may have been biased 
because even the lowest-AIC copula 
(`r BivMS_res_Loecke$InfCritRes$copname[which.max(BivMS_res_Loecke$InfCritRes$AICw)]`) 
was a poor fit ($p=$ `r BivMS_res_Loecke$GofRes_CvM` and `r BivMS_res_Loecke$GofRes_KS` for the 
Cramer-von Mises and Kolmogorov-Smirnov goodness-of-fit tests, respectively, to within the precision 
available from `BiCopGofTest`).
Our alternative nonparametric
approach, results detailed next, provides informaton about significance of tail
associations and asymmetries of tail associations.

The values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for the soil
C and N data
were compared to distributions of their values on 1000 Kendall- or Spearman-preserving normal surrogates
of the data,
separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$ 
(Fig. \ref{fig_soilCN_nonparam}). Results showed that tail associations of data are stronger in 
both the lower and upper tails than would be expected under a null hypothesis of a normal copula (Fig. \ref{fig_soilCN_nonparam}B-F).
The values of the statistics $\cor_{0,0.1}-\cor_{0.9,1}$, $\Ps_{0,0.1}-\Ps_{0.9,1}$ and
$\Dsq_{0.9,1}-\Dsq_{0,0.1}$ for the soil C and N data were also compared to distributions of their values
on surrogates, separately in two 
comparisons (Table \ref{tab_soilCN_asym}). Results showed that
upper-tail association was significantly stronger than lower-tail association. These results
are consistent with the model selection results, but go beyond them by providing information
about statistical significance. Thus our results provide an
affirmative answer to Q1 for the soil C and N data; this answer is 
represented in Fig. \ref{MasterFigure} as a solid outline around "soil C and N"
in the right-most box in the middle row.
<!--***DAN: you need to change the box around the Cedar Creek data-->

<!--bivariate summary-->
Table \ref{tab_bivar_summary} summarizes Q1 results 
for the bivariate datasets, with some details in the Appendices. The two 
variables were significantly related for all datasets 
(Table \ref{tab_bivar_summary}, row 1). A non-normal copula 
always emerged as the lowest-AIC copula 
(Table \ref{tab_bivar_summary}, row 2). 
The normal copula was a poor fit compared to the lowest-AIC copula 
(Table \ref{tab_bivar_summary}, rows 3-4), except for Cedar Creek, for which the AIC
difference between these fits was not meaningfully large.
Thus we answer the first part of Q1 in the affirmative for 3 of 4 of our bivariate datasets. 
Often, either lower- or upper-tail dependence statistics differed substantially from 0 
(Table \ref{tab_bivar_summary}, rows 7-8), and/or these statistics differed from each other (Table \ref{tab_bivar_summary}, row 9), 
helping to answering the second and third parts of Q1. Though again the model
selection results do not provide 
information on statistical significance of these differences, and are subject to the
caveat that, for some datasets, even the lowest-AIC copula was not an objectively good fit
(Table \ref{tab_bivar_summary}, rows 5 and 6), nonparametric results 
(Table \ref{tab_bivar_summary}, row 10) showed that tail associations
deviated significantly from what would be expected from a normal-copula
null model, and were also significantly asymmetric, except for the Cedar Creek data.
See Figs \ref{SM-fig_bmr_birds_nonparam}-\ref{SM-fig_cc2000_nonparam} for analogues to 
Fig. \ref{fig_soilCN_nonparam} for the bivariate datasets other than the soil C and N
data, see Tables \ref{SM-tab_bmr_birds_fit}-\ref{SM-tab_cc2000fit} for analogues to Table \ref{tab_soilCNfit},
and see Table \ref{SM-tab_3biv_npa_summary} for analogues to Table \ref{tab_soilCN_asym}.
The first three datasets had stronger upper- 
than lower-tail dependence, and the last dataset had the reverse, though non-significantly. Our 
generally affirmative (except for Cedar Creek), 
empirically based answer to Q1 for the bivariate datasets is 
represented in Fig. \ref{MasterFigure} as solid outlines around the names of 
the soil C and N and bird and mammal body mass versus metabolic rate
datasets, in the middle row of boxes.
<!--***DAN: change the box for Cedar Creek-->
We did the same analysis 
with Cedar Creek data from other available years, 1996-1999 and 2007. For 1996 and 
2007, biomass and Shannon's index were independent; for 1997-1999, results 
were similar to 2000.

We present green spruce aphid abundance (Data section; Table \ref{tab_data_info}) results
as an example multivariate analysis.
Independence was rejected for each pair of the 10 sampling locations. 
Best-fitting (lowest-AIC) copulas were non-normal
for the large majority of location pairs (`r tab_multivar_summary$Green_spruce_aphid_count$num.non_N` of `r tab_multivar_summary$Green_spruce_aphid_count$num.pairs` pairs), 
and the AIC
of the normal copula minus the minimum AIC, averaged across 
location pairs, was `r tab_multivar_summary$Green_spruce_aphid_count$Avg.N_minus_bestfit.AIC`.
Whereas `r tab_multivar_summary$Green_spruce_aphid_count$Avg.N_minus_bestfit.AIC`
would be only a marginally meaningful AIC difference for a single
location pair, it is more meaningful as an average across many
pairs. Relatedly, and illustrating the concept here, 
the chances of getting `r tab_multivar_summary$Green_spruce_aphid_count$num.non_N` or more 
non-normal location pairs if the chances were equal of getting
a normal versus a non-normal result (and taking into account that
the location pair (A,B) will necessarily 
produce the same result as the pair
(B,A)) is low, 
$`r 1-pbinom(tab_multivar_summary$Green_spruce_aphid_count$num.non_N/2-1,45,0.5)`$. 
These results help answer the first part of Q1: non-normal
copula structure appears to be meaningfully common in
these data, even if not universal.
Goodness of fit tests in every case 
failed to reject the hypothesis that the AIC-best copula 
family was also 
an objectively adequate description of 
the data; i.e., the collection of 
copula families we used was sufficiently broad to characterize 
these data. Model-averaged lower- and upper-tail
dependence statistics had $2.5^{th}$ and $97.5^{th}$ quantiles (across
location pairs)
(`r tab_multivar_summary$Green_spruce_aphid_count$LT_0.025_CI`, `r tab_multivar_summary$Green_spruce_aphid_count$LT_0.975_CI`) and 
(`r tab_multivar_summary$Green_spruce_aphid_count$UT_0.025_CI`, `r tab_multivar_summary$Green_spruce_aphid_count$UT_0.975_CI`), respectively,
thereby commonly differing 
from what a normal copula would give (i.e., 0), and helping to
answer the second
part of Q1: these data have greater tail dependence 
(lower and upper) than would be expected from a normal-copula null hypothesis. We note however that our model selection 
procedures again do not reveal whether tail-dependence 
parameters are *significantly* different from $0$, and we refer 
the reader to nonparametric results below for information
on significance. 
Model-averaged lower- minus upper-tail dependence statistics 
were positive for all but a few location pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` 
out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`). The chances
of getting `r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` 
or more positive values, here, under a null hypothesis of equal chances 
for positive and negative values (and again accounting for the fact that location pairs (A,B) and (B,A) will show the same result) was
again low, 
$`r 1-pbinom(sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)/2-1,45,0.5)`$.
Thus the spatial synchrony of rarity in the green spruce aphid is stronger 
than the spatial synchrony of outbreaks. This answers the third part of Q1.

Nonparametric statistics verified that tail associations were asymmetric for the green 
spruce aphid abundance data. Mean values across all pairs of 
locations of the Spearman and Kendall correlations and the statistics 
$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ were positive and
confidence intervals excluded $0$ (Table \ref{tab_count_resamp}). 
Mean values of the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$
also were always positive and had confidence intervals that 
excluded $0$ 
(Table \ref{tab_count_resamp}).

<!--multivariate summary-->
Table \ref{tab_multivar_summary} summarizes Q1 results for
multivariate data. Results supported the 
conclusions that non-normal copula structure, non-normal tail dependence, and asymmetric tail 
associations were common, though not universal, answering 
Q1 in the affirmative for these data. 
Most location pairs were non-independent 
(Table \ref{tab_multivar_summary}, row 2). The large majority of non-independent location 
pairs had best-fitting copulas that were not the normal copula (Table \ref{tab_multivar_summary}, 
row 3), and AIC values of best-fitting copulas were, on average
across location pairs, between $2.714$ (*Ceratium furca* abundance data)
and $5.764$ (leaf-curling plum aphid first-flight data) lower than
AIC values for the normal copula (Table \ref{tab_multivar_summary}, 
row 4). Best-fitting copulas were nearly always considered an adequate fit 
(Table \ref{tab_multivar_summary}, row 5). Some of the datasets (green spruce aphid
abundance, \textit{Ceratium furca} abundance, methane-flux) showed stronger 
lower- than upper-tail dependence 
(Table \ref{tab_multivar_summary}, rows 6-8), whereas leaf-curling plum aphid first flight
data showed the reverse. All of the values in Table \ref{tab_multivar_summary}, rows 8 deviated highly significantly
from what would have been expected under a null hypothesis 
of equal chances for positive and negative values.

Asymmetry results were generally verified by nonparametric approaches,
with the exception of the methane-flux data. 
For instance, the $95\%$ confidence intervals of the
mean over pairs of sampling locations of the statistic $\cor_l-\cor_u$ were 
(`r unname(round(stat_aphid_ff$numericdf[9,c(3,4)],3))`)
for the leaf-curling plum aphid first flight data
(Table \ref{tab_multivar_summary}, row 9), 
indicating greater upper-tail dependence, and consistent with
the results of Table \ref{tab_multivar_summary}, row 8. 
For the *Ceratium furca* abundance data, confidence intervals were
(`r unname(round(stat_plankton_north_sea$numericdf[9,c(3,4)],3))`),
indicating greater lower-tail dependence, and again 
consistent with the results of Table \ref{tab_multivar_summary} (row 9).
For the methane-flux data, confidence intervals were 
(`r unname(round(stat_methane$numericdf[9,c(3,4)],3))`);
the asymmetry in tail dependence revealed by the model selection
results for the methane-flux data was apparently not strong enough to
also be revealed by the nonparametric analyses. 
Analogues to Table \ref{tab_count_resamp} are in Tables
\ref{SM-tab_ff_resamp}-\ref{SM-tab_methane_resamp}.
Our generally
affirmative, empirically based answer to Q1 for the multivariate datasets is 
represented in Fig. \ref{MasterFigure} as solid outlines around the names
of those datasets, in the middle row of boxes.
We also carried out the same analyses for abundance and first-flight data for the 18 
aphid species for which we had data other than the green spruce and leaf-curling plum aphids,
as well as for the 21 plankton taxa for which we had abundance data other than
\textit{Ceratium furca} (results not shown); results supported the conclusion that 
non-normal copula structure and tail dependence, and asymmetry of tail associations, are 
common.

<!--bivariate results for soil CN data-->

<!--Display model selection results for soilCN data in a table-->
```{r tab_soilCNfit, echo=F, results="asis"}
#knitr::opts_knit$set(kable.force.latex = TRUE)
BivMS_res_Loecke<-readRDS("./Results/fitting_results/BivMS_soilCN.RDS")
h<-BivMS_res_Loecke$InfCritRes[,c("copname","AIC","AICw","BIC","BICw","LTdep","UTdep")]
colnames(h)<-c("Copula","AIC","AICw","BIC","BICw","LT","UT")
h_ord <- h[order(h$AIC),] #sorting data: min AIC to max AIC

#---- This line forces the formating which is not always true by using digits arg in kable function ----
h_ord$AIC<-formatC(h_ord$AIC,2,format="f")
h_ord$AICw<-formatC(h_ord$AICw,2,format="f")
h_ord$BIC<-formatC(h_ord$BIC,2,format="f")
h_ord$BICw<-formatC(h_ord$BICw,2,format="f")
h_ord$LT<-formatC(h_ord$LT,4,format="f")
h_ord$UT<-formatC(h_ord$UT,4,format="f")
#-------------------------------------------------

knitr::kable(h_ord, #longtable = T, 
             format = "latex",linesep = "",row.names = F, #digits = c(4,2,2,2,2,4,4), 
      caption.short = "Model fitting results for soil C and N dataset",
      caption = "Model fitting results for soil C and N dataset using 16 copula families: normal (N), Frank (F), Clayton (C), survival Clayton (SC), Gumbel (G), survival Gumbel (SG), Joe (J), survival Joe (SJ), BB1, survival BB1 (SBB1), BB6, survival BB6 (SBB6), BB7, survival BB7 (SBB7), BB8, and survival BB8 (SBB8). Table rows are sorted by AIC. AICw = AIC weight; BICw = BIC weight; LT = the lower-tail dependence statistic for the indicated copula family with fitted parameters; UT = the same for upper-tail dependence. \\label{tab_soilCNfit}",
      booktabs = TRUE)
```

<!--Displaying a table with npa summary results for soilCN data-->
```{r tab_soilCN_asym,echo=F,results="asis"}
tab_biv_npa_summary<-readRDS("./Results/stat_results/tab_biv_npa_summary.RDS")
tab_soilCN_asym<-tab_biv_npa_summary[1:3,2:4]
knitr::kable(tab_soilCN_asym, booktabs = T, escape=F,
             format = "latex",linesep = "",
      caption.short = "Nonparametric asymmetry tests against a normal-copula null, soil C and N",
      caption = "Soil C and N results for nonparametric tests of whether asymmetry of tail association was significant, compared to a normal-copula null hypothesis. The values of the listed statistics (cor$_{FS}$-cor$_{LS}$ = $\\cor_{0,0.1}-\\cor_{0.9,1}$, P$_{FS}$-P$_{LS}$ = $\\Ps_{0,0.1}-\\Ps_{0.9,1}$, D$_{LS}^2$-D$_{FS}^2$ = $\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$; here FS stands for `first slice' and LS for `last slice') for real data were compared to their values for 1000 Kendall- or Spearman-preserving normal surrogates, in seperate comparisons. A table entry $<X$ indicates the value of the given statistic on the data was less than its value on $X$ of the surrogates, so entries of the form $<X$ for $X$ equal to $975$ or above indicate that upper-tail dependence was significantly stronger than lower-tail dependence. Results were significant only for the $P$ and $D^2$ statistics.\\label{tab_soilCN_asym}")
```

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/stat_results/stat_soilCN/soilCN_bivfunctionplot.pdf}
\caption[Nonparametric tests against a normal-copula null, soil C and N data]{Nonparametric tests for tail association and other deviations from normal copula structure, for soil C and N data. As described in the main text, the values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data (red crosses) were compared to distributions of their values on 1000 Kendall- or Spearman-preserving normal surrogates of the data (blue and green x's show 0.025 and 0.975 quantiles), separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$. Whenever the red cross was outside the range given by the x's, text at the top of panels indicates the number of surrogate values the real-data value was greater than or less than. For instance, a value $>N$ (respectively, $<N$) means the value of the statistic on real data was greater than (respectively, less than) its value on $N$ surrogates. When the statistic was greater than 975 or less than 975 surrogate values, it indicates significance ($95\%$ confidence level). When $\cor$ or $\Ps$ values (respectively, $\Dsq$ values) were greater than surrogates, it means association in that part of the distributions was stronger than (respectively, weaker than) expected from a normal-copula null hypothesis. \label{fig_soilCN_nonparam}}
\end{center}
\end{figure}

<!--summary table : results with bivariate data-->
```{r tab_bivar_summaryl, echo=F, results='asis',message=F}
tab_bivar_summary<-readRDS("./Results/tab_bivar_summary.RDS")
tab_biv_npa_summary<-readRDS("./Results/stat_results/tab_biv_npa_summary.RDS")

# adding npa stat summary res: spearman cor. preserving corl-coru for 4 bivariate data
my_mat<- matrix(tab_biv_npa_summary[c(1,4,7,10),4],nrow=1,ncol=4)
colnames(my_mat)<-colnames(tab_bivar_summary) # this line is needed to apply rbind later: colnames should be matched

tab_bivar_summary<-rbind(tab_bivar_summary,my_mat)
row.names(tab_bivar_summary)<-NULL

library(kableExtra)
library(dplyr)
library(tibble)

dt <- tibble(Description = c("1. p-value, independence test",
           "2. Minimum-AIC copula",
           "3. AIC for best copula",
           "4. AIC for normal copula",
           "5. p-value, Cramer-von Mises goodness of fit test",
           "6. p-value, Kolmogorov-Smirnov goodness of fit test",
           "7. Model averaged lower-tail dependence",
           "8. Model averaged upper-tail dependence",
           "9. Model averaged lower- minus upper-tail dependence",
           "10. Asymmetry rank in surrogates", # this row will be commented out, that's why the numbering starts from 10 in the next line
           "10. Placement of cor$_{FS}$-cor$_{LS}$ in surrogate distribution (see caption)")
)
tab_bivar_summary<-cbind(dt,tab_bivar_summary)
tab_bivar_summary<-tab_bivar_summary[c(1:9,11),]

knitr::kable(tab_bivar_summary, 
             format="latex", escape=F,row.names=F,
             caption = "Summary of Q1 results for bivariate datasets. The $p$-values (rows 5-6) are for the minimum-AIC copula. Model averaging used for rows 7-9 was based on AIC weights. Row 10 shows values as in Table \\ref{tab_soilCN_asym}, upper right table entry. Although the result shown in row 10 for the soil C and N data is non-significant, see Table \\ref{tab_soilCN_asym} for significant results using the $P$ and $D^2$ statistics. The first 3 datasets use $(l_b,u_b)=(0,0.1)$ for the first slice (FS) and $(0.9,1)$ for the last slice (LS), whereas $(0, 0.25)$ and $(0.75,1)$ were used for Cedar Creek due to fewer data for that system. Results for Cedar Creek for the year 2000 are shown. See Table \\ref{tab_soilCNfit} for copula family abbreviations.\\label{tab_bivar_summary}",
             booktabs=T, linesep = "\\addlinespace",
             col.names = c(" ", "Soil C and N", "Bird masses and BMR", "Mammal masses and BMR", "Cedar Creek data")
             )%>%column_spec(2:5,width="1.5cm")%>%
             column_spec(1, width = "8.4cm")%>%kable_styling(latex_options = "basic",position = "center")
```

```{r tab_count_resamp,echo=F,results="markup"}
s<-stat_aphid_count$numericdf[,c(3,2,4)]
sc<-as.data.frame(c("Spearman","Kendall","cor$_{l}$","cor$_{u}$","P$_{l}$","P$_{u}$","D$_{l}^2$","D$_{u}^2$",
                    "cor$_{l}$ - cor$_{u}$","P$_{l}$ - P$_{u}$","D$_{u}^2$ - D$_{l}^2$"))
s<-cbind(sc,s)
knitr::kable(s, digits = 3,
             col.names=c("","2.5$^{th}$ quantile", "Mean", "97.5$^{th}$ quantile"),
             format = "latex",linesep = "", escape = F,align="c",
             caption.short="Results for nonparametric statistics, green spruce aphid abundance data",
             caption = "Average values of statistics across all locations pairs and confidence intervals based on spatial resampling for green spruce aphid abundance data.\\label{tab_count_resamp}",
             booktabs=TRUE)
```

<!--summary table : results with multi-variate data-->
```{r tab_multivar_summaryl, echo=F, results='asis',message=F}
tab_multivar_summaryl<-tab_multivar_summary[-c(4,12:15),] #exclude extra rows for AICw, Corl, Coru

tab1<-tab_multivar_summaryl[c(6,8,12),] # see carefully if you rearrange tab_multivar_summary's row
                                                   #in suppmat this will change
tab2<-tab_multivar_summaryl[c(7,9,13),]  # see carefully if you rearrange tab_multivar_summary's row
                                                   #in suppmat this will change

two_tab_in1 <- as.data.frame(do.call(cbind, lapply(1:ncol(tab1), function(i) paste0(" (",tab1[,i], ", ", tab2[ , i], ")") )))   # see carefully if you rearrange tab_multivar_summary's row in suppmat this will change

two_tab_in1_again <- as.data.frame(do.call(cbind, lapply(1:ncol(two_tab_in1), 
                                                         function(i) paste0(tab_multivar_summaryl[11,i], ", ", two_tab_in1[3,i]) )))

colnames(two_tab_in1)<-colnames(two_tab_in1_again)<-colnames(tab_multivar_summaryl)

compiled_tab<-rbind(two_tab_in1[1:2,],two_tab_in1_again)
rownames(compiled_tab)<-c("CI_LT","CI_UT","Avg_CI_Corl-Coru")

tab_multivar_summary_ed<-rbind(tab_multivar_summaryl[c(1:5),],
                            compiled_tab[1:2,], #CI of LT, CI of UT
                            tab_multivar_summaryl[10,],
                            compiled_tab[3,] #avg and CI of Corl-Coru
                           )

row.names(tab_multivar_summary_ed)<-NULL
library(kableExtra)
library(dplyr)
library(tibble)
dt <- tibble(
            Description = c("1. Location pairs (excluding self comparisons)",
           "2. Number of non-independent pairs",
           "3. Number of pairs with non-normal copula as best fit",
           "4. Average of AIC differences (normal AIC minus min AIC) across location pairs",
           "5. Number of well-fitting location pairs",
           "6. 2.5$^{th}$ and 97.5$^{th}$ quantiles of model-avg. lower-tail dependence",
           "7. 2.5$^{th}$ and 97.5$^{th}$ quantiles of model-avg. upper-tail dependence",
           "8. Percent pairs showing stronger lower tail dependence",
           "9. Mean with 2.5$^{th}$ and 97.5$^{th}$ quantiles of cor$_{l}$ - cor${_u}$ across all location pair")
           )

tab_multivar_summary_ed<-cbind(dt,tab_multivar_summary_ed)

knitr::kable(tab_multivar_summary_ed, 
             format="latex", escape = F,
             caption.short = "Summary of results for multivariate datasets",
             caption = "Summary of Q1 results for multivariate datasets. Rows 3-8 of the table were computed for the non-independent pairs (row 2) only. A well-fitting location pair (row 5) was one for which the best-fitting copula had $p$-values $>0.01$ for both the Cramer-von Mises and Kolmogorov-Smirnov goodness of fit tests. \\label{tab_multivar_summary}",
             booktabs=T, linesep = "\\addlinespace",
             col.names = c(" ","Green spruce aphid abundance",
                        "Leaf-curling plum aphid first flight",
                        "\\textit{Ceratium furca} abundance","Methane-flux")
             )%>%column_spec(2:5,width="2.5cm")%>%
             column_spec(1, width = "6cm")%>%kable_styling(latex_options = "basic",position = "center")
```

<!--Question 2 analyses-->
# Concepts and methods for Q2 \label{Causes}

Having demonstrated that non-normal copula structure and asymmetric tail associations are 
common in ecological variables, we 
addressed Q2 by exploring, using models, three possible mechanisms
by which these phenomena may arise. 
<!--***DAN: Maybe insert some text indicating these are first, exploratory models, and are not intended to comprehensively explore
a mechanism (that can be done in future work), but rather to get a sense
for whether observed copula structure may be partly due to the
mechanism, and whether the mechanism is worthy of more comprehensive
exploration in that connection in future work. One way these are noncomprehensive
explorations is that I chose particular limited sets of parameters.-->
The first mechanism relates to the ideas in the Introduction about 
Liebig's law of the minimum, and to nonlinear influences of environmental
variables on ecological variables.
If an environmental variables influences an ecological variable 
more strongly in one of its tails, we explored whether the ecological
variable could then exhibit asymmetric tail dependence across space.
Let $E_i(t)$ be an ecological variable 
measured at location $i$ ($i=1,2$) at time $t$.
Assume the dynamics 
$E_i(t+1)=bE_i(t)+g(\epsilon_i(t))+a \delta_i(t)$, where 
the $\delta_i(t)$ are standard-normally distributed 
and independent across time and locations, $a=0.2$,
$b=0.1, -0.1, 0.5, -0.5$ in different simulations, and 
the $(\epsilon_1(t),\epsilon_2(t))$ were drawn, independently through
time, from a bivariate normal distribution with $\var(\epsilon_i)=1$ and 
$\cov(\epsilon_1,\epsilon_2)=0.8$.
The function $g$ describes how the environment $\epsilon_i(t)$ influences
$E_i(t)$. We used $g$ equal to $g_1$ or $g_2$ in different simulations:
$g_1(\epsilon)$ equals $\epsilon$ if $\epsilon<0$ and $0$ otherwise;
$g_2(\epsilon)$ equals $\epsilon$ if $\epsilon>0$ and $0$ otherwise. Thus
$g_1$ represents environmental effects that negatively impact populations,
and that occur only below a threshold of $\epsilon=0$; 
and $g_2$ represents effects that positively impact populations and 
that occur only above a threshold of $\epsilon=0$.
Negative values of $b$ correspond to overcompensating dynamics 
and positive values to undercompensating dynamics; larger $|b|$
means slower return to equilibrium after a disturbance.

For each $b$ and $g_i$, we simulated the model for $25000$ time
steps and retained the $E_i(t)$ for the final $2500$ time 
steps. We applied our model selection algorithms and 
nonparametric statistics (see section on methods for Q1) to these 
outputs to discover if the model could
produce non-normal copula structure and asymmetric tail dependence in the 
$E_i(t)$. Because $(\epsilon_1,\epsilon_2)$
and $(\delta_1,\delta_2)$ have normal copula structure (they were drawn
from bivariate normal distributions), the Moran 
mechanism analyzed below 
does not operate here.
<!--***DAN: You may also want to have some additional commentary in the Discussion on
how the models and related models can be explored more comprehensively. In fact you may already have some of that, though you may want to beef it up. You may
want to insert a point at the end of the Methids section here again emphasizing
that these are simple models for an initial exploration, and we discuss in the 
Discussion how...-->

Our second mechanism is an extension of the well known Moran
effect, and was summarized conceptually in the Introduction. 
Let $E_i(t)$ again be an ecological variable, $i=1,2$.
We initially assume the AR(1) dynamics 
$E_i(t+1)=\beta E_{i}(t)+ \sqrt{1-\beta^2} \epsilon_i(t)$, with $\beta=0.5$. 
The environmental noises $\epsilon_i(t)$ were 
standard-normal random variables that were independent for distinct times, $t$, but 
exhibited different kinds of 
dependence across locations in different simulations (see below).
The variable $E$ is very general, and could represent deviations
of a population density from a carrying capacity, deviations of
total plant community biomass from an average value,
flux of a biogeochemical variable such as methane, or other quantities.

To explore the importance of nonlinear dynamics for our proposed mechanism, 
we also considered a
nonlinear population model $P_i(t+1)=P_i(t)\exp \left(r\left(1-P_i(t)/K  \right)+\sigma\epsilon_i(t)\right)$, using $r=0.5$, $K=100$, and $\sigma=0.1$ or 
$\sigma=1$. When $\sigma=0.1$, population dynamics stay close to the carrying capacity, 
$K$, and the nonlinearities of the model have limited influence on dynamics. When
$\sigma=1$, model dynamics are strongly nonlinear. We refer to these as
the weak-noise and strong-noise cases, though the importance of the noise here
is that, when it is strong, it brings the nonlinearities of the model into play. 
<!--***DAN: Didn't we use another value of r as well, and present results
on it? If so, probably should mention it here, in part so the exploration
seems a bit broader? Well, we mentioned that we did those and we verbally summarize
results but we did not show any. So maybe leave things as is in this respect.-->

For each of the model setups above, for each of several copula families, 
<!--***DAN: I currently use C, SC, G, SG, J, SJ. Do I really need all those? 
Surely it would be enough to just do C and SC and then state that the others 
give similar results? And if I move one of the result figure from this section
into the main text, that will mean only one such fig in the sup mat, at least
for AR(1)? I mean, right now we have 18 figure in sup mat from this analysis,
and only part of one of them in main text! See if the results are actually similar across
S/SC, G/SG, J/SJ, and if so consider keeping only the S/SC results in either the
main text or the sup mat, and just stating the others were similar. Then you would presumably
move the full C fig into the main text, replacing the current partial version.
So instead of having 18 figs in sup mat and half a fig in main text, you would
have 5 figs in sup mat and 1 in main text, and then you would just say "we did the
same thing for G, SG, J, SJ and got similar results."
So let's check, are results, broadly speaking, independent of which of the 6 copulas
I used?:
A) for AR(1) models, yes
B) for Ricker, weak noise, yes
C) for Ricker, strong noise, its more complicated. It's generally true that you get
more approximate agreement between noise and pop stats, with some rather glaring disagreement.
But which stats disagree the most depends on which cop we used!
So I guess we can still just display the C and SC results, as planned above, and explain
in text.
-->
for each $\tau=0.1,0.2,\ldots,0.9$, and for each of
$50$ replicate simulations, we generated 5000 noise pairs 
$(\epsilon_1(t),\epsilon_2(t))$ from the
bivariate random variable with the given copula family and the given Kendall
correlation $\tau$.
We then used this noise to drive the model, and retained both the
noise and population values for the final 500 
time steps. For each simulation, the following
statistics were then computed for both noises and populations: 
Pearson, Spearman and Kendall correlations, $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, 
$\Dsq_l$, $\Dsq_u$, $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$. 
Values were plotted against $\tau$ for noises and populations. If the 
hypothesis from the Introduction was correct that characteristics of 
the copula structure of spatial dependence in an
ecological variable can be inherited from the copula structure of spatial 
dependence in an environmental variables, then 
plots should be similar for populations and noises.

<!--***DAN: How much of the material from Appendix S8 should be moved into
the main text? I am thinking a fair amount of it. Currrently there
is really no results on that topic at all in the main text, and that seems
a bit ridiculous and probably is why the referee said we treated this 
material in a "breezy" way.-->
The next mechanism we investigated is evolutionary, and pertains to bivariate 
trait data across species, e.g., our bird and mammal 
data. This mechanism is a hypothetical explanation for the right tail dependence
observed in those data (Fig. \ref{fig_biv_multi_raw_cop}F,G). 
The hypothesis is that tail dependence occurs in evolutionary changes
in bivariate characters, and gives rise to tail dependence between the 
two character values across extant species.
We simulated bivariate character evolution on an estimate of 
the phylogeny, taken from @GenoudIM2018, of $817$ mammal species.
The root character state and the change 
across each branch were randomly chosen from matrices of one million 
independent draws from bivariate distributions showing one of five distinct types of copula 
structure: 1) extreme or 2) moderate left-tail dependence,
3) symmetric tail dependence, or 4) moderate or 5) extreme
right-tail dependence (Appendix \ref{SM-evol_mech}). 
All distributions had standard-normal marginals and Spearman 
correlation $0.875$ between components, so our simulations assess 
the impact of copula structure only. For 
each of the five copulas, mammalian character
evolution was simulated 100 times. For each simulation, symmetry of tail dependence 
of the two characters across phylogeny tips was assessed.
We hypothesized that cases 1 and 2 above would yield stronger left- than right-tail 
dependence in tip characters, and cases 4 and 5 would yield the reverse. 
The simulator was written in Python and used version 4.4 of the DendroPy 
package [@DendroPy].

# Results for Q2: Moran effects and asymmetric dependencies produce non-normal copula structure\label{Results_Q2}

Our model with asymmetric environmental effects 
produces outputs with visually apparent asymmetry 
of tail dependence, to an extent that depended on the value of $b$;
thus this is a mechanism that can produce non-normal copula structure and
asymmetric tail dependence across space in ecological variables.
<!--***DAN: move the following to the end of this part-->
This result is represented in Fig. \ref{MasterFigure} as the solid box around 
"Nonlinear environmental effects, Liebig's law" and the arrows labelled "A".
It is explained in the Discussion why the box and some of those arrows 
are solid, instead of dashed, although the results we 
present in this section are theoretical.
<!--***DAN: move the above to the end of this part-->
For $b=0.1$ (Fig. \ref{fig_ms_asym_b_0.1}) and $b=-0.1$ 
(Fig. \ref{SM-fig_ms_asym_b_-0.1_0.5_-0.5}A,B), 
for both $g_1$ and $g_2$,
asymmetry of tail dependence was strong; for 
$b=\pm 0.5$, asymmetry was weaker but still apparent
(Fig. \ref{SM-fig_ms_asym_b_-0.1_0.5_-0.5}C-F).
<!--***DAN: Shyamolina, can you please combine Figs S24-S26 into one single figure
with 6 panels? Please also search for citations in the main text and sup mat that 
used to refer to the three figures and replace them with a citation to the
new, combined figure.
***Shya: Done.-->
Lower-tail (respectively, upper-tail) spatial dependence in 
effects of noise ($g_1$, respectively, $g_2$) 
produced lower-tail (respectively, upper-tail) spatial 
dependence in the ecological variable.
Model selection results (Table \ref{SM-tab_summary_stat_asym_sens}) confirmed that
ecological outputs had non-normal copula structure, but did not always reflect 
asymmetry of tail dependence, perhaps because those methods measured
dependence in the extreme tails. Results using our nonparametric statistics, however,
reflected the visually apparent asymmetry (Tables \ref{SM-tab_stat_npa_asym}). 
<!--***DAN: Simplify as follows, maybe:
1) get rid of table S10 (you will have to change the methods so it no longer says we
do our model selection analysis here)
2) simplify table S11 so it only shows the cor results (also means a change to methods)
and then maybe bring the resulting table to main text?
-->

<!--***DAN: Shyamolina, can you please simplify Table S11 so it only shows the cor
results? That also means the third column will be unnecessary. Please do this as follows:
1) keep but comment out the code that produces the old table; 2) write new code (presumably
starting from a copy of the old code) that makes the new, simpler version of the table. 
That will make it very easy to produce the old version again if we ever want to.
***Shya: Done.-->

\begin{figure}[!h]
\begin{center}
\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf}
\caption{If environmental effects operate asymmetrically in their tails on
ecological variables, it can result in non-normal copula structure and asymmetric 
tail dependence across space in the ecological variables. Shown are the last
$500$ points for (A) $g=g_1$ and (B) $g=g_2$ from simulations described in the text, $b=0.1$.\label{fig_ms_asym_b_0.1}}
\end{center}
\end{figure}

The hypothesis that characteristics of 
spatial copula structure and tail dependence
in an ecological variable can be inherited from an environmental 
variable was largely confirmed; 
<!--***DAN: You are using language about confirming hypotheses, here and elsewhere,
but this may rub some people the wrong way because we have not confirmed anything
empirically, we are just doing models, and only a cursory exploration of each
model, at that! So change the language.-->
however, similarities between environmental- and 
ecological-variable copula structure were reduced when 
dynamics were strongly nonlinear.
Thus this is a mechanism that can produce non-normal copula structure and
asymmetric tail dependence 
<!--****DAN: Don't firget to replace @tail dependence@ by @tail asociation@ as appropriate-->
across space in ecological variables.
<!--***DAN: move the below to end of this part-->
This result is represented in Fig. \ref{MasterFigure} as the dashed box around 
"Moran effects" and the arrows labelled "B".
<!--***DAN: move the above to end of this part-->
For simulations using AR(1) models, copula 
statistics were always very similar for noise and populations 
(Fig \ref{fig_Cause4copula_scatter_C_kend_ar1}, 
\ref{SM-fig_Cause4copula_C_kend_ar1}-\ref{SM-fig_Cause4copula_SJ_kend_ar1}),
though there were significant but minor differences for some statistics
and simulations.
For weak noise and our nonlinear model, copula statistics
were again very similar for noise and model outputs 
(Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1}-\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1}).
Though there were again significant but small differences for some statistics
and simulations, the hypothesis was supported that copula 
structure in the noise translates
to similar but not identical copula structure in output populations. 
Since many ecological models are nonlinear, this result provides the 
expectation that there will be a Moran-effect-like correspondence 
between noise and model-output copula structure across space 
for typical ecological dynamics,
as long as noise is small enough that dynamics stay relatively close
to the model equilibrium. Results that hold for "weak noise" in this sense
are common in ecology. 
We repeated the analysis with $r=1.3$, with similar results. The deterministic 
one-habitat-patch Ricker model exhibits a monotonic approach to a stable equilibrium
at $K$ when $r<1$ (undercompensating dynamics, e.g., the value $r=0.5$ used initially), 
but exhibits an oscillatory approach to $K$ when $r>1$ (overcompensating
dynamics, e.g., $r=1.3$).

\begin{figure}[!h]
\begin{center}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
\caption{Example results showing similarity of copula statistics for 
environmental-noise inputs and ecological-variable outputs of a dynamical model.
Asymmetric tail dependence was similar in both noise inputs and model outputs
for this model. The 
AR(1) model and a Clayton copula were used (see methods section for Q2).
Each point is the mean across 50 replicate simulations for which the same 
Kendall $\tau$ value was used. Error bars are standard errors 
and panel headers give Pearson correlation results for
the points. The regression line through the points (black line) was similar
to the 1-1 line (green line). 
See Fig. \ref{SM-fig_Cause4copula_C_kend_ar1} for additional results for the AR(1) model and Clayton copula, 
Figs \ref{SM-fig_Cause4copula_SC_kend_ar1}-\ref{SM-fig_Cause4copula_SJ_kend_ar1} for AR(1) results with other copula families, and 
Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1}-\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1} for
the nonlinear models described in the methods section for Q2.\label{fig_Cause4copula_scatter_C_kend_ar1}}
\end{center}
\end{figure}

For strong noise and using our nonlinear model, 
copula statistics, generally speaking, were approximately similar 
between noise and model
outputs; however, similarity was reduced relative to previous results, 
and for a few simulations, tail dependence asymmetry statistics
had opposite signs for noise and model outputs 
(Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1}-\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1}). For instance, using a Clayton  copula with a large Kendall $\tau$,
$\cor_l - \cor_u$ was slightly positive for noise, but slightly negative for 
population model outputs (Fig. \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_1}); 
and likewise for survival Gumbel and survival Joe copulas (Figs \ref{SM-fig_Cause4copula_SG_kend_sRicker_nsd_1}, 
\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1}). Overall,
similarities between noise and model-output copula statistics
were more noticable than discrepancies, in spite of nonlinearities.
We repeated the analysis with $r=1.3$. Discrepancies between noise and population
copula structure in this case were often glaring. Apparently 
noise of standard deviation 1 interacted especially strongly with 
model nonlinearities when the model was in its overcompensatory 
regime.

For our evolutionary model, the hypothesis was 
correct that asymmetric tail dependence in evolutionary changes 
can produce similarly asymmetric tail dependence between characters across
phylogeny tips (Appendix \ref{SM-evol_mech}, Fig. \ref{SM-fig_cop_evo} 
for details). 
Thus this is a mechanism that may produce non-normal copula structure and
asymmetric tail dependence between characters.
This result is represented in Fig. \ref{MasterFigure} as the dashed box around 
"If character evolution has copula structure" and the arrow labelled "D".
This topic is revisited in the Discussion, as is the box "Asymmetric species
interactions" and the arrow labelled "C" in Fig. \ref{MasterFigure}.

<!--Question 3 analyses-->
# Concepts and methods for Q3 \label{Consequences}

The hypothesis was presented in the Introduction that
the distribution (through time) of a spatially 
averaged quantity should be influenced by 
dependencies between the local quantities being averaged, including their
copula structure and tail dependence. 
To refine this hypothesis, suppose an ecological variable 
$E_i(t)$ is measured at locations $i=1,\ldots,N$ and times $t=1,\ldots,T$, and
the spatial mean $\sum_i E_i(t)/N$ is of
interest. The $E_i(t)$ could be, for instance, local abundances of a pest or 
exploited species, or local fluxes of a greenhouse gas. If  $E_i$ and $E_j$ are 
right-tail dependent for most location pairs $i$ and $j$, then
exceptionally large values tend to occur at the same time in most locations.
We hypothesize that this 
will tend to increase the skewness of the distribution 
of the spatial mean. Similarly, left-tail dependence between local variables
should tend to decrease skewness. Strong positive skewness of the 
spatial-mean time series corresponds to "spikiness" of that time series,
which corresponds to instability through time. 
The spatial-mean time series and its skewness may be quantities of 
principal importance for 
pest or resource abundance, for which extreme values (spikes) 
in the spatial mean have large effects.

To help answer Q3, we tested the above hypothesis
using our multivariate datasets (Table \ref{tab_data_info}).
For each dataset we calculated the spatial mean time series, and then 
the skewness through time of that
mean. Then, for each dataset, we compared the value obtained 
to a distribution of values 
of the same quantity for each of 10000 surrogate datasets. Surrogate 
datasets were produced by randomizing the empirical data
in a special way to have the 
copula structure of a multivariate-normal distribution, but 
to retain exactly the same
distributions of values for each sampling location as the original data
(Appendix \ref{SM-Surrog_multi}). 
Surrogates also had very similar Spearman correlations 
between pairs of sampling locations
as the data. 
Our comparisons therefore tested the null 
hypothesis that the skewness of the spatial 
mean took values on the empirical data no different 
from what one would expect if the copula
structure of the data were multivariate N, but the data were otherwise statistically
unchanged. Significant differences indicate that non-normal copula structure in the data 
contributed to the skewness of the spatial mean time series, 
i.e., to its instability and "spikiness" through time.

For green spruce aphid abundance data, *C. furca* abundance data, and methane-flux
data, because these datasets exhibited lower-tail dependence 
(Table \ref{tab_multivar_summary}), we compared empirical and surrogate skewness values
via a one-tailed test in the left tail: the $p$-value was the fraction
of surrogate skewnesses less than the skewness for the empirical data; the test
examines whether lower-tail dependence caused the spatial average to have significantly lower
skewness than would have been expected without tail dependence. For
leaf-curling plum aphid data, because that dataset exhibited upper-tail 
dependence (Table \ref{tab_multivar_summary}), we did the analogous one-tailed test
in the right tail. The test examines whether upper-tail dependence caused the spatial
average to have significantly higher skewness than 
expected without tail dependence.

To further address Q3, we also examined a hypothesis that spatial 
tail dependence of an environmental variable can influence the 
extinction risk of a metapopulation. We hypothesized that environmental noises
exhibiting greater left-tail dependence across different habitat patches would 
cause higher metapopulation extinction risks because then very bad years for the 
component populations would occur simultaneously in many patches,
reducing rescue effects. Here we assume, for simplicity, 
that low values of the environmental
variable are "bad" for the populations and high values are "good".
We tested the reasonableness of this hypothesis using a metapopulation
extension of the Lewontin-Cohen model,
$\vec{P}(t+1) = D \lambda(t) \vec{P}(t)$,
where the $i^{th}$ component of the length-$N$ vector 
$\vec{P}(t)$ represents
population density in the $i^{th}$ habitat patch at time $t$. 
The $N \times N$ matrix $\lambda(t)$ was diagonal with $i^{th}$ diagonal entry 
$\exp(r+\epsilon_i(t))$. Here $r$ is a growth rate; we used $r=0$. The
$\epsilon_i(t)$ represent environmental noises. They were standard-normally 
distributed, were independent through time, and showed the same 
spatial correlations for every simulation, 
but were made to exhibit right- or left-tail dependence between patches in
different simulations (Appendix \ref{SM-ext_risk_model}). The $N \times N$
matrix $D$ was a dispersal matrix
modelling local or global dispersal at rate $d$, in different simulations
(Appendix \ref{SM-ext_risk_model}).
After each step, 
if the density in a patch was $<1$, it was
set to $0$. We simulated the model 10000 times for each combination of parameters,
starting from $p_0 = 50$ in each patch, 
and calculated extinction risk after $25$ time steps.

# Results for Q3: Tail dependence influences skewness of the spatial average, extinction risk, and Taylor's law

\noindent Results were consistent with the hypothesis that the skewness
of a spatial-average time series is influenced by tail dependence
between the local quantities being averaged. Thus copula 
structure and tail dependence are important for spatially averaged 
quantities. This result is represented in Fig. \ref{MasterFigure}
as the solid box around "Instability/skewness of mean or total 
time series" and the solid arrows labeled "X". For datasets that
exhibited strong lower-tail (respectively, upper-tail) 
dependence in earlier analyses (green spruce 
aphid abundance data for lower-tail, and leaf-curling plum aphid first flight data
for upper-tail, Table 
\ref{tab_multivar_summary}), skewness of the spatial average 
was less than (respectively, greater than) a significant fraction of 
surrogate skewness values 
(Fig. \ref{fig_skewness_spearman}A,B). For datasets with moderate lower-tail 
dependence (\emph{Ceratium furca} abundance and methane data), 
skewness of the spatial
average showed a non-significant or marginally significant tendency 
toward being less than surrogate skewnesses (Fig. \ref{fig_skewness_spearman}C,D).

\begin{figure}[!h]
\textbf{ \hspace{6.6 cm} (A) \hspace{4.6 cm} (B)} \\
\vspace{-0.4 cm}
\begin{center}
\includegraphics[width=5 cm]{./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10_spearman.pdf}
\includegraphics[width=5 cm]{./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11_spearman.pdf}\\
\end{center}
\vspace{0.2 cm}
\textbf{ \hspace{6.6 cm} (C) \hspace{4.6 cm} (D)} \\
\vspace{-0.4 cm}
\begin{center}
\includegraphics[width=5 cm]{./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16_spearman.pdf}
\includegraphics[width=5 cm]{./Results/skewness_results/skewness_methane/skewness_methane_spearman.pdf}
\caption{Skewness of spatially averaged green spruce aphid abundance $(A)$, leaf-curling plum 
aphid first flight dates $(B)$, \emph{Ceratium furca} abundance $(C)$, and methane-flux
$(D)$ time series compared to a multivariate 
normal-copula null hypothesis. Black dots are empirical 
skewnesses; see text 
and Appendix \ref{SM-Surrog_multi} for details of the null hypothesis. Results show a 
tendency for skewness of the spatial average to be affected as hypothesized 
by tail dependence. \label{fig_skewness_spearman}}
\end{center}
\end{figure}

Consistent with our extinction risk hypothesis, 
left-tail-dependent environmental fluctuations increased metapopulation
extinction risk for the spatial 
Lewontin-Cohen model (Fig. \ref{fig_extrisk_lcohen_local}), 
for $N=5$ and $N=25$ habitat patches, and for local 
(Fig. \ref{fig_extrisk_lcohen_local}) and global (Fig. \ref{SM-fig_extrisk_lcohen_global})
dispersal. This result is represented in Fig. \ref{MasterFigure}
as the dashed box around "Extinction risk" and the dashed arrow labelled "Y".
We also substantiated using models (Appendix \ref{SM-Consequence_TL_simulation}, \ref{SM-sect:copsurrognd}, Fig. \ref{SM-fig_spTL})
a hypothesis (Fig. \ref{MasterFigure}, arrow labelled "Z") that 
Taylor's law can be influenced by
tail dependence.

\begin{figure}[!h]
\textbf{ \hspace{6.6 cm} (A) \hspace{4.2 cm} (B)} \\
\vspace{-0.8 cm}
\begin{center}
\includegraphics[width=10 cm]{./Results/ext_risk_copula/lcohen_ext_risk_local_disp.pdf}
\caption{Extinction risk for the metapopulation extension of the Lewontin-Cohen model, after 
$25$ time steps, was higher for left-tail-dependent environmental noise. Dispersal was local,
with dispersal rate $d$ (Appendix \ref{SM-ext_risk_model}). Simulations used $N$ patches
for $N=5$ (A) and $N=25$ (B).\label{fig_extrisk_lcohen_local}}
\end{center}
\end{figure}

# Discussion \label{Discussion}

<!--what we did and why is it important to study?-->
We showed that non-normal copula structure and asymmetric tail dependence are 
common across multiple sub-disciplines
in ecology, although no research we know of has 
taken this structure into account. We hypothesized 
mechanisms that may cause non-normal copula structures and tail dependence;
we discuss below how commonly some of our
mechanisms may opperate. 
We also hypothesized and demonstrated important consequences of 
non-normal copula structure and tail-dependence for the field of ecology. 
For instance, the skewness
of a spatial-average time series is influenced by tail dependencies between its 
constituent time series: right-tail-dependent local time series 
lead to "spiky" spatially averaged time series, with large outbreaks.
Thus tail dependence could have implications for exploited and pest species.
Extinction risk can also be altered by tail dependence across space
in environmental variables. 
In our view, our results make it reasonable to suggest 
that comprehensive understanding 
of many ecological phenomena may be impossible without 
a complete copula characterization of dependence.

Copula approaches are
well developed [@joe2014_dependence; @nelsen2006_copula], and 
computer implementations exist (`copula` and 
`VineCopula` packages). Ecologists can apply these tools
immediately. We created several interrelated randomization procedures
(Appendices \ref{SM-surrog_test}, \ref{SM-Surrog_multi}, \ref{SM-sect:copsurrognd}) 
that build upon existing copula methods. 
To our knowledge, only normal copulas have 
been used in ecology [@Valpine2014; @Anderson2018; @Popovic2018].
Such use is equivalent to using normal distributions with transformed
marginals; i.e., it is equivalent to an approach not using copulas at all, and
therefore does not capitalize on the power of copula techniques.

<!--a para on how likely one is to get copula structure via the Liebig mechanism-->
Our first causal mechanism (Fig. \ref{MasterFigure}, A) probably operates commonly,
for two reasons.
First, Liebig's law and the idea of limiting 
nutrients are dominant paradigms in ecology, and 
many studies have documented nonlinear or 
threshold influences of environmental
variables on ecological quantities. 
Second, fluctuations in environmental variables through time
are very commonly correlated across space. 
Because these factors, which are the essential ingredients of
the mechanism, are so common, it is reasonable to suppose the mechanism
opperates commonly. It may be a dominant cause of 
tail depencence and non-normal copula struture of ecological dependencies 
across space.
We provide some empirical support for the mechanism 
in our discussion of green sprice aphids and
winter temperature below.

<!--para on how likely one is to get copula structure via the Moran mechanism-->
There are also reasons to expect that 
our Moran mechanism (Fig. \ref{MasterFigure}, B) 
operates commonly: Moran effects are common [@Liebhold2004], and 
non-normal and tail-dependent copula structures are often found in 
environmental variables. If intense meteorological events are 
also widespread, then
environmental variables associated with these events should take extreme
values simultaneously across large spatial areas, producing tail dependence in measurements
made through time at different locations. Non-extreme values may instead
be associated with local phenomena, and therefore may be
less correlated across large areas.
@Serinaldi2008 examined the spatial dependence 
of rainfall in Central Italy; Gumbel or Student 2-copulas 
were candidates for modelling dependence, and 
neither of these is a normal copula.
A long-term study (1950–2014) in the Loess Plateau of China [@She2018]
showed that a Gumbel copula effectively modeled 
the spatial dependence of drought variables.
Bivariate copula analysis was also used in forecasting the 
co-occurrence of extreme events (flood or drought) 
over the North Sikkim Himalayas using spatial datasets [@Goswami2018].

<!--Causal mechanism C, which we also did not test-->
We suggested in the Introduction that asymmetric competitive relationships
between species could yield tail dependence between abundance data for the 
species. This is another theoretical mechanism for non-normal copula
structure and tail dependence, represented in Fig. \ref{MasterFigure} 
by the box "Asymmetric species interactions" and the arrow labelled "C".
It could be tested by analyzing copulas of abundances of competing 
species, sampled across space or time.

<!--Holder stuff-->
Our simulations of character evolution suggest the hypothesis that
changes through evolutionary time in bird and 
mammal BMR and body size may exhibit right-tail dependence,
contrary to standard normality assumptions of character evolution models. This is
a hypothesis only, because the right-tail dependence of Fig. \ref{fig_biv_multi_raw_cop}F,G
could have come about in another, unknown way. Our simulations show that tail dependence 
in evolutionary changes is sufficient, but may not be necessary, to produce tail
dependence in characters of extant species. For instance (see below, 
Appendix \ref{SM-sect:missingdata}, Fig. \ref{SM-fig:missingdata}), 
systematically missing data can also produce tail dependence and may
have influenced results for the BMR-body mass datasets. Even if the hypothesized 
evolutionary mechanism (Fig. \ref{MasterFigure}, D)
is correct, our results only replace one question, i.e., why do we see right-tail
dependence between BMR and body mass, with another,
i.e., why might we see right-tail dependence in evolutionary changes in these traits?
See Appendix \ref{SM-evol_mech} for further thoughts on potential importance of 
copulas for character evolution.

<!--Discuss stuff here on tradeoff and life history that Bever told me about-->
Relationships between BMR and body mass relate
to a trade-off between mass-specific BMR (BMR per unit body mass) and body
mass itself. Copulas probably interrelate with 
life-history trade-off theory in additional ways beyond what we demonstrated.
For instance, it is well known that energy allocation to a life function, F (e.g., reproduction)
will reduce the energy that can be allocated to other functions, $G_1$, $G_2$, $G_3$ 
(e.g., growth, predation avoidance). This is the principle of 
allocation. But F can trade off against any or all of 
the $G_i$. Therefore, for large F, approaching absolute 
limitations, there may be a strong 
association between F and $G_1$. For small F, there may be
little association because resources not allocated to F can instead be allocated 
to any combination of the $G_i$. This constitutes 
asymmetric tail dependence between F and $G_1$. @Winemiller1992
described a three-way trade-off in fishes between age of reproductive
maturity, juvenile survivorship, and fecundity. The trade-off should, in theory,
produce a tight association between age of maturity and fecundity for 
fishes with low age of maturity, but little such association for later-maturing
fishes because those fishes may invest the resources not invested in maturing quickly into
either fecundity or juvenile survival.
These ideas suggest that copulas will be 
useful for studying multi-dimensional trade-offs. But applications will 
require careful attention to the possible consequences of biased sampling:
if the degree of completeness of a dataset is associated with one or
more of the characters, then statistical artefacts
can bias conclusions (Appendix \ref{SM-sect:missingdata}, 
Fig. \ref{SM-fig:missingdata}).

<!-- We have only given a few mechanisms, but there are likely to be others, and discoverig
additional mechanisms of copula structure is certainly worthy of future work. -->
Other mechanisms probably also operate, 
and should be considered. For instance, 
measurement error may modify copula structure. 
Our models were intentionally simple in other respects, 
too, and did not include
delayed density dependence, dispersal, population stage structure, trophic
interactions, etc., and we did not comprehensively 
explore parameter space.
We hope by enumerating a few potential mechanisms of copula structure,
we inspire additional research on the
potentially numerous mechanisms that may operate in diverse datasets,
and their relative importance under different circumstances.

<!--some consequences paragraphs next-->

<!--green spruce aphid all the way from causes to consequences-->
Our hypotheses and results cover the presence, causes, and consequences
of non-normal copula structure and tail dependence in ecological 
datasets (Fig. \ref{MasterFigure}).
We here take a closer look at the green spruce aphid, which simultaneously 
illustrates causes and consequences.
Green spruce aphid abundance, as measured in the data we use,
is strongly influenced by temperature of the previous winter (@Sheppard2016,
their supplementary fig. 6).
For each of our 10 sampling locations, we therefore 
examined (Appendix \ref{SM-sect:greenspruce}, 
Fig. \ref{SM-fig:greenspruce}, Table \ref{SM-tab:greenspruce}) the 
copula of winter temperature
and aphid abundance time series for the location, finding 
left-tail dependence. Apparently winter temperature has an asymmetric
influence on aphid abundance in that cold winters produce low abundances
but warm winters often do not yield higher abundances than moderate winters.
One of our hypothesized mechanisms (Fig. \ref{MasterFigure}, A), 
which our modelling results
supported (section on results for Q2), therefore suggests that spatial 
dependence between green spruce aphid counts
in different locations should show left-tail dependence. This is 
exactly what was observed (Table \ref{tab_multivar_summary}), providing 
empirical evidence supporting the mechanism (this is why the box around "Nonlinear 
environmental effects, Liebig's law" and some of the arrows labelled "A" are solid
instead of dashed in Fig. \ref{MasterFigure}). The consequences of such tail 
dependence for the skewness of spatially averaged aphid counts was 
described previously 
(Fig. \ref{fig_skewness_spearman}A). Thus the asymmetric influence of winter 
temperature ultimately causes spatially averaged aphid abundance time series 
to have lower skewness (i.e., less spikiness, and greater stability through time) 
than they would otherwise.

<!--a para on stability of the total biomass time series in a community, and skewness - sets up Shya's later paper as future work-->
Our results showed that the skewness of the 
spatial average of local time series is influenced 
by their tail dependence. But the same logic should also apply to any collection of time
series, whether associated with locations in space or not. 
Another potential application is time series of all species from a single community, e.g., all
plants in a quadrat surveyed repeatedly over time. A huge 
literature has focused on synchrony versus 
compensatory dynamics between such time series, and 
the influence of interspecific relationships on the 
variability of community
properties such as total biomass. Typically, variability of community biomass 
is measured with the coefficient of variation, but skewness
may also be of interest because it can help characterize "spikiness" 
through time. Future work on copula 
structure of interspecific relationships in communities and its implications for community 
variability is likely to be valuable.

<!--A para that sets up Shya's extinction risk paper-->
Although we demonstrated that tail dependence between environmental variables
can influence extinction risk, substantial work remains to determine the importance 
of this effect. First, we used 
a non-density-dependent model. Do similar results pertain
when density dependence
acts? Second, we considered metapopulation extinction risk, but the huge field 
of population viability analysis (PVA) via stochastic 
matrix modelling [@Caswell2000; @MorrisDoak2002]
uses a framework in which a single population's 
vital rates (e.g., life-stage-specific 
fecundity and survival rates) are considered to
vary stochastically through time due to environmental variation. Do relationships
between different vital rates exhibit tail dependence, and does tail dependence 
influence extinction risk in this context?
Finally, is the copula structure of environmental variables
changing through time, and, if so, how do such changes influence extinction risks?
Climate change is known to amplify the factors that lead to extreme weather events
[@Hansen2012] and hence may alter spatial tail dependence for weather variables.

# Acknowledgments

We thank the many contributors to the large datasets we used; D. Stevens and P. Verrier for data 
extraction; and Joel E. Cohen, David Tilman, Lauren Hallet, Jonathan Walter, Thomas 
Anderson, and Lei Zhao for helpful suggestions. We thank James Bell of the Rothamsted Insect Survey 
(RIS). 
The RIS, a UK Capability, is funded by the Biotechnology and Biological Sciences Research Council 
under the Core Capability Grant BBS/E/C/000J0200. SG, LWS and DCR were partly funded by US National 
Science Foundation grants 1714195 and 1442595 and the James S McDonnell Foundation.

\clearpage
\newpage

# References

\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{1pt}
\noindent
\sloppy

<div id="refs"></div>